### 2016-05-24 README.Rmd
Add README
Couldn't do this with `use_readme_rmd` for some reason... I think also that I'll want to change the figure paths.

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r echo = FALSE}
library(hurricaneexposure)
```

## Loading the package

The package currently exists in [a development version](https://github.com/geanders/hurricaneexposure) on GitHub. You can use the following code to load it:

```{r eval = FALSE}
library(devtools)
install_github("geanders/hurricaneexposure", 
               build_vignettes = TRUE)
library(hurricaneexposure)
```

Once you've loaded the dataframe, you can load the included data using the `data` function. For example:

```{r}
data("hurr_tracks")
head(hurr_tracks)
```

The following datasets are included with the package:

- `closest_dist`
- `county_centers`
- `hurr_tracks`
- `rain`

For each, you can see the helpfiles for the data for more information about the data included in each.

## Creating time series datasets of exposure

There are several functions in this package that help to take the data on rainfall and storm tracks stored in the package and use them, in conjunction with thresholds for the distance to the storm tracks and the amount of rainfall required for exposure, to create time series of exposure that can be easily integrated into time series datasets of, for example, health data.

First, the `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds for distance (storm path must come that close or closer to the county) and rainfall (total rainfall summed over the days you specify-- for example, `days_included = c(-1, 0, 1)` would use the total rainfall summed over the day before, day of, and day after the date when the storm was closest to the county). The function outputs a dataframe with all of the storms to which each of the counties was exposed.

For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over three days was 100 millimeters or more, you could run:

```{r}
county_rain(counties = c("22071", "51700"),
            start_year = 1995, end_year = 2005,
            rain_limit = 100, dist_limit = 100,
            days_included = c(-1, 0, 1))
```

In addition to giving you the names and closest dates of each storm for each county (`closest_date`-- note, this is given using the UTC timezone), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).

To get a dataframe listing the relevant storms for multi-county communities, you can use the `multi_county_rain` function in a similar way:

```{r warning = FALSE}
communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
                         fips = c("36005", "36047", "36061",
                                  "36085", "36081", "36119",
                                  "22071", "51700"))
rain_storm_df <- multi_county_rain(communities = communities,
                                   start_year = 1995, end_year = 2005,
                                   rain_limit = 100, dist_limit = 100,
                                   days_included = c(-1, 0, 1))
```

This output includes columns for the average closest distance for any of the counties in the community (`mean_dist`), the average precipitation for all the counties (`mean_precip`), the highest precipitation for any of the counties (`max_rain`), and the smallest distance between the storm track and any of the county population-weighted centers (`min_dist`).

## Creating and writing time series of exposure

To create time series of hurricane exposure for all of the storms that meet the exposure definition for study communities, you can use the function `rain_exposure`. If you have counties for your study, you can run this with a vector of the county FIPS as the `locations` argument (specify the directory path to write the files with `out_dir`):

```{r}
rain_exposure(locations = c("22071", "51700"),
              start_year = 1995, end_year = 2005,
              rain_limit = 100, dist_limit = 100,
              out_dir = "~/tmp/storms")
```

If you have multi-county communities, set `locations` instead to be a dataframe with community names (`commun` column) and FIPS codes (`fips` column):

```{r message = FALSE, warning = FALSE}
communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
                          fips = c("36005", "36047", "36061",
                          "36085", "36081", "36119",
                          "22071", "51700"))
rain_exposure(locations = communities,
              start_year = 1995, end_year = 2005,
              rain_limit = 100, dist_limit = 100, out_dir = "~/tmp/storms")
```

Both functions will output one file of exposure per county or community into the directory that you specify using the `out_dir` argument.

## Mapping hurricane exposure

This package allows you to create some different maps of hurricane exposures based on distance to the storm track and rainfall.

### Plotting county-level exposure

The `map_counties` function creates county choropleths of different storm exposure variables (right now, rainfall and distance from the storm tracks). For example, to plot rain exposure for Hurricane Floyd in 1999 (the ID for this storm is "Floyd-2012"):

```{r fig.width = 7, fig.height = 4, message = FALSE}
map_1 <- map_counties(storm = "Floyd-1999", metric = "rainfall")
map_1
```

You can also use this function to plot the closest distance between the storm and each county. For this, you use the argument `metric = "distance"`.

```{r fig.width = 7, fig.height = 4}
map_2 <- map_counties(storm = "Sandy-2012", metric = "distance")
map_2
```

You can map a binary variable of distance-based exposure using `map_distance_exposure`:

```{r fig.width = 7, fig.height = 4}
allison_map <- map_distance_exposure(storm = "Allison-2001",
                                     dist_limit = 75)
plot(allison_map)
```

You can also map a binary variable of rain exposure for the communities that were exposed, based on a certain rainfall limit and distance limit:

```{r fig.width = 7, fig.height = 4}
map_3 <- map_rain_exposure(storm = "Floyd-1999", rain_limit = 125,
                           dist_limit = 500, 
                           days_included = c(-1, 0, 1))
plot(map_3)
```

### Plotting storm tracks

The `map_tracks` function will map the hurricane tracks for one or more storms. For example, to plot the tracks of Hurricane Floyd in 1999 (the ID for this storm is "Floyd-2012"):

```{r fig.width = 6.5, fig.height = 4}
map_4 <- map_tracks(storms = "Floyd-1999")
map_4
```

There are some different options you can use for the tracks' appearance. For example, if you wanted to plot the tracks of several storms, not plot each point when the track locations were measured (typically every six hours), and use some transparency so you can see all the lines, you can use:

```{r fig.width = 5.5, fig.height = 4}
map_5 <- map_tracks(storms = c("Floyd-1999", "Sandy-2012",
                               "Katrina-2005"),
                    plot_points = FALSE,
                    alpha = 0.5)
map_5
```

You can also add these tracks to an existing `ggplot`-created US map. You do this through the `plot_object` argument. For example, to add the storm track to the plot of distance exposure for Sandy or rain exposure for Floyd, you could run:

```{r fig.width = 7, fig.height = 4}
map_6 <- map_tracks(storms = "Sandy-2012", plot_object = map_2,
                    plot_points = FALSE)
map_6
```

```{r fig.width = 7, fig.height = 4}
map_7 <- map_tracks(storms = "Floyd-1999", plot_object = map_3,
                    plot_points = FALSE)
map_7
```

### 2016-05-26 R/data.R
Fix `check` warnings and notes

@@ -9,7 +9,7 @@
 #'   \item{state_fips}{state FIPS code}
 #'   \item{county_fips}{county 3-digit FIPS code}
 #'   \item{fips}{county 5-digit FIPS code}
-#'   \item{count_name}{county name}
+#'   \item{county_name}{county name}
 #'   \item{state_name}{state name}
 #'   \item{population}{population}
 #'   \item{latitude}{latitude of county's center of population}
@@ -34,12 +34,11 @@
 #'
 #' @format A data frame with 12,206 rows and 6 variables:
 #' \describe{
+#'   \item{storm_id}{Unique storm identifier}
 #'   \item{date}{Date and time of storm track recording}
-#'   \item{status}{Storm status}
 #'   \item{latitude}{Latitude of storm center}
 #'   \item{longitude}{Longitude of storm center}
-#'   \item{wind}{[what is this measuring?]}
-#'   \item{storm_id}{Unique storm identifier}
+#'   \item{wind}{Maximum storm wind speed}
 #' }
 #'
 #' @note The original dataset included more separate `status` categories
@@ -62,11 +61,13 @@
 #'
 #' @format A list with 991,530 rows and 3 variables:
 #' \describe{
-#'   \item{storm_id}{unique identifier for each tropical storm}
 #'   \item{fips}{county 5-digit FIPS code}
-#'   \item{tot_precip}{Total precipitation, in millimeters [?], for a
-#'       one-week window centered at the date the storm was closest to
-#'       the county}
+#'   \item{storm_id}{unique identifier for each tropical storm}
+#'   \item{lag}{Number of days from date when storm was closest to the county}
+#'   \item{precip}{Average daily precipitation, in millimeters, for NLDAS grid
+#'    points with the county for the given lag day}
+#'   \item{precip_max}{Maximum daily precipitation, in millimeters, for NLDAS grid
+#'    points with the county for the given lag day}
 #' }
 #'
 #' @references

### 2016-05-26 R/distance_exposure.R

@@ -23,12 +23,15 @@
 #' @export
 #'
 #' @importFrom dplyr %>%
-county_distance <- function(counties, start_year, end_year,
-                           dist_limit){
+county_distance <- function(counties, start_year, end_year, dist_limit){
 
-        distance_df <- dplyr::mutate(closest_dist,
-                                     closest_date = lubridate::ymd_hm(closest_date)) %>%
-                dplyr::filter(fips %in% counties &
+        dots <- stats::setNames(list(lazyeval::interp(~ lubridate::ymd_hm(x),
+                                                      x = quote(closest_date))),
+                                    "closest_date")
+
+        distance_df <- hurricaneexposure::closest_dist %>%
+                dplyr::mutate_(.dots = dots) %>%
+                dplyr::filter_(~ fips %in% counties &
                                     lubridate::year(closest_date) >= start_year &
                                     lubridate::year(closest_date) <= end_year &
                                     storm_dist <= dist_limit)
@@ -63,20 +66,23 @@ county_distance <- function(counties, start_year, end_year,
 multi_county_distance <- function(communities, start_year, end_year,
                          dist_limit){
 
-        communities <- dplyr::mutate(communities, fips = as.character(fips))
+        communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
 
-        distance_df <- dplyr::mutate(closest_dist,
-                                     closest_date = lubridate::ymd_hm(closest_date)) %>%
-                dplyr::filter(fips %in% communities$fips &
-                                      lubridate::year(closest_date) >= start_year &
-                                      lubridate::year(closest_date) <= end_year) %>%
+        distance_df <- hurricaneexposure::closest_dist %>%
+                dplyr::mutate_(closest_date =
+                                       ~ lubridate::ymd_hm(closest_date)) %>%
+                dplyr::filter_(~ fips %in% communities$fips &
+                                      lubridate::year(closest_date) >=
+                                       start_year &
+                                      lubridate::year(closest_date) <=
+                                       end_year) %>%
                 dplyr::left_join(communities, by = "fips") %>%
-                dplyr::group_by(commun, storm_id) %>%
-                dplyr::mutate(min_dist = min(storm_dist)) %>%
-                dplyr::filter(min_dist <= dist_limit) %>%
-                dplyr::summarize(closest_date = first(closest_date),
-                                 mean_dist = mean(storm_dist),
-                                 min_dist = first(min_dist))
[…]

### 2016-05-26 R/map_exposure.R

@@ -12,22 +12,23 @@
 #' @examples
 #' default_map()
 default_map <- function(){
+
+        eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
+                            "district of columbia", "florida", "georgia", "illinois",
+                            "indiana", "iowa", "kansas", "kentucky", "louisiana",
+                            "maine", "maryland", "massachusetts", "michigan",
+                            "mississippi", "missouri", "new hampshire", "new jersey",
+                            "new york", "north carolina", "ohio", "oklahoma",
+                            "pennsylvania", "rhode island", "south carolina",
+                            "tennessee", "texas", "vermont", "virginia",
+                            "west virginia", "wisconsin")
+
         map_data <- ggplot2::map_data("state")
-        map_data <- dplyr::filter(map_data,
-                                  region %in% c("alabama", "arkansas",
-                                                "connecticut", "delaware",
-                                                "district of columbia", "florida",
-                                                "georgia", "illinois", "indiana",
-                                                "iowa", "kansas", "kentucky", "louisiana",
-                                                "maine", "maryland", "massachusetts",
-                                                "michigan", "mississippi",
-                                                "missouri", "new hampshire", "new jersey",
-                                                "new york", "north carolina", "ohio",
-                                                "oklahoma", "pennsylvania", "rhode island",
-                                                "south carolina", "tennessee", "texas",
-                                                "vermont", "virginia", "west virginia",
-                                                "wisconsin"))
-        out <- ggplot2::ggplot(map_data, ggplot2::aes(x = long, y = lat, group = group)) +
+        map_data <- map_data %>%
+                dplyr::filter_(~ region %in% eastern_states)
+        out <- ggplot2::ggplot(map_data,
+                               ggplot2::aes_(x = ~ long, y = ~ lat,
+                                             group = ~ group)) +
                 ggplot2::geom_polygon(fill = "lightgray", color = "white") +
                 ggmap::theme_nothing(legend = TRUE)
         return(out)
@@ -52,6 +53,7 @@ default_map <- function(){
 #'    as well as lines, when plotting the hurricane tracks.
 #' @param alpha Numerical value designating the amount of transparency to
[…]

### 2016-05-26 R/rain_exposure.R

@@ -31,32 +31,38 @@
 #'    period centered at the date of closest approach to the county.
 #'
 #' @examples
+#' \dontrun{
 #' county_rain(counties = c("22071", "51700"),
 #'             start_year = 1995, end_year = 2005,
 #'             rain_limit = 100, dist_limit = 100)
-#'
-#' @export
+#'}
 #'
 #' @importFrom dplyr %>%
+#'
+#' @export
 county_rain <- function(counties, start_year, end_year,
                            rain_limit, dist_limit,
                         days_included = c(-1, 0, 1)){
 
-        rain_storm_df <- dplyr::mutate(closest_dist,
-                                       closest_date = lubridate::ymd_hm(closest_date)) %>%
-                dplyr::filter(fips %in% counties &
+        dots <- stats::setNames(list(lazyeval::interp(~ lubridate::ymd_hm(x),
+                                                      x = quote(closest_date))),
+                                "closest_date")
+
+        rain_storm_df <- hurricaneexposure::closest_dist %>%
+                dplyr::mutate_(.dots = dots) %>%
+                dplyr::filter_(~ fips %in% counties &
                                     lubridate::year(closest_date) >= start_year &
                                     lubridate::year(closest_date) <= end_year &
                                     storm_dist <= dist_limit) %>%
-                dplyr::left_join(rain,
+                dplyr::left_join(hurricaneexposure::rain,
                                  by = c("storm_id", "fips")) %>%
-                dplyr::filter(lag %in% days_included) %>%
-                dplyr::group_by(storm_id, fips) %>%
-                dplyr::summarize(closest_date = first(closest_date),
-                                 storm_dist = first(storm_dist),
-                                 tot_precip = sum(precip)) %>%
+                dplyr::filter_(~ lag %in% days_included) %>%
+                dplyr::group_by_(~ storm_id, ~ fips) %>%
+                dplyr::summarize_(closest_date = ~ dplyr::first(closest_date),
+                                 storm_dist = ~ dplyr::first(storm_dist),
+                                 tot_precip = ~ sum(precip)) %>%
                 dplyr::ungroup() %>%
-                dplyr::filter(tot_precip >= rain_limit)
+                dplyr::filter_(~ tot_precip >= rain_limit)
         return(rain_storm_df)
 }
 
@@ -93,33 +99,36 @@ multi_county_rain <- function(communities, start_year, end_year,
                          rain_limit, dist_limit,
                          days_included = c(-1, 0, 1)){
 
[…]

### 2016-06-10 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add draft for paper

---
title: Assessing exposure to Atlantic Basin tropical storms in United States counties
author:
  - name: Joshua Ferreri
    email: alice@example.com
    affiliation: Colorado State University
  - name: Meilin Yan
    email: alice@example.com
    affiliation: Colorado State University
  - name: Mohammad Z. Al-Hamdan
    email: alice@example.com
    affiliation: NASA Marshall Space Flight Center
  - name: William L. Crosson
    email: alice@example.com
    affiliation: NASA Marshall Space Flight Center
  - name: Seth Guikema
    email: alice@example.com
    affiliation: University of Michigan
  - name: Roger D. Peng
    email: alice@example.com
    affiliation: Johns Hopkins Bloomberg School of Public Health
  - name: G. Brooke Anderson
    email: brooke.anderson@colostate.edu
    affiliation: Colorado State University
    footnote: Corresponding Author
address:
  - code: Colorado State University
    address: Department of Environmental and Radiological Health Sciences, Lake Street, Fort Collins, CO, Zip
  - code: NASA Marshall Space Flight Center
    address: Universities Space Research Association, Street, Huntsville, AL, Zip
  - code: University of Michigan
    address: Department, Street, City, State, Zip
  - code: Johns Hopkins Bloomberg School of Public Health
    address: Department of Biostatistics, Street, Baltimore, MD, Zip
abstract: |
  There are many important applications for having county-level estimates of exposure to tropical storms over many years. For example, ... . Current approaches include ... but have the following limitations ... .

  Here, we present open source software we have developed to explore county-level exposure to tropical storms in United States counties between 1988 and 2011. Further, we explore the differences in exposure classification when using different metrics (e.g., wind speed, rainfall, distance).

bibliography: mybibfile.bib
output: rticles::elsevier_article
---

```{r}
library(hurricaneexposure)
```

Introduction
==========================

#### What it means to assess county-level tropical storm exposure

#### Why it's important to measure TS exposure for counties

#### Examples of using TS exposure estimates for multiple storms

#### Examples of other datasets at the county level

If exposure to tropical storms over multiple storms and years can be assessed, these exposure datasets can be joined with other time series to explore the impacts of tropical storms. For example, daily counts of human health outcomes in environmental epidemiology studies are often available aggregated at the county level, and such data has often been paired with time series of environmental exposures (e.g., air pollution, temperature) to determine associated risks.

Data and Methods
==================

#### Distance-based exposure

We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database. ... This data gives time stamps for each observation in UTC.

These data typically give measurements of storm center location at 6-hour intervals. We interpolated these location values to every 15 minutes during the period when the storm was active, using a linear interpolation between each measured point.

We calculated the distance between each county's center and each of the 15-minute-interval estimates of a storm's location. To do this, we used the Great Circle method to calculate the distance between pairs of latitude and longitude coordinates, using the R package `sp`. For each county, we used the population mean center, based on the U.S. Census's 2010 Decennial Census, as the center location of the county. From these distance calculations, we identified the date-time and distance between the storm track and the county center at the 15-minute interval when the storm was closest to the county center. We used the `countytimezones` R package to convert each date-time of the closest storm approach for a county to the county's local time zone and identified the local date when the storm was closest to the county. This "closest date" was used to pair the distance estimates with rainfall estimates.

#### Rain-based exposure

#### Wind-based exposure

Results
============

```{r echo = FALSE}
library(hurricaneexposure)
map_rain_exposure("Floyd-1999", rain_limit = 75, dist_limit = 500,
                      days_included = -1:1)
```

Discussion
===================

#### Distance-based exposure

There are some sources of uncertainty for storm locations from the best track hurricane data. These include ...

However, these best tracks should be fairly reliable for more recent years of storms, as we use here. Many of the uncertainties related to storm positions in best track data are more of a concern for the years before ... (e.g., pre-19[xx]).

While there are often several different "best tracks" datasets, from different sources [?-- weather services?] for other ocean basins, the "best tracks" data from ..., which is the basis of the tracks we use here, are the undisputed primary source of tropical storm track data for the Atlantic basin.

There are a number of limitations to using distance to assess exposure to tropical storms. Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could mis-classify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).

#### Rain-based exposure

It can be very difficult to reliabily measure rain during extreme rain events, including tropical storms. For example, a heavy rain can wash away [?] rain monitors [?]. It can also be very hard to measure rain during heavy wind, as the rain does not fall straight into the monitor [?].

Some of the other possible sources for estimating rain during tropical storms include ...

The estimated rainfall amounts from our data are likely underestimates. This data source, however, should be internally consistent and so useful for comparing across different storms when all exposure estimates are based on this rain data.

Rainfall estimates are likely underestimates for a few reasons. First, they are based on averaging hourly measurements to a daily mean estimate. This averaging would smooth over shorter periods of very extreme rainfall. Further, this data is averaged over multiple grid points within each county and so would not fully reflect very extreme local precipitation (although this might be less of a concern for classifying exposure to a large-scale storm system, like a tropical storm, compared to more fine-scale storms). Finally, this NLDAS data provides a re-analysis that incorporates measured rainfall, using models, etc., to incorporate that observed data into a spatially and temporally continuous dataset of rainfall. However, during extreme storms, the problems with measuring rainfall using [rain monitors] would propogate into the NLDAS data, so although NLDAS would prevent missing values during the storm if monitors are not able to provide data, if monitors are out, rainfall estimates from NLDAS will be based more on models than on observations.

Exposure classification based on rainfall has some advantages. For example, it allows the identification of exposed counties that are inland, rather than coastal, but that were affected by heavy rainfall during the storm. Often, storm-related deaths are associated with inland flooding. Storms can produce a lot of rain especially in certain topographies, like near mountains, so counties that are well inland can sometimes experience more extreme rain that other counties at similar distances from the storm's track.

In comparison to the storm track, the areas of extreme rain and extreme wind can vary. For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015].

Some storms can be of lower intensity (i.e., on the Saffir-Simpson scale) yet bring dangerous rainfall, including well inland of the storm's landfall. For example, storms including Floyd in 1999, Gaston in 2004, Irene in 2011, and Lee in 2011, have had severe inland impacts, often through extreme rainfall, as post-tropical storms [Halverson2015]. This rainfall can be particularly severe in the Appalachians [Halverson2015].

For example,

```{r echo = FALSE}
map_rain_exposure("Floyd-1999", rain_limit = 75, dist_limit = 500,
                      days_included = -1:1)
```

#### Wind-based exposure

Wind suffers from similar challenges for measuring during tropical storms. In particular, the strong winds of tropical storms can break or blow away the anemometers used to measure wind speed.

Here, we used wind speed models, rather than observed wind speed, to estimate exposure to tropical storms based on winds.

A variety of other wind speed models exist besides the one used here. In particular, there are options for wind speed models as far as ...

#### Example uses of exposure datasets

References {#references .unnumbered}
==========

### 2016-06-10 writing/DraftExposurePaper/DraftExposurePaper.tex
Add draft for paper

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
%\makeatletter
%\def\ps@pprintTitle{%
% \let\@oddhead\@empty
% \let\@evenhead\@empty
% \def\@oddfoot{\it \hfill\today}%
% \let\@evenfoot\@oddfoot}
%\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Assessing exposure to Atlantic Basin tropical storms in United States counties},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\usepackage[nomarkers]{endfloat}

\begin{document}
\begin{frontmatter}

  \title{Assessing exposure to Atlantic Basin tropical storms in United States
counties}
    \author[Colorado State University]{Joshua Ferreri}
   \ead{alice@example.com}

    \author[Colorado State University]{Meilin Yan}
   \ead{alice@example.com}

    \author[NASA Marshall Space Flight Center]{Mohammad Z. Al-Hamdan}
   \ead{alice@example.com}

    \author[NASA Marshall Space Flight Center]{William L. Crosson}
   \ead{alice@example.com}

    \author[University of Michigan]{Seth Guikema}
   \ead{alice@example.com}

    \author[Johns Hopkins Bloomberg School of Public Health]{Roger D. Peng}
   \ead{alice@example.com}

    \author[Colorado State University]{G. Brooke Anderson\corref{c1}}
   \ead{brooke.anderson@colostate.edu} 
   \cortext[c1]{Corresponding Author}
      \address[Colorado State University]{Department of Environmental and Radiological Health Sciences, Lake
Street, Fort Collins, CO, Zip}
    \address[NASA Marshall Space Flight Center]{Universities Space Research Association, Street, Huntsville, AL, Zip}
    \address[University of Michigan]{Department, Street, City, State, Zip}
    \address[Johns Hopkins Bloomberg School of Public Health]{Department of Biostatistics, Street, Baltimore, MD, Zip}

  \begin{abstract}
  There are many important applications for having county-level estimates
  of exposure to tropical storms over many years. For example, \ldots{} .
  Current approaches include \ldots{} but have the following limitations
  \ldots{} .

  Here, we present open source software we have developed to explore
  county-level exposure to tropical storms in United States counties
  between 1988 and 2011. Further, we explore the differences in exposure
  classification when using different metrics (e.g., wind speed, rainfall,
  distance).
  \end{abstract}

 \end{frontmatter}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(hurricaneexposure)}
\end{Highlighting}
\end{Shaded}

\section{Introduction}\label{introduction}

\paragraph{What it means to assess county-level tropical storm
exposure}\label{what-it-means-to-assess-county-level-tropical-storm-exposure}

\paragraph{Why it's important to measure TS exposure for
counties}\label{why-its-important-to-measure-ts-exposure-for-counties}

\paragraph{Examples of using TS exposure estimates for multiple
storms}\label{examples-of-using-ts-exposure-estimates-for-multiple-storms}

\paragraph{Examples of other datasets at the county
level}\label{examples-of-other-datasets-at-the-county-level}

If exposure to tropical storms over multiple storms and years can be
assessed, these exposure datasets can be joined with other time series
to explore the impacts of tropical storms. For example, daily counts of
human health outcomes in environmental epidemiology studies are often
available aggregated at the county level, and such data has often been
paired with time series of environmental exposures (e.g., air pollution,
temperature) to determine associated risks.

\section{Data and Methods}\label{data-and-methods}

\paragraph{Distance-based exposure}\label{distance-based-exposure}

We collected ``best tracks'' data on hurricane tracks for Atlantic basin
storms between 1988 and 2014 from the extended best tracks database.
\ldots{} This data gives time stamps for each observation in UTC.

These data typically give measurements of storm center location at
6-hour intervals. We interpolated these location values to every 15
minutes during the period when the storm was active, using a linear
interpolation between each measured point.

We calculated the distance between each county's center and each of the
15-minute-interval estimates of a storm's location. To do this, we used
the Great Circle method to calculate the distance between pairs of
latitude and longitude coordinates, using the R package \texttt{sp}. For
each county, we used the population mean center, based on the U.S.
Census's 2010 Decennial Census, as the center location of the county.
From these distance calculations, we identified the date-time and
distance between the storm track and the county center at the 15-minute
interval when the storm was closest to the county center. We used the
\texttt{countytimezones} R package to convert each date-time of the
closest storm approach for a county to the county's local time zone and
identified the local date when the storm was closest to the county. This
``closest date'' was used to pair the distance estimates with rainfall
estimates.

\paragraph{Rain-based exposure}\label{rain-based-exposure}

\paragraph{Wind-based exposure}\label{wind-based-exposure}

\section{Results}\label{results}

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-2-1.pdf}

\section{Discussion}\label{discussion}

\paragraph{Distance-based exposure}\label{distance-based-exposure-1}

There are some sources of uncertainty for storm locations from the best
track hurricane data. These include \ldots{}

However, these best tracks should be fairly reliable for more recent
years of storms, as we use here. Many of the uncertainties related to
storm positions in best track data are more of a concern for the years
before \ldots{} (e.g., pre-19{[}xx{]}).

While there are often several different ``best tracks'' datasets, from
different sources {[}?-- weather services?{]} for other ocean basins,
the ``best tracks'' data from \ldots{}, which is the basis of the tracks
we use here, are the undisputed primary source of tropical storm track
data for the Atlantic basin.

There are a number of limitations to using distance to assess exposure
to tropical storms. Tropical storms vary in size and intensity, and a
measure of distance from the storm track will not incorporate these
differences and so could mis-classify exposure both in terms of
generating false positives (counties close to the storm track of very
mild or small-radius storms) and false negatives (e.g., during very
large or intense storms).

\paragraph{Rain-based exposure}\label{rain-based-exposure-1}

It can be very difficult to reliabily measure rain during extreme rain
events, including tropical storms. For example, a heavy rain can wash
away {[}?{]} rain monitors {[}?{]}. It can also be very hard to measure
rain during heavy wind, as the rain does not fall straight into the
monitor {[}?{]}.

Some of the other possible sources for estimating rain during tropical
storms include \ldots{}

The estimated rainfall amounts from our data are likely underestimates.
This data source, however, should be internally consistent and so useful
for comparing across different storms when all exposure estimates are
based on this rain data.

Rainfall estimates are likely underestimates for a few reasons. First,
they are based on averaging hourly measurements to a daily mean
estimate. This averaging would smooth over shorter periods of very
extreme rainfall. Further, this data is averaged over multiple grid
points within each county and so would not fully reflect very extreme
local precipitation (although this might be less of a concern for
classifying exposure to a large-scale storm system, like a tropical
storm, compared to more fine-scale storms). Finally, this NLDAS data
provides a re-analysis that incorporates measured rainfall, using
models, etc., to incorporate that observed data into a spatially and
temporally continuous dataset of rainfall. However, during extreme
storms, the problems with measuring rainfall using {[}rain monitors{]}
would propogate into the NLDAS data, so although NLDAS would prevent
missing values during the storm if monitors are not able to provide
data, if monitors are out, rainfall estimates from NLDAS will be based
more on models than on observations.

Exposure classification based on rainfall has some advantages. For
example, it allows the identification of exposed counties that are
inland, rather than coastal, but that were affected by heavy rainfall
during the storm. Often, storm-related deaths are associated with inland
flooding. Storms can produce a lot of rain especially in certain
topographies, like near mountains, so counties that are well inland can
sometimes experience more extreme rain that other counties at similar
distances from the storm's track.

In comparison to the storm track, the areas of extreme rain and extreme
wind can vary. For example, storms undergoing an extratropical
transition can bring heavy rains to the left of the storm's center
track, while extreme winds are more common to the right of the track
(Halverson 2015).

Some storms can be of lower intensity (i.e., on the Saffir-Simpson
scale) yet bring dangerous rainfall, including well inland of the
storm's landfall. For example, storms including Floyd in 1999, Gaston in
2004, Irene in 2011, and Lee in 2011, have had severe inland impacts,
often through extreme rainfall, as post-tropical storms
{[}Halverson2015{]}. This rainfall can be particularly severe in the
Appalachians {[}Halverson2015{]}.

For example,

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-3-1.pdf}

\paragraph{Wind-based exposure}\label{wind-based-exposure-1}

Wind suffers from similar challenges for measuring during tropical
storms. In particular, the strong winds of tropical storms can break or
blow away the anemometers used to measure wind speed.

Here, we used wind speed models, rather than observed wind speed, to
estimate exposure to tropical storms based on winds.

A variety of other wind speed models exist besides the one used here. In
particular, there are options for wind speed models as far as \ldots{}

\paragraph{Example uses of exposure
datasets}\label{example-uses-of-exposure-datasets}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Halverson2015}{}
Halverson, Jeffrey B. 2015. ``Second Wind: The Dealdy and Destructive
Inland Phase of East Coast Hurricanes.'' \emph{Weatherwise} 68 (2):
20--27.
doi:\href{https://doi.org/10.1080/00431672.2015.997562}{10.1080/00431672.2015.997562}.

\end{document}

### 2016-06-10 writing/rain_plots.R

@@ -1,0 +1,18 @@
+all_storms <- unique(closest_dist$storm_id)
+# ex_storms <- all_storms[-grep("Unnamed", all_storms)]
+
+plot_both <- function(storm, metric = "distance"){
+        map_1 <- map_counties(storm = storm, metric = metric)
+        map_2 <- map_tracks(storms = storm, plot_object = map_1,
+                            plot_points = FALSE)
+        map_2 <- map_2 + ggplot2::ggtitle(storm)
+        return(map_2)
+}
+
+pdf(file = "writing/rain_plots.pdf", width = 7, height = 4)
+lapply(ex_storms[1:100], plot_both, metric = "rainfall")
+dev.off()
+
+pdf(file = "writing/distance_plots.pdf", width = 7, height = 4)
+lapply(ex_storms[1:100], plot_both, metric = "distance")
+dev.off()

### 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add text to paper draft

@@
 
-We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database. ... This data gives time stamps for each observation in UTC.
+We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database.
+This dataset is based on a poststorm assessment of each storm conducted by the United States National Hurricane Center (NHC) and incorporates data from a variety of sources, including satellite data and, when available, aircraft reconnaisance data [@Landsea2013].
+This data gives time stamps for each observation in Coordinated Universal Time (UTC; also known as Zulu Time, sometimes indicated by "Z").
 
-These data typically give measurements of storm center location at 6-hour intervals.
+These data typically give measurements of storm center location at 6-hour intervals, at synoptic times (i.e., 6:00 am, 12:00 pm, 6:00 pm, and 12:00 am UTC); some landfalling storms have an additional observation at the time of landfall [@Landsea2013].
+These positions are given to within 0.1 degrees latitude / longitude at these synoptic times [@Landsea2013].
 We interpolated these location values to every 15 minutes during the period when the storm was active, using a linear interpolation between each measured point.
@@
 #### Wind-based exposure
+
+These wind models use estimates of maximum wind speed from the best tracks dataset.
+These maximum wind speed estimates in the best tracks dataset are rounded to 5-knot intervals and is the maximum 1-minute-average windspeed at 10 meters above the ground [@Landsea2013].
+This windspeed is typically determined using the Dvorak technique to estimate windspeed from satellite measurements, although data from aircraft reconnaisance in also sometimes incorporated in the estimate [@Levinson2010].
 
@@
 Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could mis-classify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).
+
+There is some uncertainty in the position estimates from the best tracks hurricane dataset, since the estimation of storm position for best tracks data involes a poststorm subjective smoothing and integration of many different types of data [@Landsea2013].
+In a survey of researchers who perform the poststorm data aggregation to create the best tracks datasets, uncertainty in the center position of a US landfalling storm in the "best tracks" dataset was estimated at approximately 8 nautical miles for major hurricanes, 11 nautical miles for Category 1 and 2 hurricanes, and 15 nautical miles for tropical storms [@Landsea2013].
+Uncertainty in estimates of a storm's position also varies by time of day, with more certain estimates during daylight than during the night [@Landsea2013].
 
@@
 
+There is some uncertainty in the maximum wind speed values estimated at each synoptic time for each storm.
+For US landfalling storms, the estimate of storm intensity in the best tracks dataset is estimated to have an uncertainty of around 8 knots for tropical storms, 10 knots for Category 1 and 2 hurricanes, and 13 knots for major hurricanes [@Landsea2013].
+Since the wind model used here uses these best tracks maximum wind speeds as an input, this uncertaintly would propogate into our estimates of windspeed within each county for each storm.
+
 #### Example uses of exposure datasets

### 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.tex
Add text to paper draft

@@
 storms between 1988 and 2014 from the extended best tracks database.
-\ldots{} This data gives time stamps for each observation in UTC.
+This dataset is based on a poststorm assessment of each storm conducted
+by the United States National Hurricane Center (NHC) and incorporates
+data from a variety of sources, including satellite data and, when
+available, aircraft reconnaisance data (Landsea and Franklin 2013). This
+data gives time stamps for each observation in Coordinated Universal
+Time (UTC; also known as Zulu Time, sometimes indicated by ``Z'').
 
 These data typically give measurements of storm center location at
-6-hour intervals. We interpolated these location values to every 15
-minutes during the period when the storm was active, using a linear
-interpolation between each measured point.
+6-hour intervals, at synoptic times (i.e., 6:00 am, 12:00 pm, 6:00 pm,
+and 12:00 am UTC); some landfalling storms have an additional
+observation at the time of landfall (Landsea and Franklin 2013). These
+positions are given to within 0.1 degrees latitude / longitude at these
+synoptic times (Landsea and Franklin 2013). We interpolated these
+location values to every 15 minutes during the period when the storm was
+active, using a linear interpolation between each measured point.
 
@@
 \paragraph{Wind-based exposure}\label{wind-based-exposure}
+
+These wind models use estimates of maximum wind speed from the best
+tracks dataset. These maximum wind speed estimates in the best tracks
+dataset are rounded to 5-knot intervals and is the maximum
+1-minute-average windspeed at 10 meters above the ground (Landsea and
+Franklin 2013). This windspeed is typically determined using the Dvorak
+technique to estimate windspeed from satellite measurements, although
+data from aircraft reconnaisance in also sometimes incorporated in the
+estimate (Levinson et al. 2010).
 
@@
 large or intense storms).
+
+There is some uncertainty in the position estimates from the best tracks
+hurricane dataset, since the estimation of storm position for best
+tracks data involes a poststorm subjective smoothing and integration of
+many different types of data (Landsea and Franklin 2013). In a survey of
+researchers who perform the poststorm data aggregation to create the
+best tracks datasets, uncertainty in the center position of a US
+landfalling storm in the ``best tracks'' dataset was estimated at
+approximately 8 nautical miles for major hurricanes, 11 nautical miles
+for Category 1 and 2 hurricanes, and 15 nautical miles for tropical
+storms (Landsea and Franklin 2013). Uncertainty in estimates of a
+storm's position also varies by time of day, with more certain estimates
+during daylight than during the night (Landsea and Franklin 2013).
 
@@
 
+There is some uncertainty in the maximum wind speed values estimated at
+each synoptic time for each storm. For US landfalling storms, the
+estimate of storm intensity in the best tracks dataset is estimated to
+have an uncertainty of around 8 knots for tropical storms, 10 knots for
+Category 1 and 2 hurricanes, and 13 knots for major hurricanes (Landsea
+and Franklin 2013). Since the wind model used here uses these best
+tracks maximum wind speeds as an input, this uncertaintly would
+propogate into our estimates of windspeed within each county for each
+storm.
+
 \paragraph{Example uses of exposure
@@
 
+\hypertarget{ref-Landsea2013}{}
+Landsea, Christopher W., and James L. Franklin. 2013. ``Atlantic
+Hurricane Database Uncertainty and Presentation of a New Database
+Format.'' \emph{Monthly Weather Review} 141: 3576--92.
+
+\hypertarget{ref-Levinson2010}{}
+Levinson, David H., Howard J. Diamond, Kenneth R. Knapp, Michael C.
+Kruk, and Ethan J. Gibney. 2010. ``Toward a Homogenous Global Tropical
+Cyclone Best-Track Dataset.'' \emph{American Meteorological Society} x:
+377--80.
+doi:\href{https://doi.org/10.1175/2010BAMS2930.1}{10.1175/2010BAMS2930.1}.
+
 \end{document}

### 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.Rmd
made two quick spelling edits

@@
   - name: Joshua Ferreri
-    email: alice@example.com
+    email: joshua.m.ferreri@gmail.com
     affiliation: Colorado State University
@@
 
-It can be very difficult to reliabily measure rain during extreme rain events, including tropical storms.
+It can be very difficult to reliably measure rain during extreme rain events, including tropical storms.
 For example, a heavy rain can wash away [?] rain monitors [?].

### 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Added info on distance and rain in paper draft. Added references

@@
 
+Currently, distance parameters involved with assessing risk of a particular storm have been rather arbitrary, contributing to the necessity in understanding how such parameters influence a county's exposure status.
+In assessing the risk of a given storm based on distance, recent studies have defined the distance from the storm track affected by a given storm differently.
+[@Czajkowski2011] assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties.
+Such a defenition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011].
 #### Rain-based exposure
@@
 Often, storm-related deaths are associated with inland flooding.
+In fact, over half of the hurricane-related fatalities from 1970 to 1999 were a result of freshwater flooding, accounting for the vast majority of deaths in inland counties [@Rapport2000].
+In another analysis, researchers found that 79\% of freshwater-drowning fatalities occured in inland counties,[@Czajkowski2011] further stressing the importance of rainfall in hurricane risk analysis.
 Storms can produce a lot of rain especially in certain topographies, like near mountains, so counties that are well inland can sometimes experience more extreme rain that other counties at similar distances from the storm's track.
@@
 Some storms can be of lower intensity (i.e., on the Saffir-Simpson scale) yet bring dangerous rainfall, including well inland of the storm's landfall.
-For example, storms including Floyd in 1999, Gaston in 2004, Irene in 2011, and Lee in 2011, have had severe inland impacts, often through extreme rainfall, as post-tropical storms [Halverson2015].
-This rainfall can be particularly severe in the Appalachians [Halverson2015].
+For example, storms including Floyd in 1999, Gaston in 2004, Irene in 2011, and Lee in 2011, have had severe inland impacts, often through extreme rainfall, as post-tropical storms [@Halverson2015].
+This rainfall can be particularly severe in the Appalachians [@Halverson2015].
+Further, heavier rainfall is more common for storms moving at slower wind speeds [@Chang2013].
+Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].
 

### 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Added references. More info on distance parameter

@@
 Such a defenition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011].
+Such a distance is slightly greater than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].
+In other cases, distance has been defined on the basis of maximum sustained wind speeds and their decay, with exposure defined as those counties exposed to wind speeds greater than or equal to tropical storm strength (39--73 miles per hour) [@Zandbergen2009] and may not provide an accurate respresentation of the communities exposed to a given storm.
 #### Rain-based exposure

### 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add to introduction. Add references

--- 2016-06-13 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+Current hurricane exposure assignments are often nonspecific and may lead to missclassification of exposure at the county-level [@Grabich2016].
+Such missclassification can have important public health and economic implications.
+
+Studies have begun to assess county-level hurricane exposure using large historical datasets [@Zandbergen2009] ... making the `hurricaneexposure` package a powerful tool in future exposure analysis.
+
 #### Examples of using TS exposure estimates for multiple storms
+
+[@Zandbergen2009] developed an exposure factor based on location (longitude and latitude), county shape, county size, and distance from the coast.
 
@@
 Such a defenition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011].
+Such a distance is slightly greater than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].
+In other cases, distance has been defined on the basis of maximum sustained wind speeds and their decay, with exposure defined as those counties exposed to wind speeds greater than or equal to tropical storm strength (39--73 miles per hour) [@Zandbergen2009] and may not provide an accurate respresentation of the communities exposed to a given storm (e.g., a large number of inland counties are likely underrepresented under this definition).
 #### Rain-based exposure

### 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Results of Grabich referenced

@@
 Current hurricane exposure assignments are often nonspecific and may lead to missclassification of exposure at the county-level [@Grabich2016].
+Researchers have found heterogeneity in exposure assignments of counties in assessing multiple classification methods, both within and accross storms (i.e., storm dependent) [@Grabich2016].
 Such missclassification can have important public health and economic implications.
@@
 In comparison to the storm track, the areas of extreme rain and extreme wind can vary.
-For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015].
+For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016].
 

### 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Hurricane tracks now available in R
Kruk reference

@@
 
+Currently, it seems that most hurricane track mapping capabilities are restricted to geographical information system software.
+The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.
 #### Examples of using TS exposure estimates for multiple storms
@@
 
+A recent study suggests that nearly all states east of the Rocky Mountains have experienced wind exposure associated with either tropical or post-tropical storms [@Kruk2010].
+
 #### Example uses of exposure datasets

### 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add Currie reference
added Villarini reference

@@
 
+[@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones. However, the Stage IV dataset only provides data for ...
+
 #### Wind-based exposure
@@
 In other cases, distance has been defined on the basis of maximum sustained wind speeds and their decay, with exposure defined as those counties exposed to wind speeds greater than or equal to tropical storm strength (39--73 miles per hour) [@Zandbergen2009] and may not provide an accurate respresentation of the communities exposed to a given storm (e.g., a large number of inland counties are likely underrepresented under this definition).
+
+In a study assessing the association between hurricanes and undesireable birth outcomes, researchers found that results were not sensitive to the omission of residences 100 km from the storm path, and that results varied insignificantly from 30-75 km [@Currie2013].
+
 #### Rain-based exposure

### 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Stage IV data info and Lin reference

@@
 The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.
+
 #### Examples of using TS exposure estimates for multiple storms
@@
 [@Zandbergen2009] developed an exposure factor based on location (longitude and latitude), county shape, county size, and distance from the coast.
+The study assumed symmetrical activity of storm related factors,[@Zandbergen2009] which does not accurately represent storm characteristics once landfall has been made.[@Kruk2010, @Halverson2015]
 
@@
 
-[@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones. However, the Stage IV dataset only provides data for ...
+Given the its high resolution and consistancy with rain gage values, [@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
+However, the Stage IV dataset only provides data since 2001 [@Lin2005] and is therefore inadequate for use in this analysis.
 
@@
 
-Some of the other possible sources for estimating rain during tropical storms include ...
+Some of the other possible sources for estimating rain during tropical storms include ...[Stage IV, TMPA, NEXRAD]
 
@@
 However, during extreme storms, the problems with measuring rainfall using [rain monitors] would propogate into the NLDAS data, so although NLDAS would prevent missing values during the storm if monitors are not able to provide data, if monitors are out, rainfall estimates from NLDAS will be based more on models than on observations.
+More on NLDAS bias low here [@Villarini2011].
+Such bias, though unimportant in our analysis, may play a larger role in the extrapolation to other studies or exposure analysese where the precipitation cutoffs used here may provide a threshold that results in the missclassification of less affected counties as exposed.
 
@@
 This rainfall can be particularly severe in the Appalachians [@Halverson2015].
-Further, heavier rainfall is more common for storms moving at slower wind speeds [@Chang2013].
+Further, heavier rainfall is more common for storms moving at slower forward speeds [@Chang2013].
 Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].

### 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add Atallah reference (rain dist'n)

@@
 
-Given the its high resolution and consistancy with rain gage values, [@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
+Given its high resolution and consistancy with rain gage values, [@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
 However, the Stage IV dataset only provides data since 2001 [@Lin2005] and is therefore inadequate for use in this analysis.
+
+NLDAS does not provide information on precipitation over the ocean, preventing its use in assessing the develpment of the storm prior to landfall [@Villarini2011].
+This point is moot for this analysis.
 
@@
 For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016].
+In contrast, storm that interact with a downstream ridge tend to bring heavy rains to the right of the storm's center track [@Atallah2007].
 

### 2016-06-15 writing/DraftExposurePaper/DraftExposurePaper.Rmd
try to update draft
table with literature and parameters

@@
 library(hurricaneexposure)
+library(knitr)
 ```
@@
 Currently, it seems that most hurricane track mapping capabilities are restricted to geographical information system software.
-The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.
+The introduction of this package expands the mapping potential of hurricane tracks, and therefore the assessment of hurricane exposure and impacts, to R.
 
@@
 
+Most distance exposure definitions in the past have been based on maximum wind radii based on maximum sustained winds [@Kruck2010].
+
 #### Rain-based exposure
@@
 
+#### Table including various exposure papers and the parameters used
+
+```{r}
+studies <- c("Czajkowski 2010", "Rezempour 2014", "Zandbergen 2009", 
+             "Currie 2013", "Grabich 2016", "Zaharan 2010", "Kruck 2010", 
+             "Villarini 2011", "Atallah 2007")
+distance_km <- c("120", "", "Buffer", "25, 30, 60, 75", "30, 60, 100", 
+                 "2 counties", "Radial Max Wind", "100-500", "")
+rain_mm <- c("", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
+wind_kt <- c("", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
+exposure_parameters <- data.frame(studies, distance_km, rain_mm, wind_kt)
+kable(exposure_parameters, "markdown")
+```
+SSWS = Staffir-Simpson Hurricane Wind Scale (TS, 1, 2, 3, 4, and 5). Wind Cutoffs at 34, 64, 83, 96, 113, and 137 kt.  
 References {#references .unnumbered}

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.tex
Close #36

@@
     \author[Colorado State University]{Joshua Ferreri}
-   \ead{alice@example.com}
+   \ead{joshua.m.ferreri@gmail.com}
 
@@
 
+Current hurricane exposure assignments are often nonspecific and may
+lead to missclassification of exposure at the county-level (Grabich et
+al. 2016). Researchers have found heterogeneity in exposure assignments
+of counties in assessing multiple classification methods, both within
+and accross storms (i.e., storm dependent) (Grabich et al. 2016). Such
+missclassification can have important public health and economic
+implications.
+
+Studies have begun to assess county-level hurricane exposure using large
+historical datasets (Zandbergen 2009) \ldots{} making the
+\texttt{hurricaneexposure} package a powerful tool in future exposure
+analysis.
+
+Currently, it seems that most hurricane track mapping capabilities are
+restricted to geographical information system software. The introduction
+of this package expands the mapping potential of hurricane tracks and
+therefore the assessment of hurricane exposure and impacts to R.
+
 \paragraph{Examples of using TS exposure estimates for multiple
 storms}\label{examples-of-using-ts-exposure-estimates-for-multiple-storms}
+
+(Zandbergen 2009) developed an exposure factor based on location
+(longitude and latitude), county shape, county size, and distance from
+the coast. The study assumed symmetrical activity of storm related
+factors,(Zandbergen 2009) which does not accurately represent storm
+characteristics once landfall has been made.(Kruk et al. 2010, Halverson
+(2015))
 
@@
 \paragraph{Rain-based exposure}\label{rain-based-exposure}
+
+Given its high resolution and consistancy with rain gage values,
+(Villarini et al. 2011) recommends the use of Stage IV precipitation
+data for the analysis of rainfall distribution in landfalling tropical
+cyclones. However, the Stage IV dataset only provides data since 2001
+(Lin and Mitchell 2005) and is therefore inadequate for use in this
+analysis.
+
+NLDAS does not provide information on precipitation over the ocean,
+preventing its use in assessing the develpment of the storm prior to
+landfall (Villarini et al. 2011). This point is moot for this analysis.
 
@@
 
+Currently, distance parameters involved with assessing risk of a
+particular storm have been rather arbitrary, contributing to the
+necessity in understanding how such parameters influence a county's
+exposure status. In assessing the risk of a given storm based on
+distance, recent studies have defined the distance from the storm track
+affected by a given storm differently. (Czajkowski, Simmons, and Sutter
+2011) assessed county-level risk and exposure based on a three-tierd
+definition, with primary counties being those closest to the storm track
+on either side, secondary counties being adjacent to primary coutnies,
+and tertiary counties adjacent to secondary counties. Such a defenition
+resulted in an exposure defenition based on an average distance radius
+of 120 km on either side of the storm track (Czajkowski, Simmons, and
+Sutter 2011). Such a distance is slightly greater than that commonly
+used by public health departments (i.e., 100 km) (Czajkowski, Simmons,
+and Sutter 2011). In other cases, distance has been defined on the basis
+of maximum sustained wind speeds and their decay, with exposure defined
+as those counties exposed to wind speeds greater than or equal to
+tropical storm strength (39--73 miles per hour) (Zandbergen 2009) and
+may not provide an accurate respresentation of the communities exposed
+to a given storm (e.g., a large number of inland counties are likely
+underrepresented under this definition).
+
+In a study assessing the association between hurricanes and undesireable
+birth outcomes, researchers found that results were not sensitive to the
+omission of residences 100 km from the storm path, and that results
+varied insignificantly from 30-75 km (Currie and Rossin-Slater 2013).
+
 \paragraph{Rain-based exposure}\label{rain-based-exposure-1}
 
-It can be very difficult to reliabily measure rain during extreme rain
+It can be very difficult to reliably measure rain during extreme rain
 events, including tropical storms. For example, a heavy rain can wash
@@
 Some of the other possible sources for estimating rain during tropical
-storms include \ldots{}
+storms include \ldots{}{[}Stage IV, TMPA, NEXRAD{]}
 
@@
 data, if monitors are out, rainfall estimates from NLDAS will be based
-more on models than on observations.
+more on models than on observations. More on NLDAS bias low here
+(Villarini et al. 2011). Such bias, though unimportant in our analysis,
+may play a larger role in the extrapolation to other studies or exposure
+analysese where the precipitation cutoffs used here may provide a
+threshold that results in the missclassification of less affected
+counties as exposed.
 
@@
 during the storm. Often, storm-related deaths are associated with inland
-flooding. Storms can produce a lot of rain especially in certain
-topographies, like near mountains, so counties that are well inland can
-sometimes experience more extreme rain that other counties at similar
-distances from the storm's track.
+flooding. In fact, over half of the hurricane-related fatalities from
+1970 to 1999 were a result of freshwater flooding, accounting for the
+vast majority of deaths in inland counties ({\textbf{???}}). In another
+analysis, researchers found that 79\% of freshwater-drowning fatalities
+occured in inland counties,(Czajkowski, Simmons, and Sutter 2011)
+further stressing the importance of rainfall in hurricane risk analysis.
+Storms can produce a lot of rain especially in certain topographies,
+like near mountains, so counties that are well inland can sometimes
+experience more extreme rain that other counties at similar distances
+from the storm's track.
 
@@
 track, while extreme winds are more common to the right of the track
-(Halverson 2015).
+(Halverson 2015, Grabich et al. (2016)). In contrast, storm that
+interact with a downstream ridge tend to bring heavy rains to the right
+of the storm's center track (Atallah, Bosart, and Aiyyer 2007).
 
@@
 2004, Irene in 2011, and Lee in 2011, have had severe inland impacts,
-often through extreme rainfall, as post-tropical storms
-{[}Halverson2015{]}. This rainfall can be particularly severe in the
-Appalachians {[}Halverson2015{]}.
+often through extreme rainfall, as post-tropical storms (Halverson
+2015). This rainfall can be particularly severe in the Appalachians
+(Halverson 2015). Further, heavier rainfall is more common for storms
+moving at slower forward speeds (Chang, Yang, and Kuo 2013). Storms with
+a slower forward speed and larger rainfall area contribute to a longer
+duration of rainfall, and therefore an increased chance of flood-related
+damage (Rezapour and Baldock 2014).
 
@@
 
+A recent study suggests that nearly all states east of the Rocky
+Mountains have experienced wind exposure associated with either tropical
+or post-tropical storms (Kruk et al. 2010).
+
 \paragraph{Example uses of exposure
 datasets}\label{example-uses-of-exposure-datasets}
+
+\paragraph{Table including various exposure papers and the parameters
+used}\label{table-including-various-exposure-papers-and-the-parameters-used}
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Atallah2007}{}
+Atallah, Eyad, Lance F. Bosart, and Anantha R. Aiyyer. 2007.
+``Precipitation Distribution Associated with Landfalling Tropical
+Cyclones over the Eastern United States.'' \emph{Monthly Weather Review}
+135: 2185--2206.
+doi:\href{https://doi.org/10.1175/MWR3382.1}{10.1175/MWR3382.1}.
+
+\hypertarget{ref-Chang2013}{}
+Chang, Chih-Pei, Yi-Ting Yang, and Hung-Chi Kuo. 2013. ``Large
+Increasing Trend of Tropical Cyclone Rainfall in Taiwan and the Roles of
+Terrain.'' \emph{Journal of Climate} 26: 4136--47.
+doi:\href{https://doi.org/10.1175/JCLI-D-12-00463.1}{10.1175/JCLI-D-12-00463.1}.
+
+\hypertarget{ref-Currie2013}{}
+Currie, Janet, and Maya Rossin-Slater. 2013. ``Weathering the Storm:
+Hurricanes and Birth Outcomes.'' \emph{Journal of Health Economics} 32:
+487--503.
+doi:\href{https://doi.org/10.1016/j.jhealeco.2013.01.004}{10.1016/j.jhealeco.2013.01.004}.
+
+\hypertarget{ref-Czajkowski2011}{}
+Czajkowski, Jeffrey, Kevin Simmons, and Daniel Sutter. 2011. ``An
+Analysis of Coastal and Inland Fatalities in Landfalling US
+Hurricanes.'' \emph{Natural Hazards} 59: 1513--31.
+doi:\href{https://doi.org/10.1007/s11069-011-9849-x}{10.1007/s11069-011-9849-x}.
+
+\hypertarget{ref-Grabich2016}{}
+Grabich, S. C., J. Horney, C. Konrad, and D. T. Lobdell. 2016.
+``Measuring the Storm: Methods of Quantifying Hurricane Exposure with
+Pregnancy Outcomes.'' \emph{Natural Hazards Review} 17 (1):
+06015002--1--06015002--7.
+doi:\href{https://doi.org/10.1061/(ASCE)NH.1527-6996.0000204}{10.1061/(ASCE)NH.1527-6996.0000204}.
+
 \hypertarget{ref-Halverson2015}{}
-Halverson, Jeffrey B. 2015. ``Second Wind: The Dealdy and Destructive
+Halverson, Jeffrey B. 2015. ``Second Wind: The Deadly and Destructive
 Inland Phase of East Coast Hurricanes.'' \emph{Weatherwise} 68 (2):
@@
 doi:\href{https://doi.org/10.1080/00431672.2015.997562}{10.1080/00431672.2015.997562}.
+
+\hypertarget{ref-Kruk2010}{}
+Kruk, Michael C., Ethan J. Gibney, David H. Levinson, and Michael
+Squires. 2010. ``A Climatology of Inland Winds from Tropical Cyclones
+for the Eastern United States.'' \emph{Journal of Applied Meteorology
+and Climatology} 49: 1538--47.
+doi:\href{https://doi.org/10.1175/2010JAMC2389.1}{10.1175/2010JAMC2389.1}.
 
@@
 
+\hypertarget{ref-Lin2005}{}
+Lin, Ying, and Kenneth E. Mitchell. 2005. ``The NCEP Stage II/IV Hourly
+Precipitation Analyses: Development and Applications.'' \emph{19th
+Conference of Hydrology, American Meteorlogical Society}.
+
+\hypertarget{ref-Rezapour2014}{}
+Rezapour, Mehdi, and Tom E. Baldock. 2014. ``Classification of Hurricane
+Hazards: The Importance of Rainfall.'' \emph{Weather and Forcasting} 29:
+1319--31.
+
+\hypertarget{ref-Villarini2011}{}
+Villarini, Gabriele, James A. Smith, Mary Lynn Baeck, Timothy Marchok,
+and Gabriel A. Vecchi. 2011. ``Characterization of Rainfall Distribution
+and Flooding Associated with U.S. Landfalling Tropical Cyclones:
+Analysis of Hurricanes Frances, Ivan, and Jeanne (2004).'' \emph{Journal
+of Geophysical Research} 116: D23116.
+
+\hypertarget{ref-Zandbergen2009}{}
+Zandbergen, Paul A. 2009. ``Exposure of US Counties to Atlantic Tropical
+Storms and Hurricanes, 1851--2003.'' \emph{Natural Hazards} 48: 83--99.
+doi:\href{https://doi.org/10.1007/s11069-008-9250-6}{10.1007/s11069-008-9250-6}.
+
 \end{document}

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Close #36
Add to paper draft

--- 2016-06-14 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r}
+```{r echo = FALSE}
 library(hurricaneexposure)
@@
 
-Current hurricane exposure assignments are often nonspecific and may lead to missclassification of exposure at the county-level [@Grabich2016].
-Researchers have found heterogeneity in exposure assignments of counties in assessing multiple classification methods, both within and accross storms (i.e., storm dependent) [@Grabich2016].
-Such missclassification can have important public health and economic implications.
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights).
+Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009].
+Some studies suggest methods to determine exposure, including inland exposure, to tropical storms at the county-level over many years [@Zandbergen2009].
+Here, we offer tools, through the open-source `hurricaneexposure` package, to determine county-level exposure to tropical storms using several different metrics (distance, rainfall, wind).
+In this paper, we describe how this software package assesses exposure to tropical storms at the county level, provide an analysis of the advantages and disadvantages of different strategies for assessing county-level exposure, and explore variation in exposure assessments based on using different exposure metrics.
 
-Studies have begun to assess county-level hurricane exposure using large historical datasets [@Zandbergen2009] ... making the `hurricaneexposure` package a powerful tool in future exposure analysis.
+Current hurricane exposure assignments are often nonspecific [GBA-- I found this wording a bit unclear] and may lead to missclassification of exposure at the county-level [@Grabich2016].
+Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016].
+Such missclassification can have important implications of assessments of the public health and economic impacts of tropical storms.
 
-Currently, it seems that most hurricane track mapping capabilities are restricted to geographical information system software.
-The introduction of this package expands the mapping potential of hurricane tracks and therefore the assessment of hurricane exposure and impacts to R.
+Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies].
+Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.
 
-#### Examples of using TS exposure estimates for multiple storms
+#### Examples of estimating TS for multiple storms
 
-[@Zandbergen2009] developed an exposure factor based on location (longitude and latitude), county shape, county size, and distance from the coast.
-The study assumed symmetrical activity of storm related factors,[@Zandbergen2009] which does not accurately represent storm characteristics once landfall has been made.[@Kruk2010, @Halverson2015]
+When assessing exposure to a single tropical storm, a detailed analysis of ...
+can be performed for the storm.
+However, this level of detailed analysis is unreasonable when assessing exposure to all storms over an extended time period; further, such a long-term exposure assessment requires consistency in the data being used to determine levels of exposure to factors like wind and rain.
+
+A few studies have sought to determine exposure to tropical storms over a  study period that includes multiple storms.
+For example, ...
+assessed the exposure of counties in Florida to hurricanes during the 2004 season [citation].
+The largest county-level exposure study is likely that of @Zandbergen2009, which estimated exposure in US counties to all [US landfalling?
+or all storms?] Atlantic basin tropical storms between 1851 and 2003, using both distance and an exposure metric that incorporated distance and windspeed.
+They used this exposure evaluation to create maps of total exposure to tropical storms within US counties, as well as to explore associations between a county's long-term exposure to tropical storms and its location, distance from the coast, size, and shape [@Zandbergen2009].
 
@@
 There are a number of limitations to using distance to assess exposure to tropical storms.
-Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could mis-classify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).
+Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could misclassify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).
+In particular, the forces of a storm tend to be distributed around the center in a non-symmetrical way [citation], while distance-based exposure metrics that use buffers tend to use an equal buffer distance on each side of the storm track (e.g., [citations]).
+
+Other distance-based exposure metrics have defined "exposure" to occur only when a storm's center track passes through a county's boundary (e.g., @Zandbergen2009, who describe this metric of exposure as a "hit" on the county).
+This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track, especially since hurricanes can have a width ([typical storm width; citation]) that is much larger than a typical county's width.
+Further, exposure assessed using this method has been found to be significantly [?
+double-check] associated with the size and shape of a county [@Zandbergen2009].
+By contrast, assessing distance-based exposure based on measuring distance between the storm track and each county's center would be less susceptible to differences in county size and shape and so might be a more reliable metric of exposure than determining if a storm track ever crossed a county boundary.
+Further, assessing the minimum distance between county center and the storm track, as we do here, allows for a continuous, rather than binary, metric of distance-based exposure to a storm.
+
+The @Zandbergen2009 study assumed symmetrical activity of storm related factors,which does not accurately represent storm characteristics once landfall has been made [@Kruk2010, @Halverson2015] [GBA-- I did not understand this thought.
+I agree that storms are asymmetric, but I don't think that's just after landfall.
+Is this something about characteristics of storms as they transform to extratropical systems?
+Let's discuss.]
 
@@
 [@Czajkowski2011] assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties.
-Such a defenition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011].
+Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011].
 Such a distance is slightly greater than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].
-In other cases, distance has been defined on the basis of maximum sustained wind speeds and their decay, with exposure defined as those counties exposed to wind speeds greater than or equal to tropical storm strength (39--73 miles per hour) [@Zandbergen2009] and may not provide an accurate respresentation of the communities exposed to a given storm (e.g., a large number of inland counties are likely underrepresented under this definition).
+
+Another study used distance buffers of 30, 60, and 100 kilometers for different metrics of distance-based exposure [@Grabich2016] [check the paper-- was this distance to any part of a county, or to the center of the county?].
 
@@
 Often, storm-related deaths are associated with inland flooding.
-In fact, over half of the hurricane-related fatalities from 1970 to 1999 were a result of freshwater flooding, accounting for the vast majority of deaths in inland counties [@Rapport2000].
+In fact, over half of the hurricane-related fatalities from 1970 to 1999 were a result of freshwater flooding, accounting for the vast majority of deaths in inland counties [@Rappaport2000].
 In another analysis, researchers found that 79\% of freshwater-drowning fatalities occured in inland counties,[@Czajkowski2011] further stressing the importance of rainfall in hurricane risk analysis.
@@
 In comparison to the storm track, the areas of extreme rain and extreme wind can vary.
-For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016].
-In contrast, storm that interact with a downstream ridge tend to bring heavy rains to the right of the storm's center track [@Atallah2007].
+For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
+The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference].
+Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation].
+In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007].
 
@@
 
-A variety of other wind speed models exist besides the one used here. In particular, there are options for wind speed models as far as ...
+A variety of other wind speed models exist besides the one used here.
+In particular, there are options for wind speed models as far as ...
+One study used a very simple model of constant winds up to a distance of 30 km from the storm track and a simple model of wind decay beyond that distance [@Zandbergen2009], based on a model developed by ...
+[see references for this paper].
+Another study used wind data from a NOAA public database based on real-time hurricane wind analysis, rather than modeling windspeeds themselves [@Grabich2016] [see the Powell reference in this paper for more on that data].
 
@@
 
+Exposure metrics based on wind speed may not provide an accurate representation of the communities exposed to a given storm, since, while wind speeds rapidly decay after landfall [citation], extreme rain can persist,for some  storms affecting a large number of inland counties, which would be underrepresented using an exposure metric the depends on windspeed.
+
+Examples of studies that used windspeed as a sole criteria when determining tropical storm exposure include ...
+[citation].
+In other studies, wind criteria have been combined with distance criteria, and with windspeed modeled based on maximum sustained windspeed estimates at each "best tracks" observation, combined with a simple model of decay in wind speed at distances further from the storm's center (wind speed equals the storm's maximum sustained windspeed up to 30 kilometers from the storm's center then follows a simple decay function beyond that distance) [@Zandbergen2009] [GBA-- this paper references another paper with more details on this decay model for windspeed-- the Gray and Klotzbach reference they have].
+In this study, wind speed cutoffs of 40 mph, 75 mph, and 115 mph were used to determine if a county was exposed to a storm (tropical, hurricane-force, and intense hurricane-force, respectively) [@Zandbergen2009].
+
+One study found that using a distance-based exposure metric that required the storm track to cross a county boundary assessed a number of counties as "unexposed" that suffered tropical storm- or hurricane-force winds from the storm and were assessed as "exposed" using a metric that combined distance and windspeed [@Zandbergen2009]; in fact, almost twice as many counties were assessed as having been exposed to at least one tropical storm between 1851 and 2003 when exposure was determined based on a combined wind-distance metric compared to a "hit" distance-based metric [@Zandbergen2009].
+This suggests that some distance-based metrics might result in a number of false negatives for county-level storm exposure.
+
+Studies that have used wind-based exposure metrics have used a variety of windspeed cutoff values to determine exposure.
+One study, for example, assigned a binary exposure value based on windspeed thresholds of 63 and 119 kilometers per hour, as well as a continuous exposure metric based on the exact value of the windspeed in the county during the storm and a categorical exposure metric based on the Saffir-Simpson storm severity categories [@Grabich2016].
+
+#### Other ways of assessing tropical storm exposure
+
+One study used FEMA disaster declarations as a binary indicator of tropical storm exposure [@Grabich2016].
+They found that this metric tended to assess a lot more counties as "exposed" to a storm than distance- or wind-based metrics [@Grabich2016], although they only compared the two exposure metrics as applied to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in 2004).
+
 #### Example uses of exposure datasets
+
+#### Table including various exposure papers and the parameters used
 

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.tex
Add to paper draft

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\ImportTok}[1]{{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
-\newcommand{\BuiltInTok}[1]{{#1}}
-\newcommand{\ExtensionTok}[1]{{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
 \usepackage{graphicx}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(hurricaneexposure)}
-\end{Highlighting}
-\end{Shaded}
-
 \section{Introduction}\label{introduction}
@@
 
-Current hurricane exposure assignments are often nonspecific and may
-lead to missclassification of exposure at the county-level (Grabich et
-al. 2016). Researchers have found heterogeneity in exposure assignments
-of counties in assessing multiple classification methods, both within
-and accross storms (i.e., storm dependent) (Grabich et al. 2016). Such
-missclassification can have important public health and economic
-implications.
+Many outcomes of interest are available at county-level aggregations
+(e.g., counts of health outcomes: Grabich et al. (2016): premature
+births and low birth weights). Further, decisions and policies to
+prepare for and respond to storms are often undertaken at the county
+level (Zandbergen 2009). Some studies suggest methods to determine
+exposure, including inland exposure, to tropical storms at the
+county-level over many years (Zandbergen 2009). Here, we offer tools,
+through the open-source \texttt{hurricaneexposure} package, to determine
+county-level exposure to tropical storms using several different metrics
+(distance, rainfall, wind). In this paper, we describe how this software
+package assesses exposure to tropical storms at the county level,
+provide an analysis of the advantages and disadvantages of different
+strategies for assessing county-level exposure, and explore variation in
+exposure assessments based on using different exposure metrics.
 
-Studies have begun to assess county-level hurricane exposure using large
-historical datasets (Zandbergen 2009) \ldots{} making the
-\texttt{hurricaneexposure} package a powerful tool in future exposure
-analysis.
+Current hurricane exposure assignments are often nonspecific {[}GBA-- I
+found this wording a bit unclear{]} and may lead to missclassification
+of exposure at the county-level (Grabich et al. 2016). Researchers have
+found heterogeneity in exposure assignments of counties, both within and
+between storms, when different metrics of exposure are used (Grabich et
+al. 2016). Such missclassification can have important implications of
+assessments of the public health and economic impacts of tropical
+storms.
 
-Currently, it seems that most hurricane track mapping capabilities are
-restricted to geographical information system software. The introduction
-of this package expands the mapping potential of hurricane tracks and
-therefore the assessment of hurricane exposure and impacts to R.
+Many of the studies that have assessed exposure to tropical storms in
+the US have used geographical information system software (e.g., ArcGIS)
+{[}example studies{]}. Here, we offer methods to map and output historic
+exposure to tropical storms that does not require the use of proprietary
+software but instead uses a package written in the R statistical
+programming language, which is free and open-source.
 
-\paragraph{Examples of using TS exposure estimates for multiple
-storms}\label{examples-of-using-ts-exposure-estimates-for-multiple-storms}
+\paragraph{Examples of estimating TS for multiple
+storms}\label{examples-of-estimating-ts-for-multiple-storms}
 
-(Zandbergen 2009) developed an exposure factor based on location
-(longitude and latitude), county shape, county size, and distance from
-the coast. The study assumed symmetrical activity of storm related
-factors,(Zandbergen 2009) which does not accurately represent storm
-characteristics once landfall has been made.(Kruk et al. 2010, Halverson
-(2015))
+When assessing exposure to a single tropical storm, a detailed analysis
+of \ldots{} can be performed for the storm. However, this level of
+detailed analysis is unreasonable when assessing exposure to all storms
+over an extended time period; further, such a long-term exposure
+assessment requires consistency in the data being used to determine
+levels of exposure to factors like wind and rain.
+
+A few studies have sought to determine exposure to tropical storms over
+a study period that includes multiple storms. For example, \ldots{}
+assessed the exposure of counties in Florida to hurricanes during the
+2004 season {[}citation{]}. The largest county-level exposure study is
+likely that of Zandbergen (2009), which estimated exposure in US
+counties to all {[}US landfalling? or all storms?{]} Atlantic basin
+tropical storms between 1851 and 2003, using both distance and an
+exposure metric that incorporated distance and windspeed. They used this
+exposure evaluation to create maps of total exposure to tropical storms
+within US counties, as well as to explore associations between a
+county's long-term exposure to tropical storms and its location,
+distance from the coast, size, and shape (Zandbergen 2009).
 
@@
 measure of distance from the storm track will not incorporate these
-differences and so could mis-classify exposure both in terms of
+differences and so could misclassify exposure both in terms of
 generating false positives (counties close to the storm track of very
 mild or small-radius storms) and false negatives (e.g., during very
-large or intense storms).
+large or intense storms). In particular, the forces of a storm tend to
+be distributed around the center in a non-symmetrical way
+{[}citation{]}, while distance-based exposure metrics that use buffers
+tend to use an equal buffer distance on each side of the storm track
+(e.g., {[}citations{]}).
+
+Other distance-based exposure metrics have defined ``exposure'' to occur
+only when a storm's center track passes through a county's boundary
+(e.g., Zandbergen (2009), who describe this metric of exposure as a
+``hit'' on the county). This way of measuring exposure can exclude
+nearby counties that suffered extreme conditions from the storm but were
+not directly on the hurricane's track, especially since hurricanes can
+have a width ({[}typical storm width; citation{]}) that is much larger
+than a typical county's width. Further, exposure assessed using this
+method has been found to be significantly {[}? double-check{]}
+associated with the size and shape of a county (Zandbergen 2009). By
+contrast, assessing distance-based exposure based on measuring distance
+between the storm track and each county's center would be less
+susceptible to differences in county size and shape and so might be a
+more reliable metric of exposure than determining if a storm track ever
+crossed a county boundary. Further, assessing the minimum distance
+between county center and the storm track, as we do here, allows for a
+continuous, rather than binary, metric of distance-based exposure to a
+storm.
+
+The Zandbergen (2009) study assumed symmetrical activity of storm
+related factors,which does not accurately represent storm
+characteristics once landfall has been made (Kruk et al. 2010, Halverson
+(2015)) {[}GBA-- I did not understand this thought. I agree that storms
+are asymmetric, but I don't think that's just after landfall. Is this
+something about characteristics of storms as they transform to
+extratropical systems? Let's discuss.{]}
 
@@
 on either side, secondary counties being adjacent to primary coutnies,
-and tertiary counties adjacent to secondary counties. Such a defenition
+and tertiary counties adjacent to secondary counties. Such a definition
 resulted in an exposure defenition based on an average distance radius
@@
 used by public health departments (i.e., 100 km) (Czajkowski, Simmons,
-and Sutter 2011). In other cases, distance has been defined on the basis
-of maximum sustained wind speeds and their decay, with exposure defined
-as those counties exposed to wind speeds greater than or equal to
-tropical storm strength (39--73 miles per hour) (Zandbergen 2009) and
-may not provide an accurate respresentation of the communities exposed
-to a given storm (e.g., a large number of inland counties are likely
-underrepresented under this definition).
+and Sutter 2011).
+
+Another study used distance buffers of 30, 60, and 100 kilometers for
+different metrics of distance-based exposure (Grabich et al. 2016)
+{[}check the paper-- was this distance to any part of a county, or to
+the center of the county?{]}.
 
@@
 track, while extreme winds are more common to the right of the track
-(Halverson 2015, Grabich et al. (2016)). In contrast, storm that
-interact with a downstream ridge tend to bring heavy rains to the right
-of the storm's center track (Atallah, Bosart, and Aiyyer 2007).
+(Halverson 2015, Grabich et al. (2016) GBA-- I think it would be better
+to just use meteorological papers as references for this point. The
+Halverson paper is great, but the Grabich paper I think might just be a
+secondary, rather than primary, reference for this point, so I think we
+could probably find a more direct reference). Wind speeds are typically
+more severe to the right of the storm's track because here the
+counter-clockwise cyclonic {[}right word?{]} winds are moving in the
+same direction as the forward motion of the storm, creating an additive
+effect for total windspeeds {[}citation{]}. In contrast, rains can be
+heaviest to the left of the storm's track, especially in cases when the
+storm interacts with a downstream ridge (Atallah, Bosart, and Aiyyer
+2007).
 
@@
 particular, there are options for wind speed models as far as \ldots{}
+One study used a very simple model of constant winds up to a distance of
+30 km from the storm track and a simple model of wind decay beyond that
+distance (Zandbergen 2009), based on a model developed by \ldots{}
+{[}see references for this paper{]}. Another study used wind data from a
+NOAA public database based on real-time hurricane wind analysis, rather
+than modeling windspeeds themselves (Grabich et al. 2016) {[}see the
+Powell reference in this paper for more on that data{]}.
 
@@
 or post-tropical storms (Kruk et al. 2010).
+
+Exposure metrics based on wind speed may not provide an accurate
+representation of the communities exposed to a given storm, since, while
+wind speeds rapidly decay after landfall {[}citation{]}, extreme rain
+can persist,for some storms affecting a large number of inland counties,
+which would be underrepresented using an exposure metric the depends on
+windspeed.
+
+Examples of studies that used windspeed as a sole criteria when
+determining tropical storm exposure include \ldots{} {[}citation{]}. In
+other studies, wind criteria have been combined with distance criteria,
+and with windspeed modeled based on maximum sustained windspeed
+estimates at each ``best tracks'' observation, combined with a simple
+model of decay in wind speed at distances further from the storm's
+center (wind speed equals the storm's maximum sustained windspeed up to
+30 kilometers from the storm's center then follows a simple decay
+function beyond that distance) (Zandbergen 2009) {[}GBA-- this paper
+references another paper with more details on this decay model for
+windspeed-- the Gray and Klotzbach reference they have{]}. In this
+study, wind speed cutoffs of 40 mph, 75 mph, and 115 mph were used to
+determine if a county was exposed to a storm (tropical, hurricane-force,
+and intense hurricane-force, respectively) (Zandbergen 2009).
+
+One study found that using a distance-based exposure metric that
+required the storm track to cross a county boundary assessed a number of
+counties as ``unexposed'' that suffered tropical storm- or
+hurricane-force winds from the storm and were assessed as ``exposed''
+using a metric that combined distance and windspeed (Zandbergen 2009);
+in fact, almost twice as many counties were assessed as having been
+exposed to at least one tropical storm between 1851 and 2003 when
+exposure was determined based on a combined wind-distance metric
+compared to a ``hit'' distance-based metric (Zandbergen 2009). This
+suggests that some distance-based metrics might result in a number of
+false negatives for county-level storm exposure.
+
+Studies that have used wind-based exposure metrics have used a variety
+of windspeed cutoff values to determine exposure. One study, for
+example, assigned a binary exposure value based on windspeed thresholds
+of 63 and 119 kilometers per hour, as well as a continuous exposure
+metric based on the exact value of the windspeed in the county during
+the storm and a categorical exposure metric based on the Saffir-Simpson
+storm severity categories (Grabich et al. 2016).
+
+\paragraph{Other ways of assessing tropical storm
+exposure}\label{other-ways-of-assessing-tropical-storm-exposure}
+
+One study used FEMA disaster declarations as a binary indicator of
+tropical storm exposure (Grabich et al. 2016). They found that this
+metric tended to assess a lot more counties as ``exposed'' to a storm
+than distance- or wind-based metrics (Grabich et al. 2016), although
+they only compared the two exposure metrics as applied to four storms
+(Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in 2004).
 

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd

@@
 
+```{r}
+studies <- c("Czajkowski 2010", "Rezempour 2014", "Zandbergen 2009", 
+             "Currie 2013", "Grabich 2016", "Zaharan 2010", "Kruck 2010", 
+             "Villarini 2011", "Atallah 2007")
+distance_km <- c("120", "", "Buffer", "25, 30, 60, 75", "30, 60, 100", 
+                 "2 counties", "Radial Max Wind", "100-500", "")
+rain_mm <- c("", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
+wind_kt <- c("", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
+exposure_parameters <- data.frame(studies, distance_km, rain_mm, wind_kt)
+kable(exposure_parameters, "markdown")
+```
+SSWS = Staffir-Simpson Hurricane Wind Scale (TS, 1, 2, 3, 4, and 5). Wind Cutoffs at 34, 64, 83, 96, 113, and 137 kt.  
 References {#references .unnumbered}

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.tex

@@
 1970 to 1999 were a result of freshwater flooding, accounting for the
-vast majority of deaths in inland counties ({\textbf{???}}). In another
+vast majority of deaths in inland counties (Rappaport 2000). In another
 analysis, researchers found that 79\% of freshwater-drowning fatalities
@@
 
+\hypertarget{ref-Rappaport2000}{}
+Rappaport, Edward N. 2000. ``Loss of Life in the United States
+Associated with Recent Atlantic Tropical Cyclones.'' \emph{Bulletin of
+the American Meteorological Society} x: 2065--73.
+
 \hypertarget{ref-Rezapour2014}{}

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add to writing draft.

--- 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights).
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates).
 Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009].
@@
 Current hurricane exposure assignments are often nonspecific [GBA-- I found this wording a bit unclear] and may lead to missclassification of exposure at the county-level [@Grabich2016].
-Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016].
+Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015].
 Such missclassification can have important implications of assessments of the public health and economic impacts of tropical storms.
 
-Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies].
+Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009].
 Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.
@@
 A few studies have sought to determine exposure to tropical storms over a  study period that includes multiple storms.
-For example, ...
-assessed the exposure of counties in Florida to hurricanes during the 2004 season [citation].
-The largest county-level exposure study is likely that of @Zandbergen2009, which estimated exposure in US counties to all [US landfalling?
+For example, @Grabich2016 assessed the exposure of counties in Florida to land-falling hurricanes during the 2004 season.
+The largest-scale county-level exposure study is likely that of @Zandbergen2009, which estimated exposure in US counties to all [US landfalling?
 or all storms?] Atlantic basin tropical storms between 1851 and 2003, using both distance and an exposure metric that incorporated distance and windspeed.
 They used this exposure evaluation to create maps of total exposure to tropical storms within US counties, as well as to explore associations between a county's long-term exposure to tropical storms and its location, distance from the coast, size, and shape [@Zandbergen2009].
+
+Other studies have used county-level exposure criteria to determine exposure to specific storms.
+For example, @Grabich2015 determined county-level exposure to two severe storms in 2004 in Florida counties for an epidemiological study.
 
@@
 Another study used distance buffers of 30, 60, and 100 kilometers for different metrics of distance-based exposure [@Grabich2016] [check the paper-- was this distance to any part of a county, or to the center of the county?].
+Another study used a distance criterion of 60 kilometers from the storm track [@Grabich2015] [again, check to see if this was to any point in the county or to some center point of the county].
 
 In a study assessing the association between hurricanes and undesireable birth outcomes, researchers found that results were not sensitive to the omission of residences 100 km from the storm path, and that results varied insignificantly from 30-75 km [@Currie2013].
+
+Most distance exposure definitions in the past have been based on maximum wind radii based on maximum sustained winds [@Kruk2010].
 
@@
 One study, for example, assigned a binary exposure value based on windspeed thresholds of 63 and 119 kilometers per hour, as well as a continuous exposure metric based on the exact value of the windspeed in the county during the storm and a categorical exposure metric based on the Saffir-Simpson storm severity categories [@Grabich2016].
+Another study used a criterion of wind speed at or above 74 miles per hour [@Grabich2015].
 
@@
 
+```{r echo = FALSE}
+studies <- c("Czajkowski 2010", "Rezempour 2014", "Zandbergen 2009", 
+             "Currie 2013", "Grabich 2016", "Zaharan 2010", "Kruk 2010", 
+             "Villarini 2011", "Atallah 2007")
+distance_km <- c("120", "", "Buffer", "25, 30, 60, 75", "30, 60, 100", 
+                 "2 counties", "Radial Max Wind", "100-500", "")
+rain_mm <- c("", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
+wind_kt <- c("", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
+exposure_parameters <- data.frame(studies, distance_km, rain_mm, wind_kt)
+kable(exposure_parameters, "markdown")
+```
+
+SSWS = Staffir-Simpson Hurricane Wind Scale (TS, 1, 2, 3, 4, and 5). Wind Cutoffs at 34, 64, 83, 96, 113, and 137 kt.
+
 References {#references .unnumbered}

### 2016-06-16 writing/DraftExposurePaper/DraftExposurePaper.tex
Add to writing draft.

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage{longtable}
 \usepackage{graphicx}
@@
 (e.g., counts of health outcomes: Grabich et al. (2016): premature
-births and low birth weights). Further, decisions and policies to
-prepare for and respond to storms are often undertaken at the county
-level (Zandbergen 2009). Some studies suggest methods to determine
-exposure, including inland exposure, to tropical storms at the
-county-level over many years (Zandbergen 2009). Here, we offer tools,
-through the open-source \texttt{hurricaneexposure} package, to determine
-county-level exposure to tropical storms using several different metrics
-(distance, rainfall, wind). In this paper, we describe how this software
-package assesses exposure to tropical storms at the county level,
-provide an analysis of the advantages and disadvantages of different
-strategies for assessing county-level exposure, and explore variation in
-exposure assessments based on using different exposure metrics.
+births and low birth weights; Grabich et al. (2015): birth rates).
+Further, decisions and policies to prepare for and respond to storms are
+often undertaken at the county level (Zandbergen 2009). Some studies
+suggest methods to determine exposure, including inland exposure, to
+tropical storms at the county-level over many years (Zandbergen 2009).
+Here, we offer tools, through the open-source \texttt{hurricaneexposure}
+package, to determine county-level exposure to tropical storms using
+several different metrics (distance, rainfall, wind). In this paper, we
+describe how this software package assesses exposure to tropical storms
+at the county level, provide an analysis of the advantages and
+disadvantages of different strategies for assessing county-level
+exposure, and explore variation in exposure assessments based on using
+different exposure metrics.
 
@@
 between storms, when different metrics of exposure are used (Grabich et
-al. 2016). Such missclassification can have important implications of
-assessments of the public health and economic impacts of tropical
-storms.
+al. 2016; Grabich et al. 2015). Such missclassification can have
+important implications of assessments of the public health and economic
+impacts of tropical storms.
 
@@
 the US have used geographical information system software (e.g., ArcGIS)
-{[}example studies{]}. Here, we offer methods to map and output historic
-exposure to tropical storms that does not require the use of proprietary
-software but instead uses a package written in the R statistical
-programming language, which is free and open-source.
+(example studies: Grabich et al. 2016; Zandbergen 2009). Here, we offer
+methods to map and output historic exposure to tropical storms that does
+not require the use of proprietary software but instead uses a package
+written in the R statistical programming language, which is free and
+open-source.
 
@@
 A few studies have sought to determine exposure to tropical storms over
-a study period that includes multiple storms. For example, \ldots{}
-assessed the exposure of counties in Florida to hurricanes during the
-2004 season {[}citation{]}. The largest county-level exposure study is
-likely that of Zandbergen (2009), which estimated exposure in US
-counties to all {[}US landfalling? or all storms?{]} Atlantic basin
-tropical storms between 1851 and 2003, using both distance and an
-exposure metric that incorporated distance and windspeed. They used this
-exposure evaluation to create maps of total exposure to tropical storms
-within US counties, as well as to explore associations between a
-county's long-term exposure to tropical storms and its location,
-distance from the coast, size, and shape (Zandbergen 2009).
+a study period that includes multiple storms. For example, Grabich et
+al. (2016) assessed the exposure of counties in Florida to land-falling
+hurricanes during the 2004 season. The largest-scale county-level
+exposure study is likely that of Zandbergen (2009), which estimated
+exposure in US counties to all {[}US landfalling? or all storms?{]}
+Atlantic basin tropical storms between 1851 and 2003, using both
+distance and an exposure metric that incorporated distance and
+windspeed. They used this exposure evaluation to create maps of total
+exposure to tropical storms within US counties, as well as to explore
+associations between a county's long-term exposure to tropical storms
+and its location, distance from the coast, size, and shape (Zandbergen
+2009).
+
+Other studies have used county-level exposure criteria to determine
+exposure to specific storms. For example, Grabich et al. (2015)
+determined county-level exposure to two severe storms in 2004 in Florida
+counties for an epidemiological study.
 
@@
 {[}check the paper-- was this distance to any part of a county, or to
-the center of the county?{]}.
+the center of the county?{]}. Another study used a distance criterion of
+60 kilometers from the storm track (Grabich et al. 2015) {[}again, check
+to see if this was to any point in the county or to some center point of
+the county{]}.
 
@@
 varied insignificantly from 30-75 km (Currie and Rossin-Slater 2013).
+
+Most distance exposure definitions in the past have been based on
+maximum wind radii based on maximum sustained winds (Kruk et al. 2010).
 
@@
 the storm and a categorical exposure metric based on the Saffir-Simpson
-storm severity categories (Grabich et al. 2016).
+storm severity categories (Grabich et al. 2016). Another study used a
+criterion of wind speed at or above 74 miles per hour (Grabich et al.
+2015).
 
@@
 used}\label{table-including-various-exposure-papers-and-the-parameters-used}
+
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+studies & distance\_km & rain\_mm & wind\_kt\tabularnewline
+\midrule
+\endhead
+Czajkowski 2010 & 120 & &\tabularnewline
+Rezempour 2014 & & 2/hr, 48/d &\tabularnewline
+Zandbergen 2009 & Buffer & & SSWS\tabularnewline
+Currie 2013 & 25, 30, 60, 75 & &\tabularnewline
+Grabich 2016 & 30, 60, 100 & & SSWS\tabularnewline
+Zaharan 2010 & 2 counties & &\tabularnewline
+Kruk 2010 & Radial Max Wind & & 34, 50, 64\tabularnewline
+Villarini 2011 & 100-500 & &\tabularnewline
+Atallah 2007 & & 200-500 (Floyd) &\tabularnewline
+\bottomrule
+\end{longtable}
+
+SSWS = Staffir-Simpson Hurricane Wind Scale (TS, 1, 2, 3, 4, and 5).
+Wind Cutoffs at 34, 64, 83, 96, 113, and 137 kt.
 
@@
 ``Measuring the Storm: Methods of Quantifying Hurricane Exposure with
-Pregnancy Outcomes.'' \emph{Natural Hazards Review} 17 (1):
-06015002--1--06015002--7.
+Pregnancy Outcomes.'' \emph{Natural Hazards Review} 17 (1): 06015002.
 doi:\href{https://doi.org/10.1061/(ASCE)NH.1527-6996.0000204}{10.1061/(ASCE)NH.1527-6996.0000204}.
+
+\hypertarget{ref-Grabich2015}{}
+Grabich, S. C., Whitney R. Robinson, Stephanie M. Engel, Charles E.
+Konrad, C. Konrad, and J. Horney. 2015. ``County-Level Hurricane
+Exposure and Birth Rates: Application of Difference-in-Differences
+Analysis for Confounding Control.'' \emph{Emerging Themes in
+Epidemiology} 12: 19.
+doi:\href{https://doi.org/10.1186/s12982-015-0042-7}{10.1186/s12982-015-0042-7}.
 

### 2016-06-17 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Edit paper draft

@@
   - name: Meilin Yan
-    email: alice@example.com
+    email: meilin.yan@colostate.edu
     affiliation: Colorado State University
   - name: Mohammad Z. Al-Hamdan
-    email: alice@example.com
+    email: mohammad.alhamdan@nasa.gov
     affiliation: NASA Marshall Space Flight Center
   - name: William L. Crosson
-    email: alice@example.com
+    email: bcrosson@usra.edu
     affiliation: NASA Marshall Space Flight Center
   - name: Seth Guikema
-    email: alice@example.com
+    email: sguikema@umich.edu
     affiliation: University of Michigan
   - name: Roger D. Peng
-    email: alice@example.com
+    email: rdpeng@jhu.edu
     affiliation: Johns Hopkins Bloomberg School of Public Health
@@
   - code: NASA Marshall Space Flight Center
-    address: Universities Space Research Association, Street, Huntsville, AL, Zip
+    address: Universities Space Research Association, 320 Sparkman Dr., Huntsville, AL, 35805
   - code: University of Michigan
-    address: Department, Street, City, State, Zip
+    address: Department of Industrial and Operations Engineering, 1205 Beal Ave., Ann Arbor, MI, 48109
   - code: Johns Hopkins Bloomberg School of Public Health
-    address: Department of Biostatistics, Street, Baltimore, MD, Zip
+    address: Department of Biostatistics, 615 North Wolfe Street, Baltimore, MD, 21205
 abstract: |
@@
 library(knitr)
+library(gridExtra)
 ```
@@
 
-Given its high resolution and consistancy with rain gage values, [@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
-However, the Stage IV dataset only provides data since 2001 [@Lin2005] and is therefore inadequate for use in this analysis.
+Given its better consistancy with rain gage values compared to NLDAS precipitation data [GBA-- you had something in here about resolution, as well.
+Does it have a better resolution than the NLDAS data?
+I think we probably want to focus on comparing this source with NLDAS, and we don't need to put much about advantages that it has that are very similar to NLDAS], [@Villarini2011] recommends Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
+However, the Stage IV dataset has only been archived since 2002 [@Lin2005] and cannot be used to analyze exposure to tropical storms for studies that include storms before 2002.
 
@@
 
-```{r echo = FALSE}
-library(hurricaneexposure)
-map_rain_exposure("Floyd-1999", rain_limit = 75, dist_limit = 500,
-                      days_included = -1:1)
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+a <- map_counties(storm = "Floyd-1999", metric = "rainfall")
+a <- map_tracks(storms = "Floyd-1999", plot_object = a, plot_points = FALSE)
+b <- map_counties(storm = "Floyd-1999", metric = "distance")
+b <- map_tracks(storms = "Floyd-1999", plot_object = b, plot_points = FALSE)
+grid.arrange(a, b, ncol = 1)
 ```
@@
 In comparison to the storm track, the areas of extreme rain and extreme wind can vary.
-For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
+For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track [@Halverson2015], while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
 The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference].
@@
 Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].
-
-For example,
-
-```{r echo = FALSE}
-map_rain_exposure("Floyd-1999", rain_limit = 75, dist_limit = 500,
-                      days_included = -1:1)
-```
 

### 2016-06-17 writing/DraftExposurePaper/DraftExposurePaper.tex
Edit paper draft

@@
     \author[Colorado State University]{Meilin Yan}
-   \ead{alice@example.com}
+   \ead{meilin.yan@colostate.edu}
 
     \author[NASA Marshall Space Flight Center]{Mohammad Z. Al-Hamdan}
-   \ead{alice@example.com}
+   \ead{mohammad.alhamdan@nasa.gov}
 
     \author[NASA Marshall Space Flight Center]{William L. Crosson}
-   \ead{alice@example.com}
+   \ead{bcrosson@usra.edu}
 
     \author[University of Michigan]{Seth Guikema}
-   \ead{alice@example.com}
+   \ead{sguikema@umich.edu}
 
     \author[Johns Hopkins Bloomberg School of Public Health]{Roger D. Peng}
-   \ead{alice@example.com}
+   \ead{rdpeng@jhu.edu}
 
@@
 Street, Fort Collins, CO, Zip}
-    \address[NASA Marshall Space Flight Center]{Universities Space Research Association, Street, Huntsville, AL, Zip}
-    \address[University of Michigan]{Department, Street, City, State, Zip}
-    \address[Johns Hopkins Bloomberg School of Public Health]{Department of Biostatistics, Street, Baltimore, MD, Zip}
+    \address[NASA Marshall Space Flight Center]{Universities Space Research Association, 320 Sparkman Dr., Huntsville,
+AL, 35805}
+    \address[University of Michigan]{Department of Industrial and Operations Engineering, 1205 Beal Ave., Ann
+Arbor, MI, 48109}
+    \address[Johns Hopkins Bloomberg School of Public Health]{Department of Biostatistics, 615 North Wolfe Street, Baltimore, MD,
+21205}
 
@@
 
-For example,
-
-\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-3-1.pdf}
-
 \paragraph{Wind-based exposure}\label{wind-based-exposure-1}

### 2016-06-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on paper draft

@@
 library(knitr)
+library(ggplot2)
 library(gridExtra)
@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates).
-Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009].
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000: direct hurricane-related deaths).
+Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000].
 Some studies suggest methods to determine exposure, including inland exposure, to tropical storms at the county-level over many years [@Zandbergen2009].
@@
 Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015].
-Such missclassification can have important implications of assessments of the public health and economic impacts of tropical storms.
+Such missclassification can have important consequences in assessments of the public health and economic impacts of tropical storms.
 
@@
 For example, daily counts of human health outcomes in environmental epidemiology studies are often available aggregated at the county level, and such data has often been paired with time series of environmental exposures (e.g., air pollution, temperature) to determine associated risks.
+Databases of storm impacts, including direct deaths [@Rappaport2000], ..., and ..., are often aggregated at the county level.
 
@@
 We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database.
-This dataset is based on a poststorm assessment of each storm conducted by the United States National Hurricane Center (NHC) and incorporates data from a variety of sources, including satellite data and, when available, aircraft reconnaisance data [@Landsea2013].
+This dataset is based on a post-storm assessment of each storm conducted by the United States National Hurricane Center (NHC) and incorporates data from a variety of sources, including satellite data and, when available, aircraft reconnaisance data [@Landsea2013].
 This data gives time stamps for each observation in Coordinated Universal Time (UTC; also known as Zulu Time, sometimes indicated by "Z").
@@
 From these distance calculations, we identified the date-time and distance between the storm track and the county center at the 15-minute interval when the storm was closest to the county center.
-We used the `countytimezones` R package to convert each date-time of the closest storm approach for a county to the county's local time zone and identified the local date when the storm was closest to the county.
-This "closest date" was used to pair the distance estimates with rainfall estimates.
+We identified the date when the storm was closest to each county and converted this date-time  to the county's local time zone using the `countytimezones` R package [citation].
+This "closest date" was used to pair the distance estimates with rainfall estimates and can also be used to merge exposure datasets with daily aggregated outcomes like daily county-level mortality and morbidity counts.
 
@@
 
-NLDAS does not provide information on precipitation over the ocean, preventing its use in assessing the develpment of the storm prior to landfall [@Villarini2011].
+NLDAS does not provide information on precipitation over the ocean, preventing its use in assessing the development of the storm prior to landfall [@Villarini2011].
 This point is moot for this analysis.
@@
 
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-a <- map_counties(storm = "Floyd-1999", metric = "rainfall")
-a <- map_tracks(storms = "Floyd-1999", plot_object = a, plot_points = FALSE)
-b <- map_counties(storm = "Floyd-1999", metric = "distance")
-b <- map_tracks(storms = "Floyd-1999", plot_object = b, plot_points = FALSE)
+#### Are exposure patterns different when using distance-, rain-, and wind-based metrics?
+
+One interesting contrast is between storms that were noteworthy for direct wind-related deaths (Anderew in 1992, Hugo in 1989) versus storms noteworthy for direct freshwater flooding-related deaths (Floyd in 1999, Alberto in 1994, Charley in 1998) for the 1970--1999 period [@Rappaport2000].
+Here are two examples, one of each of these two types:
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+storm <- "Hugo-1989"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "distance")) + 
+        ggtitle(paste(storm, "distance"))
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle(paste(storm, "rainfall"))
+
 grid.arrange(a, b, ncol = 1)
 ```
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+storm <- "Floyd-1999"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "distance")) + 
+        ggtitle(paste(storm, "distance"))
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle(paste(storm, "rainfall"))
+
+grid.arrange(a, b, ncol = 1)
+```
+
+#### Which storms exposed the most counties when using distance-, rain-, and wind-based metrics? What were the top county-storm exposures for rain- and wind-based metrics?
+
+#### Are storms with heavy rain exposure likely to occur early in the hurricane season?
 
@@
 Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could misclassify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).
-In particular, the forces of a storm tend to be distributed around the center in a non-symmetrical way [citation], while distance-based exposure metrics that use buffers tend to use an equal buffer distance on each side of the storm track (e.g., [citations]).
+While a number of storm characteristics are strongly associated with distance from the storm's center (e.g., wind and, at the coast, storm surge and waves; @Rappaport2000), other characteristics like rain can occur at dangerous levels well away from the storm's central track [@Rappaport2000].
+Although storm surge was a major cause of hurricane-related deaths in the past, more recently freshwater flooding has become a much more common cause of deaths directly resulting from tropical storms in the US [@Rappaport2000].
+
+Further, the forces of a storm tend to be distributed around the center in a non-symmetrical way [citation], while distance-based exposure metrics that use buffers tend to use an equal buffer distance on each side of the storm track (e.g., [citations]).
 
@@
 
+Rain from tropical storms can bring important risks to human health and property, even in counties well inland from the coast.
+For example, most of the direct hurricane-related deaths in the US from 1970 to 1999 were caused by freshwater flooding and were in inland, rather than coastal, counties [@Rappaport2000].
+Rains from tropical storms can flood roads [@Rappaport2000].
+
 It can be very difficult to reliably measure rain during extreme rain events, including tropical storms.
@@
 Often, storm-related deaths are associated with inland flooding.
-In fact, over half of the hurricane-related fatalities from 1970 to 1999 were a result of freshwater flooding, accounting for the vast majority of deaths in inland counties [@Rappaport2000].
+In the United States, over half of hurricane-related direct deaths from 1970 to 1999 from Atlantic basin storms were a result of freshwater flooding; as a result, most of these deaths were in inland, rather than coastal, counties [@Rappaport2000].
 In another analysis, researchers found that 79\% of freshwater-drowning fatalities occured in inland counties,[@Czajkowski2011] further stressing the importance of rainfall in hurricane risk analysis.
@@
 
+Several storm characteristics are associated with the likelihood of heavy rain and flooding from the storm, including the storm's forward speed and size [@Rappaport2000].
+There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season, when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000].
+
 #### Wind-based exposure
+
+A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000].
+While severe winds pose an important threat, risks of a tropical storm can exist without severe wind; for example, one study found that most of the direct hurricane-related deaths in the US between 1970 and 1999 occurred in cases when wind was below hurricane strength (i.e., < 64 knots) [@Rappaport2000].
 

### 2016-06-18 writing/DraftExposurePaper/DraftExposurePaper.tex
Work on paper draft

@@
 
-Given its high resolution and consistancy with rain gage values,
-(Villarini et al. 2011) recommends the use of Stage IV precipitation
-data for the analysis of rainfall distribution in landfalling tropical
-cyclones. However, the Stage IV dataset only provides data since 2001
-(Lin and Mitchell 2005) and is therefore inadequate for use in this
-analysis.
+Given its better consistancy with rain gage values compared to NLDAS
+precipitation data {[}GBA-- you had something in here about resolution,
+as well. Does it have a better resolution than the NLDAS data? I think
+we probably want to focus on comparing this source with NLDAS, and we
+don't need to put much about advantages that it has that are very
+similar to NLDAS{]}, (Villarini et al. 2011) recommends Stage IV
+precipitation data for the analysis of rainfall distribution in
+landfalling tropical cyclones. However, the Stage IV dataset has only
+been archived since 2002 (Lin and Mitchell 2005) and cannot be used to
+analyze exposure to tropical storms for studies that include storms
+before 2002.
 
@@
 
+One interesting contrast is between storms that were noteworthy for
+direct wind-related deaths (Anderew in 1992, Hugo in 1989) versus storms
+noteworthy for direct freshwater flooding-related deaths (Floyd in 1999,
+Alberto in 1994, Charley in 1998) for the 1970--1999 period (Rappaport
+2000). Here are two examples, one of each of these two types:
+
 \includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-3-1.pdf}
 
@@
 wind can vary. For example, storms undergoing an extratropical
-transition can bring heavy rains to the left of the storm's center
-track, while extreme winds are more common to the right of the track
-(Halverson 2015, Grabich et al. (2016) GBA-- I think it would be better
-to just use meteorological papers as references for this point. The
-Halverson paper is great, but the Grabich paper I think might just be a
-secondary, rather than primary, reference for this point, so I think we
-could probably find a more direct reference). Wind speeds are typically
-more severe to the right of the storm's track because here the
+transition can bring heavy rains to the left of the storm's center track
+(Halverson 2015), while extreme winds are more common to the right of
+the track (Halverson 2015, Grabich et al. (2016) GBA-- I think it would
+be better to just use meteorological papers as references for this
+point. The Halverson paper is great, but the Grabich paper I think might
+just be a secondary, rather than primary, reference for this point, so I
+think we could probably find a more direct reference). Wind speeds are
+typically more severe to the right of the storm's track because here the
 counter-clockwise cyclonic {[}right word?{]} winds are moving in the

### 2016-06-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Start adding more research questions for us to answer
I'm adding these in Results as I think of them. We can add more, think about whether the ones we have are useful, and think about how to answer ones that we think are useful
Add some results

@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, message = FALSE, warning = FALSE}
 library(hurricaneexposure)
@@
 library(gridExtra)
+library(dplyr)
 ```
@@
 
+Tropical storms bring a number of threats, including high winds, heavy rain and flooding, storm surge, tornadoes, ...
+.
+To determine whether a county is exposed to a particular tropical storm, a researcher could use criteria based on measurements of any of these values during the storm, use a metric based on the distance from the county to the storm's track, or use a metric that combines several of these measurements.
+
+While the risks to life from some tropical storm threats (e.g., storm surge) have been greatly reduced over the past century, other risks (e.g., inland flooding) have remained more steady [@Rappaport2000].
+Determining the metrics of storm exposure that are most associated with loss of life and property damage may help identify and emphasize the important threats that remain from tropical storms in the United States, which could allow similar future success in reducing these threats through community planning, warning systems, etc.
+
 #### Why it's important to measure TS exposure for counties
@@
 Some studies suggest methods to determine exposure, including inland exposure, to tropical storms at the county-level over many years [@Zandbergen2009].
+
 Here, we offer tools, through the open-source `hurricaneexposure` package, to determine county-level exposure to tropical storms using several different metrics (distance, rainfall, wind).
@@
 
-Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009].
+Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009; @Czajkowski2011].
 Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.
@@
 #### Are exposure patterns different when using distance-, rain-, and wind-based metrics?
+In particular, (1) do some exposure metrics result in more "exposed" classifications than others and (2) are geographic patterns of exposure different among the different metrics?
+
+Map of county exposures to the top 1000 county-storm pairs by distance-based exposure:
+
+```{r echo = FALSE}
+# Code for this map here
+```
+
+Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:
+
+```{r echo = FALSE}
+# Code for this map here
+```
+
+Map of county exposures to the top 1000 county-storm pairs by rain-based exposure:
+
+```{r echo = FALSE}
+# Code for this map here
+```
+
+Map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this? What seems reasonable?] kilometers or less:
+
+```{r echo = FALSE}
+# Code for this map here
+```
+
+Map of total count of "exposures" by county, based on "exposed" = rainfall over a three-day window of 75 millimeters or more and distance of 500 kilometers or less:
+
+```{r echo = FALSE}
+# Code for this map here
+```
 
@@
 
+Here is a table of the top 10 storms by number of counties exposed, based on a distance metric of within 100 kilometers or less of the storm track:
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+data("closest_dist")
+all_counties <- unique(closest_dist$fips)
+
+all_distance <- county_distance(counties = all_counties, start_year = 1988, 
+                        end_year = 2011, dist_limit = 60)
+
+distance_exposed <- all_distance %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
+distance_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of distance.")
+```
+
+Here are plots of the top four storms based on number of counties exposed by distance:
+
+```{r echo = FALSE, fig.width = 10, fig.height = 8}
+top_distance_exposed <- function(storm_id){
+        my_map <- map_distance_exposure(storm_id, dist_limit = 60)
+        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
+                             plot_points = FALSE) + 
+                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
+}
+
+n <- 4
+distance_plots <- lapply(distance_exposed$storm_id[1:n], top_distance_exposed)
+names(distance_plots) <- letters[1:n]
+grid.arrange(distance_plots$a, distance_plots$b, distance_plots$c,
+             distance_plots$d, ncol = 2)
+```
+
+Here is a table of the top 10 storms by number of counties exposed, based on a rain metric of 75 millimeters or more and within 500 kilometers or less of the storm track:
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+data("closest_dist")
+all_counties <- unique(closest_dist$fips)
+
+all_rain <- county_rain(counties = all_counties, start_year = 1988, 
+                        end_year = 2011, rain_limit = 75, dist_limit = 500,
+                        days_included = -1:1)
+
+rain_exposed <- all_rain %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
+rain_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of rainfall.")
+```
+
+Here are plots of the exposed counties for the top 4 storms based on that rain metric:
+
+```{r echo = FALSE, fig.width = 10, fig.height = 8}
+top_rain_exposed <- function(storm_id){
+        my_map <- map_rain_exposure(storm_id, rain_limit = 75, 
+                                    dist_limit = 500)
+        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
+                             plot_points = FALSE) + 
+                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
+}
+
+n <- 4
+rain_plots <- lapply(rain_exposed$storm_id[1:n], top_rain_exposed)
+names(rain_plots) <- letters[1:n]
+grid.arrange(rain_plots$a, rain_plots$b, rain_plots$c, rain_plots$d, ncol = 2)
+```
+
 #### Are storms with heavy rain exposure likely to occur early in the hurricane season?
@@
 Many of the uncertainties related to storm positions in best track data are more of a concern for the years before ...
-(e.g., pre-19[xx]).
+(e.g., pre-19[xx], before aircraft reconnaisance of Atlantic basin tropical storms).
 
@@
 Rains from tropical storms can flood roads [@Rappaport2000].
+Examples of historical tropical storms that caused many direct deaths from flooding in inland counties include Hurricane Diane in 1955 and Hurricane Camille in 1969 [@Rappaport2000]; more recent examples include Hurricane Floyd in 1999, Hurricane Alberto in 1994, Hurricane Charley in 1998, and Hurricane Fran in 1996 [@Rappaport2000].
 
@@
 Several storm characteristics are associated with the likelihood of heavy rain and flooding from the storm, including the storm's forward speed and size [@Rappaport2000].
-There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season, when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000].
+There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season (e.g., June--August), when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000].
 
@@
 A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000].
-While severe winds pose an important threat, risks of a tropical storm can exist without severe wind; for example, one study found that most of the direct hurricane-related deaths in the US between 1970 and 1999 occurred in cases when wind was below hurricane strength (i.e., < 64 knots) [@Rappaport2000].
+These types of wind-related damage can occur when winds are below hurricane strength (i.e., < 64 knots) [@Rappaport2000].
+While severe winds pose an important threat, risks of a tropical storm can exist without severe wind; for example, one study found that most of the direct hurricane-related deaths in the US between 1970 and 1999 occurred in cases when wind was below hurricane strength, including for Tropical Storms Charley in 1998 and Alberto in 1994 [@Rappaport2000].
+Further, wind-based exposure metrics are directly linked to a storm's strength, and some research suggests that hurricane strength and landfall is not strongly associated with the number of direct deaths ultimately caused by the storm [@Rappaport2000].
 

### 2016-06-18 writing/DraftExposurePaper/DraftExposurePaper.tex
Add some results

@@
 
+Tropical storms bring a number of threats, including high winds, heavy
+rain and flooding, storm surge, tornadoes, \ldots{} . To determine
+whether a county is exposed to a particular tropical storm, a researcher
+could use criteria based on measurements of any of these values during
+the storm, use a metric based on the distance from the county to the
+storm's track, or use a metric that combines several of these
+measurements.
+
+While the risks to life from some tropical storm threats (e.g., storm
+surge) have been greatly reduced over the past century, other risks
+(e.g., inland flooding) have remained more steady (Rappaport 2000).
+Determining the metrics of storm exposure that are most associated with
+loss of life and property damage may help identify and emphasize the
+important threats that remain from tropical storms in the United States,
+which could allow similar future success in reducing these threats
+through community planning, warning systems, etc.
+
 \paragraph{Why it's important to measure TS exposure for
@@
 (e.g., counts of health outcomes: Grabich et al. (2016): premature
-births and low birth weights; Grabich et al. (2015): birth rates).
-Further, decisions and policies to prepare for and respond to storms are
-often undertaken at the county level (Zandbergen 2009). Some studies
+births and low birth weights; Grabich et al. (2015): birth rates;
+Rappaport (2000): direct hurricane-related deaths). Further, decisions
+and policies to prepare for and respond to storms are often undertaken
+at the county level (Zandbergen 2009; Rappaport 2000). Some studies
 suggest methods to determine exposure, including inland exposure, to
 tropical storms at the county-level over many years (Zandbergen 2009).
+
 Here, we offer tools, through the open-source \texttt{hurricaneexposure}
@@
 al. 2016; Grabich et al. 2015). Such missclassification can have
-important implications of assessments of the public health and economic
+important consequences in assessments of the public health and economic
 impacts of tropical storms.
@@
 the US have used geographical information system software (e.g., ArcGIS)
-(example studies: Grabich et al. 2016; Zandbergen 2009). Here, we offer
-methods to map and output historic exposure to tropical storms that does
-not require the use of proprietary software but instead uses a package
-written in the R statistical programming language, which is free and
-open-source.
+(example studies: Grabich et al. 2016; Zandbergen 2009; Czajkowski,
+Simmons, and Sutter 2011). Here, we offer methods to map and output
+historic exposure to tropical storms that does not require the use of
+proprietary software but instead uses a package written in the R
+statistical programming language, which is free and open-source.
 
@@
 paired with time series of environmental exposures (e.g., air pollution,
-temperature) to determine associated risks.
+temperature) to determine associated risks. Databases of storm impacts,
+including direct deaths (Rappaport 2000), \ldots{}, and \ldots{}, are
+often aggregated at the county level.
 
@@
 storms between 1988 and 2014 from the extended best tracks database.
-This dataset is based on a poststorm assessment of each storm conducted
+This dataset is based on a post-storm assessment of each storm conducted
 by the United States National Hurricane Center (NHC) and incorporates
@@
 distance between the storm track and the county center at the 15-minute
-interval when the storm was closest to the county center. We used the
-\texttt{countytimezones} R package to convert each date-time of the
-closest storm approach for a county to the county's local time zone and
-identified the local date when the storm was closest to the county. This
-``closest date'' was used to pair the distance estimates with rainfall
-estimates.
+interval when the storm was closest to the county center. We identified
+the date when the storm was closest to each county and converted this
+date-time to the county's local time zone using the
+\texttt{countytimezones} R package {[}citation{]}. This ``closest date''
+was used to pair the distance estimates with rainfall estimates and can
+also be used to merge exposure datasets with daily aggregated outcomes
+like daily county-level mortality and morbidity counts.
 
@@
 NLDAS does not provide information on precipitation over the ocean,
-preventing its use in assessing the develpment of the storm prior to
+preventing its use in assessing the development of the storm prior to
 landfall (Villarini et al. 2011). This point is moot for this analysis.
@@
 
+\paragraph{\texorpdfstring{Are exposure patterns different when using
+distance-, rain-, and wind-based metrics? In particular, (1) do some
+exposure metrics result in more ``exposed'' classifications than others
+and (2) are geographic patterns of exposure different among the
+different
+metrics?}{Are exposure patterns different when using distance-, rain-, and wind-based metrics?
+In particular, (1) do some exposure metrics result in more exposed classifications than others and (2) are geographic patterns of exposure different among the different metrics?}}\label{are-exposure-patterns-different-when-using-distance--rain--and-wind-based-metrics-in-particular-1-do-some-exposure-metrics-result-in-more-exposed-classifications-than-others-and-2-are-geographic-patterns-of-exposure-different-among-the-different-metrics}
+
+Map of county exposures to the top 1000 county-storm pairs by
+distance-based exposure:
+
+Map of county exposures to the top 1000 county-storm pairs by wind-based
+exposure:
+
+Map of county exposures to the top 1000 county-storm pairs by rain-based
+exposure:
+
+Map of total count of ``exposures'' by county, based on ``exposed'' =
+distance of 100 {[}or do we want a different distance limit for this?
+What seems reasonable?{]} kilometers or less:
+
+Map of total count of ``exposures'' by county, based on ``exposed'' =
+rainfall over a three-day window of 75 millimeters or more and distance
+of 500 kilometers or less:
+
 One interesting contrast is between storms that were noteworthy for
@@
 
-\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-7-1.pdf}
 
-\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\paragraph{Which storms exposed the most counties when using distance-,
+rain-, and wind-based metrics? What were the top county-storm exposures
+for rain- and wind-based
+metrics?}\label{which-storms-exposed-the-most-counties-when-using-distance--rain--and-wind-based-metrics-what-were-the-top-county-storm-exposures-for-rain--and-wind-based-metrics}
+
+Here is a table of the top 10 storms by number of counties exposed,
+based on a distance metric of within 100 kilometers or less of the storm
+track:
+
+\begin{longtable}[]{@{}lr@{}}
+\caption{Top 10 storms based on the number of counties exposed, using a
+metric of distance.}\tabularnewline
+\toprule
+Storm & \# of exposed counties\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Storm & \# of exposed counties\tabularnewline
+\midrule
+\endhead
+Beryl (1994) & 195\tabularnewline
+Frances (2004) & 194\tabularnewline
+Cindy (2005) & 189\tabularnewline
+Dennis (2005) & 187\tabularnewline
+Chris (1988) & 181\tabularnewline
+Erin (1995) & 175\tabularnewline
+Allison (2001) & 170\tabularnewline
+Isidore (2002) & 167\tabularnewline
+Jeanne (2004) & 167\tabularnewline
+Ivan (2004) & 159\tabularnewline
+\bottomrule
+\end{longtable}
+
+Here are plots of the top four storms based on number of counties
+exposed by distance:
+
+\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+Here is a table of the top 10 storms by number of counties exposed,
+based on a rain metric of 75 millimeters or more and within 500
+kilometers or less of the storm track:
+
+\begin{longtable}[]{@{}lr@{}}
+\caption{Top 10 storms based on the number of counties exposed, using a
+metric of rainfall.}\tabularnewline
+\toprule
+Storm & \# of exposed counties\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Storm & \# of exposed counties\tabularnewline
+\midrule
+\endhead
+Frances (2004) & 454\tabularnewline
+Ivan (2004) & 405\tabularnewline
+Isidore (2002) & 307\tabularnewline
+Floyd (1999) & 294\tabularnewline
+Lee (2011) & 279\tabularnewline
+Opal (1995) & 266\tabularnewline
+Jeanne (2004) & 255\tabularnewline
+Gustav (2008) & 252\tabularnewline
+Ike (2008) & 236\tabularnewline
+Katrina (2005) & 232\tabularnewline
+\bottomrule
+\end{longtable}
+
+Here are plots of the exposed counties for the top 4 storms based on
+that rain metric:
+
+\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-12-1.pdf}
+
+\paragraph{Are storms with heavy rain exposure likely to occur early in
+the hurricane
+season?}\label{are-storms-with-heavy-rain-exposure-likely-to-occur-early-in-the-hurricane-season}
 
@@
 storm positions in best track data are more of a concern for the years
-before \ldots{} (e.g., pre-19{[}xx{]}).
+before \ldots{} (e.g., pre-19{[}xx{]}, before aircraft reconnaisance of
+Atlantic basin tropical storms).
 
@@
 mild or small-radius storms) and false negatives (e.g., during very
-large or intense storms). In particular, the forces of a storm tend to
-be distributed around the center in a non-symmetrical way
-{[}citation{]}, while distance-based exposure metrics that use buffers
-tend to use an equal buffer distance on each side of the storm track
-(e.g., {[}citations{]}).
+large or intense storms). While a number of storm characteristics are
+strongly associated with distance from the storm's center (e.g., wind
+and, at the coast, storm surge and waves; Rappaport (2000)), other
+characteristics like rain can occur at dangerous levels well away from
+the storm's central track (Rappaport 2000). Although storm surge was a
+major cause of hurricane-related deaths in the past, more recently
+freshwater flooding has become a much more common cause of deaths
+directly resulting from tropical storms in the US (Rappaport 2000).
+
+Further, the forces of a storm tend to be distributed around the center
+in a non-symmetrical way {[}citation{]}, while distance-based exposure
+metrics that use buffers tend to use an equal buffer distance on each
+side of the storm track (e.g., {[}citations{]}).
 
@@
 
+Rain from tropical storms can bring important risks to human health and
+property, even in counties well inland from the coast. For example, most
+of the direct hurricane-related deaths in the US from 1970 to 1999 were
+caused by freshwater flooding and were in inland, rather than coastal,
+counties (Rappaport 2000). Rains from tropical storms can flood roads
+(Rappaport 2000). Examples of historical tropical storms that caused
+many direct deaths from flooding in inland counties include Hurricane
+Diane in 1955 and Hurricane Camille in 1969 (Rappaport 2000); more
+recent examples include Hurricane Floyd in 1999, Hurricane Alberto in
+1994, Hurricane Charley in 1998, and Hurricane Fran in 1996 (Rappaport
+2000).
+
 It can be very difficult to reliably measure rain during extreme rain
@@
 during the storm. Often, storm-related deaths are associated with inland
-flooding. In fact, over half of the hurricane-related fatalities from
-1970 to 1999 were a result of freshwater flooding, accounting for the
-vast majority of deaths in inland counties (Rappaport 2000). In another
-analysis, researchers found that 79\% of freshwater-drowning fatalities
-occured in inland counties,(Czajkowski, Simmons, and Sutter 2011)
-further stressing the importance of rainfall in hurricane risk analysis.
-Storms can produce a lot of rain especially in certain topographies,
-like near mountains, so counties that are well inland can sometimes
-experience more extreme rain that other counties at similar distances
-from the storm's track.
+flooding. In the United States, over half of hurricane-related direct
+deaths from 1970 to 1999 from Atlantic basin storms were a result of
+freshwater flooding; as a result, most of these deaths were in inland,
+rather than coastal, counties (Rappaport 2000). In another analysis,
+researchers found that 79\% of freshwater-drowning fatalities occured in
+inland counties,(Czajkowski, Simmons, and Sutter 2011) further stressing
+the importance of rainfall in hurricane risk analysis. Storms can
+produce a lot of rain especially in certain topographies, like near
+mountains, so counties that are well inland can sometimes experience
+more extreme rain that other counties at similar distances from the
+storm's track.
 
@@
 
+Several storm characteristics are associated with the likelihood of
+heavy rain and flooding from the storm, including the storm's forward
+speed and size (Rappaport 2000). There is some evidence that the worst
+tropical storms, in terms of heavy rainfall and freshwater flood-related
+deaths in the US, tend to occur early in the hurricane season (e.g.,
+June--August), when it is more common to get the meteorological
+conditions (weak steering currents) that lead to slow-moving storms
+(Rappaport 2000).
+
 \paragraph{Wind-based exposure}\label{wind-based-exposure-1}
+
+A tropical storm's high winds can bring a number of dangers, including
+health risks and property damage caused by structural damage of houses
+and other buildings, falling trees, and wind-borne debris (Rappaport
+2000). These types of wind-related damage can occur when winds are below
+hurricane strength (i.e., \textless{} 64 knots) (Rappaport 2000). While
+severe winds pose an important threat, risks of a tropical storm can
+exist without severe wind; for example, one study found that most of the
+direct hurricane-related deaths in the US between 1970 and 1999 occurred
+in cases when wind was below hurricane strength, including for Tropical
+Storms Charley in 1998 and Alberto in 1994 (Rappaport 2000). Further,
+wind-based exposure metrics are directly linked to a storm's strength,
+and some research suggests that hurricane strength and landfall is not
+strongly associated with the number of direct deaths ultimately caused
+by the storm (Rappaport 2000).
 

### 2016-06-18 R/rain_exposure.R

@@ -1,6 +1,6 @@
 #' Hurricane exposure by rain for counties
 #'
-#' #' This function takes a list of US counties,based on their 5-digit Federal
+#' This function takes a list of US counties,based on their 5-digit Federal
 #' Information Processing Standard (FIPS) codes, boundaries on
 #' the range of years to be considered, thresholds for distance between each
 #' county and the storm track as well as minimum rainfall over a certain time

### 2016-06-19 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Update paper draft

@@
 all_distance <- county_distance(counties = all_counties, start_year = 1988, 
-                        end_year = 2011, dist_limit = 60)
+                        end_year = 2011, dist_limit =100)
 
@@
 top_distance_exposed <- function(storm_id){
-        my_map <- map_distance_exposure(storm_id, dist_limit = 60)
+        my_map <- map_distance_exposure(storm_id, dist_limit = 100)
         my_map <- map_tracks(storm = storm_id, plot_object = my_map, 

### 2016-06-19 writing/DraftExposurePaper/DraftExposurePaper.tex
Update paper draft

@@
 \endhead
-Beryl (1994) & 195\tabularnewline
-Frances (2004) & 194\tabularnewline
-Cindy (2005) & 189\tabularnewline
-Dennis (2005) & 187\tabularnewline
-Chris (1988) & 181\tabularnewline
-Erin (1995) & 175\tabularnewline
-Allison (2001) & 170\tabularnewline
-Isidore (2002) & 167\tabularnewline
-Jeanne (2004) & 167\tabularnewline
-Ivan (2004) & 159\tabularnewline
+Beryl (1994) & 346\tabularnewline
+Frances (2004) & 321\tabularnewline
+Cindy (2005) & 317\tabularnewline
+Jeanne (2004) & 296\tabularnewline
+Dennis (2005) & 290\tabularnewline
+Chris (1988) & 289\tabularnewline
+Allison (2001) & 283\tabularnewline
+Ivan (2004) & 283\tabularnewline
+Erin (1995) & 280\tabularnewline
+Isidore (2002) & 267\tabularnewline
 \bottomrule

### 2016-06-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Update to use hurricaneexposuredata package

@@
 library(hurricaneexposure)
+# install hurricane exposure data package if you don't already have it
+# library(devtools)
+# install_github("geanders/hurricaneexposuredata")
+library(hurricaneexposuredata)
 library(knitr)

### 2016-06-20 README.Rmd
Fix `check` warnings and notes
Update to use hurricaneexposuredata package

@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+# If you need to install the hurricaneexposuredata package: 
+# library(devtools)
+# install_github("geanders/hurricaneexposuredata")
+library(hurricaneexposuredata)
 library(hurricaneexposure)
@@
 
-The package currently exists in [a development version](https://github.com/geanders/hurricaneexposure) on GitHub. You can use the following code to load it:
+The package currently exists in [a development version](https://github.com/geanders/hurricaneexposure) on GitHub, and all the associated data is in its own [development version package](https://github.com/geanders/hurricaneexposuredata), also on GitHub.
+You can use the following code to load both:
 
@@
 library(devtools)
+install_github("geanders/hurricaneexposuredata")
 install_github("geanders/hurricaneexposure", 
                build_vignettes = TRUE)
+library(hurricaneexposuredata)
 library(hurricaneexposure)
@@
 
-```{r fig.width = 6.5, fig.height = 4}
+```{r fig.width = 5.5, fig.height = 4}
 map_4 <- map_tracks(storms = "Floyd-1999")

### 2016-06-20 R/distance_exposure.R

@@ -99,7 +99,7 @@ multi_county_distance <- function(communities, start_year, end_year,
 
         communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
 
-        distance_df <- hurricaneexposure::closest_dist %>%
+        distance_df <- hurricaneexposuredata::closest_dist %>%
                 dplyr::mutate_(closest_date =
                                        ~ lubridate::ymd(closest_date)) %>%
                 dplyr::filter_(~ fips %in% communities$fips &

### 2016-06-20 R/filter_storm_data.R

@@ -44,7 +44,7 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
                               include_rain = FALSE, days_included = NULL,
                               output_vars = c("fips")){
 
-        closest_dist <- data.table::data.table(hurricaneexposure::closest_dist)
+        closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
 
         if(!is.null(counties)){
                 closest_dist <- closest_dist[get("fips") %in% counties]
@@ -73,7 +73,7 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
         }
 
         if(include_rain){
-                rain <- data.table::data.table(hurricaneexposure::rain)
+                rain <- data.table::data.table(hurricaneexposuredata::rain)
                 rain <- rain[get("lag") %in% days_included]
                 rain <- rain[ , .(tot_precip = sum(get("precip"))),
                               by = .(fips, storm_id)]

### 2016-06-20 R/map_exposure.R

@@ -86,7 +86,7 @@ map_tracks <- function(storms, plot_object = NULL,
         map_dim <- apply(map_data[ , c("long", "lat")],
                          MARGIN = 2,
                          function(x) range(x) + c(-1, 1) * padding)
-        tracks <- hurricaneexposure::hurr_tracks %>%
+        tracks <- hurricaneexposuredata::hurr_tracks %>%
                 dplyr::select_(~ latitude, ~ longitude, ~ storm_id) %>%
                 dplyr::filter_(~ as.character(storm_id) %in% storms &
                               longitude > map_dim[1, 1] &

### 2016-06-20 R/rain_exposure.R

@@ -175,13 +175,13 @@ multi_county_rain <- function(communities, start_year, end_year,
         dots <- stats::setNames(list(lazyeval::interp(~ lubridate::ymd(x),
                                                       x = quote(closest_date))),
                                 "closest_date")
-        rain_storm_df <- hurricaneexposure::closest_dist %>%
+        rain_storm_df <- hurricaneexposuredata::closest_dist %>%
                 dplyr::mutate_(.dots = dots) %>%
                 dplyr::filter_(~ fips %in% communities$fips &
                                       lubridate::year(closest_date) >= start_year &
                                       lubridate::year(closest_date) <= end_year) %>%
                 dplyr::left_join(communities, by = "fips") %>%
-                dplyr::left_join(hurricaneexposure::rain,
+                dplyr::left_join(hurricaneexposuredata::rain,
                                  by = c("storm_id", "fips")) %>%
                 dplyr::filter_(~ lag %in% days_included) %>%
                 dplyr::group_by_(~ storm_id, ~ fips) %>%

### 2016-06-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on paper draft

@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000: direct hurricane-related deaths).
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
 Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000].
@@
 For example, @Grabich2015 determined county-level exposure to two severe storms in 2004 in Florida counties for an epidemiological study.
+
+One of the challenges of trying to assess exposure to multiple storms is that sometimes storms can closely follow each other, which can make it difficult to tease apart separate exposures to different storms in a systematic way.
+For example, when assessing streamflow responses to tropical storms, it can be difficult to separate delayed responses to assign impacts to separate storms [@Chen2015].
 
@@
 #### Rain-based exposure
+
+In this study, we use precipitation data from the North American Land Data Assimilation System, phase 2 (NLDAS-2) to characterize rain-based exposure to tropical storms.
+The NLDAS-2 data integrates satellite-based and land-based monitoring and applies a land-surface model to create a reanalysis dataset that is spatially and temporally complete across the continental United States.
+This data has been used in previous research to investigate tropical storms, including in a study of the relationship between rainfall and streamflow responses for tropical storms in a North Carolina watershed [@Chen2015].
+
+The NLDAS-2 precipitation data is originally provided hourly for a 1/8 degree grid.
+To generate county-level daily rainfall estimates, we first aggregated data at each grid point, after converting the timestamp of each observation to local time, to create a daily estimate at the grid point.
+We then averaged all grid points within a county's boundaries to generate a county-level average of daily precipitation.
 
@@
 Tropical storms vary in size and intensity, and a measure of distance from the storm track will not incorporate these differences and so could misclassify exposure both in terms of generating false positives (counties close to the storm track of very mild or small-radius storms) and false negatives (e.g., during very large or intense storms).
+This point is especially important given the dramatic range possible in the size of storms; US storms have been observed with radii to maximum winds as small as 20 kilometers to over 200 kilometers [@Mallin2006].
+
 While a number of storm characteristics are strongly associated with distance from the storm's center (e.g., wind and, at the coast, storm surge and waves; @Rappaport2000), other characteristics like rain can occur at dangerous levels well away from the storm's central track [@Rappaport2000].
 Although storm surge was a major cause of hurricane-related deaths in the past, more recently freshwater flooding has become a much more common cause of deaths directly resulting from tropical storms in the US [@Rappaport2000].
+Further, fatal tropical storm tornadoes most often occur between 200 kilometers and 500 kilometers from the storm's center, and are more common as a storm decays inland than at landfall, and were linked to over 300 deaths in the US between 1995 and 2009 [@Moore2012].
 
@@
 
+Most distance-based exposure assessments have used equal distance criteria on either side of the storm's track to classify exposure.
+However, the hazards associated with a storm can display marked asymmetry.
+For example, fatal tornadoes associated with US tropical storms between 1995 and 2009 occurred almost exclusively to the right of the storm's track, mostly in the right-front quandrant (northeast) of the storm [@Moore2012].
+
 #### Rain-based exposure
@@
 Examples of historical tropical storms that caused many direct deaths from flooding in inland counties include Hurricane Diane in 1955 and Hurricane Camille in 1969 [@Rappaport2000]; more recent examples include Hurricane Floyd in 1999, Hurricane Alberto in 1994, Hurricane Charley in 1998, and Hurricane Fran in 1996 [@Rappaport2000].
+Storms notable for heavy rains include Allison in 2001, Alberto in 1994, and Claudette in 1979.
 
 It can be very difficult to reliably measure rain during extreme rain events, including tropical storms.
-For example, a heavy rain can wash away [?] rain monitors [?].
-It can also be very hard to measure rain during heavy wind, as the rain does not fall straight into the monitor [?].
+For example, a heavy rain can wash away [?] rain gauges [?].
+It can also be very hard to measure rain during heavy wind, as the rain does not fall straight into the gauge [@Medlin2007].
+Bias in tipping-bucket rain gauges during tropical storms are caused both by high winds, with higher bias as windspeed increases, and by the bucket not being able to tip collected rain and right itself quickly enough during extreme rain to capture all rainfall ("tipping bucket error") [@Medlin2007].
+Rain can also be measured by radar, but only within a certain distance of the radar (around 100 kilometers [double-check that this is still approximately correct, or if things have improved since Danny in 1997]) [@Medlin2007], and radar measurements can underestimate rain for a hurricane [@Medlin2007].
+Satellites can also provide some tropical storm rain data [@Medlin2007].
+On occassion, data related to rainfall for tropical storms also comes from aircraft reconnaissance [@Medlin2007].
+
+One difficulty in measuring rain exposure consistently across many different storms is how many days to include in the rain total.
+Some storms, especially ones associated with dangerous rainfalls, can move very slowly, meaning a window of several days needs to be used to fully capture storm-related rainfall.
+For example, one study used a period of over four days to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997 [@Medlin2007].
+Single-storm studies often measure rainfall as the total rainfall over the full "event" (e.g., Medlin2007); however, the timing of the event is determine in a customized way for each storm, making it difficult to extend to multi-storm analyses.
+The figure below shows rain exposure for Danny based on including the day before, day of, and day after the storm (top) versus including from the day before through three days after the storm (bottom).
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+storm <- "Danny-1997"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle(paste(storm, "rainfall, -1 to 1 days")) 
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall",
+                                           days_included = -1:3)) + 
+        ggtitle(paste(storm, "rainfall, -1 to 3 days"))
+
+grid.arrange(a, b, ncol = 1)
+```
+
+The NLDAS-2 data can give rainfall estimates well below those published for specific monitors.
+For example, our analysis of Hurricane Danny (1997) found a maximum county-level aggregated rainfall of under 175 millimeters, even when including study days out to the third day after the storm came closest to the county, while monitor-based measurements indicated rainfalls of over 500 millimeters at a number of monitors in Mobile County (FIPS: 01097) and Baldwin County (FIPS: 01003) [@Medlin2007]:
+
+```{r echo = FALSE}
+rain %>% 
+        filter(storm_id %in% "Danny-1997" &
+                       fips %in% c("01097", "01003")) %>%
+        kable()
+```
 
@@
 This rainfall can be particularly severe in the Appalachians [@Halverson2015].
-Further, heavier rainfall is more common for storms moving at slower forward speeds [@Chang2013].
+Further, heavier rainfall is more common for storms moving at slower forward speeds [@Chang2013; @Medlin2007].
 Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].
@@
 Several storm characteristics are associated with the likelihood of heavy rain and flooding from the storm, including the storm's forward speed and size [@Rappaport2000].
-There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season (e.g., June--August), when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000].
+There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season (e.g., June--August), when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000] (although, conversely, some have suggested that a storm with the same amount of rain is likely to cause less flooding earlier in the season, since more water can evaporate in the hotter conditions more common early in the hurricane season [@Mallin2006]).
+
+Rain is a hazard of tropical storms mostly because it causes flooding.
+The likelihood and extent of flooding during a tropical storm is related to the amount of rain from the storm, but it is also related to other factors, and storms with similar amounts of rainfall can cause different flooding risks [@Chen2015].
+For example, one study of a North Carolina watershed found that streamflow following tropical storms was related not only to rainfall from the storm, but also to top soil saturation and timing of the storm during the hurricane season [@Chen2015].
+Therefore, an exposure metric based on rainfall may not perfectly characterize flooding-related risks, and future research could explore either streamflow-based exposure metrics or metrics that combine rain with remote measurements of factors like soil moisture, vegetation states, and meteorological conditions that might modify the amount of flooding caused by a tropical storm's rains [@Chen2015].
 
@@
 A tropical storm's high winds can bring a number of dangers, including health risks and property damage caused by structural damage of houses and other buildings, falling trees, and wind-borne debris [@Rappaport2000].
-These types of wind-related damage can occur when winds are below hurricane strength (i.e., < 64 knots) [@Rappaport2000].
+One study of US hurricanes between 1925 and 1995 found that property damage associated with a storm was strongly associated with the intensity of the storm, with most damage occuring during category 3, 4, or 5 storms, even though storms of these categories made up under a quarter of all hurricanes in the study [@Pielke1998].
+Wind-based hazards are also closely associated with power outages [citation], which can introduce a number of threats to human and ecological health and property, including water quality risks if the outage affects wastewater treatment plants [@Mallin2006].
+
+However, these types of wind-related damage can occur when winds are below hurricane strength (i.e., < 64 knots) [@Rappaport2000].
 While severe winds pose an important threat, risks of a tropical storm can exist without severe wind; for example, one study found that most of the direct hurricane-related deaths in the US between 1970 and 1999 occurred in cases when wind was below hurricane strength, including for Tropical Storms Charley in 1998 and Alberto in 1994 [@Rappaport2000].
@@
 
+Since rainfall is not perfectly correlated with flooding risks from tropical storms [@Chen2015], it may also be useful in future research to explore exposure metrics based on more direct measurements of flooding.
+An associated risk from hurricanes is degraded water quality [@Mallin2006]; other metrics, including dissolved oxygen, biochemical oxygen demand, and fecal coliform bacteria, might help classify exposure to these hurricane risks [@Mallin2006], which can occur when flooding or power outages affect wastewater treatment plants, landfills, or concentrated animal feedlot operations [@Mallin2006].
+These water quality issues can threaten both human and ecological health; for example, some hurricanes (e.g., Andrew, Hugo, Fran, Bonnie, and Floyd) have been linked to large fish kills [@Mallin2006].
+The severity of these impacts can be associated with factors we do not incorporate into our exposure measurements here, including whether a storm track follows a tributary, which can increase impacts to water quality [@Mallin2006].
+
+Storm impacts can also be associated with the forward speed of a storm, since a slow-moving storm can linger and cause a large accumulation of rain [@Medlin2007] as well as persist over multiple high tides, increasing coastal impacts [@Mallin2006].
+
+Data on storm impacts, including property damage or insurance claims data, are sometimes available through storm databases or publications (e.g., NOAA's Storm Events data, *Monthly Weather Review*).
+While such data could potentially be used to create an exposure metric, tropical storm damage data requires some normalization, to incorporate both changes in dollar value over time and also changes in development in at-risk areas, to be comparable over extended time periods [@Pielke1998].
+
 #### Example uses of exposure datasets

### 2016-06-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on paper draft

@@
 Tropical storms bring a number of threats, including high winds, heavy rain and flooding, storm surge, tornadoes, ...
-.
+[@Rezapour2014: wind, storm surge, rain].
 To determine whether a county is exposed to a particular tropical storm, a researcher could use criteria based on measurements of any of these values during the storm, use a metric based on the distance from the county to the storm's track, or use a metric that combines several of these measurements.
@@
 
-Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009; @Czajkowski2011].
+Many of the studies that have assessed exposure to tropical storms in the US have used geographical information system software (e.g., ArcGIS) [example studies: @Grabich2016; @Zandbergen2009; @Czajkowski2011; @Kruk2010].
 Here, we offer methods to map and output historic exposure to tropical storms that does not require the use of proprietary software but instead uses a package written in the R statistical programming language, which is free and open-source.
@@
 They used this exposure evaluation to create maps of total exposure to tropical storms within US counties, as well as to explore associations between a county's long-term exposure to tropical storms and its location, distance from the coast, size, and shape [@Zandbergen2009].
+
+@Kruk2010 explored exposure to hurricane-related winds in the United States, including inland areas, for 1900-2008.
 
@@
 
-We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database.
+We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database [@Demuth2006].
 This dataset is based on a post-storm assessment of each storm conducted by the United States National Hurricane Center (NHC) and incorporates data from a variety of sources, including satellite data and, when available, aircraft reconnaisance data [@Landsea2013].
 This data gives time stamps for each observation in Coordinated Universal Time (UTC; also known as Zulu Time, sometimes indicated by "Z").
+This dataset has been used extensively to study Atlantic basin tropical storms (e.g., @Rezapour2014, ...).
+The extended best tracks data extension of HURDAT [@Demuth2006], which we use in this study, has been used in studies of hurricane-related inland wind exposure [@Kruk2010], ...
+[citations].
 
@@
 One interesting contrast is between storms that were noteworthy for direct wind-related deaths (Anderew in 1992, Hugo in 1989) versus storms noteworthy for direct freshwater flooding-related deaths (Floyd in 1999, Alberto in 1994, Charley in 1998) for the 1970--1999 period [@Rappaport2000].
-Here are two examples, one of each of these two types:
+(Another study notes the following storms were notable for inland impacts-- Hugo, 1989 (wind and rain); Floyd, 1999 (flooding); Allison, 2001 (flooding); Fran, 1996 (wind) [@Kruk2010].) Here are two examples, one of each of these two types:
 
@@
 
+#### Differences between exposure classification for different exposure metrics
+
+Some storm hazards might be inversely associated.
+For example, while a slow-moving storm can bring more rain [citation] and cause more damage because of sustained hazardous conditions [@Rezapour2014], fast-moving storms bring higher risks of dangerous winds inland [@Kruk2010].
+
 #### Distance-based exposure
@@
 
-While a number of storm characteristics are strongly associated with distance from the storm's center (e.g., wind and, at the coast, storm surge and waves; @Rappaport2000), other characteristics like rain can occur at dangerous levels well away from the storm's central track [@Rappaport2000].
+While a number of storm characteristics are strongly associated with distance from the storm's center (e.g., wind and, at the coast, storm surge and waves; @Rappaport2000, @Kruk2010), other characteristics like rain can occur at dangerous levels well away from the storm's central track [@Rappaport2000], and while winds typically weaken inland, dangerous wind exposures from tropical storms are still observed in the eastern United States well inland from the coast [@Kruk2010].
 Although storm surge was a major cause of hurricane-related deaths in the past, more recently freshwater flooding has become a much more common cause of deaths directly resulting from tropical storms in the US [@Rappaport2000].
@@
 Other distance-based exposure metrics have defined "exposure" to occur only when a storm's center track passes through a county's boundary (e.g., @Zandbergen2009, who describe this metric of exposure as a "hit" on the county).
-This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track, especially since hurricanes can have a width ([typical storm width; citation]) that is much larger than a typical county's width.
-Further, exposure assessed using this method has been found to be significantly [?
-double-check] associated with the size and shape of a county [@Zandbergen2009].
-By contrast, assessing distance-based exposure based on measuring distance between the storm track and each county's center would be less susceptible to differences in county size and shape and so might be a more reliable metric of exposure than determining if a storm track ever crossed a county boundary.
+This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track [@Kruk2010], especially since hurricanes can have a width ([typical storm width; citation]) that is much larger than a typical county's width and since a hurricane's maximum winds can extend well away from the storm's center.
+Further, exposure assessed using this method is associated with the size [@Zandbergen2009; @Kruk2010] and shape [@Zandbergen2009] of a county.
+(While some studies have avoided this problem by determining exposure for equal-area grid cells (e.g., [@Kruk2010]), in studies where exposure needs to be matched with a county-level outcome, some process of aggregation from grid cells to counties would still need to be performed.) By contrast, assessing distance-based exposure based on measuring distance between the storm track and each county's center would be less susceptible to differences in county size and shape and so might be a more reliable metric of exposure than determining if a storm track ever crossed a county boundary.
 Further, assessing the minimum distance between county center and the storm track, as we do here, allows for a continuous, rather than binary, metric of distance-based exposure to a storm.
 
-The @Zandbergen2009 study assumed symmetrical activity of storm related factors,which does not accurately represent storm characteristics once landfall has been made [@Kruk2010, @Halverson2015] [GBA-- I did not understand this thought.
-I agree that storms are asymmetric, but I don't think that's just after landfall.
-Is this something about characteristics of storms as they transform to extratropical systems?
-Let's discuss.]
+The @Zandbergen2009 study assumed symmetrical activity of storm-related factors,which is not a realistic assumption for storm wind and rain patterns, particularly at and following landfall [@Kruk2010, @Halverson2015, @Rezapour2014].
 
@@
 
-Most distance exposure definitions in the past have been based on maximum wind radii based on maximum sustained winds [@Kruk2010].
+Most distance exposure definitions in the past have been based on maximum wind radii based on maximum sustained winds, [@Kruk2010] [GBA-- this is what our wind model uses as an input, too.
+Also I had a hard time finding this specific point in this reference.], although wind radii are available through the extended best tracks data [@Demuth2006].
 
@@
 
-Some of the other possible sources for estimating rain during tropical storms include ...[Stage IV, TMPA, NEXRAD]
+Some of the other possible sources for estimating rain during tropical storms include ...[Stage IV, TMPA, NEXRAD].
+Rain data for tropical storms can potentially come from satellite based observations (e.g., Tropical Rainfall Measuring Mission [TRMM], Global Precipitation Measurement [GPM], Precipitation Estimation from Remotely Sensed Imagery Using Artificial Neural Network [PERSIANN]) and weather forecast models (e.g., European Centre for Medium-Range Weather Forecasts [ECMWF], National Centers for Environmental Prediction [NCEP]) [@Rezapour2014] [GBA-- are the "weather forecast models" analogous with re-analysis data like NLDAS?].
+Precipitation data can be measured by ground-based monitor, satellite, or radar [@Hou2014].
+Monitor-based measurements can be difficult to calibrate and compare across large regions [@Hou2014], and certain types of terrain (e.g., mountainous) can make precipitation measurements difficult by radar [@Hou2014].
+There is no "gold standard" of precipitation measurement; all three of these ways of measuring precipitation are prone to their own uncertainties [@Hou2014].
 
@@
 This rainfall can be particularly severe in the Appalachians [@Halverson2015].
-Further, heavier rainfall is more common for storms moving at slower forward speeds [@Chang2013; @Medlin2007].
-Storms with a slower forward speed and larger rainfall area contribute to a longer duration of rainfall, and therefore an increased chance of flood-related damage [@Rezapour2014].
+Further, heavier rainfall is more common for storms moving at slower forward speeds [@Medlin2007].
+Storms with a slower forward speed and larger rainfall area contribute to a longer duration of hazardous conditions, and therefore an increased chance of storm-related damage [@Rezapour2014].
 
@@
 Further, wind-based exposure metrics are directly linked to a storm's strength, and some research suggests that hurricane strength and landfall is not strongly associated with the number of direct deaths ultimately caused by the storm [@Rappaport2000].
+A number of hurricane hazards, including dangerous rainfall, are not captured by exposure metrics that depend solely on storm winds [@Rezapour2014].
 
@@
 
+The H*Wind database [@Powell1998] also has been used in studies (e.g., @Rezapour2014, @Grabich2016), but was discontinued as a NOAA product after 2013 (see \url{http://www.hwind.co/legacy_data/} for more on this dataset and its current status).
+These data are publicly available for storms between 1993 and 2013 (\url{http://www.aoml.noaa.gov/hrd/data_sub/wind.html}).
+It appears that gridded data and shapefiles are available for certain times during the storms for this database.
+
 There is some uncertainty in the maximum wind speed values estimated at each synoptic time for each storm.
@@
 
-A recent study suggests that nearly all states east of the Rocky Mountains have experienced wind exposure associated with either tropical or post-tropical storms [@Kruk2010].
+A recent study found a history of at least some exposure to high winds related to tropical or post-tropical storms in most eastern US states, including in many inland locations [@Kruk2010].
 

### 2016-06-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on draft

@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
 Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000].
@@
 
+We acknowledge that in some cases it will be important to estimate hurricane exposure at a different level of aggregation.
+For example, one study investigated the association between hurricanes and adverse health outcomes by comparing women within the same county who were "exposed" and "unexposed" to storms during pregnancy [@Currie2013]; this type of study design would require a finer resolution of hurricane exposure estimates, as it requires different exposure levels within a county.
+
 #### Examples of estimating TS for multiple storms
@@
 
+Another study explored trends in direct hurricane fatalities for all landfalling US hurricanes between 1970 and 2007 in counties that were "exposed" to storms, based on a distance-based metric (whether the county was along the path of the storm [primary exposure], neighboring a county along the storm's path [secondary counties], or neighboring a secondary county [tertiary county]), determining wind- and rain-related hazards for all "exposed" counties and exploring associated with fatality counts [@Czajkowski2011].
+
 @Kruk2010 explored exposure to hurricane-related winds in the United States, including inland areas, for 1900-2008.
+
+@Currie2013, while studying exposures at a sub-county rather than county resolution, investigated risks of adverse birth outcomes associated with hurricanes in Texas between 1996 and 2008.
 
@@
 
-In this study, we use precipitation data from the North American Land Data Assimilation System, phase 2 (NLDAS-2) to characterize rain-based exposure to tropical storms.
-The NLDAS-2 data integrates satellite-based and land-based monitoring and applies a land-surface model to create a reanalysis dataset that is spatially and temporally complete across the continental United States.
+In this study, we use precipitation data from the North American Land Data Assimilation System, phase 2 (NLDAS-2) [@Rui2014] to characterize rain-based exposure to tropical storms.
+The NLDAS-2 data integrates satellite-based and land-based monitoring and applies a land-surface model to create a reanalysis dataset that is spatially and temporally complete across the continental United States [@Rui2014; @AlHamdan2014].
 This data has been used in previous research to investigate tropical storms, including in a study of the relationship between rainfall and streamflow responses for tropical storms in a North Carolina watershed [@Chen2015].
 
-The NLDAS-2 precipitation data is originally provided hourly for a 1/8 degree grid.
+The NLDAS-2 precipitation data is originally provided hourly for a 1/8 degree grid [@Rui2014; @AlHamdan2014].
 To generate county-level daily rainfall estimates, we first aggregated data at each grid point, after converting the timestamp of each observation to local time, to create a daily estimate at the grid point.
-We then averaged all grid points within a county's boundaries to generate a county-level average of daily precipitation.
+We then averaged all grid points within a county's boundaries to generate a county-level average of daily precipitation [@AlHamdan2014].
+This county-level precipitation data is publicly available through the US Centers for Disease Control's Wide-ranging Online Data for Epidemiological Research (WONDER) database [citation for WONDER; @AlHamdan2014].
 
@@
 For example, while a slow-moving storm can bring more rain [citation] and cause more damage because of sustained hazardous conditions [@Rezapour2014], fast-moving storms bring higher risks of dangerous winds inland [@Kruk2010].
+Storms with weak winds can bring dangerous rains and flooding [@Atallah2007].
+
+Some studies have done sensitivity analyses to consider how results changed using different exposure metrics.
+For example, @Czajkowski2011 performed a sensitivity analysis using different assumptions about wind decay away from the center of a storm, while @Currie2013, who used a distance-based exposure metric, considered different buffer distances as a sensitivity analysis.
+
+#### Restricting to certain types of storms
+
+Some studies of extended time periods have restricted their analysis to certain types of tropical storms.
+For example, some studies have only included landfalling hurricanes [studies] [citations].
+One study of risks from hurricane exposures in Texas limited their analysis to storms causing $10 million or more in damage [@Currie2013], based on data in the Weather Underground Hurricane Archive.
 
@@
 While there are often several different "best tracks" datasets, from different sources [?-- weather services?] for other ocean basins, the "best tracks" data from ..., which is the basis of the tracks we use here, are the undisputed primary source of tropical storm track data for the Atlantic basin.
+One study [@Currie2013] did, however, use hurricane track data from the Weather Underground Hurricane Archive [GBA-- is Weather Underground just using the NOAA best tracks?
+That's my guess.
+In that case, this source would be the same].
 
@@
 Further, fatal tropical storm tornadoes most often occur between 200 kilometers and 500 kilometers from the storm's center, and are more common as a storm decays inland than at landfall, and were linked to over 300 deaths in the US between 1995 and 2009 [@Moore2012].
+However, one study found that damage to buildings was strongly associated with distance from the storm's center [check Deryugina reference in @Czajkowski2011].
 
-Further, the forces of a storm tend to be distributed around the center in a non-symmetrical way [citation], while distance-based exposure metrics that use buffers tend to use an equal buffer distance on each side of the storm track (e.g., [citations]).
+Further, the forces of a storm tend to be distributed around the center in a non-symmetrical way [citation], while distance-based exposure metrics that use buffers tend to use an equal buffer distance on each side of the storm track (e.g., @Czajkowski2011, ...).
 
@@
 
-The @Zandbergen2009 study assumed symmetrical activity of storm-related factors,which is not a realistic assumption for storm wind and rain patterns, particularly at and following landfall [@Kruk2010, @Halverson2015, @Rezapour2014].
+Other studies have compared a storm's track to specific points to determine closest exposure, as we do here.
+For example, one study of adverse birth outcomes and hurricanes in Texas determined distance-based exposure by measuring the closest distance between any point on a storm track and a woman's residence [@Currie2013].
+
+Some studies (e.g., @Zandbergen2009, @Czajkowski2011) assumed symmetrical activity of storm-related factors,which is not a realistic assumption for storm wind and rain patterns, particularly at and following landfall [@Kruk2010, @Halverson2015, @Rezapour2014].
 
@@
 In assessing the risk of a given storm based on distance, recent studies have defined the distance from the storm track affected by a given storm differently.
-[@Czajkowski2011] assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties.
-Such a definition resulted in an exposure defenition based on an average distance radius of 120 km on either side of the storm track [@Czajkowski2011].
-Such a distance is slightly greater than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].
+@Czajkowski2011 assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties.
+Such a definition resulted in an exposure defenition based on an average distance radius of 120 km [75 miles-- double check that this conversion is correct] on either side of the storm track [@Czajkowski2011].
+Such a distance is slightly larger than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].
 
-Another study used distance buffers of 30, 60, and 100 kilometers for different metrics of distance-based exposure [@Grabich2016] [check the paper-- was this distance to any part of a county, or to the center of the county?].
+Other studies used distance buffers of 30 kilometers [@Currie2013] or of 30, 60, and 100 kilometers as thresholds for identifying distance-based exposure [@Grabich2016] [check the Grabich paper-- was this distance to any part of a county, or to the center of the county?].
 Another study used a distance criterion of 60 kilometers from the storm track [@Grabich2015] [again, check to see if this was to any point in the county or to some center point of the county].
@@
 In a study assessing the association between hurricanes and undesireable birth outcomes, researchers found that results were not sensitive to the omission of residences 100 km from the storm path, and that results varied insignificantly from 30-75 km [@Currie2013].
+
+One study used distance to determine exposure to landfalling US hurricanes by assigning counties along the storm's tracks as "primary" exposures, counties neighboring primary counties as "secondary" counties, and counties neighboring secondary counties as "tertiary" counties; this method correspond to a distance buffer of approximately 75 miles ([x] kilometers) from the storm track for a county to qualify as "exposed" to a storm [@Czajkowski2011].
+The study found that about 30 direct hurricane fatalities over their study period (1970--2007) occurred outside of this distance range from the storm center [@Czajkowski2011], suggesting that this distance choice can cause some counties with dangerous exposures to a storm to be mis-identified as "unexposed".
 
@@
 
+When using a distance-based metric, often the majority of counties classified as "exposed" to a storm are inland [@Czajkowski2011].
+This is because the distance-based metric is linked to distance from the storm track, which can extend well inland.
+However, many of a storm's hazards decrease rapidly after landfall (e.g., maximum winds) or are almost exclusively experienced at the coast (e.g., storm surge, waves).
+
+Distance-based measures that are based on a storm track passing through a county can also miss exposures from storms when they are offshore but near enough to the coast to cause important hazards to health and property (e.g., Agnes in 1972, FLoyd in 1999, Chantal in 1989) [@Czajkowski2011].
+By basing the distance criteria on distance from the storm's track to county centers, more of these offshore storm exposures could be captured.
+
 #### Rain-based exposure
@@
 For example, most of the direct hurricane-related deaths in the US from 1970 to 1999 were caused by freshwater flooding and were in inland, rather than coastal, counties [@Rappaport2000].
+Direct fatalities from hurricanes tend to increase with increasing rainfall from the storm [@Czajkowski2011].
 Rains from tropical storms can flood roads [@Rappaport2000].
@@
 On occassion, data related to rainfall for tropical storms also comes from aircraft reconnaissance [@Medlin2007].
+Generating county-level rain estimates from surface monitors requires some method of spatial interpolation, to create a spatially continuous estimate of rainfall values, and aggregating across counties (e.g., [@Czajkowski2011]).
 
@@
 
+Gridded, re-analysis precipitation data (e.g., NCEP UPD) have been found to underestimate totl rainfall from tropical storms, but correctly identify the qualitative distribution of low to high rainfall for storms (i.e., it can identify the areas with the highest rainfall, but underestimates the exact amount of rainfall) [@Atallah2007].
+
 Exposure classification based on rainfall has some advantages.
@@
 In comparison to the storm track, the areas of extreme rain and extreme wind can vary.
-For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track [@Halverson2015], while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
+For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track [@Halverson2015; @Atallah2007], while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
 The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference].
@@
 Therefore, an exposure metric based on rainfall may not perfectly characterize flooding-related risks, and future research could explore either streamflow-based exposure metrics or metrics that combine rain with remote measurements of factors like soil moisture, vegetation states, and meteorological conditions that might modify the amount of flooding caused by a tropical storm's rains [@Chen2015].
+The topography of a location (e.g., elevation change in a county) could also affect whether rains from a tropical storm cause a flood, particularly a flash flood [@Czajkowski2011 mentions this, but is not a primary source for this point].
+
+Other studies, like ours, do not attempt to identify and remove any precipitation over the aggregating period associated with weather systems other than the tropical storm (e.g., @Czajkowski2011).
 
@@
 Wind-based hazards are also closely associated with power outages [citation], which can introduce a number of threats to human and ecological health and property, including water quality risks if the outage affects wastewater treatment plants [@Mallin2006].
+Direct fatalities from hurricanes tend to increase with increasing wind exposure from the storm [@Czajkowski2011].
 
@@
 [see references for this paper].
+Another study used a wind model that assumed that all counties along a storm's path had the maximum wind speed observed as the track passed the county, while counties neighboring this "primary" county had a windspeed of 75% of this maximum sustained wind value, and counties neighboring these "secondary" counties had a windspeed of 50% of the maximum sustained wind [@Czajkowski2011].
 Another study used wind data from a NOAA public database based on real-time hurricane wind analysis, rather than modeling windspeeds themselves [@Grabich2016] [see the Powell reference in this paper for more on that data].

### 2016-06-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on paper draft

@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Kinney2008: autism prevalence; @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
 Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000].
@@
 We collected "best tracks" data on hurricane tracks for Atlantic basin storms between 1988 and 2014 from the extended best tracks database [@Demuth2006].
-This dataset is based on a post-storm assessment of each storm conducted by the United States National Hurricane Center (NHC) and incorporates data from a variety of sources, including satellite data and, when available, aircraft reconnaisance data [@Landsea2013].
-This data gives time stamps for each observation in Coordinated Universal Time (UTC; also known as Zulu Time, sometimes indicated by "Z").
+This dataset is based on a post-storm assessment of each storm conducted by the United States National Hurricane Center (NHC) and incorporates data from a variety of sources, including satellite data and, when available, aircraft reconnaisance data [@Landsea2013; @Jarvinen1988].
+This data gives time stamps for each observation in Coordinated Universal Time (UTC; also known as Zulu Time, sometimes indicated by "Z") [@Jarvinen1988; this source (1978) says GMT].
 This dataset has been used extensively to study Atlantic basin tropical storms (e.g., @Rezapour2014, ...).
@@
 
-Given its better consistancy with rain gage values compared to NLDAS precipitation data [GBA-- you had something in here about resolution, as well.
+Given its better consistancy with streamflow gage measurements compared to NLDAS precipitation data [GBA-- you had something in here about resolution, as well.
 Does it have a better resolution than the NLDAS data?
-I think we probably want to focus on comparing this source with NLDAS, and we don't need to put much about advantages that it has that are very similar to NLDAS], [@Villarini2011] recommends Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
+I think we probably want to focus on comparing this source with NLDAS, and we don't need to put much about advantages that it has that are very similar to NLDAS], @Villarini2011 recommends Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
 However, the Stage IV dataset has only been archived since 2002 [@Lin2005] and cannot be used to analyze exposure to tropical storms for studies that include storms before 2002.
@@
 These wind models use estimates of maximum wind speed from the best tracks dataset.
-These maximum wind speed estimates in the best tracks dataset are rounded to 5-knot intervals and is the maximum 1-minute-average windspeed at 10 meters above the ground [@Landsea2013].
+These maximum wind speed estimates in the best tracks dataset are rounded to 5-knot intervals and is the maximum 1-minute-average windspeed at 10 meters above the ground [@Landsea2013; @Jarvinen1988].
 This windspeed is typically determined using the Dvorak technique to estimate windspeed from satellite measurements, although data from aircraft reconnaisance in also sometimes incorporated in the estimate [@Levinson2010].
+Sometimes, the maximum windspeed in this dataset is estimated from the central pressure of the storm [@Jarvinen1988].
 
@@
 One study of risks from hurricane exposures in Texas limited their analysis to storms causing $10 million or more in damage [@Currie2013], based on data in the Weather Underground Hurricane Archive.
+A study of the association between autism prevalence and hurricane exposure in Louisiana between 1980 and 1995 used only storms they identified as "severe", based on an assessment [?] of National Weather Service data [@Kinney2008].
 
@@
 Many of the uncertainties related to storm positions in best track data are more of a concern for the years before ...
-(e.g., pre-19[xx], before aircraft reconnaisance of Atlantic basin tropical storms).
+(e.g., pre-1944, before aircraft reconnaisance of Atlantic basin tropical storms) [@Jarvinen1988].
 
-While there are often several different "best tracks" datasets, from different sources [?-- weather services?] for other ocean basins, the "best tracks" data from ..., which is the basis of the tracks we use here, are the undisputed primary source of tropical storm track data for the Atlantic basin.
+While there are often several different "best tracks" datasets, from different sources [?-- weather services?] for other ocean basins [@Knapp2010], the "best tracks" data from NOAA's National Hurricane Center [@Jarvinen1988], which are the basis of the extended hurricane tracks data we use here, are the undisputed primary source of tropical storm track data for the Atlantic basin [@Knapp2010].
 One study [@Currie2013] did, however, use hurricane track data from the Weather Underground Hurricane Archive [GBA-- is Weather Underground just using the NOAA best tracks?
@@
 
-Other distance-based exposure metrics have defined "exposure" to occur only when a storm's center track passes through a county's boundary (e.g., @Zandbergen2009, who describe this metric of exposure as a "hit" on the county).
+Other distance-based exposure metrics have defined "exposure" to occur only when a storm's center track passes through a county's boundary (e.g., @Zandbergen2009, who describe this metric of exposure as a "hit" on the county; @Kinney2008 is another example that considered highest exposures for counties directly on the storm's path).
 This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track [@Kruk2010], especially since hurricanes can have a width ([typical storm width; citation]) that is much larger than a typical county's width and since a hurricane's maximum winds can extend well away from the storm's center.
@@
 The estimated rainfall amounts from our data are likely underestimates.
-This data source, however, should be internally consistent and so useful for comparing across different storms when all exposure estimates are based on this rain data.
+@Villarini2011, for example, found that NLDAS-2 precipitation underestimated total precipitation for Hurricanes Frances, Ivan, and Jeanne in 2004 compared to streamgage results.
+NLDAS-2 precipitation data, however, should be internally consistent (i.e., correctly identify areas and storms with greatest rainfall, although underestimating exact rainfall totals) and so useful for comparing across different storms when all exposure estimates are based on this rain data.
 
@@
 However, during extreme storms, the problems with measuring rainfall using [rain monitors] would propogate into the NLDAS data, so although NLDAS would prevent missing values during the storm if monitors are not able to provide data, if monitors are out, rainfall estimates from NLDAS will be based more on models than on observations.
-More on NLDAS bias low here [@Villarini2011].
+[More on NLDAS bias to underestimate in @Villarini2011].
 Such bias, though unimportant in our analysis, may play a larger role in the extrapolation to other studies or exposure analysese where the precipitation cutoffs used here may provide a threshold that results in the missclassification of less affected counties as exposed.
@@
 In particular, there are options for wind speed models as far as ...
-One study used a very simple model of constant winds up to a distance of 30 km from the storm track and a simple model of wind decay beyond that distance [@Zandbergen2009], based on a model developed by ...
-[see references for this paper].
+One study used a simple, symmetrical model that assumed constant winds (equal to the maximum wind speed from best tracks data, [converted to wind gusts?]) up to 30 km from the storm track and symmetrical wind decay beyond that distance [@Zandbergen2009], based on a model developed by @Klotzbach.
 Another study used a wind model that assumed that all counties along a storm's path had the maximum wind speed observed as the track passed the county, while counties neighboring this "primary" county had a windspeed of 75% of this maximum sustained wind value, and counties neighboring these "secondary" counties had a windspeed of 50% of the maximum sustained wind [@Czajkowski2011].

### 2016-06-24 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Added code for Rain exposure map. need to fix legend and set own cut points

@@
 library(gridExtra)
+library(plyr)
 library(dplyr)
+library(choroplethr)
 ```
@@
 
-```{r echo = FALSE}
-# Code for this map here
+```{r echo = FALSE, warning=FALSE}
+data(closest_dist)
+my_fips <- closest_dist$fips
+my_fips <- unique(my_fips)
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500,
+                             days_include = c(-1, 0, 1))
+rain_1000 <- rain_exposure %>%
+        arrange(desc(tot_precip)) %>%
+        group_by(fips) %>%
+        summarise(value = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), value = as.factor(value)) %>%
+        arrange(desc(value)) %>%
+        slice(1:1000)
+eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
+                    "district of columbia", "florida", "georgia", "illinois",
+                    "indiana", "iowa", "kansas", "kentucky", "louisiana",
+                    "maine", "maryland", "massachusetts", "michigan",
+                    "mississippi", "missouri", "new hampshire", "new jersey",
+                    "new york", "north carolina", "ohio", "oklahoma",
+                    "pennsylvania", "rhode island", "south carolina",
+                    "tennessee", "texas", "vermont", "virginia",
+                    "west virginia", "wisconsin")
+rain_1000$value <- as.numeric(rain_1000$value)
+map_1000 <- county_choropleth(rain_1000,
+                              title = "Hits for top 1000 Storm-County Pairs\nfor Rain > 75 mm",
+                              legend = "Hits per County", num_colors = 9,
+                              state_zoom = eastern_states)
+map_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
 ```

### 2016-06-24 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Added code for distance count plot/ rain count plot

@@
 ```{r echo = FALSE}
-# Code for this map here
+distance_exposure <- county_distance(counties = my_fips,
+                                 start_year = 1988, end_year = 2011,
+                                 dist_limit = 100)
+distance_1000 <- distance_exposure %>%
+        arrange(desc(storm_dist)) %>%
+        group_by(fips) %>%
+        summarise(value = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), value = as.numeric(value)) %>%
+        arrange(desc(value)) %>%
+        slice(1:1000)
+map_dist_1000 <- county_choropleth(distance_1000, 
+                              title = "Hits by County\nfor Distance < 100 mm",
+                              legend = "Hits per County", num_colors = 9,
+                              state_zoom = eastern_states)
+map_dist_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
 ```
@@
 ```{r echo = FALSE}
-# Code for this map here
+distance_exposure <- county_distance(counties = my_fips,
+                                 start_year = 1988, end_year = 2011,
+                                 dist_limit = 100)
+distance_1000 <- distance_exposure %>%
+        arrange(desc(storm_dist)) %>%
+        group_by(fips) %>%
+        summarise(value = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), value = as.numeric(value)) %>%
+        arrange(desc(value))
+map_dist_count <- county_choropleth(distance_1000, 
+                              title = "Hits by County\nfor Distance < 100 mm",
+                              legend = "Hits per County", num_colors = 9,
+                              state_zoom = eastern_states)
+map_dist_count + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
 ```
@@
 ```{r echo = FALSE}
-# Code for this map here
+rain_count <- rain_exposure %>%
+              group_by(fips) %>%
+              summarise(count = n()) %>%
+              rename(region = fips, value = count) %>%
+              mutate(region = as.numeric(region), value = as.numeric(value))
+map_counts <- county_choropleth(rain_count,
+                                title = "Hits by County\nfor Rain > 75 mm",
+                                legend = "Hits per County", num_colors = 9,
+                                state_zoom = eastern_states)
+map_counts + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
 ```

### 2016-06-24 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Rearrange code

--- 2016-06-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-Map of county exposures to the top 1000 county-storm pairs by distance-based exposure:
+Map of the count of county exposures to the top 1000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
 
 ```{r echo = FALSE}
-# Code for this map here
+data(closest_dist)
+my_fips <- closest_dist$fips
+my_fips <- unique(my_fips)
+distance_exposure <- county_distance(counties = my_fips,
+                                 start_year = 1988, end_year = 2011,
+                                 dist_limit = 100)
+distance_1000 <- distance_exposure %>%
+        arrange(desc(storm_dist)) %>%
+        group_by(fips) %>%
+        summarise(value = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), value = as.numeric(value)) %>%
+        arrange(desc(value)) %>%
+        slice(1:1000)
+map_dist_1000 <- county_choropleth(distance_1000, 
+                              title = "Hits by County\nfor Distance < 100 mm",
+                              legend = "Hits per County", num_colors = 9,
+                              state_zoom = eastern_states)
+map_dist_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
 ```
@@
 
-```{r echo = FALSE}
-# Code for this map here
+```{r echo = FALSE, warning=FALSE}
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500,
+                             days_include = c(-1, 0, 1))
+rain_1000 <- rain_exposure %>%
+        arrange(desc(tot_precip)) %>%
+        group_by(fips) %>%
+        summarise(value = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), value = as.factor(value)) %>%
+        arrange(desc(value)) %>%
+        slice(1:1000)
+eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
+                    "district of columbia", "florida", "georgia", "illinois",
+                    "indiana", "iowa", "kansas", "kentucky", "louisiana",
+                    "maine", "maryland", "massachusetts", "michigan",
+                    "mississippi", "missouri", "new hampshire", "new jersey",
+                    "new york", "north carolina", "ohio", "oklahoma",
+                    "pennsylvania", "rhode island", "south carolina",
+                    "tennessee", "texas", "vermont", "virginia",
+                    "west virginia", "wisconsin")
+rain_1000$value <- as.numeric(rain_1000$value)
+map_1000 <- county_choropleth(rain_1000,
+                              title = "Hits for top 1000 Storm-County Pairs\nfor Rain > 75 mm",
+                              legend = "Hits per County", num_colors = 9,
+                              state_zoom = eastern_states)
+map_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
 ```

### 2016-06-24 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Edit code for distance figures

@@
 ```{r echo = FALSE}
-distance_exposure <- county_distance(counties = my_fips,
-                                 start_year = 1988, end_year = 2011,
-                                 dist_limit = 100)
-distance_1000 <- distance_exposure %>%
+distance_count <- distance_exposure %>%
         arrange(desc(storm_dist)) %>%
@@
         arrange(desc(value))
-map_dist_count <- county_choropleth(distance_1000, 
-                              title = "Hits by County\nfor Distance < 100 mm",
+map_dist_count <- county_choropleth(distance_count, 
+                              title = "Hits by County\nfor Distance < 100 km",
                               legend = "Hits per County", num_colors = 9,

### 2016-06-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Zandbergen 2009 used only landfalling storms

@@
 For example, @Grabich2016 assessed the exposure of counties in Florida to land-falling hurricanes during the 2004 season.
-The largest-scale county-level exposure study is likely that of @Zandbergen2009, which estimated exposure in US counties to all [US landfalling?
-or all storms?] Atlantic basin tropical storms between 1851 and 2003, using both distance and an exposure metric that incorporated distance and windspeed.
+The largest-scale county-level exposure study is likely that of @Zandbergen2009, which estimated exposure in US counties to all US landfalling Atlantic basin tropical storms between 1851 and 2003, using both distance and an exposure metric that incorporated distance and windspeed.
 They used this exposure evaluation to create maps of total exposure to tropical storms within US counties, as well as to explore associations between a county's long-term exposure to tropical storms and its location, distance from the coast, size, and shape [@Zandbergen2009].

### 2016-06-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
compare WU and NOAA best tracks data
Respond to GBA. Update count figures. Add to bib.

@@
 
-Current hurricane exposure assignments are often nonspecific [GBA-- I found this wording a bit unclear] and may lead to missclassification of exposure at the county-level [@Grabich2016].
+The most often used hurricane exposure criteria at the county-level include FEMA disaster declarations and area within the storm track, which can be somewhat nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016].
+FEMA declaration is largely subjective, requiring state governors and/or tribal chiefs to submit a proposal for government aid on the basis that local resources are overwhelmed [@FEMA].
+As a result, using FEMA disaster declarations as exposure classification to hurricanes restricts exposure to 1) those states whose resources are overwhelmed and 2) those states who apply for FEMA aid.
+As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane.
 Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015].
@@
 
-Another study explored trends in direct hurricane fatalities for all landfalling US hurricanes between 1970 and 2007 in counties that were "exposed" to storms, based on a distance-based metric (whether the county was along the path of the storm [primary exposure], neighboring a county along the storm's path [secondary counties], or neighboring a secondary county [tertiary county]), determining wind- and rain-related hazards for all "exposed" counties and exploring associated with fatality counts [@Czajkowski2011].
+Another study explored trends in direct hurricane fatalities for all landfalling US hurricanes between 1970 and 2007 in counties that were "exposed" to storms, based on a distance-based metric (whether the county was along the path of the storm [primary exposure], neighboring a county along the storm's path [secondary counties], or neighboring a secondary county [tertiary county]), determining wind- and rain-related hazards for all "exposed" counties and exploring associations with fatality counts [@Czajkowski2011].
 
@@
 We then averaged all grid points within a county's boundaries to generate a county-level average of daily precipitation [@AlHamdan2014].
-This county-level precipitation data is publicly available through the US Centers for Disease Control's Wide-ranging Online Data for Epidemiological Research (WONDER) database [citation for WONDER; @AlHamdan2014].
+This county-level precipitation data is publicly available through the US Centers for Disease Control's Wide-ranging Online Data for Epidemiological Research (WONDER) database [@WONDER; @AlHamdan2014].
 
@@
 Does it have a better resolution than the NLDAS data?
-I think we probably want to focus on comparing this source with NLDAS, and we don't need to put much about advantages that it has that are very similar to NLDAS], @Villarini2011 recommends Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
+I think we probably want to focus on comparing this source with NLDAS, and we don't need to put much about advantages that it has that are very similar to NLDAS; JMF-- Page 5 in Villarini 2011 disucsses how the spacial resolution is much finer for Stage IV than NLDAS and TMPA, which they argue causes a smoothing of maximum rainfall data for the Appalachians], @Villarini2011 recommends Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones.
 However, the Stage IV dataset has only been archived since 2002 [@Lin2005] and cannot be used to analyze exposure to tropical storms for studies that include storms before 2002.
@@
         slice(1:1000)
+eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
+                    "district of columbia", "florida", "georgia", "illinois",
+                    "indiana", "iowa", "kansas", "kentucky", "louisiana",
+                    "maine", "maryland", "massachusetts", "michigan",
+                    "mississippi", "missouri", "new hampshire", "new jersey",
+                    "new york", "north carolina", "ohio", "oklahoma",
+                    "pennsylvania", "rhode island", "south carolina",
+                    "tennessee", "texas", "vermont", "virginia",
+                    "west virginia", "wisconsin")
 map_dist_1000 <- county_choropleth(distance_1000, 
-                              title = "Hits by County\nfor Distance < 100 mm",
+                              title = "Hits by County\nfor Distance < 100 km",
                               legend = "Hits per County", num_colors = 9,
                               state_zoom = eastern_states)
-map_dist_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
+map_dist_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGn", na.value = "white")
 ```
@@
         slice(1:1000)
-eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
-                    "district of columbia", "florida", "georgia", "illinois",
-                    "indiana", "iowa", "kansas", "kentucky", "louisiana",
-                    "maine", "maryland", "massachusetts", "michigan",
-                    "mississippi", "missouri", "new hampshire", "new jersey",
-                    "new york", "north carolina", "ohio", "oklahoma",
-                    "pennsylvania", "rhode island", "south carolina",
-                    "tennessee", "texas", "vermont", "virginia",
-                    "west virginia", "wisconsin")
 rain_1000$value <- as.numeric(rain_1000$value)
@@
 
-Map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this? What seems reasonable?] kilometers or less:
+Map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?
+What seems reasonable?; JMF-- Based on what I have read so far, I think this seems reasonable, unless we want to push it out to 120 km to loosely reflect the Czajkowski study which seems to have found a great number of fatalities in tertiary counties] kilometers or less:
 
@@
                               state_zoom = eastern_states)
-map_dist_count + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
+map_dist_count + scale_fill_brewer("Hits per\nCounty", palette = "YlGn", na.value = "white")
 ```
@@
 That's my guess.
-In that case, this source would be the same].
+In that case, this source would be the same.
+JMF-- they have an archive from 1851-2016; I think it is the same source.
+I'll keep digging though].
 

### 2016-06-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add reference for rain exposure from the central track

@@
 
-While a number of storm characteristics are strongly associated with distance from the storm's center (e.g., wind and, at the coast, storm surge and waves; @Rappaport2000, @Kruk2010), other characteristics like rain can occur at dangerous levels well away from the storm's central track [@Rappaport2000], and while winds typically weaken inland, dangerous wind exposures from tropical storms are still observed in the eastern United States well inland from the coast [@Kruk2010].
+While a number of storm characteristics are strongly associated with distance from the storm's center (e.g., wind and, at the coast, storm surge and waves; @Rappaport2000, @Kruk2010), other characteristics like rain can occur at dangerous levels well away from the storm's central track [@Rappaport2000; @Atallah2007], and while winds typically weaken inland, dangerous wind exposures from tropical storms are still observed in the eastern United States well inland from the coast [@Kruk2010].
 Although storm surge was a major cause of hurricane-related deaths in the past, more recently freshwater flooding has become a much more common cause of deaths directly resulting from tropical storms in the US [@Rappaport2000].

### 2016-06-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Responding to comments

@@
 As a result, using FEMA disaster declarations as exposure classification to hurricanes restricts exposure to 1) those states whose resources are overwhelmed and 2) those states who apply for FEMA aid.
-As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane.
+As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane].
 Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015].
@@
 Other distance-based exposure metrics have defined "exposure" to occur only when a storm's center track passes through a county's boundary (e.g., @Zandbergen2009, who describe this metric of exposure as a "hit" on the county; @Kinney2008 is another example that considered highest exposures for counties directly on the storm's path).
-This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track [@Kruk2010], especially since hurricanes can have a width ([typical storm width; citation]) that is much larger than a typical county's width and since a hurricane's maximum winds can extend well away from the storm's center.
+This way of measuring exposure can exclude nearby counties that suffered extreme conditions from the storm but were not directly on the hurricane's track [@Kruk2010], especially since hurricanes can have a width ([typical storm width; citation])[JMF-- Weatherford and Gray paper referenced in Grabich 2016 may be a good place to start.
+Suggest 30-60km width of the eye of the storm itself] that is much larger than a typical county's width and since a hurricane's maximum winds can extend well away from the storm's center.
 Further, exposure assessed using this method is associated with the size [@Zandbergen2009; @Kruk2010] and shape [@Zandbergen2009] of a county.
@@
 @Czajkowski2011 assessed county-level risk and exposure based on a three-tierd definition, with primary counties being those closest to the storm track on either side, secondary counties being adjacent to primary coutnies, and tertiary counties adjacent to secondary counties.
-Such a definition resulted in an exposure defenition based on an average distance radius of 120 km [75 miles-- double check that this conversion is correct] on either side of the storm track [@Czajkowski2011].
+Such a definition resulted in an exposure defenition based on an average distance radius of 120 km [75 miles] on either side of the storm track [@Czajkowski2011].
 Such a distance is slightly larger than that commonly used by public health departments (i.e., 100 km) [@Czajkowski2011].
 
-Other studies used distance buffers of 30 kilometers [@Currie2013] or of 30, 60, and 100 kilometers as thresholds for identifying distance-based exposure [@Grabich2016] [check the Grabich paper-- was this distance to any part of a county, or to the center of the county?].
-Another study used a distance criterion of 60 kilometers from the storm track [@Grabich2015] [again, check to see if this was to any point in the county or to some center point of the county].
+Other studies used distance buffers of 30 kilometers [@Currie2013] or of 30, 60, and 100 kilometers as thresholds for identifying distance-based exposure [@Grabich2016] [check the Grabich paper-- was this distance to any part of a county, or to the center of the county?; JMF-- I think any part of the county].
+Another study used a distance criterion of 60 kilometers from the storm track, classifying all counties for which the 60 km buffer fully or partially passed through the county's boundaries as exposed [@Grabich2015] [again, check to see if this was to any point in the county or to some center point of the county][JMF-- it is not abundantly clear in the Grabich 2016 article; based on their graphs it seems like they considered a county exposed if the definition they used crossed any county boundary.
+The Grabich 2015 paper specifies both full and partial counties being included].
 
@@
 
-One study used distance to determine exposure to landfalling US hurricanes by assigning counties along the storm's tracks as "primary" exposures, counties neighboring primary counties as "secondary" counties, and counties neighboring secondary counties as "tertiary" counties; this method correspond to a distance buffer of approximately 75 miles ([x] kilometers) from the storm track for a county to qualify as "exposed" to a storm [@Czajkowski2011].
+One study used distance to determine exposure to landfalling US hurricanes by assigning counties along the storm's tracks as "primary" exposures, counties neighboring primary counties as "secondary" counties, and counties neighboring secondary counties as "tertiary" counties; this method correspond to a distance buffer of approximately 75 miles (120 kilometers) from the storm track for a county to qualify as "exposed" to a storm [@Czajkowski2011].
 The study found that about 30 direct hurricane fatalities over their study period (1970--2007) occurred outside of this distance range from the storm center [@Czajkowski2011], suggesting that this distance choice can cause some counties with dangerous exposures to a storm to be mis-identified as "unexposed".
 
-Most distance exposure definitions in the past have been based on maximum wind radii based on maximum sustained winds, [@Kruk2010] [GBA-- this is what our wind model uses as an input, too.
-Also I had a hard time finding this specific point in this reference.], although wind radii are available through the extended best tracks data [@Demuth2006].
+Most distance exposure definitions in the past have been based on maximum wind radii using maximum sustained winds, [@Kruk2010] [GBA-- this is what our wind model uses as an input, too.
+Also I had a hard time finding this specific point in this reference.
+JMF-- I am not sure where I thought I found this.], although wind radii are available through the extended best tracks data [@Demuth2006][JMF-- Delete this section?].
 

### 2016-06-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add table to compare total counts

@@
 library(dplyr)
+library(tidyr)
 library(choroplethr)
@@
 map_counts + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
+```
+
+```{r}
+exposure_name <- c("Rain_Exposure", "Distance_Exposure")
+count <- c(sum(rain_count$value), sum(distance_count$value))
+exposure_table <- cbind(exposure_name, count)
+exposure_table <- data.frame(exposure_table)
+exposure_table %>%
+        mutate(exposure_name = gsub("_", " ", exposure_name)) %>%
+        kable(col.names = c("Exposure Criteria", "Total # exposed counties"), 
+              caption = "Total number of county hits based on rain, distance, and wind exposures")
 ```

### 2016-06-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Try to remove warning message

@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning=FALSE}
 data(closest_dist)
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning=FALSE}
 distance_count <- distance_exposure %>%
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning=FALSE}
 rain_count <- rain_exposure %>%

### 2016-06-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
remove message comments for figures

--- 2016-06-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 data(closest_dist)
@@
 
-```{r echo = FALSE, warning=FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 rain_exposure <- county_rain(counties = my_fips,
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 distance_count <- distance_exposure %>%
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 rain_count <- rain_exposure %>%

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Set custom scale for distance_1000 map

@@
         arrange(desc(value)) %>%
-        slice(1:1000)
+        slice(1:1000) %>%
+        rename(count = value)
+distance_1000$value <- cut(distance_1000$count, c(0, 5, 8, 12, 15, 18, 21, 24),
+                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18", "19-21", "22-24"),
+                           include.lowest = TRUE)
 eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
@@
                               title = "Hits by County\nfor Distance < 100 km",
-                              legend = "Hits per County", num_colors = 9,
+                              legend = "Hits per County", num_colors = 7,
                               state_zoom = eastern_states)

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
set custom cutoffs for distance and rainfall count maps

--- 2016-06-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r echo = FALSE, warning=FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 data(closest_dist)
@@
         group_by(fips) %>%
-        summarise(value = n()) %>%
+        summarise(count = n()) %>%
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), value = as.numeric(value)) %>%
-        arrange(desc(value)) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+        arrange(desc(count)) %>%
         slice(1:1000)
+distance_1000$value <- cut(distance_1000$count, c(0, 5, 8, 12, 15, 18, 21, 24),
+                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18", "19-21", "22-24"),
+                           include.lowest = TRUE)
 eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
@@
         group_by(fips) %>%
-        summarise(value = n()) %>%
+        summarise(count = n()) %>%
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), value = as.factor(value)) %>%
-        arrange(desc(value)) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+        arrange(desc(count)) %>%
         slice(1:1000)
-rain_1000$value <- as.numeric(rain_1000$value)
+rain_1000$value <- cut(rain_1000$count, c(0, 3, 6, 9, 12, 15, 17), 
+                   labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
+                   include.lowest = TRUE)
 map_1000 <- county_choropleth(rain_1000,
                               title = "Hits for top 1000 Storm-County Pairs\nfor Rain > 75 mm",
-                              legend = "Hits per County", num_colors = 9,
+                              legend = "Hits per County", num_colors = 6,
                               state_zoom = eastern_states)
@@
 
-```{r echo = FALSE, warning=FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 distance_count <- distance_exposure %>%
@@
         group_by(fips) %>%
-        summarise(value = n()) %>%
+        summarise(count = n()) %>%
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), value = as.numeric(value)) %>%
-        arrange(desc(value))
+        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+        arrange(desc(count))
+distance_count$value <- cut(distance_count$count, c(0, 5, 8, 12, 15, 18, 21, 24),
+                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18", "19-21", "22-24"),
+                           include.lowest = TRUE)
 map_dist_count <- county_choropleth(distance_count, 
                               title = "Hits by County\nfor Distance < 100 km",
-                              legend = "Hits per County", num_colors = 9,
+                              legend = "Hits per County", num_colors = 7,
                               state_zoom = eastern_states)
@@
 
-```{r echo = FALSE, warning=FALSE}
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 rain_count <- rain_exposure %>%
@@
               summarise(count = n()) %>%
-              rename(region = fips, value = count) %>%
-              mutate(region = as.numeric(region), value = as.numeric(value))
+              rename(region = fips) %>%
+              mutate(region = as.numeric(region), count = as.numeric(count))
+rain_count$value <- cut(rain_count$count, c(0, 3, 6, 9, 12, 15, 17), 
+                   labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
+                   include.lowest = TRUE)
 map_counts <- county_choropleth(rain_count,
                                 title = "Hits by County\nfor Rain > 75 mm",
-                                legend = "Hits per County", num_colors = 9,
+                                legend = "Hits per County", num_colors = 6,
                                 state_zoom = eastern_states)
@@
 exposure_name <- c("Rain_Exposure", "Distance_Exposure")
-count <- c(sum(rain_count$value), sum(distance_count$value))
+count <- c(sum(rain_count$count), sum(distance_count$count))
 exposure_table <- cbind(exposure_name, count)

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
change title of distance_1000 map

@@
 map_dist_1000 <- county_choropleth(distance_1000, 
-                              title = "Hits by County\nfor Distance < 100 km",
+                              title = "Hits for top 1000 Storm-County Pairs\nfor Distance < 100 km",
                               legend = "Hits per County", num_colors = 7,

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add middle initial

@@
 author:
-  - name: Joshua Ferreri
+  - name: Joshua M. Ferreri
     email: joshua.m.ferreri@gmail.com

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Floyd example for mechanisms of flooding

@@
 The topography of a location (e.g., elevation change in a county) could also affect whether rains from a tropical storm cause a flood, particularly a flash flood [@Czajkowski2011 mentions this, but is not a primary source for this point].
+However, @Colle2003 performed an extensive evaluation of Hurricane Floyd (1999) in an effort to understand the mechanisms that led to the development of heavy precipitation and flooding.
+Though differences in the terrain, namely the Appalachian Mountains, contributed to the development of heavy precipitation in this event, @Colle2003 concluded that flooding would have occured in this region regardless of the presence of the Appalachians and other coastal hills.
+Instead, the heavy precipitation is thought to have been primarily caused by strong frontogenesis in combination with slant-wise neutrality [@Colle2003].
+[Look at Bosart and Dean reference in Colle 2003 reference for more on conditions leading to severe flooding]
 

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Chang reference regarding terrain effects in Taiwan

@@
 Therefore, an exposure metric based on rainfall may not perfectly characterize flooding-related risks, and future research could explore either streamflow-based exposure metrics or metrics that combine rain with remote measurements of factors like soil moisture, vegetation states, and meteorological conditions that might modify the amount of flooding caused by a tropical storm's rains [@Chen2015].
-The topography of a location (e.g., elevation change in a county) could also affect whether rains from a tropical storm cause a flood, particularly a flash flood [@Czajkowski2011 mentions this, but is not a primary source for this point].
+The topography of a location (e.g., elevation change in a county) could also affect how extreme rainfall is from a particular storm, as well as whether rains from a tropical storm cause a flood, particularly a flash flood [@Czajkowski2011 mentions this, but is not a primary source for this point].
+For example, topographical characteristics in Taiwan play a key role in determining both the storm's track once landfall has been made, as well as rainfall intensity and location [@Chang2013].
 However, @Colle2003 performed an extensive evaluation of Hurricane Floyd (1999) in an effort to understand the mechanisms that led to the development of heavy precipitation and flooding.

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Info on Appalachian flooding

--- 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 Rain is a hazard of tropical storms mostly because it causes flooding.
-The likelihood and extent of flooding during a tropical storm is related to the amount of rain from the storm, but it is also related to other factors, and storms with similar amounts of rainfall can cause different flooding risks [@Chen2015].
-For example, one study of a North Carolina watershed found that streamflow following tropical storms was related not only to rainfall from the storm, but also to top soil saturation and timing of the storm during the hurricane season [@Chen2015].
+The likelihood and extent of flooding during a tropical storm is related to the amount of rain from the storm, but it is also related to other factors [@Chen2015; @Rees2001], and storms with similar amounts of rainfall can cause different flooding risks [@Chen2015].
+For example, one study of a North Carolina watershed found that streamflow following tropical storms was related not only to rainfall from the storm, but also to top soil saturation [JMF--more on this in the Wood 1990 reference in the Rees paper] and timing of the storm during the hurricane season [@Chen2015].
+The structure of a water basin's drainage network can also play a key role in regional flood response [@Rees2001].
 Therefore, an exposure metric based on rainfall may not perfectly characterize flooding-related risks, and future research could explore either streamflow-based exposure metrics or metrics that combine rain with remote measurements of factors like soil moisture, vegetation states, and meteorological conditions that might modify the amount of flooding caused by a tropical storm's rains [@Chen2015].
-The topography of a location (e.g., elevation change in a county) could also affect whether rains from a tropical storm cause a flood, particularly a flash flood [@Czajkowski2011 mentions this, but is not a primary source for this point].
-However, @Colle2003 performed an extensive evaluation of Hurricane Floyd (1999) in an effort to understand the mechanisms that led to the development of heavy precipitation and flooding.
+
+Storm-motion, as well as the direction from which a storm approaches, can play significant roles in determining the extent of precipitation produced by a storm [@Rees2001; @Chang2013].
+(Examples include the south-to-north storm circulation for Hurricane Fran (1996) with an east-to-west movement, causing an amplification of rainband convection, which played a key role in runoff distribution [@Rees2001] and increased precipitation intensity for storms approaching Taiwan from the east [@Chang2013]).
+
+The topography of a location (e.g., elevation change in a county) could also affect how extreme rainfall is from a particular storm, as well as whether rains from a tropical storm cause a flood, particularly a flash flood [@Rees2001 on orographic enhancement; @Czajkowski2011 mentions this, but is not a primary source for this point].
+For example, topographical characteristics in Taiwan play a key role in regulating rainfall intensity and its location [@Chang2013].
+@Chang2013 found that the most intense rainfall was found in areas with the greatest wind-terrain interaction.
+Storms embodying such characteristics are increasing in both frequency and duration in Taiwan [@Chang2013].[JMF--I am not sure how much of this detail is needed beyond discussing topographical impact, or if mentioning Taiwan distracts from our main focus].
+More specifically, the Appalachian mountains provide topography that both enhance precipitation during tropical storms as well as provide ample hydrological conditions for severe flooding [@Rees2001].
+Maximum rainfall for Hurricane Fran (1996) occurred in the Appalachians, with flood-peaks reaching recurrance intervals greater than 100 years in many river basins [@Rees2001].
+[JMF--This seems to be true for Floyd, too, even though the following study argues severe flooding would have occurred without the mountain regions] However, @Colle2003 performed an extensive evaluation of Hurricane Floyd (1999) in an effort to understand the mechanisms that led to the development of heavy precipitation and flooding.
 Though differences in the terrain, namely the Appalachian Mountains, contributed to the development of heavy precipitation in this event, @Colle2003 concluded that flooding would have occured in this region regardless of the presence of the Appalachians and other coastal hills.

### 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Info on precipitation over ocean versus land

@@
 
+When assessing rainfall while a hurricane is over the ocean, @Cerveny2000 found that tropical cyclones occurring later in the hurricane season (e.g., October--January) were associated with significantly greater rates of precipitation (rain per day) than earlier in the season when evaluating storms in the North Pacific and Atlantic basins from 1979--1995.
+This obeservation is likely due to the high frequency of intense storms in the later months for the Pacific basin, seeing as the most intense storms often occur in September for the Atlantic basin [@Cerveny2000].
+The rate of rainfall was found to be positively correlated to the wind speed (i.e., maximum surface wind speed) of a particular tropical cyclone, and therefore storm intensity [@Cerveny2000].
+This is somewhat contradictory for the Atlantic basin, which often experiences its most intense storms early in the hurricane season (e.g., September) [@Cerveny2000].[JMF--All of the above results are based solely off observations over the ocean and therefore may not reflect storm characteristics once landfall has been made].
+The relationship between precipitation and wind speed does not remain as strong once the hurricane has made landfall [@Jiang2008].
+
 Rain is a hazard of tropical storms mostly because it causes flooding.

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Rainfall potential parameter

@@
 The relationship between precipitation and wind speed does not remain as strong once the hurricane has made landfall [@Jiang2008].
+In fact, though the intensity of a storm can be considerably reduced once landfall has been made, peak rainfalls overland tend to be greater than those over the ocean [@Jiang2008].
+The area of peak rainfall overland seem to be close to the storm's center, about 50 km [@Jiang2008].
+@Jiang2008 also found statistically significant relationships between decreasing rain rates from the center of the storm out to about 400 km, with no further significant relationship beyond this distance.
+
+Researchers have attempted to predict the expected severity of rainfall for a given storm once it has made landfall using a rainfall potential parameter [@Griffith1978].
+The rainfall potetnial parameter is developed by analyzing characteristics of the storm's evolution over the ocean, including: storm size, translational speed, and average rain rates derived from infrared satellite observations [@Griffith1978].
 

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
precipitation prediction index

--- 2016-06-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 The relationship between precipitation and wind speed does not remain as strong once the hurricane has made landfall [@Jiang2008].
+In fact, though the intensity of a storm can be considerably reduced once landfall has been made, peak rainfalls overland tend to be greater than those over the ocean [@Jiang2008].[JMF--for hurricanes specifically].
+The area of peak rainfall overland seem to be close to the storm's center, within about 50 km [@Jiang2008].
+@Jiang2008 also found statistically significant relationships between decreasing rain rates from the center of the storm out to 400 km, with no further significant relationship beyond this distance.
+
+Researchers have attempted to predict the expected severity of rainfall for a given storm once it has made landfall using a rainfall potential parameter [@Griffith1978].
+The rainfall potetnial parameter is developed by analyzing characteristics of the storm's evolution over the ocean, including: storm size, translational speed, and average rain rates derived from infrared satellite observations [@Griffith1978].
+@Jiang2008 found that the rainfall potential parameters evaluated before landfall were highly correlated with the actual total maximum rainfall of a storm overland, which they used to develop a prediction index based on the rain potential for one day before landfall.
+The proposed index includes five index categories, ranging from less than 149 mm to greater than 505 mm of maximum storm total rain [@Jiang2008].
+Researchers were able to predict land rainfall values accurately for three of six storms in 2005 including Arlene, Katrina, and Rita.
+The predictive model overestimated rainfall for both Cindy and Wilma due to the inability of the model to adjust for changes in the magnitude of rain rate as well as rapid shifts in translational speed before and after landfall [@Jiang2008].
 

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
more references for ET rain distribution

@@
 Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation].
-In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007].
+In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007] or becomes extratropical [JMF-- Look at Elsberry 2002 and Atallah 2002 in the Jiang2008 paper].
 

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
More on ET

--- 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 In comparison to the storm track, the areas of extreme rain and extreme wind can vary.
-For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track [@Halverson2015; @Atallah2007], while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
+For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track [@Halverson2015; @Atallah2007; @Elsberry2002], while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point.
 The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference].
 Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation].
-In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007].
+In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2003; @Atallah2007] or becomes extratropical [@Elsberry2002; JMF-- Look at Elsberry 2002 and Atallah 2002 in the Jiang2008 paper].
 

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
PTSD/MDD and hurricanes

@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Kinney2008: autism prevalence; @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Kinney2008: autism prevalence; @Kilpatrick2007: risk of PTSD and Major Depression in predisposed individuals; @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
 Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000].

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
difficulty in predicting flood risk due to data resolution

--- 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Kinney2008: autism prevalence; @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
+Many outcomes of interest are available at county-level aggregations (e.g., counts of health outcomes: @Grabich2016: premature births and low birth weights; @Grabich2015: birth rates; @Kinney2008: autism prevalence; @Kilpatrick2007: risk of PTSD and Major Depression in predisposed individuals (This is more of a cross-sectional study, I think); @Rappaport2000 and @Czajkowski2011: direct hurricane-related deaths; also, other impacts can be available aggregated by county, including fish kills [@Mallin2006]).
 Further, decisions and policies to prepare for and respond to storms are often undertaken at the county level [@Zandbergen2009; @Rappaport2000].
@@
 
+Currently, precipitation measurements have not become resolute enough to quantitatively predict flooding or estimate flooding risks for a storm [@Elsberry2002; @Emanuel2006].[JMF--I need to keep digging to see if this is still true]
+
 Storm-motion, as well as the direction from which a storm approaches, can play significant roles in determining the extent of precipitation produced by a storm [@Rees2001; @Chang2013].

### 2016-07-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add more on tropical cyclone-induced flooding

@@
 There is some evidence that the worst tropical storms, in terms of heavy rainfall and freshwater flood-related deaths in the US, tend to occur early in the hurricane season (e.g., June--August), when it is more common to get the meteorological conditions (weak steering currents) that lead to slow-moving storms [@Rappaport2000] (although, conversely, some have suggested that a storm with the same amount of rain is likely to cause less flooding earlier in the season, since more water can evaporate in the hotter conditions more common early in the hurricane season [@Mallin2006]).
+@Smith2011 conducted an analysis of 572 USGS sream gauging stations with more than 75 years of data in the Eastern United States in an effort to better understand the etiology of severe flooding in the region, a major component of which are those induced by tropical cyclones.
+A large portion of these include extratropical storms in the winter--spring season (e.g.
+March--April), with extratropical systems occurring earlier in the season being less of a risk for flooding events due to mechanistic differences [@Smith2011].
+Among tropical cyclone-induced floods is significant spacial heterogeneity, as well as substantial variability in frequency each year with no obvious long-term trend [@Smith2011].
 

### 2016-07-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
More from Smith 2011

@@
 Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation].
-In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2003; @Atallah2007] or becomes extratropical [@Elsberry2002; JMF-- Look at Elsberry 2002 and Atallah 2002 in the Jiang2008 paper].
+In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2003; @Atallah2007] or becomes extratropical [@Elsberry2002].
 
@@
 March--April), with extratropical systems occurring earlier in the season being less of a risk for flooding events due to mechanistic differences [@Smith2011].
-Among tropical cyclone-induced floods is significant spacial heterogeneity, as well as substantial variability in frequency each year with no obvious long-term trend [@Smith2011].
+Among tropical cyclone-induced floods is significant spacial heterogeneity, largely due to orographic mechanisms and other parameters associated with the encouragement of flooding, as well as substantial variability in frequency each year with no obvious long-term trend [@Smith2011].
+Maximum values were located along the Appalacian Mountains [@Smith2011].
+Further, @Smith2011 suggests that the clustering of Atlantic basin storms may play an important role in the induction of floods in this region.
+Maximum rainfall values were found to be most prevalent east of the Appalachians [@Smith2011].
 

### 2016-07-05 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add comment on rainfall and storm intensity

@@
 The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference].
-Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation].
+Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?--I think so, Medlin 2007 uses cyclonically so this seems correct to me] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation].
 In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2003; @Atallah2007] or becomes extratropical [@Elsberry2002].
@@
 Storms with a slower forward speed and larger rainfall area contribute to a longer duration of hazardous conditions, and therefore an increased chance of storm-related damage [@Rezapour2014].
+
+Tropical storm Claudette (1979) and Hurricane Danny (1997), a minimal hurricane, are thought to be associated with the highest accumulation of rainfall in history thus far [@Medlin2007], suggesting storm intensity alone is likely not appropriate in assessing or predicting the impact of precipitation.
 

### 2016-07-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add some edits to Intro, Methods, and start of Results.

@@
 
-While the risks to life from some tropical storm threats (e.g., storm surge) have been greatly reduced over the past century, other risks (e.g., inland flooding) have remained more steady [@Rappaport2000].
+While the risks to life from some tropical storm threats (e.g., storm surge) have been greatly reduced over the past century, other risks (e.g., inland flooding) have remained more steady [@Rappaport2000], and costs related to property damage have increased [citation].
 Determining the metrics of storm exposure that are most associated with loss of life and property damage may help identify and emphasize the important threats that remain from tropical storms in the United States, which could allow similar future success in reducing these threats through community planning, warning systems, etc.
@@
 
-The most often used hurricane exposure criteria at the county-level include FEMA disaster declarations and area within the storm track, which can be somewhat nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016].
+[GBA-- Somewhere (maybe just for helping us draft the paper), let's come up with a list of the "ideal" characteristics of an exposure metric for hurricanes.
+I think that will be helpful for us to have a framework of a "gold standard" to compare advantages and disadvantages of different metrics used.
+To start, we could create a list of "good characteristics" and another of "bad characteristics", and read through the text we have and pull out characteristics we describe as good or bad aspects of different metrics.]
+
+The most often used hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used?
+We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure] and whether a county's perimeter intercepts a swath describing a buffer a certain distance from the storm's central track.
+These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016].
 FEMA declaration is largely subjective, requiring state governors and/or tribal chiefs to submit a proposal for government aid on the basis that local resources are overwhelmed [@FEMA].
 As a result, using FEMA disaster declarations as exposure classification to hurricanes restricts exposure to 1) those states whose resources are overwhelmed and 2) those states who apply for FEMA aid.
-As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane].
+As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane; GBA-- let's think about why that might have had "false positives".
+I think that would be an interesting point to include here.].
 Researchers have found heterogeneity in exposure assignments of counties, both within and between storms, when different metrics of exposure are used [@Grabich2016; @Grabich2015].
@@
 
-When assessing exposure to a single tropical storm, a detailed analysis of ...
-can be performed for the storm.
+When assessing exposure to a single tropical storm, a detailed *ad hoc* analysis can be performed for the single storm.
 However, this level of detailed analysis is unreasonable when assessing exposure to all storms over an extended time period; further, such a long-term exposure assessment requires consistency in the data being used to determine levels of exposure to factors like wind and rain.
@@
 Databases of storm impacts, including direct deaths [@Rappaport2000], ..., and ..., are often aggregated at the county level.
+[Let's see if we can find other examples of county-level outcomes that researchers might want to pair with hurricane exposure.
+For example, property damage, crop damage, injuries, ...
+These are all given in NOAA's Storm Events database, but let's also see if we can find studies that have analyzed these outcomes in relation to tropical storms, if we can.]
 
@@
 These positions are given to within 0.1 degrees latitude / longitude at these synoptic times [@Landsea2013].
-We interpolated these location values to every 15 minutes during the period when the storm was active, using a linear interpolation between each measured point.
+We interpolated these location values to every 15 minutes during the period when the storm was active, using a linear interpolation [should we somehow be using great circle distance to do this?] between each measured point.
 
 We calculated the distance between each county's center and each of the 15-minute-interval estimates of a storm's location.
-To do this, we used the Great Circle method to calculate the distance between pairs of latitude and longitude coordinates, using the R package `sp`.
+To do this, we used the Great Circle method to calculate the distance between pairs of latitude and longitude coordinates [citation-- R package `sp`].
 For each county, we used the population mean center, based on the U.S.
@@
 From these distance calculations, we identified the date-time and distance between the storm track and the county center at the 15-minute interval when the storm was closest to the county center.
-We identified the date when the storm was closest to each county and converted this date-time  to the county's local time zone using the `countytimezones` R package [citation].
-This "closest date" was used to pair the distance estimates with rainfall estimates and can also be used to merge exposure datasets with daily aggregated outcomes like daily county-level mortality and morbidity counts.
+We identified the date when the storm was closest to each county and converted this date-time  to the county's local time zone [citation-- `countytimezones` R package].
+This "closest date" was used to pair the distance estimates with rainfall estimates for a specific storm and county.
+Futher, this "closest date" provides a county-specific estimate of the date of storm exposure and so can also be used to merge exposure datasets with daily aggregated outcomes like daily county-level mortality and morbidity counts.
 
@@
 
+### Analysis of hurricane exposure classification methods
+
+[We'll want to describe the methods for the analyses we perform here, as well as which set of storms we use in our analyses.]
+
 Results
 ============
+
+Our analysis included [x] Atlantic basin tropical storms over the years 1988 to 2011. This dataset includes several noteworthy storms, including ... .
+
+[As a summary of our available dataset, do we want to include a US map showing the storm tracks for all of the storms included in our analysis?]
+
+```{r}
+# Include code for map of tracks of all storms in study here. 
+# I think there's a function in `hurricaneexposure` we can use for this.
+```
 
@@
 
-Map of the count of county exposures to the top 1000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
+Map of the count of county exposures to the top 1000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less [GBA-- These are **awesome**.
+I think we want to tweak the color scale and the legend a bit so that they're consistent between the three plots showing the top 1000 storm-county pairs for the three metrics (distance, rain, and wind).
+That will make it easier to compare across the three maps, which is what we want readers to be able to do here.]:
 
@@
 data(closest_dist)
+
+# Generate a list of unique study fips (all in the eastern US)
 my_fips <- closest_dist$fips
 my_fips <- unique(my_fips)
-distance_exposure <- county_distance(counties = my_fips,
-                                 start_year = 1988, end_year = 2011,
-                                 dist_limit = 100)
+
+# Get distance-based exposure for all study counties, all years, using a 
+# distance limit that will ensure at least 1000 storm-county pairs
+distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2011, dist_limit = 100)
+
 distance_1000 <- distance_exposure %>%
@@
         slice(1:1000)
+
 distance_1000$value <- cut(distance_1000$count, c(0, 5, 8, 12, 15, 18, 21, 24),
-                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18", "19-21", "22-24"),
+                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18",
+                                      "19-21", "22-24"),
                            include.lowest = TRUE)

### 2016-07-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Make some tweaks to the "5,000" plots for rain and distance

@@
 Map of the count of county exposures to the top 1000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less [GBA-- These are **awesome**.
-I think we want to tweak the color scale and the legend a bit so that they're consistent between the three plots showing the top 1000 storm-county pairs for the three metrics (distance, rain, and wind).
+I think we want to tweak the color scale and the legend a bit so that they're consistent between the three plots showing the top 5000 storm-county pairs for the three metrics (distance, rain, and wind).
 That will make it easier to compare across the three maps, which is what we want readers to be able to do here.]:
@@
 
-distance_1000 <- distance_exposure %>%
-        arrange(desc(storm_dist)) %>%
+# distance_1000 <- distance_exposure %>%
+#         arrange(desc(storm_dist)) %>%
+#         group_by(fips) %>%
+#         summarise(count = n()) %>%
+#         rename(region = fips) %>%
+#         mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+#         arrange(desc(count)) %>%
+#         slice(1:1000)
+
+distance_5000 <- distance_exposure %>%
+        arrange(storm_dist) %>% # For distance, I don't think we want `desc`
+        slice(1:5000) %>% # I think we need to slice out the top 1000 before 
+                          # counting up by county. We want the top 1000
+                          # county-storm pairs, not to count up the numbers for
+                          # all storms within 100 km and then take the top 1000
+                          # counties from that (which is what the old code did)
         group_by(fips) %>%
@@
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-        arrange(desc(count)) %>%
-        slice(1:1000)
+        mutate(region = as.numeric(region), count = as.numeric(count))
 
-distance_1000$value <- cut(distance_1000$count, c(0, 5, 8, 12, 15, 18, 21, 24),
-                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18",
-                                      "19-21", "22-24"),
+distance_5000$value <- cut(distance_5000$count, c(0, 3, 6, 9, 12, 15, 17), 
+                   labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
                    include.lowest = TRUE)
@@
                     "west virginia", "wisconsin")
-map_dist_1000 <- county_choropleth(distance_1000, 
-                              title = "Hits for top 1000 Storm-County Pairs\nfor Distance < 100 km",
-                              legend = "Hits per County", num_colors = 7,
+map_dist_5000 <- county_choropleth(distance_5000, 
+                              title = "Hits for top 5000 Storm-County Pairs\nfor Distance < 100 km",
+                              legend = "Hits per County", # num_colors = 7,
                               state_zoom = eastern_states)
-map_dist_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGn", na.value = "white")
+map_dist_5000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu",
+                                  na.value = "white")
 ```
@@
 ```{r echo = FALSE, warning=FALSE, message=FALSE}
+# Get the rain-based exposure for all counties over the study period, 
+# using a rain limit of 75 mm and a distance limit of 500 km
 rain_exposure <- county_rain(counties = my_fips,
@@
                              days_include = c(-1, 0, 1))
-rain_1000 <- rain_exposure %>%
+
+# rain_1000 <- rain_exposure %>%
+#         arrange(desc(tot_precip)) %>%
+#         group_by(fips) %>%
+#         summarise(count = n()) %>%
+#         rename(region = fips) %>%
+#         mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+#         arrange(desc(count)) %>%
+#         slice(1:1000)
+rain_5000 <- rain_exposure %>%
         arrange(desc(tot_precip)) %>%
+        slice(1:5000) %>%
         group_by(fips) %>%
@@
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-        arrange(desc(count)) %>%
-        slice(1:1000)
-rain_1000$value <- cut(rain_1000$count, c(0, 3, 6, 9, 12, 15, 17), 
+        mutate(region = as.numeric(region), count = as.numeric(count))
+
+rain_5000$value <- cut(rain_5000$count, c(0, 3, 6, 9, 12, 15, 17), 
                    labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
                    include.lowest = TRUE)
-map_1000 <- county_choropleth(rain_1000,
-                              title = "Hits for top 1000 Storm-County Pairs\nfor Rain > 75 mm",
+map_5000 <- county_choropleth(rain_5000,
+                              title = "Hits for top 5000 Storm-County Pairs\nfor Rain > 75 mm",
                               legend = "Hits per County", num_colors = 6,
                               state_zoom = eastern_states)
-map_1000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
+map_5000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu",
+                             na.value = "white")
 ```

### 2016-07-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on maps in Results

@@
 library(choroplethr)
+library(RColorBrewer)
 ```
@@
 
-Map of the count of county exposures to the top 1000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less [GBA-- These are **awesome**.
-I think we want to tweak the color scale and the legend a bit so that they're consistent between the three plots showing the top 5000 storm-county pairs for the three metrics (distance, rain, and wind).
+Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less [GBA-- These are **awesome**.
+I think we want to tweak the color scale and the legend a bit so that they're consistent between the three plots showing the top 2,000 storm-county pairs for the three metrics (distance, rain, and wind).
 That will make it easier to compare across the three maps, which is what we want readers to be able to do here.]:
@@
 
-distance_5000 <- distance_exposure %>%
+distance_2000 <- distance_exposure %>%
         arrange(storm_dist) %>% # For distance, I don't think we want `desc`
-        slice(1:5000) %>% # I think we need to slice out the top 1000 before 
+        slice(1:2000) %>% # I think we need to slice out the top 1000 before 
                           # counting up by county. We want the top 1000
@@
 
-distance_5000$value <- cut(distance_5000$count, c(0, 3, 6, 9, 12, 15, 17), 
-                   labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
+distance_2000$value <- cut(distance_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
                    include.lowest = TRUE)
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+
 eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
@@
                     "west virginia", "wisconsin")
-map_dist_5000 <- county_choropleth(distance_5000, 
-                              title = "Hits for top 5000 Storm-County Pairs\nfor Distance < 100 km",
-                              legend = "Hits per County", # num_colors = 7,
-                              state_zoom = eastern_states)
-map_dist_5000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu",
+
+map_dist_2000 <- CountyChoropleth$new(distance_2000)
+map_dist_2000$set_zoom(eastern_states)
+map_dist_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
                                                 na.value = "white")
+map_dist_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance < 100 km"
+a <- map_dist_2000$render()
 ```
-
-Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:
-
-```{r echo = FALSE}
-# Code for this map here
-```
-
-Map of county exposures to the top 1000 county-storm pairs by rain-based exposure:
 
@@
 #         slice(1:1000)
-rain_5000 <- rain_exposure %>%
+rain_2000 <- rain_exposure %>%
         arrange(desc(tot_precip)) %>%
-        slice(1:5000) %>%
+        slice(1:2000) %>%
         group_by(fips) %>%
@@
 
-rain_5000$value <- cut(rain_5000$count, c(0, 3, 6, 9, 12, 15, 17), 
-                   labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
+rain_2000$value <- cut(rain_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
                    include.lowest = TRUE)
-map_5000 <- county_choropleth(rain_5000,
-                              title = "Hits for top 5000 Storm-County Pairs\nfor Rain > 75 mm",
-                              legend = "Hits per County", num_colors = 6,
-                              state_zoom = eastern_states)
-map_5000 + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu",
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+
+map_rain_2000 <- CountyChoropleth$new(rain_2000)
+map_rain_2000$set_zoom(eastern_states)
+map_rain_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
                                                 na.value = "white")
+map_rain_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain > 75 mm"
+b <- map_rain_2000$render()
 ```
 
-Map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?
-What seems reasonable?; JMF-- Based on what I have read so far, I think this seems reasonable, unless we want to push it out to 120 km to loosely reflect the Czajkowski study which seems to have found a great number of fatalities in tertiary counties] kilometers or less:
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+grid.arrange(a, b, ncol = 1)
+```
+
+Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:
+
+```{r echo = FALSE}
+# Code for this map here
+```
+
+Here are a map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?
+What seems reasonable?; JMF-- Based on what I have read so far, I think this seems reasonable, unless we want to push it out to 120 km to loosely reflect the Czajkowski study which seems to have found a great number of fatalities in tertiary counties] kilometers or less and a map of total count of "exposures" by county, based on "exposed" = rainfall over a three-day window of 75 millimeters or more and distance of 500 kilometers or less:
 
@@
 distance_count <- distance_exposure %>%
-        arrange(desc(storm_dist)) %>%
         group_by(fips) %>%
@@
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-        arrange(desc(count))
-distance_count$value <- cut(distance_count$count, c(0, 5, 8, 12, 15, 18, 21, 24),
-                           labels = c("0-5", "6-8", "9-12", "13-15", "16-18", "19-21", "22-24"),
+        mutate(region = as.numeric(region), count = as.numeric(count)) 
+distance_count$value <- cut(distance_count$count, c(0, 5, 8, 12, 15,
+                                                    18, 21, 24),
+                           labels = c("0-5", "6-8", "9-12", "13-15",
+                                      "16-18", "19-21", "22-24"),
                            include.lowest = TRUE)
-map_dist_count <- county_choropleth(distance_count, 
-                              title = "Hits by County\nfor Distance < 100 km",
-                              legend = "Hits per County", num_colors = 7,
-                              state_zoom = eastern_states)
-map_dist_count + scale_fill_brewer("Hits per\nCounty", palette = "YlGn", na.value = "white")
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+
+map_dist_count <- CountyChoropleth$new(distance_count)
+map_dist_count$set_zoom(eastern_states)
+map_dist_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_dist_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance < 100 km"
+a <- map_dist_count$render()
 ```
-
-Map of total count of "exposures" by county, based on "exposed" = rainfall over a three-day window of 75 millimeters or more and distance of 500 kilometers or less:
 
@@
               mutate(region = as.numeric(region), count = as.numeric(count))
-rain_count$value <- cut(rain_count$count, c(0, 3, 6, 9, 12, 15, 17), 
-                   labels = c("0-3", "4-6", "7-9", "10-12", "13-15", "16-17"), 
+rain_count$value <- cut(rain_count$count, c(0, 5, 8, 12, 15,
+                                                    18, 21, 24),
+                           labels = c("0-5", "6-8", "9-12", "13-15",
+                                      "16-18", "19-21", "22-24"),
                            include.lowest = TRUE)
-map_counts <- county_choropleth(rain_count,
-                                title = "Hits by County\nfor Rain > 75 mm",
-                                legend = "Hits per County", num_colors = 6,
-                                state_zoom = eastern_states)
-map_counts + scale_fill_brewer("Hits per\nCounty", palette = "YlGnBu", na.value = "white")
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+
+map_rain_count <- CountyChoropleth$new(rain_count)
+map_rain_count$set_zoom(eastern_states)
+map_rain_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_rain_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance > 75 mm"
+b <- map_rain_count$render()
+```
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+grid.arrange(a, b, ncol = 1)
 ```

### 2016-07-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add more code for figures and results

@@
 
-```{r}
-# Include code for map of tracks of all storms in study here. 
-# I think there's a function in `hurricaneexposure` we can use for this.
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+# Map of tracks of all storms in study.
+data("closest_dist")
+storms <- unique(closest_dist$storm_id)
+storms <- storms[which(gsub("*.+-", "", storms) <= 2011)]
+map_tracks(storms, plot_points = FALSE, alpha = 0.3)
 ```
@@
 
-```{r}
+```{r echo = FALSE}
 exposure_name <- c("Rain_Exposure", "Distance_Exposure")
@@
 grid.arrange(rain_plots$a, rain_plots$b, rain_plots$c, rain_plots$d, ncol = 2)
+```
+
+Here are the top storms when measured by summing all rainfall for all counties exposed to the storm (where "exposed" means the county was within 500 km of the storm track and had 75 mm or more of rainfall over the three-day window when the storm was closest to the county).
+
+```{r echo = FALSE}
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500,
+                             days_include = c(-1, 0, 1))
+
+rain_exposure %>%
+        group_by(storm_id) %>%
+        summarise(sum_rain = round(sum(tot_precip), -2)) %>%
+        arrange(desc(sum_rain)) %>%
+        slice(1:10) %>%
+        kable(col.names = c("Storm", "Summed county-level rainfall"),
+              caption = "Top 10 storms based on summed rainfall across all counties exposed, using a metric of rainfall.")
+```
+
+Here is a map of the rain totals for Frances:
+
+```{r echo = FALSE}
+map_tracks(storms = "Frances-2004", plot_points = FALSE, 
+                plot_object = map_counties(storm = "Frances-2004", 
+                                           metric = "rainfall")) + 
+        ggtitle(paste("Frances-2004", "rainfall"))
 ```

### 2016-07-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work some on text

@@
 Determining the metrics of storm exposure that are most associated with loss of life and property damage may help identify and emphasize the important threats that remain from tropical storms in the United States, which could allow similar future success in reducing these threats through community planning, warning systems, etc.
+
+It is critical to characterize the risks of cyclonic storms in order to help individuals and communities prepare for and respond to storms.
+By characterizing the health risks of other climate-related disasters, communities have improved their responses in ways tailored to protect human health.
+For example, in response to the health threats of heat, several US cities—including Philadelphia, PA; Washington, DC; Phoenix, AZ; Dallas-Fort Worth, TX; Chicago, IL; and New Orleans, LA—have implemented heat watch-warning systems specifically designed to limit all-cause heat-related health outcomes [@Sheridan2004].
+An alert of these heat-health warning systems can instigate a number of health-protecting responses including: halting suspensions of electric and water utilities; increasing the number of Fire Department Emergency Medical Service staff on duty; and issuing media announcements [@Sheridan2004].
+
+According to the Intergovernmental Panel on Climate Change, climate change may alter patterns of climate-related natural disasters [@Seneviratne2012 -- double-check formatting for this reference], while future population movement may continue to increase the number of people at risk from natural disasters (e.g., movement of retirees to Florida [@Smith1992]).
+Research assessing the risks of tropical storms will aid city- and national-level planning for storms, both in the short-term, to prepare for specific current storms, and in the long-term, to prepare for the health risks of climate change.
 
@@
                                                 na.value = "white")
-map_rain_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance > 75 mm"
+map_rain_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain > 75 mm"
 b <- map_rain_count$render()

### 2016-07-10 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Make some small changes to figures and tables in paper draft

@@
                                                 na.value = "white")
-map_dist_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance < 100 km"
+map_dist_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance Metric"
 a <- map_dist_2000$render()
@@
                                                 na.value = "white")
-map_rain_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain > 75 mm"
+map_rain_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain Metric"
 b <- map_rain_2000$render()
@@
                                                 na.value = "white")
-map_dist_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance < 100 km"
+map_dist_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance <= 100 km"
 a <- map_dist_count$render()
@@
                                                 na.value = "white")
-map_rain_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain > 75 mm"
+map_rain_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain > 75 mm and Distance <= 500 km"
 b <- map_rain_count$render()
@@
 
-Here are the top storms when measured by summing all rainfall for all counties exposed to the storm (where "exposed" means the county was within 500 km of the storm track and had 75 mm or more of rainfall over the three-day window when the storm was closest to the county).
+Here are the top storms when measured by summing all rainfall for all eastern US counties exposed within 500 km of the storm track (rainfall totals are for the three-day window when the storm was closest to each county).
 
@@
                              start_year = 1988, end_year = 2011,
-                             rain_limit = 75, dist_limit = 500,
+                             rain_limit = 0, dist_limit = 500,
                              days_include = c(-1, 0, 1))
@@
         kable(col.names = c("Storm", "Summed county-level rainfall"),
-              caption = "Top 10 storms based on summed rainfall across all counties exposed, using a metric of rainfall.")
+              caption = "Top 10 storms based on summed rainfall across all eastern US counties within 500 km of the storm track.")
 ```

### 2016-07-11 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add text to paper draft

@@
 
+Flash floods cause most of the flood-related deaths in the eastern US [@Ashley2008], and one study found that most flash flooding in the eastern half of the US results from extreme rainfall over a period of six hours or less [@Ashley2008], which suggests that are exposure window of the three days around a storm may be wide enough to capture the most dangerous exposures.
+
 Other studies, like ours, do not attempt to identify and remove any precipitation over the aggregating period associated with weather systems other than the tropical storm (e.g., @Czajkowski2011).

### 2016-07-12 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add code to develope 2x2 table

--- 2016-07-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```
-
+```{r}
+rain_exposure$county_storm <- transmute(rain_exposure, county_storm = paste0(fips, " ", storm_id))
+distance_exposure$county_storm <- transmute(distance_exposure, county_storm = paste0(fips, " ", storm_id))
+all_exposure <- full_join(rain_exposure, distance_exposure, by = "county_storm")
+both_exposure <- subset(all_exposure, tot_precip >= 75 & storm_dist.x <= 100) #4040 observations
+rain_only <- subset(all_exposure, tot_precip >= 75 & storm_dist.x > 100) #4200 obs
+distance_only <- subset(all_exposure, tot_precip < 75 & storm_dist.x <= 100)
+sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x <= 100, na.rm = TRUE)
+sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x > 100, na.rm = TRUE)
+sum(all_exposure$tot_precip < 75 & all_exposure$storm_dist.x <= 100, na.rm = TRUE) # It appears that in all county-
+                                                                                   # storm pairs, there is ample
+                                                                                   # rain to meet criteria
+```
 One interesting contrast is between storms that were noteworthy for direct wind-related deaths (Anderew in 1992, Hugo in 1989) versus storms noteworthy for direct freshwater flooding-related deaths (Floyd in 1999, Alberto in 1994, Charley in 1998) for the 1970--1999 period [@Rappaport2000].

### 2016-07-12 writing/DraftExposurePaper/DraftExposurePaper.Rmd
exposure table problems

--- 2016-07-11 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```
+```{r}
+rain_exposure$county_storm <- transmute(rain_exposure, county_storm = paste0(fips, " ", storm_id))
+distance_exposure$county_storm <- transmute(distance_exposure, county_storm = paste0(fips, " ", storm_id))
+all_exposure <- full_join(rain_exposure, distance_exposure, by = "county_storm")
+both_exposure <- subset(all_exposure, tot_precip >= 75 & storm_dist.x <= 100) #4040 observations
+rain_only <- subset(all_exposure, tot_precip >= 75 & storm_dist.x > 100) #4200 obs
+distance_only <- subset(all_exposure, tot_precip < 75 & storm_dist.x <= 100) #0 obs
+sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x <= 100, na.rm = TRUE)
+sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x > 100, na.rm = TRUE)
+sum(all_exposure$tot_precip < 75 & all_exposure$storm_dist.x <= 100, na.rm = TRUE) # It appears that in all county-
+                                                                                   # storm pairs, there is ample
+                                                                                   # rain to meet criteria        
+sum(all_exposure$tot_precip < 75, na.rm = TRUE) # This also returns 0
+```
+
+JMF-- I was surprised to see that there were no observations in the data frame that had total precipitation less than 75 mm. I am wondering if I am missing something in the `full_join` code
 

### 2016-07-12 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Run checks on exposure table

@@
 sum(all_exposure$tot_precip < 75, na.rm = TRUE) # This also returns 0
+#Test
+rain_exposure_2 <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 100,
+                             days_include = c(-1, 0, 1)) # Creates a data frame with 4040 obs (expected)
+distance_rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 0, dist_limit = 100,
+                             days_include = c(-1, 0, 1)) 
+sum(distance_rain_exposure$tot_precip >= 75) # 4040 obs (meet both criteria)
+sum(distance_rain_exposure$tot_precip < 75 & distance_rain_exposure$storm_dist <= 100) #7103 obs meet distance but 
+                                                                                       #not rain criteria
 ```

### 2016-07-12 R/distance_exposure.R
Close #39

@@ -169,7 +169,8 @@ distance_exposure <- function(locations, start_year, end_year,
                                   start_year = start_year,
                                   end_year = end_year,
                                   dist_limit = dist_limit) %>%
-                        dplyr::rename_(loc = ~ commun)
+                        dplyr::rename_(loc = ~ commun) %>%
+                        dplyr::ungroup()
         } else {
                 df <- county_distance(counties = locations,
                                   start_year = start_year,

### 2016-07-12 R/rain_exposure.R

@@ -271,7 +271,8 @@ rain_exposure <- function(locations, start_year, end_year,
                                   rain_limit = rain_limit,
                                   dist_limit = dist_limit,
                                   days_included = days_included) %>%
-                        dplyr::rename_(loc = ~ commun)
+                        dplyr::rename_(loc = ~ commun) %>%
+                        dplyr::ungroup()
         } else {
                 df <- county_rain(counties = locations,
                                   start_year = start_year,
@@ -286,7 +287,7 @@ rain_exposure <- function(locations, start_year, end_year,
         for(i in 1:length(locs)){
                 out_df <- df %>%
                         dplyr::filter_(~ loc == locs[i]) %>%
-                        dplyr::select_('-loc')
+                        dplyr::select_(.dots = list(quote(-loc)))
                 out_file <- paste0(out_dir, "/", locs[i], ".", out_type)
                 if(out_type == "rds"){
                         saveRDS(out_df, file = out_file)

### 2016-07-12 writing/DraftExposurePaper/DraftExposurePaper.Rmd

--- 2016-07-12 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 JMF-- I was surprised to see that there were no observations in the data frame that had total precipitation less than 75 mm. I am wondering if I am missing something in the `full_join` code
+
+Below, I tried recreating the above data set using different parameters to see if I could get the exposure counts to match up okay:
+
+```{r}
+#Test
+rain_exposure_2 <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 100,
+                             days_include = c(-1, 0, 1)) # Creates a data frame with 4040 obs (expected)
+distance_rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 0, dist_limit = 100,
+                             days_include = c(-1, 0, 1)) 
+sum(distance_rain_exposure$tot_precip >= 75) # 4040 obs (meet both criteria)
+sum(distance_rain_exposure$tot_precip < 75 & distance_rain_exposure$storm_dist <= 100) #7103 obs meet distance but 
+                                                                                       #not rain criteria
+```
+
+JMF-- In the above, using the `distance_rain_exposure` data frame, I found that there were 7103 storm-county pairs that did not meet the precipitation criteria ($\leq 75 mm$) but did meet the distance createria ($\leq 100 km$), contrary to the 0 I found in the code before this with the `all_exposure` data frame.
 

### 2016-07-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add to draft

@@
 ```
+
+#### Are our methods of assessing distance reasonable
+
+[Spline vs. linear interpolation? Is 15 minutes adequate? Should we use shapefiles instead? IBTrACS probably has this (`rnoaa`).]
+
+#### Are our methods of assessing rainfall reasonable
+
+[Is three-day window reasonable?
+Probably should be.
+If we go larger, there's the problem of picking up things that aren't related to the storm.
+Storms to check-- Floyd (ET), Irene (ET), Lee (slow moving), Ike (for timing), Faye, Allison]
+
+#### Are our methods of assessing wind speed reasonable
 

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
df for wind exposure
Data sets for top storm county pairs for wind

@@
 ```{r echo = FALSE}
+#TS winds and above
+TS_wind_exposure <- county_wind(counties = my_fips,
+                             start_year = 1988,
+                             end_year = 2011,
+                             wind_limit = 18,
+                             wind_var = "max_sust")
+#HU exposures (32.1 mps and above)
+HU_wind_exposure <- county_wind(counties = my_fips,
+                                start_year = 1988,
+                                end_year = 2011,
+                                wind_limit = 32.1,
+                                wind_var = "max_sust")
+TS_wind_2000 <- TS_wind_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max is 1 for hits for top Storm-County
+                                                                       # pairs
+test <- TS_wind_exposure %>%         
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+        arrange(desc(count)) %>%
+        slice (1:2000) #This gives a max of 17 for number of counts for the top 2000 counties
+HU_wind_2000 <- HU_wind_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max of 1 again for hits for top Storm-
+                                                                       # County pairs
 # Code for this map here 

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Plot wind exposure

--- 2016-07-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r echo = FALSE}
-# Code for this map here
+#TS winds and above
+TS_wind_exposure <- county_wind(counties = my_fips,
+                             start_year = 1988,
+                             end_year = 2011,
+                             wind_limit = 18,
+                             wind_var = "max_sust")
+#HU exposures (32.1 mps and above)
+HU_wind_exposure <- county_wind(counties = my_fips,
+                                start_year = 1988,
+                                end_year = 2011,
+                                wind_limit = 32.1,
+                                wind_var = "max_sust")
+TS_wind_2000 <- TS_wind_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max is 1 for hits for top Storm-County
+                                                                       # pairs
+test <- TS_wind_exposure %>%         
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+        arrange(desc(count)) %>%
+        slice (1:2000) #This gives a max of 17 for number of counts for the top 2000 counties
+HU_wind_2000 <- HU_wind_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max of 1 again for hits for top Storm-
+                                                                       # County pairs
+TS_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+                   include.lowest = TRUE)
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+map_TS_2000 <- CountyChoropleth$new(TS_wind_2000)
+map_TS_2000$set_zoom(eastern_states)
+map_TS_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_TS_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor TS Wind Metric"
+a <- map_TS_2000$render()
+HU_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+                   include.lowest = TRUE)
+map_HU_2000 <- CountyChoropleth$new(HU_wind_2000)
+map_HU_2000$set_zoom(eastern_states)
+map_HU_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_HU_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor HU Wind Metric"
+b <- map_HU_2000$render()
+```
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+grid.arrange(a, b, ncol = 1)
 ```

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Maps for different wind exposure categories

--- 2016-07-16 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r echo = FALSE}
-# Code for this map here
+#TS winds and above
+TS_wind_exposure <- county_wind(counties = my_fips,
+                             start_year = 1988,
+                             end_year = 2011,
+                             wind_limit = 18,
+                             wind_var = "max_sust")
+#HU exposures (32.1 mps and above)
+HU_wind_exposure <- county_wind(counties = my_fips,
+                                start_year = 1988,
+                                end_year = 2011,
+                                wind_limit = 32.1,
+                                wind_var = "max_sust")
+TS_wind_2000 <- TS_wind_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max is 1 for hits for top Storm-County
+                                                                       # pairs
+test <- TS_wind_exposure %>%         
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
+        arrange(desc(count)) %>%
+        slice (1:2000) #This gives a max of 17 for number of counts for the top 2000 counties
+HU_wind_2000 <- HU_wind_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max of 1 again for hits for top Storm-
+                                                                       # County pairs
+TS_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+                   include.lowest = TRUE)
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+map_TS_2000 <- CountyChoropleth$new(TS_wind_2000)
+map_TS_2000$set_zoom(eastern_states)
+map_TS_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_TS_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor TS Wind Metric"
+a <- map_TS_2000$render()
+HU_wind_2000$value <- cut(HU_wind_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+                   include.lowest = TRUE)
+map_HU_2000 <- CountyChoropleth$new(HU_wind_2000)
+map_HU_2000$set_zoom(eastern_states)
+map_HU_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_HU_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor HU Wind Metric"
+b <- map_HU_2000$render()
 ```
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r}
+#Category 3 and above 
+cat3_exposure <- county_wind(counties = my_fips,
+                                start_year = 1988,
+                                end_year = 2011,
+                                wind_limit = 50,
+                                wind_var = "max_sust")
+cat3_2000 <- cat3_exposure %>% 
+        arrange(desc(max_sust)) %>%
+        slice (1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count))
+cat3_2000$value <- cut(cat3_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
+                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+                   include.lowest = TRUE)
+map_cat3_2000 <- CountyChoropleth$new(cat3_2000)
+map_cat3_2000$set_zoom(eastern_states)
+map_cat3_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_cat3_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Category 3 and Above Wind Metric"
+c <- map_cat3_2000$render()
+```
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE}
+grid.arrange(a, b, c, ncol = 1)
+```
+
+JMF-- I do not think that using the top 2000 Storm-County pairs will be too descriptive (All 3 maps have the identical 2000 pairs), any ideas?
 

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add wind exposure table

@@
 
+Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 18 m/s or greater (e.g., tropical storm strength or greater):
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+data("closest_dist")
+all_counties <- unique(closest_dist$fips)
+
+all_wind <- county_wind(counties = all_counties, start_year = 1988, 
+                        end_year = 2011, wind_limit = 18)
+
+wind_exposed <- all_wind %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
+wind_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of wind.")
+```
+#### Is our 3-day window for rian exposure reasonable?
+
+```{r}
+
+```
+
 #### Are storms with heavy rain exposure likely to occur early in the hurricane season?

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Map wind exposure

@@
 ```
+
+Here are plots of the top four storms based on number of counties exposed by wind:
+
+```{r echo = FALSE, fig.width = 10, fig.height = 8}
+top_wind_exposed <- function(storm_id){
+        my_map <- map_wind_exposure(storm_id, wind_limit = 18)
+        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
+                             plot_points = FALSE) + 
+                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
+}
+
+n <- 4
+wind_plots <- lapply(wind_exposed$storm_id[1:n], top_wind_exposed)
+names(wind_plots) <- letters[1:n]
+grid.arrange(wind_plots$a, wind_plots$b, wind_plots$c,
+             wind_plots$d, ncol = 2)
+```
 #### Is our 3-day window for rian exposure reasonable?

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
wind tables

@@
 ```
+
+Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 33 m/s or greater (e.g., Hurricane strength, category 1 or greater):
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+data("closest_dist")
+all_counties <- unique(closest_dist$fips)
+
+all_wind <- county_wind(counties = all_counties, start_year = 1988, 
+                        end_year = 2011, wind_limit = 33)
+
+hu_exposed <- all_wind %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
+hu_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of wind.")
+```
+
+Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 50 m/s or greater (e.g., Intense Hurricane strength, category 3 or greater):
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+data("closest_dist")
+all_counties <- unique(closest_dist$fips)
+
+all_wind <- county_wind(counties = all_counties, start_year = 1988, 
+                        end_year = 2011, wind_limit = 50)
+
+cat3_exposed <- all_wind %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
+cat3_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of wind.")
+```
+
+For all of the above exposure categories, Lili (1996), Karl (2004), and Ophelia (2005) remained the top 3 storms by number of counties exposed based on the various wind exposure metrics, each with 2396 counties exposed [JMF-- I think something weird is going on here].
+
 #### Is our 3-day window for rian exposure reasonable?

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Draft code for precipitation validation

@@
 ```{r}
-
+my_hurricanes <- c("Floyd-1999", 
+                   #"Irene-2011", 
+                   #"Lee-2011", 
+                   #"Ike-2008", 
+                   #"Faye-2014",
+                   #"Allison-2001")
+my_day <- c(-3, -2, -1, 0, 1, 2, 3)
+for(day in my_day){
+        exposed <- map_counties(storm = "Floyd-1999", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Floyd-1999", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Irene-2011", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Irene-2011", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Lee-2011", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Lee-2011", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Ike-2008", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Ike-2008", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Allison-2001", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Allison-2001", 
+                        plot_object = exposed)
+        print(tracks)
+}
 ```

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
draft precip validation plots

--- 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r}
-
+my_hurricanes <- c("Floyd-1999", 
+                   "Irene-2011", 
+                   "Lee-2011", 
+                   "Ike-2008",
+                   "Allison-2001")
+pdf("validate_precip.pdf")
+for(hurricane in my_hurricanes){
+        a <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -3)
+        a <- map_tracks(storm = hurricane, 
+                        plot_object = a)
+        b <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -2)
+        b <- map_tracks(storm = hurricane, 
+                        plot_object = b)
+        c <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -1)
+        c <- map_tracks(storm = hurricane, 
+                        plot_object = c)
+        d <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 0)
+        d <- map_tracks(storm = hurricane, 
+                        plot_object = d)
+        e <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 1)
+        e <- map_tracks(storm = hurricane, 
+                        plot_object = e)
+        f <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 2)
+        f <- map_tracks(storm = hurricane, 
+                        plot_object = f)
+        g <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 3)
+        g <- map_tracks(storm = hurricane, 
+                        plot_object = g)
+        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2)
+        print(precip_plot)
+}
+dev.off()
+```
+
+```{r}
+for(day in my_day){
+        exposed <- map_counties(storm = "Floyd-1999", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Floyd-1999", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Irene-2011", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Irene-2011", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Lee-2011", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Lee-2011", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Ike-2008", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Ike-2008", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Allison-2001", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Allison-2001", 
+                        plot_object = exposed)
+        print(tracks)
+}
 ```

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Comment out loop for precip validation plots

--- 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r}
-
+my_hurricanes <- c("Floyd-1999", 
+                   "Irene-2011", 
+                   "Lee-2011", 
+                   "Ike-2008",
+                   "Allison-2001")
+#pdf("validate_precip.pdf")
+#for(hurricane in my_hurricanes){
+        a <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -3)
+        a <- map_tracks(storm = hurricane, 
+                        plot_object = a)
+        b <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -2)
+        b <- map_tracks(storm = hurricane, 
+                        plot_object = b)
+        c <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -1)
+        c <- map_tracks(storm = hurricane, 
+                        plot_object = c)
+        d <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 0)
+        d <- map_tracks(storm = hurricane, 
+                        plot_object = d)
+        e <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 1)
+        e <- map_tracks(storm = hurricane, 
+                        plot_object = e)
+        f <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 2)
+        f <- map_tracks(storm = hurricane, 
+                        plot_object = f)
+        g <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 3)
+        g <- map_tracks(storm = hurricane, 
+                        plot_object = g)
+        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2)
+        print(precip_plot)
+}
+#dev.off()
+```
+
+```{r}
+for(day in my_day){
+        exposed <- map_counties(storm = "Floyd-1999", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Floyd-1999", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Irene-2011", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Irene-2011", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Lee-2011", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Lee-2011", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Ike-2008", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Ike-2008", 
+                        plot_object = exposed)
+        print(tracks)
+}
+for(day in my_day){
+        exposed <- map_counties(storm = "Allison-2001", metric = "rainfall", 
+                        days_included = day)
+        tracks <- map_tracks(storm = "Allison-2001", 
+                        plot_object = exposed)
+        print(tracks)
+}
 ```

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
move precip validation plot to top of document

@@
 Storms to check-- Floyd (ET), Irene (ET), Lee (slow moving), Ike (for timing), Faye, Allison]
+
+```{r}
+my_hurricanes <- c("Floyd-1999", 
+                   "Irene-2011", 
+                   "Lee-2011", 
+                   "Ike-2008",
+                   "Allison-2001")
+#pdf("validate_precip.pdf")
+#for(hurricane in my_hurricanes){
+        a <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -3)
+        a <- map_tracks(storm = hurricane, 
+                        plot_object = a)
+        b <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -2)
+        b <- map_tracks(storm = hurricane, 
+                        plot_object = b)
+        c <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -1)
+        c <- map_tracks(storm = hurricane, 
+                        plot_object = c)
+        d <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 0)
+        d <- map_tracks(storm = hurricane, 
+                        plot_object = d)
+        e <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 1)
+        e <- map_tracks(storm = hurricane, 
+                        plot_object = e)
+        f <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 2)
+        f <- map_tracks(storm = hurricane, 
+                        plot_object = f)
+        g <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 3)
+        g <- map_tracks(storm = hurricane, 
+                        plot_object = g)
+        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2)
+        print(precip_plot)
+}
+#dev.off()
+```
 
@@
 
-#### Is our 3-day window for rian exposure reasonable?
-
-```{r}
-my_hurricanes <- c("Floyd-1999", 
-                   "Irene-2011", 
-                   "Lee-2011", 
-                   "Ike-2008",
-                   "Allison-2001")
-#pdf("validate_precip.pdf")
-#for(hurricane in my_hurricanes){
-        a <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -3)
-        a <- map_tracks(storm = hurricane, 
-                        plot_object = a)
-        b <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -2)
-        b <- map_tracks(storm = hurricane, 
-                        plot_object = b)
-        c <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -1)
-        c <- map_tracks(storm = hurricane, 
-                        plot_object = c)
-        d <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 0)
-        d <- map_tracks(storm = hurricane, 
-                        plot_object = d)
-        e <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 1)
-        e <- map_tracks(storm = hurricane, 
-                        plot_object = e)
-        f <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 2)
-        f <- map_tracks(storm = hurricane, 
-                        plot_object = f)
-        g <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 3)
-        g <- map_tracks(storm = hurricane, 
-                        plot_object = g)
-        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2)
-        print(precip_plot)
-}
-#dev.off()
-```
-
-```{r}
-for(day in my_day){
-        exposed <- map_counties(storm = "Floyd-1999", metric = "rainfall", 
-                        days_included = day)
-        tracks <- map_tracks(storm = "Floyd-1999", 
-                        plot_object = exposed)
-        print(tracks)
-}
-for(day in my_day){
-        exposed <- map_counties(storm = "Irene-2011", metric = "rainfall", 
-                        days_included = day)
-        tracks <- map_tracks(storm = "Irene-2011", 
-                        plot_object = exposed)
-        print(tracks)
-}
-for(day in my_day){
-        exposed <- map_counties(storm = "Lee-2011", metric = "rainfall", 
-                        days_included = day)
-        tracks <- map_tracks(storm = "Lee-2011", 
-                        plot_object = exposed)
-        print(tracks)
-}
-for(day in my_day){
-        exposed <- map_counties(storm = "Ike-2008", metric = "rainfall", 
-                        days_included = day)
-        tracks <- map_tracks(storm = "Ike-2008", 
-                        plot_object = exposed)
-        print(tracks)
-}
-for(day in my_day){
-        exposed <- map_counties(storm = "Allison-2001", metric = "rainfall", 
-                        days_included = day)
-        tracks <- map_tracks(storm = "Allison-2001", 
-                        plot_object = exposed)
-        print(tracks)
-}
-```
-
 #### Are storms with heavy rain exposure likely to occur early in the hurricane season?

### 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add text on precip validation

@@
 ```
+
+For the most part, using a three day window of the day before, day of, and day after a tropical cyclone in assessing precipitation exposure seemed to capture the effects of the storm while excluding precipitation from other weather systems.
+However, for storms that have tracks that make a loop, such as Allison (2001), this three day window may not be adequate.
 

### 2016-07-19 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Changes to draft

@@
                    "Allison-2001")
-#pdf("validate_precip.pdf")
-#for(hurricane in my_hurricanes){
+pdf("validate_precip.pdf", height = 8, width = 12)
+for(hurricane in my_hurricanes){
         a <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -3)
+                          days_included = -3) + theme(legend.position="none")
         a <- map_tracks(storm = hurricane, 
-                        plot_object = a)
+                        plot_object = a, plot_points = FALSE)
         b <- map_counties(storm = hurricane, metric = "rainfall",
@@
                         plot_object = g)
-        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2)
+        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2,
+                                    top = textGrob("Title"))
         print(precip_plot)
 }
-#dev.off()
+dev.off()
 ```

### 2016-07-19 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add code to load grid library

@@
 library(ggplot2)
+library(grid)
 library(gridExtra)

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add text
fix precip validation plots

--- 2016-07-18 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 library(RColorBrewer)
+library(grid)
 ```
@@
                    "Allison-2001")
-#pdf("validate_precip.pdf")
-#for(hurricane in my_hurricanes){
+pdf("validate_precip.pdf", height = 8, width = 12)
+for(hurricane in my_hurricanes){
         a <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -3)
+                          days_included = -3) + theme(legend.position="none")
         a <- map_tracks(storm = hurricane, 
-                        plot_object = a)
+                        plot_object = a, plot_points = FALSE)
         b <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -2)
+                          days_included = -2) + theme(legend.position = "none")
         b <- map_tracks(storm = hurricane, 
-                        plot_object = b)
+                        plot_object = b, plot_points = FALSE)
         c <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -1)
+                          days_included = -1) + theme(legend.position = "none")
         c <- map_tracks(storm = hurricane, 
-                        plot_object = c)
+                        plot_object = c, plot_points - FALSE)
         d <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 0)
+                          days_included = 0) + theme(legend.position = FALSE)
         d <- map_tracks(storm = hurricane, 
-                        plot_object = d)
+                        plot_object = d, plot_points = FALSE)
         e <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 1)
+                          days_included = 1) + theme(legend.position = "none")
         e <- map_tracks(storm = hurricane, 
-                        plot_object = e)
+                        plot_object = e, plot_points = FALSE)
         f <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 2)
+                          days_included = 2) + theme(legend.position = "none")
         f <- map_tracks(storm = hurricane, 
-                        plot_object = f)
+                        plot_object = f, plot_points = FALSE)
         g <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 3)
+                          days_included = 3) + theme(legend.position = "none")
         g <- map_tracks(storm = hurricane, 
-                        plot_object = g)
-        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2)
+                        plot_object = g, plot_points = FALSE)
+        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2,
+                                    top = textGrob(paste0("7-Day Precipitation For\n Storm ", hurricane)))
         print(precip_plot)
 }
-#dev.off()
+dev.off()
 ```
@@
 #### Are our methods of assessing wind speed reasonable
+
+[Validate windspeedmodel functions using Extended Best Tracks Data]
 

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Updated precip validation. No legend and bigger figures

@@
         c <- map_tracks(storm = hurricane, 
-                        plot_object = c, plot_points - FALSE)
+                        plot_object = c, plot_points = FALSE)
         d <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 0) + theme(legend.position = FALSE)
+                          days_included = 0) + theme(legend.position = "none")
         d <- map_tracks(storm = hurricane, 

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on wind plot

@@
         rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) # Max is 1 for hits for top Storm-County
-                                                                       # pairs
-test <- TS_wind_exposure %>%         
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-        arrange(desc(count)) %>%
-        slice (1:2000) #This gives a max of 17 for number of counts for the top 2000 counties
+        mutate(region = as.numeric(region), count = as.numeric(count)) # Max count of 178 is a NA value
+TS_wind_2000 <- subset(TS_wind_2000, !is.na(TS_wind_2000$region)) # Remove NA observation
 HU_wind_2000 <- HU_wind_exposure %>% 
@@
                                                                        # County pairs
-TS_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
-                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+TS_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                    include.lowest = TRUE)
-exposure_palette <- brewer.pal(7, name = "YlGnBu")
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
 map_TS_2000 <- CountyChoropleth$new(TS_wind_2000)

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
precipitation lag plot

@@
 
+```{r}
+data(closest_dist)
+
+# Generate a list of unique study fips (all in the eastern US)
+my_fips <- closest_dist$fips
+my_fips <- unique(my_fips)
+storm_lag <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 150, dist_limit = 500,
+                             days_include = -3:3)
+storm_lag <- storm_lag %>%
+        mutate(storm_county = paste0(storm_id, "_", fips))
+data(rain)
+rain_lag <- rain %>%
+        mutate(storm_county = paste0(storm_id, "_", fips)) %>%
+        filter(storm_county %in% storm_lag$storm_county)
+rain_lag_plot <- 
+```
+
 #### Are our methods of assessing wind speed reasonable
@@
 
-distance_2000$value <- cut(distance_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
-                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+distance_2000$value <- cut(distance_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                    include.lowest = TRUE)
-exposure_palette <- brewer.pal(7, name = "YlGnBu")
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
 
@@
 
-rain_2000$value <- cut(rain_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
-                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+rain_2000$value <- cut(rain_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                    include.lowest = TRUE)
-exposure_palette <- brewer.pal(7, name = "YlGnBu")
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
 
@@
 a <- map_TS_2000$render()
-HU_wind_2000$value <- cut(HU_wind_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
-                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+HU_wind_2000$value <- cut(HU_wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                    include.lowest = TRUE)
@@
         mutate(region = as.numeric(region), count = as.numeric(count))
-cat3_2000$value <- cut(cat3_2000$count, c(0, 1, 2, 3, 4, 5, 6, 7), 
-                   labels = c("1", "2", "3", "4", "5", "6", "7"), 
+cat3_2000$value <- cut(cat3_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
                    include.lowest = TRUE)
@@
 ```
-
-JMF-- I do not think that using the top 2000 Storm-County pairs will be too descriptive (All 3 maps have the identical 2000 pairs), any ideas?
 
@@
 
+```{r}
+wind_count <- TS_wind_exposure %>%
+              group_by(fips) %>%
+              summarise(count = n()) %>%
+              rename(region = fips) %>%
+              mutate(region = as.numeric(region), count = as.numeric(count)) #Same size df as before (only 474 obs)
+wind_count$value <- cut(wind_count$count, c(0, 5, 8, 12, 15,
+                                                    18, 21, 24),
+                           labels = c("0-5", "6-8", "9-12", "13-15",
+                                      "16-18", "19-21", "22-24"),
+                           include.lowest = TRUE)
+exposure_palette <- brewer.pal(7, name = "YlGnBu")
+
+map_wind_count <- CountyChoropleth$new(wind_count)
+map_wind_count$set_zoom(eastern_states)
+map_wind_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_wind_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Wind > 18 m/s"
+c <- map_wind_count$render()
+```
+
 ```{r fig.width = 6, fig.height = 8, echo = FALSE}
-grid.arrange(a, b, ncol = 1)
+grid.arrange(a, b, c, ncol = 1)
 ```
@@
 ```{r echo = FALSE}
-exposure_name <- c("Rain_Exposure", "Distance_Exposure")
-count <- c(sum(rain_count$count), sum(distance_count$count))
+exposure_name <- c("Distance_Exposure", "Rain_Exposure", "Wind_Exposure")
+count <- c(sum(distance_count$count), sum(rain_count$count), sum(wind_count$count))
 exposure_table <- cbind(exposure_name, count)
@@
 ```
+
 ```{r}

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
rain lag plot code

@@
         filter(storm_county %in% storm_lag$storm_county)
-rain_lag_plot <- 
+rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = precip, group = storm_county)) +
+        geom_line(alpha = 0.2, size = 0.5)
+rain_lag_plot
 ```

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
change alpha and size for rain lag plot

--- 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
         filter(storm_county %in% storm_lag$storm_county)
-rain_lag_plot <- 
+rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = precip, group = storm_county)) +
+        geom_line(alpha = 0.1, size = 0.3) + theme_minimal()
+rain_lag_plot
 ```

### 2016-07-20 R/event_exposure.R
Add code for events

@@ -1,0 +1,78 @@
+#' Find events for storms by county
+#'
+#' @note Of the event types that this function can pull, only "tornado" and
+#'    "wind" are available for years before 1996. Therefore, this function will
+#'    pull listed tornado events or wind events for all years, but for any other
+#'    event type, you should not use a start year prior to 1996, as events of
+#'    other types were not recorded in the database before 1996.
+#'
+#' @examples
+#' county_events(counties = c("37031", "37053", "37055"),
+#'               start_year = 1988, end_year = 1999,
+#'               event_type = "flood")
+county_events <- function(counties, start_year, end_year, event_type){
+        events <- hurricaneexposuredata::storm_events
+
+        event_years <- as.numeric(gsub(".+-", "", names(events)))
+        events <- events[event_years >= start_year & event_years <= end_year]
+        events <- events[lapply(events, nrow) > 0]
+        events <- purrr::map2(events, names(events), ~ cbind(.x, storm_id = .y))
+
+        events <- lapply(events, function(x) subset(x, x$fips %in% counties))
+
+        events <- do.call("rbind", events) %>%
+                dplyr::mutate(flood = grepl("Flood", events),
+                              tornado = grepl("Tornado", events) ,
+                              tropical_storm = grepl("Hurricane", events) |
+                                      grepl("Tropical Storm", events) |
+                                      grepl("Tropical Depression", events),
+                              wind = tropical_storm | grepl("Wind", events))
+        events <- events[events[ , event_type], c("fips", "storm_id")]
+        return(events)
+}
+
+#' Map county-level exposure based on reported events
+#'
+#' @examples
+#' map_event_exposure(storm_id = "Floyd-1999", event_type = "flood")
+#' map_event_exposure(storm_id = "Floyd-1999", event_type = "tornado")
+#' map_event_exposure(storm_id = "Floyd-1999", event_type = "wind")
+#' map_event_exposure(storm_id = "Floyd-1999", event_type = "tropical_storm")
+map_event_exposure <- function(storm_id, event_type){
+        storm <- storm_id
+        storm_year <- gsub("*.+-", "", storm_id)
+        counties <- hurricaneexposuredata::closest_dist %>%
+                dplyr::filter(storm_id == storm) %>%
+                dplyr::select(fips, storm_dist)
+        map_data <- county_events(counties = counties$fips,
+                                start_year = storm_year,
+                                end_year = storm_year,
+                                event_type = event_type) %>%
[…]

### 2016-07-20 writing/rain_plots.R
Update code for mapping all storms for EDA

@@ -10,6 +10,16 @@ plot_both <- function(storm, metric = "distance"){
         return(map_2)
 }
 
+plot_events <- function(storm, event_type){
+        print(storm)
+        map_1 <- map_event_exposure(storm_id = storm,
+                                    event_type = event_type)
+        map_2 <- map_tracks(storms = storm, plot_object = map_1,
+                            plot_points = FALSE)
+        map_2 <- map_2 + ggplot2::ggtitle(storm)
+        return(map_2)
+}
+
 pdf(file = "writing/rain_plots.pdf", width = 7, height = 4)
 # Only can run these through 2011
 lapply(all_storms[1:125], plot_both, metric = "rainfall")
@@ -21,4 +31,17 @@ dev.off()
 
 pdf(file = "writing/wind_plots.pdf", width = 7, height = 4)
 lapply(all_storms, plot_both, metric = "wind")
+dev.off()
+
+pdf(file = "writing/wind_events_plots.pdf", width = 7, height = 4)
+lapply(all_storms, plot_events, event_type = "wind")
+dev.off()
+
+pdf(file = "writing/tornado_plots.pdf", width = 7, height = 4)
+lapply(all_storms, plot_events, event_type = "tornado")
+dev.off()
+
+pdf(file = "writing/flood_plots.pdf", width = 7, height = 4)
+# only since 1996
+lapply(all_storms[31:length(all_storms)], plot_events, event_type = "flood")
 dev.off()

### 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Change precip plot to cumulative

--- 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 library(ggplot2)
+library(ggthemes)
 library(grid)
@@
 
-```{r}
+```{r eval = FALSE}
 my_hurricanes <- c("Floyd-1999", 
@@
         a <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -3) + theme(legend.position="none")
+                          days_included = -3) + 
+                theme(legend.position="none") +
+                ggtitle("Three days before")
         a <- map_tracks(storm = hurricane, 
@@
         b <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -2) + theme(legend.position = "none")
+                          days_included = -2) + 
+                theme(legend.position = "none") +
+                ggtitle("Two days before")
         b <- map_tracks(storm = hurricane, 
@@
         c <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -1) + theme(legend.position = "none")
+                          days_included = -1) + 
+                theme(legend.position = "none")  +
+                ggtitle("One day before")
         c <- map_tracks(storm = hurricane, 
@@
         d <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 0) + theme(legend.position = "none")
+                          days_included = 0) + 
+                theme(legend.position = "none")  +
+                ggtitle("Day of")
         d <- map_tracks(storm = hurricane, 
@@
         e <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 1) + theme(legend.position = "none")
+                          days_included = 1) + 
+                theme(legend.position = "none") +
+                ggtitle("One day after")
         e <- map_tracks(storm = hurricane, 
@@
         f <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 2) + theme(legend.position = "none")
+                          days_included = 2) + 
+                theme(legend.position = "none") +
+                ggtitle("Two days after")
         f <- map_tracks(storm = hurricane, 
@@
         g <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 3) + theme(legend.position = "none")
+                          days_included = 3) + 
+                theme(legend.position = "none") +
+                ggtitle("Three days after")
         g <- map_tracks(storm = hurricane, 
@@
 
+```{r fig.height=12, fig.width=10}
+hurricane <- "Lee-2011"
+        a <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -3) + 
+                theme(legend.position="none") +
+                ggtitle("Three days before")
+        a <- map_tracks(storm = hurricane, 
+                        plot_object = a, plot_points = FALSE)
+        b <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -2) + 
+                theme(legend.position = "none") +
+                ggtitle("Two days before")
+        b <- map_tracks(storm = hurricane, 
+                        plot_object = b, plot_points = FALSE)
+        c <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -1) + 
+                theme(legend.position = "none")  +
+                ggtitle("One day before")
+        c <- map_tracks(storm = hurricane, 
+                        plot_object = c, plot_points = FALSE)
+        d <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 0) + 
+                theme(legend.position = "none")  +
+                ggtitle("Day of")
+        d <- map_tracks(storm = hurricane, 
+                        plot_object = d, plot_points = FALSE)
+        e <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 1) + 
+                theme(legend.position = "none") +
+                ggtitle("One day after")
+        e <- map_tracks(storm = hurricane, 
+                        plot_object = e, plot_points = FALSE)
+        f <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 2) + 
+                theme(legend.position = "none") +
+                ggtitle("Two days after")
+        f <- map_tracks(storm = hurricane, 
+                        plot_object = f, plot_points = FALSE)
+        g <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 3) + 
+                theme(legend.position = "none") +
+                ggtitle("Three days after")
+        g <- map_tracks(storm = hurricane, 
+                        plot_object = g, plot_points = FALSE)
+
+precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 4,
+                                    top = textGrob(paste0("7-Day Precipitation For\n Storm ", hurricane)))
+print(precip_plot)
+```
+
 For the most part, using a three day window of the day before, day of, and day after a tropical cyclone in assessing precipitation exposure seemed to capture the effects of the storm while excluding precipitation from other weather systems.
@@
 
-```{r}
+```{r warning = FALSE}
 data(closest_dist)
@@
         mutate(storm_county = paste0(storm_id, "_", fips)) %>%
-        filter(storm_county %in% storm_lag$storm_county)
-rain_lag_plot <- 
+        filter(storm_county %in% storm_lag$storm_county) %>%
+        group_by(storm_county) %>%
+        mutate(cumulative_rain = cumsum(precip),
+               difference_rain = lead(cumulative_rain) - cumulative_rain,
+               next_lag = lead(lag),
+               next_cumulative_rain = lead(cumulative_rain),
+               most_rain = difference_rain == max(difference_rain,
+                                                  na.rm = TRUE))
+to_highlight <- data.frame(x = c(-1.5, -1.5, 1.5, 1.5), y = c(0, 400, 400, 0))
+
+rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = cumulative_rain)) +
+        theme_tufte() + 
+        geom_segment(aes(xend = next_lag, yend = next_cumulative_rain,
+                         color = most_rain),
+                     alpha = 0.1, size = 0.3) +
+        geom_polygon(data = to_highlight, aes(x = x, y = y),
+                     alpha = 0.2, color = "red", fill = NA) +
+        xlab("Days from storm's closest approach to county") + 
+        ylab("Cumulative rainfall in \ncounty from storm (mm)") + 
+        scale_color_manual(values = c("darkgray", "blue"),
+                           guide = FALSE) 
+rain_lag_plot
 ```

### 2016-07-20 R/map_exposure.R
Add lag -2 into default for rain calculation

@@ -127,7 +127,7 @@ map_tracks <- function(storms, plot_object = NULL,
 #'
 #' @examples
 #' floyd_map <- map_counties("Floyd-1999", metric = "rainfall",
-#'                            days_included = c(-1, 0, 1))
+#'                            days_included = c(-2, -1, 0, 1))
 #' floyd_map
 #'
 #' beryl_map <- map_counties("Beryl-1988", metric = "wind")
@@ -136,7 +136,7 @@ map_tracks <- function(storms, plot_object = NULL,
 #'
 #' @importFrom dplyr %>%
 map_counties <-function(storm, metric = "distance",
-                        days_included = c(-1, 0, 1)){
+                        days_included = c(-2, -1, 0, 1)){
         if(metric == "distance"){
                 map_data <- filter_storm_data(storm = storm,
                                               output_vars = c("fips",
@@ -188,7 +188,7 @@ map_counties <-function(storm, metric = "distance",
 #'
 #' @export
 map_rain_exposure <- function(storm, rain_limit, dist_limit,
-                              days_included = c(-1, 0, 1)){
+                              days_included = c(-2, -1, 0, 1)){
 
         map_data <- filter_storm_data(storm = storm,
                                            days_included = days_included,

### 2016-07-20 R/rain_exposure.R

@@ -73,7 +73,7 @@
 #' @export
 county_rain <- function(counties, start_year, end_year,
                            rain_limit, dist_limit,
-                        days_included = c(-1, 0, 1)){
+                        days_included = c(-2, -1, 0, 1)){
 
         rain_storm_df <- filter_storm_data(counties = counties,
                                            year_range = c(start_year, end_year),
@@ -168,7 +168,7 @@ county_rain <- function(counties, start_year, end_year,
 #' @export
 multi_county_rain <- function(communities, start_year, end_year,
                          rain_limit, dist_limit,
-                         days_included = c(-1, 0, 1)){
+                         days_included = c(-2, -1, 0, 1)){
 
         communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
 
@@ -257,7 +257,7 @@ multi_county_rain <- function(communities, start_year, end_year,
 #' @export
 rain_exposure <- function(locations, start_year, end_year,
                           rain_limit, dist_limit,
-                          days_included = c(-1, 0, 1),
+                          days_included = c(-2, -1, 0, 1),
                           out_dir, out_type = "csv"){
 
         if(!dir.exists(out_dir)){

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Some work on paper draft

--- 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r}
+```{r eval = FALSE, echo = FALSE}
 my_hurricanes <- c("Floyd-1999", 
@@
 
+```{r fig.height=12, fig.width=10, echo = FALSE}
+hurricane <- "Lee-2011"
+        a <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -3) + 
+                theme(legend.position="none") +
+                ggtitle("Three days before")
+        a <- map_tracks(storm = hurricane, 
+                        plot_object = a, plot_points = FALSE)
+        b <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -2) + 
+                theme(legend.position = "none") +
+                ggtitle("Two days before")
+        b <- map_tracks(storm = hurricane, 
+                        plot_object = b, plot_points = FALSE)
+        c <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -1) + 
+                theme(legend.position = "none")  +
+                ggtitle("One day before")
+        c <- map_tracks(storm = hurricane, 
+                        plot_object = c, plot_points = FALSE)
+        d <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 0) + 
+                theme(legend.position = "none")  +
+                ggtitle("Day of")
+        d <- map_tracks(storm = hurricane, 
+                        plot_object = d, plot_points = FALSE)
+        e <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 1) + 
+                theme(legend.position = "none") +
+                ggtitle("One day after")
+        e <- map_tracks(storm = hurricane, 
+                        plot_object = e, plot_points = FALSE)
+        f <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 2) + 
+                theme(legend.position = "none") +
+                ggtitle("Two days after")
+        f <- map_tracks(storm = hurricane, 
+                        plot_object = f, plot_points = FALSE)
+        g <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = 3) + 
+                theme(legend.position = "none") +
+                ggtitle("Three days after")
+        g <- map_tracks(storm = hurricane, 
+                        plot_object = g, plot_points = FALSE)
+
+precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 4,
+                                    top = textGrob(paste0("7-Day Precipitation For\n Storm ", hurricane)))
+print(precip_plot)
+```
+
 For the most part, using a three day window of the day before, day of, and day after a tropical cyclone in assessing precipitation exposure seemed to capture the effects of the storm while excluding precipitation from other weather systems.
@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 data(closest_dist)
@@
         mutate(storm_county = paste0(storm_id, "_", fips)) %>%
-        filter(storm_county %in% storm_lag$storm_county)
-rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = precip, group = storm_county)) +
-        geom_line(alpha = 0.2, size = 0.5)
+        filter(storm_county %in% storm_lag$storm_county) %>%
+        group_by(storm_county) %>%
+        mutate(cumulative_rain = cumsum(precip),
+               difference_rain = lead(cumulative_rain) - cumulative_rain,
+               next_lag = lead(lag),
+               next_cumulative_rain = lead(cumulative_rain),
+               most_rain = difference_rain == max(difference_rain,
+                                                  na.rm = TRUE))
+to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1), y = c(0, 400, 400, 0))
+
+rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = cumulative_rain)) +
+        theme_tufte() + 
+        geom_segment(aes(xend = next_lag, yend = next_cumulative_rain,
+                         color = most_rain),
+                     alpha = 0.1, size = 0.3) +
+        geom_polygon(data = to_highlight, aes(x = x, y = y),
+                     alpha = 0.2, color = "red", fill = NA) +
+        xlab("Days from storm's closest approach to county") + 
+        ylab("Cumulative rainfall in \ncounty from storm (mm)") + 
+        scale_color_manual(values = c("darkgray", "blue"),
+                           guide = FALSE) 
 rain_lag_plot
@@
 
-```{r fig.width = 6, fig.height = 8, echo = FALSE}
+```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
 grid.arrange(a, b, ncol = 1)
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, message = FALSE, warning = FALSE}
 #TS winds and above
@@
 
-```{r}
+```{r echo = FALSE, message = FALSE, warning = FALSE}
 #Category 3 and above 
@@
 
-```{r}
+```{r echo = FALSE, warning = FALSE, message = FALSE}
 wind_count <- TS_wind_exposure %>%
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, warning = FALSE, message = FALSE}
 exposure_name <- c("Distance_Exposure", "Rain_Exposure", "Wind_Exposure")
@@
 
-```{r}
+```{r echo = FALSE, message = FALSE, warning = FALSE}
 #Test
@@
 
+Our exposure metrics restricts "exposed" counties to be within 500 km of the storm track.
+In a few cases, hurricane best tracks stop when the storm transitions from a tropical to extratropical system [let's check that I got the wording right on this].
+In these cases, extreme rain from the storm system might have affected counties further along the storm's path, but that are not counted as being near the storm's track.
+For example, Lee in 2011 was only tracked through the best tracks until it was near the Georgia-Tennessee border.
+However, remnants of the system brought extreme rain further north, into the Mid-Atlantic and New England [I believe that the Monthly Weather Review for the 2011 season includes a description of these rain impacts.
+We likely want to look over that and incorporate it here]:
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Lee-2011"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle("Rainfall during Lee, 2011")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Lee, 2011")
+grid.arrange(a, b, ncol = 1)
+```
+
+Other possible examples of this during our study period include Nicole (2010) [check storm history in MWR to confirm that Mid-Atlantic and New England rain during this time was related to storm], Ida (2009), Humberto (2007), Katrina (2005), Grace (2003) [confirm rain was from storm remnants], Frances (1998), Marco (1990)
+
+For rain-based exposure, we use a distance constraint of 500 km.
+With storms where the best track extends throughout the storm, this should typically be a large enough swath to include all extreme rain exposures from the storm.
+There are only a few very large storms where this might not be the case.
+One example is Ike (2008), in which case the 500 km constraint excludes some of the counties in northern Texas that got heavy rain from the system:
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Ike-2008"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle("Rainfall during Ike, 2008")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Ike, 2008")
+grid.arrange(a, b, ncol = 1)
+```
+
+Other large systems where this might be a concern include Wilma (2005), Hermine (2004) [confirm rain in SC, NC, and VA during this time was from this storm], Arlene (1993) [confirm rain through to LA is from this storm]
+
 Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 18 m/s or greater (e.g., tropical storm strength or greater):

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add code to compare gauge to nldas

--- 2016-07-20 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+Justification of NLDAS data: comparison of NLDAS and gauge data for hurricane Alex (2010).
+
+```{r}
+gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+nldas <- county_rain(counties = unique(gauge$fips),
+                             start_year = 2010, end_year = 2010,
+                             rain_limit = 0, dist_limit = 500,
+                             days_include = c(-2, 0, 1))
+nldas <- nldas %>%
+        filter(storm_id == "Alex-2010")
+gauge$fips <- as.character(gauge$fips)
+compare_rain <- full_join(nldas, gauge, by = "fips")
+#compare_rain <- full_join(nldas, gauge, by = "fips")
+a <- ggplot(compare_rain, aes(rain_mm, tot_precip, color = fips)) + geom_point()
+a
+```
+
 #### Are our methods of assessing wind speed reasonable

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on draft

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-Our analysis included [x] Atlantic basin tropical storms over the years 1988 to 2011. This dataset includes several noteworthy storms, including ... .
-
-[As a summary of our available dataset, do we want to include a US map showing the storm tracks for all of the storms included in our analysis?]
+Our analysis included [x] Atlantic basin tropical storms over the years 1988 to 2011.
+This dataset includes several noteworthy storms, including Hugo (1989), Andrew (1992), Floyd (1999), Katrina (2005), ...
+.
 
@@
 
-[Is three-day window reasonable?
-Probably should be.
-If we go larger, there's the problem of picking up things that aren't related to the storm.
-Storms to check-- Floyd (ET), Irene (ET), Lee (slow moving), Ike (for timing), Faye, Allison]
+For rainfall, we are using the cumulative rainfall from two days before to one day after the storm is closest to the county.
+This is based on analysis of the precipitation data for storms in our study period.
+If we go larger, there is a danger of picking up rainfall that isn't related to the storm.
+However, some storms move slowly, so if we limit to just the day the storm is closest to the county, we would miss some of the associated rainfall.
+
+To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").
 
@@
 for(hurricane in my_hurricanes){
+        storm_precip_totals <- hurricaneexposuredata::rain %>%
+                dplyr::filter(storm_id == hurricane & 
+                                      lag %in% -4:3) %>%
+                dplyr::group_by(lag) %>%
+                dplyr::summarize(sum_rain = prettyNum(round(sum(precip)),
+                                                      big.mark = ",")) %>%
+                dplyr::mutate(sum_rain = paste(sum_rain, "mm"))
+        h <- map_counties(storm = hurricane, metric = "rainfall",
+                          days_included = -4) + 
+                theme(legend.position="none") +
+                ggtitle(paste0("Four days before \n(", 
+                               storm_precip_totals[1, 2], ")"))
+        h <- map_tracks(storm = hurricane, 
+                        plot_object = h, plot_points = FALSE)
         a <- map_counties(storm = hurricane, metric = "rainfall",
@@
                 theme(legend.position="none") +
-                ggtitle("Three days before")
+                ggtitle(paste0("Three days before \n(", 
+                               storm_precip_totals[2, 2], ")"))
         a <- map_tracks(storm = hurricane, 
@@
                 theme(legend.position = "none") +
-                ggtitle("Two days before")
+                ggtitle(paste0("Two days before \n(", 
+                               storm_precip_totals[3, 2], ")"))
         b <- map_tracks(storm = hurricane, 
@@
                 theme(legend.position = "none")  +
-                ggtitle("One day before")
+                ggtitle(paste0("One day before \n(", 
+                               storm_precip_totals[4, 2], ")"))
         c <- map_tracks(storm = hurricane, 
@@
                 theme(legend.position = "none")  +
-                ggtitle("Day of")
+                ggtitle(paste0("Day of \n(", 
+                               storm_precip_totals[5, 2], ")"))
         d <- map_tracks(storm = hurricane, 
@@
                 theme(legend.position = "none") +
-                ggtitle("One day after")
+                ggtitle(paste0("One day after \n(", 
+                               storm_precip_totals[6, 2], ")"))
         e <- map_tracks(storm = hurricane, 
@@
                 theme(legend.position = "none") +
-                ggtitle("Two days after")
+                ggtitle(paste0("Two days after \n(", 
+                               storm_precip_totals[7, 2], ")"))
         f <- map_tracks(storm = hurricane, 
@@
                 theme(legend.position = "none") +
-                ggtitle("Three days after")
+                ggtitle(paste0("Three days after \n(", 
+                               storm_precip_totals[8, 2], ")"))
         g <- map_tracks(storm = hurricane, 
                         plot_object = g, plot_points = FALSE)
-        precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 2,
-                                    top = textGrob(paste0("7-Day Precipitation For\n Storm ", hurricane)))
+        precip_plot <- grid.arrange(h, a, b, c, d, e, f, g, nrow = 2,
+                top = textGrob(paste("7-Day Precipitation For", hurricane)))
         print(precip_plot)
@@
 
-```{r fig.height=12, fig.width=10, echo = FALSE}
-hurricane <- "Lee-2011"
-        a <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -3) + 
-                theme(legend.position="none") +
-                ggtitle("Three days before")
-        a <- map_tracks(storm = hurricane, 
-                        plot_object = a, plot_points = FALSE)
-        b <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -2) + 
-                theme(legend.position = "none") +
-                ggtitle("Two days before")
-        b <- map_tracks(storm = hurricane, 
-                        plot_object = b, plot_points = FALSE)
-        c <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = -1) + 
-                theme(legend.position = "none")  +
-                ggtitle("One day before")
-        c <- map_tracks(storm = hurricane, 
-                        plot_object = c, plot_points = FALSE)
-        d <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 0) + 
-                theme(legend.position = "none")  +
-                ggtitle("Day of")
-        d <- map_tracks(storm = hurricane, 
-                        plot_object = d, plot_points = FALSE)
-        e <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 1) + 
-                theme(legend.position = "none") +
-                ggtitle("One day after")
-        e <- map_tracks(storm = hurricane, 
-                        plot_object = e, plot_points = FALSE)
-        f <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 2) + 
-                theme(legend.position = "none") +
-                ggtitle("Two days after")
-        f <- map_tracks(storm = hurricane, 
-                        plot_object = f, plot_points = FALSE)
-        g <- map_counties(storm = hurricane, metric = "rainfall",
-                          days_included = 3) + 
-                theme(legend.position = "none") +
-                ggtitle("Three days after")
-        g <- map_tracks(storm = hurricane, 
-                        plot_object = g, plot_points = FALSE)
-
-precip_plot <- grid.arrange(a, b, c, d, e, f, g, nrow = 4,
-                                    top = textGrob(paste0("7-Day Precipitation For\n Storm ", hurricane)))
-print(precip_plot)
-```
-
-For the most part, using a three day window of the day before, day of, and day after a tropical cyclone in assessing precipitation exposure seemed to capture the effects of the storm while excluding precipitation from other weather systems.
-However, for storms that have tracks that make a loop, such as Allison (2001), this three day window may not be adequate.
+For the most part, using a four-day window from two days before to one day after a tropical cyclone in assessing precipitation exposure seemed to capture the effects of the storm while excluding potential false positives from precipitation from other weather systems.
+However, for storms that have tracks that make a loop, such as Allison (2001), this four-day window may not be adequate, and lead to an underestimation of total rainfall associated with the storm relative to estimates for faster-moving storms.
 
@@
                              rain_limit = 150, dist_limit = 500,
-                             days_include = -3:3)
+                             days_include = -5:3)
 storm_lag <- storm_lag %>%
@@
                most_rain = difference_rain == max(difference_rain,
-                                                  na.rm = TRUE))
-to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1), y = c(0, 400, 400, 0))
+                                                  na.rm = TRUE),
+               over_50 = lead(precip) > 50)
+to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1), y = c(0, 500, 500, 0))
 
@@
         geom_segment(aes(xend = next_lag, yend = next_cumulative_rain,
-                         color = most_rain),
+                         color = over_50),
                      alpha = 0.1, size = 0.3) +
@@
 ```
+
+Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.
+Although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm.
+
+```{r echo = FALSE, fig.width = 6, fig.height = 4}
+ggplot(rain_lag, aes(x = lag, y = precip, group = lag)) + 
+        geom_boxplot() + theme_tufte() + 
+        xlab("Days from storm's closest approach to county") + 
+        ylab("Rainfall in county (mm)") 
+```
+
+Our exposure metrics restricts "exposed" counties to be within 500 km of the storm track.
+In a few cases, hurricane best tracks stop when the storm transitions from a tropical to extratropical system [let's check that I got the wording right on this].
+In these cases, extreme rain from the storm system might have affected counties further along the storm's path, but that are not counted as being near the storm's track.
+For example, Lee in 2011 was only tracked through the best tracks until it was near the Georgia-Tennessee border.
+However, remnants of the system brought extreme rain further north, into the Mid-Atlantic and New England [I believe that the Monthly Weather Review for the 2011 season includes a description of these rain impacts.
+We likely want to look over that and incorporate it here]:
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Lee-2011"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle("Rainfall during Lee, 2011")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Lee, 2011")
+grid.arrange(a, b, ncol = 1)
+```
+
+Other possible examples of this during our study period include Nicole (2010) [check storm history in MWR to confirm that Mid-Atlantic and New England rain during this time was related to storm], Ida (2009), Humberto (2007), Katrina (2005), Grace (2003) [confirm rain was from storm remnants], Frances (1998), and Marco (1990).
+
+For rain-based exposure, we use a distance constraint of 500 km.
+With storms where the best track extends throughout the storm, this should typically be a large enough swath to include all extreme rain exposures from the storm.
+There are only a few very large storms where this might not be the case.
+One example is Ike (2008), in which case the 500 km constraint excludes some of the counties in northern Texas that got heavy rain from the system:
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Ike-2008"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_counties(storm = storm, 
+                                           metric = "rainfall")) + 
+        ggtitle("Rainfall during Ike, 2008")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Ike, 2008")
+grid.arrange(a, b, ncol = 1)
+```
+
+Other large systems where this might be a concern include Wilma (2005), Hermine (2004) [confirm rain in SC, NC, and VA during this time was from this storm], Arlene (1993) [confirm rain through to LA is from this storm].
+
+For storms in 1996 or later, when flooding reports are included in NOAA's Storm Events data, exposure patterns based on floods reports generally pick up similar regions as rain-based exposure, and in some cases (e.g., Floyd, 1999), counties classified as exposed using rain-based exposure are very similar to the counties in which flood events were recorded.
+For example, here are comparisons for some of the worst storms in terms of rain-based exposures:
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Frances-2004"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Frances, 2004")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Frances, 2004")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Ivan-2004"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Ivan, 2004")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Ivan, 2004")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Isidore-2002"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Isidore, 2002")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Isidore, 2002")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Floyd-1999"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Floyd, 1999")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Floyd, 1999")
+grid.arrange(a, b, ncol = 1)
+```
+
+For all storms, we divided counties into four categories:
+
+1. "Exposed" based on both rainfall and reported flood event;
+2. "Exposed" based on rainfall, "unexposed" based on flood event; 
+3. "Unexposed" based on rainfall, "exposed" based on flood event; and
+4. "Unexposed" based on both rainfall and reported flood event.
+
+All counties in the first and fourth categories are consistently classified based on using rainfall and flood events as exposure metrics, while those in categories 2 and 3 are inconsistent, in that one metric marks the county as exposed to the storm while the other metric classifies it as unexposed.
+
+```{r echo = FALSE}
+data(closest_dist)
+
+# Generate a list of unique study fips (all in the eastern US)
+my_fips <- closest_dist$fips
+my_fips <- unique(my_fips)
+
+all_storm_counties <- closest_dist %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::filter(as.numeric(gsub(".*-", "", storm_id)) >= 1996 & 
+                              as.numeric(gsub(".*-", "", storm_id)) <= 2011)
+
+rain_exposure <- county_rain(counties = my_fips, start_year = 1996,
+                             end_year = 2011, rain_limit = 75, 
+                             dist_limit = 500) %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(rain_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, rain_exposed)
+flood_exposure <- county_events(counties = my_fips, start_year = 1996,
+                             end_year = 2011, event_type = "flood") %>%
+        dplyr::mutate(flood_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, flood_exposed)
+
+all_exposed <- rain_exposure %>%
+        dplyr::full_join(flood_exposure, by = "storm_county") %>%
+        dplyr::full_join(all_storm_counties, by = "storm_county") %>%
+        dplyr::mutate(rain_exposed = !is.na(rain_exposed),
+                      flood_exposed = !is.na(flood_exposed))
+
+storm_exposed <- all_exposed %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(both_exposed = sum(rain_exposed & flood_exposed),
+                         rain_only = sum(rain_exposed & !flood_exposed),
+                         flood_only = sum(!rain_exposed & flood_exposed),
+                         neither_exposed = sum(!rain_exposed & 
+                                                       !flood_exposed)) 
+storm_exposed %>% 
+        dplyr::mutate(percent_agreement = round(100 * (both_exposed + 
+                                                         neither_exposed) / 
+                              (both_exposed + rain_only + flood_only +
+                              neither_exposed))) %>%
+        dplyr::filter(both_exposed + rain_only + flood_only > 0) %>%
+        dplyr::arrange(neither_exposed, desc(both_exposed), 
+                       desc(rain_only)) %>%
+        kable()
+```
+
+Among the storms with a large number of counties exposed under one or both metrics, Irene (2011), Floyd (1999), and Fran (1996) are notable for very high agreement between the two metrics (98%, 96%, and 95%, respectively).
+Agreement is comparatively lower (although in no cases is the agreement below [x]%, because so many counties are labeled as unexposed under both metrics for most counties) for Frances (2004; 82% agreement), Ivan (2004, 86% agreement), Lee (2011, 88% agreement), Katrina (2005, 89% agreement), and Isidore (2002; 89% agreement).
+In all these cases, most of the discordant classifications were for counties that were classified as exposed based on rain but did not have a reported flood event.
+
+Here are histograms showing the number of counties classified as exposed for both rain and flood metrics, for flood but not rain, and for rain but not flood.
+This is limited to storms between 1996 and 2011 and only includes storm-county pairs where the county was classified as exposed to the storm under at least one of these two metrics (rain and flood).
+It was more common for a county to be classified as exposed based on rain but unexposed by flood then to be assessed as exposed by both metrics or by flood but not rain.
+In a handful of storms, more than 150 counties were assessed as exposed to the storm based on both rain and flood exposure metrics.
+In some storms, quite a number of counties (in some cases, well over 200) were classified as exposed based on the rain metric but unexposed based on the flood metric.
+In almost every storm, however, fewer than 100 counties (out of 2,396 considered for each storm) were classified as exposed based on the flood metric without also being classified as exposed based on the rain metric.
+This suggests that the rain metric might be less stringent than the flood metric and, in some cases, the two metrics identify different risks (i.e., flood risk is not completely determined by rainfall, as pre-existing local conditions can make a flood more or less likely).
+
+```{r echo = FALSE, fig.height = 6, fig.width = 5}
+storm_exposed %>%
+        dplyr::select(-neither_exposed) %>%
+        dplyr::filter(both_exposed > 0 & rain_only > 0 & flood_only > 0) %>%
+        tidyr::gather(classification, n_counties, -storm_id) %>%
+        ggplot(aes(x = n_counties)) + 
+        geom_histogram(bins = 30) + 
+        facet_wrap(~ classification, ncol = 1) + 
+        theme_tufte() + 
+        xlab("Number of counties classified as exposed") + 
+        ylab("Number of storms")
+```
+
+Here is a table showing, for all storms where 200 or more counties were classified as exposed based on at least one of the two metrics, what percent were classified in a way that agreed (i.e., classified as exposed by both metrics):
+
+```{r echo = FALSE}
+storm_exposed %>%
+        dplyr::mutate(any_exposed = both_exposed + rain_only + flood_only,
+                      both_of_any = round(100 * both_exposed /
+                                                  any_exposed)) %>%
+        dplyr::filter(any_exposed >= 200) %>%
+        dplyr::select(-neither_exposed, -any_exposed) %>%
+        dplyr::arrange(desc(both_of_any)) %>%
+        kable()
+```
+
+Only two storms (Irene in 2011 and Floyd and 1999) have over 50% agreement among counties classified as exposed by at least one of the two metrics.
+Throughout, in almost every storm, among the discordant counties (classified as exposed by one but not both metrics), there were many more counties classified as exposed based on the rain metric but where no flood events were reported in the NOAA Storm Events database.
+The only exceptions, where more discordant counties were classified as exposed based on flood than rain, were Allison (2001) and Cindy (2005).
 
@@
 
-```{r echo = FALSE, message = FALSE, warning = FALSE}
+```{r message = FALSE, warning = FALSE}
 #Test
@@
 a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_counties(storm = storm, 
-                                           metric = "distance")) + 
-        ggtitle(paste(storm, "distance"))
+                plot_object = map_rain_exposure(storm = storm, 
+                                          rain_limit = 75,
+                                          dist_limit = 500)) + 
+        ggtitle(paste0(storm, ", Rain-based exposure"))
 b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_counties(storm = storm, 
-                                           metric = "rainfall")) + 
-        ggtitle(paste(storm, "rainfall"))
+                plot_object = map_wind_exposure(storm = storm, 
+                                           wind_limit = 15)) + 
+        ggtitle(paste0(storm, ", Wind-based exposure"))
 
@@
 ```{r fig.width = 6, fig.height = 8, echo = FALSE}
-storm <- "Floyd-1999"
+storm <- "Allison-2001"
 a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_counties(storm = storm, 
-                                           metric = "distance")) + 
-        ggtitle(paste(storm, "distance"))
+                plot_object = map_rain_exposure(storm = storm, 
+                                          rain_limit = 75,
+                                          dist_limit = 500)) + 
+        ggtitle(paste0(storm, ", Rain-based exposure"))
 b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_counties(storm = storm, 
-                                           metric = "rainfall")) + 
-        ggtitle(paste(storm, "rainfall"))
+                plot_object = map_wind_exposure(storm = storm, 
+                                           wind_limit = 15)) + 
+        ggtitle(paste0(storm, ", Wind-based exposure"))
 
@@
 
-Our exposure metrics restricts "exposed" counties to be within 500 km of the storm track.
-In a few cases, hurricane best tracks stop when the storm transitions from a tropical to extratropical system [let's check that I got the wording right on this].
-In these cases, extreme rain from the storm system might have affected counties further along the storm's path, but that are not counted as being near the storm's track.
-For example, Lee in 2011 was only tracked through the best tracks until it was near the Georgia-Tennessee border.
-However, remnants of the system brought extreme rain further north, into the Mid-Atlantic and New England [I believe that the Monthly Weather Review for the 2011 season includes a description of these rain impacts.
-We likely want to look over that and incorporate it here]:
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Lee-2011"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_counties(storm = storm, 
-                                           metric = "rainfall")) + 
-        ggtitle("Rainfall during Lee, 2011")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Lee, 2011")
-grid.arrange(a, b, ncol = 1)
-```
-
-Other possible examples of this during our study period include Nicole (2010) [check storm history in MWR to confirm that Mid-Atlantic and New England rain during this time was related to storm], Ida (2009), Humberto (2007), Katrina (2005), Grace (2003) [confirm rain was from storm remnants], Frances (1998), Marco (1990)
-
-For rain-based exposure, we use a distance constraint of 500 km.
-With storms where the best track extends throughout the storm, this should typically be a large enough swath to include all extreme rain exposures from the storm.
-There are only a few very large storms where this might not be the case.
-One example is Ike (2008), in which case the 500 km constraint excludes some of the counties in northern Texas that got heavy rain from the system:
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Ike-2008"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_counties(storm = storm, 
-                                           metric = "rainfall")) + 
-        ggtitle("Rainfall during Ike, 2008")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Ike, 2008")
-grid.arrange(a, b, ncol = 1)
-```
-
-Other large systems where this might be a concern include Wilma (2005), Hermine (2004) [confirm rain in SC, NC, and VA during this time was from this storm], Arlene (1993) [confirm rain through to LA is from this storm]
-
 Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 18 m/s or greater (e.g., tropical storm strength or greater):
@@
 For all of the above exposure categories, Lili (1996), Karl (2004), and Ophelia (2005) remained the top 3 storms by number of counties exposed based on the various wind exposure metrics, each with 2396 counties exposed [JMF-- I think something weird is going on here].
+
+#### Comparing across all exposure metrics
+
+Here is a comparison of pair-wise metric agreement for all combinations of metrics.
+This analysis is restricted to 1996 to 2011, since that's the period when we have data required for all exposure metrics.
+For most storms and most pairs of exposure metrics, less than half of the storms that are classified as exposed to the storm using one of the metrics are also classified as exposed by the other metric.
+For example, if hurricane exposure classification is based on the "flood" metric, in all storms less than half of the counties with wind events reported in NOAA's Storm Data would be included as exposed counties (see the "wind_event / flood" graph).
+The few exceptions are:
+
+- In many storms, over half of the counties classified as exposed based on the flood metric are also classified as exposed based on the rain metric.
+- In about half of storms, 50% or more of the counties classified as exposed based on rain are also classified as exposed based on distance (i.e., are within 100 km of the storm's track).
+
+```{r echo = FALSE}
+data(closest_dist)
+
+# Generate a list of unique study fips (all in the eastern US)
+my_fips <- closest_dist$fips
+my_fips <- unique(my_fips)
+
+all_storm_counties <- closest_dist %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::filter(as.numeric(gsub(".*-", "", storm_id)) >= 1996 & 
+                              as.numeric(gsub(".*-", "", storm_id)) <= 2011)
+
+distance_exposure <- county_distance(counties = my_fips, start_year = 1996,
+                             end_year = 2011, dist_limit = 100) %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(distance_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, distance_exposed)
+rain_exposure <- county_rain(counties = my_fips, start_year = 1996,
+                             end_year = 2011, rain_limit = 75, 
+                             dist_limit = 500) %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(rain_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, rain_exposed)
+wind_exposure <- county_wind(counties = my_fips, start_year = 1996,
+                             end_year = 2011, wind_limit = 15) %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(wind_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, wind_exposed)
+flood_exposure <- county_events(counties = my_fips, start_year = 1996,
+                             end_year = 2011, event_type = "flood") %>%
+        dplyr::mutate(flood_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, flood_exposed)
+tornado_exposure <- county_events(counties = my_fips, start_year = 1996,
+                             end_year = 2011, event_type = "tornado") %>%
+        dplyr::mutate(tornado_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, tornado_exposed)
+wind_event_exposure <- county_events(counties = my_fips, start_year = 1996,
+                             end_year = 2011, event_type = "wind") %>%
+        dplyr::mutate(wind_event_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, wind_event_exposed)
+
+all_exposed <- distance_exposure %>%
+        dplyr::full_join(rain_exposure, by = "storm_county") %>%
+        dplyr::full_join(wind_exposure, by = "storm_county") %>%
+        dplyr::full_join(flood_exposure, by = "storm_county") %>%
+        dplyr::full_join(tornado_exposure, by = "storm_county") %>%
+        dplyr::full_join(wind_event_exposure, by = "storm_county") %>%
+        dplyr::full_join(all_storm_counties, by = "storm_county") %>%
+        dplyr::mutate(distance_exposed = !is.na(distance_exposed),
+                      rain_exposed = !is.na(rain_exposed),
+                      wind_exposed = !is.na(wind_exposed),
+                      flood_exposed = !is.na(flood_exposed),
+                      tornado_exposed = !is.na(tornado_exposed),
+                      wind_event_exposed = !is.na(wind_event_exposed)) %>%
+        tidyr::gather(exposure, exposed, -storm_county, -storm_id, -fips) %>%
+        dplyr::mutate(exposure = gsub("_exposed", "", exposure))
+
+calc_agreement <- function(all_exposed, exposure_1, exposure_2){
+        df <- all_exposed %>%
+                dplyr::filter(exposure %in% c(exposure_1, exposure_2)) %>%
+                tidyr::spread(exposure, exposed)
+        df$exposure_1 <- df[ , exposure_1]
+        df$exposure_2 <- df[ , exposure_2]
+        df <- df %>%
+                dplyr::filter(exposure_1) %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(exposure_1 = sum(exposure_1),
+                                 exposure_2 = sum(exposure_2),
+                                 agree = 100 * exposure_2 / exposure_1)
+      return(df)
+}
+
+agreement_hist <- function(all_exposed, exposure_1, exposure_2){
+        df <- calc_agreement(all_exposed, exposure_1, exposure_2)
+        a <- ggplot(df, aes(x = agree)) + geom_histogram(bins = 20) + 
+                xlim(c(0, 100)) + theme_tufte() + 
+                ylim(c(0, 20)) + 
+                ylab("# of storms") + 
+                xlab("% of counties") + 
+                ggtitle(paste(exposure_1, "/", exposure_2))
+        return(a)
+}
+```
+
+```{r echo = FALSE, fig.width = 12, fig.height = 12}
+a <- agreement_hist(all_exposed, "distance", "rain")
+b <- agreement_hist(all_exposed, "distance", "wind")
+c <- agreement_hist(all_exposed, "distance", "flood")
+d <- agreement_hist(all_exposed, "distance", "tornado")
+e <- agreement_hist(all_exposed, "distance", "wind_event")
+
+f <- agreement_hist(all_exposed, "rain", "distance")
+g <- agreement_hist(all_exposed, "rain", "wind")
+h <- agreement_hist(all_exposed, "rain", "flood")
+i <- agreement_hist(all_exposed, "rain", "tornado")
+j <- agreement_hist(all_exposed, "rain", "wind_event")
+
+k <- agreement_hist(all_exposed, "flood", "distance")
+l <- agreement_hist(all_exposed, "flood", "rain")
+m <- agreement_hist(all_exposed, "flood", "wind")
+n <- agreement_hist(all_exposed, "flood", "tornado")
+o <- agreement_hist(all_exposed, "flood", "wind_event")
+
+p <- agreement_hist(all_exposed, "tornado", "distance")
+q <- agreement_hist(all_exposed, "tornado", "rain")
+r <- agreement_hist(all_exposed, "tornado", "wind")
+s <- agreement_hist(all_exposed, "tornado", "flood")
+t <- agreement_hist(all_exposed, "tornado", "wind_event")
+
+u <- agreement_hist(all_exposed, "wind_event", "distance")
+v <- agreement_hist(all_exposed, "wind_event", "rain")
+w <- agreement_hist(all_exposed, "wind_event", "wind")
+x <- agreement_hist(all_exposed, "wind_event", "flood")
+y <- agreement_hist(all_exposed, "wind_event", "tornado")
+
+grid.arrange(a, b, c, d, e,
+             f, g, h, i, j,
+             k, l, m, n, o,
+             p, q, r, s, t, 
+             u, v, w, x, y,
+             ncol = 5)
+```
 

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
work on precip validation

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r}
-gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
-nldas <- county_rain(counties = unique(gauge$fips),
+alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+alex_nldas <- county_rain(counties = unique(gauge$fips),
                              start_year = 2010, end_year = 2010,
@@
                              days_include = c(-2, 0, 1))
-nldas <- nldas %>%
+alex_nldas <- alex_nldas %>%
         filter(storm_id == "Alex-2010")
-gauge$fips <- as.character(gauge$fips)
-compare_rain <- full_join(nldas, gauge, by = "fips")
-#compare_rain <- full_join(nldas, gauge, by = "fips")
+alex_gauge$fips <- as.character(alex_gauge$fips)
+compare_rain <- full_join(alex_nldas, alex_gauge, by = "fips")
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
 a <- ggplot(compare_rain, aes(rain_mm, tot_precip, color = fips)) + geom_point()

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
resolved merge conflict

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+Justification of NLDAS data: comparison of NLDAS and gauge data for hurricane Alex (2010).
+
+```{r}
+alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+alex_nldas <- county_rain(counties = unique(gauge$fips),
+                             start_year = 2010, end_year = 2010,
+                             rain_limit = 0, dist_limit = 500,
+                             days_include = c(-2, 0, 1))
+alex_nldas <- alex_nldas %>%
+        filter(storm_id == "Alex-2010")
+alex_gauge$fips <- as.character(alex_gauge$fips)
+compare_rain <- full_join(alex_nldas, alex_gauge, by = "fips")
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+a <- ggplot(compare_rain, aes(rain_mm, tot_precip, color = fips)) + geom_point()
+a
+```
+
 Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
alex_gauge data set

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+Justification of NLDAS data: comparison of NLDAS and gauge data for hurricane Alex (2010).
+
+```{r}
+#library(data.table)
+#alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+#save(alex_gauge, file = "alex_gauge.RData")
+data("alex_gauge")
+alex_nldas <- county_rain(counties = unique(gauge$fips),
+                             start_year = 2010, end_year = 2010,
+                             rain_limit = 0, dist_limit = 500,
+                             days_include = c(-2, 0, 1))
+alex_nldas <- alex_nldas %>%
+        filter(storm_id == "Alex-2010")
+alex_gauge$fips <- as.character(alex_gauge$fips)
+compare_rain <- full_join(alex_nldas, alex_gauge, by = "fips")
+#alex_nldas <- as.data.table(alex_nldas)
+#alex_gauge <- as.data.table(alex_gauge)
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+a <- ggplot(compare_rain, aes(rain_mm, tot_precip, color = fips)) + geom_point()
+a
+```
+
 Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
comment out pulling gauge data in

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r}
-alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+#library(data.table)
+#alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+#save(alex_gauge, file = "alex_gauge.RData")
+#data("~/Documents/hurricaneexposure/alex_gauge.RData")
 alex_nldas <- county_rain(counties = unique(gauge$fips),
@@
 compare_rain <- full_join(alex_nldas, alex_gauge, by = "fips")
+#alex_nldas <- as.data.table(alex_nldas)
+#alex_gauge <- as.data.table(alex_gauge)
 #compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
gauge data in package

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r}
-alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+#library(data.table)
+#alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
+#save(alex_gauge, file = "alex_gauge.RData")
+data(alex_gauge)
 alex_nldas <- county_rain(counties = unique(gauge$fips),

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Move `alex_gauge` file into writing draft directory

@@
 #alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
-#save(alex_gauge, file = "alex_gauge.RData")
+#save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
 data(alex_gauge)

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Change to read in "alex_gauge.RData"

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 #alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
-#save(alex_gauge, file = "alex_gauge.RData")
-data(alex_gauge)
+#save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
+load("alex_gauge.Rdata")
 alex_nldas <- county_rain(counties = unique(gauge$fips),

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work some on code for plots comparing with MWR data

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r}
+```{r echo = FALSE, fig.height = 4, fig.width = 6}
 #library(data.table)
 #alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
-#save(alex_gauge, file = "alex_gauge.RData")
-data(alex_gauge)
-alex_nldas <- county_rain(counties = unique(gauge$fips),
+#save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
+load("alex_gauge.Rdata")
+alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
                              start_year = 2010, end_year = 2010,
@@
 #compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
-a <- ggplot(compare_rain, aes(rain_mm, tot_precip, color = fips)) + geom_point()
+a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
 a
+```
+
+```{r echo = FALSE, fig.width = 4, fig.height = 3}
+compare_rain %>% group_by(fips) %>%
+        summarize(ave_rain_mm = mean(rain_mm),
+                  ave_tot_precip = mean(tot_precip),
+                  n = n()) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, size = n)) + 
+        geom_point(alpha = 0.5) + 
+        theme_bw() +
+        xlim(c(75, 225)) + ylim(c(75, 225)) + 
+        geom_abline(intercept = 0, slope = 1) + 
+        xlab("Rainfall based on NLDAS-2 data (mm)") + 
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)")
 ```

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Fix missing FIPS for Alex comparison

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r}
+```{r echo = FALSE, fig.height = 4, fig.width = 6}
 #library(data.table)
@@
 #save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
-data(alex_gauge)
-alex_nldas <- county_rain(counties = unique(gauge$fips),
+load("alex_gauge.Rdata")
+alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
                              start_year = 2010, end_year = 2010,
-                             rain_limit = 0, dist_limit = 500,
-                             days_include = c(-2, 0, 1))
+                             rain_limit = 0, dist_limit = 1500,
+                             days_include = c(-2, -1, 0, 1))
 alex_nldas <- alex_nldas %>%
@@
 alex_gauge$fips <- as.character(alex_gauge$fips)
-compare_rain <- full_join(alex_nldas, alex_gauge, by = "fips")
+compare_rain <- right_join(alex_nldas, alex_gauge, by = "fips")
 #alex_nldas <- as.data.table(alex_nldas)
@@
 #compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
-a <- ggplot(compare_rain, aes(rain_mm, tot_precip, color = fips)) + geom_point()
+a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
 a
+```
+
+```{r echo = FALSE, fig.width = 4, fig.height = 3}
+compare_rain %>% group_by(fips) %>%
+        summarize(ave_rain_mm = mean(rain_mm),
+                  ave_tot_precip = mean(tot_precip),
+                  n = n()) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, size = n)) + 
+        geom_point(alpha = 0.5) + 
+        theme_bw() +
+        xlim(c(10, 225)) + ylim(c(10, 225)) + 
+        geom_abline(intercept = 0, slope = 1) + 
+        xlab("Rainfall based on NLDAS-2 data (mm)") + 
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)")
 ```

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Fix `days_included` argument in draft code

@@
                              rain_limit = 0, dist_limit = 1500,
-                             days_include = c(-2, -1, 0, 1))
+                             days_included = c(-2, -1, 0, 1))
 alex_nldas <- alex_nldas %>%

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add color to comparison MWR plot to show within 500 km

@@
                   ave_tot_precip = mean(tot_precip),
-                  n = n()) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, size = n)) + 
+                  n = n(),
+                  within_500_km = first(storm_dist) <= 500) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
+                   size = n, color = within_500_km)) + 
         geom_point(alpha = 0.5) + 
@@
         xlab("Rainfall based on NLDAS-2 data (mm)") + 
-        ylab("Rainfall reported in Monthly \nWeather Report (mm)")
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        scale_color_manual(values = c("blue", "black"), guide = "none") 
 ```

### 2016-07-21 writing/Additional_Analysis/allison_casestudy.R
add file with allison precipitation work

@@ -59,3 +59,55 @@ allison <- data.table::fread("data-raw/nasa_precip_export_2.txt",
         dplyr::filter(storm_id == "Allison-2001")
 allison[allison$fips == 12025, "fips"] <- 12086
 
+#Create function to plot out total precipitation for Allison using different lag days (out to -10 and 10)
+
+filter_precip_data <- function(df = df, counties = NULL,
+                              distance_limit = NULL, rain_limit = NULL, include_rain = TRUE,
+                              days_included = NULL,
+                              output_vars = c("fips", "tot_precip")){
+
+        closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
+
+        if(!is.null(counties)){
+                closest_dist <- closest_dist[get("fips") %in% counties]
+        }
+
+        if(!is.null(distance_limit)){
+                closest_dist <- closest_dist[get("storm_dist") <=
+                                                     distance_limit]
+        }
+
+        if(include_rain == TRUE){
+                rain <- data.table::data.table(df)
+                rain <- rain[get("lag") %in% days_included]
+                rain <- rain[ , .(tot_precip = sum(get("precip"))),
+                             by = .(fips, storm_id)]
+                closest_dist <- merge(closest_dist, rain, all.x = TRUE,
+                                      by = c("storm_id", "fips"))
+                if(!is.null(rain_limit)){
+                        closest_dist <- closest_dist[get("tot_precip") >=
+                                                             rain_limit]
+                }
+        }
+
+        closest_dist <- closest_dist[ , .SD, , .SDcols = output_vars]
+        return(closest_dist)
+}
+
+new_function <- function(df = df, days_included = days_included)
+storm_rain <-function(df, days_included = c(-2, -1, 0, 1)){
+                map_data <- filter_data(df = df,
+                                        days_included = days_included,
+                                        output_vars = c("fips",
+                                                        "tot_precip")) %>%
+                        #filter_storm_data(storm = storm, include_rain = TRUE,
+                                              #days_included = days_included,
+                                              #output_vars = c("fips",
+                                                              #"tot_precip")) %>%
+                        dplyr::rename_(region = ~ fips, value = ~ tot_precip)
+        map_data <- map_data %>%
+                dplyr::mutate_(region = ~ as.numeric(region)) %>%
+                dplyr::tbl_df()
+        out <- hurr_choroplethr(map_data, metric = metric)
+        return(out$render())
+}

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add reference lines for 75 mm cutoff to MWR plot

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-```{r echo = FALSE, fig.width = 4, fig.height = 3}
+```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
 compare_rain %>% group_by(fips) %>%
@@
                   ave_tot_precip = mean(tot_precip),
-                  n = n()) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, size = n)) + 
+                  n = n(),
+                  within_500_km = first(storm_dist) <= 500) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
+                   size = n, color = within_500_km)) + 
+        geom_hline(yintercept = 75, color = "gray") + 
+        geom_vline(xintercept = 75, color = "gray") + 
         geom_point(alpha = 0.5) + 

### 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Answer a question in the draft

@@
 exposure_name <- c("Distance_Exposure", "Rain_Exposure", "Wind_Exposure")
-count <- c(sum(distance_count$count), sum(rain_count$count), sum(wind_count$count))
+count <- c(sum(distance_count$count), sum(rain_count$count),
+           sum(wind_count$count))
 exposure_table <- cbind(exposure_name, count)
@@
 ```{r}
-rain_exposure$county_storm <- transmute(rain_exposure, county_storm = paste0(fips, " ", storm_id))
-distance_exposure$county_storm <- transmute(distance_exposure, county_storm = paste0(fips, " ", storm_id))
-all_exposure <- full_join(rain_exposure, distance_exposure, by = "county_storm")
-both_exposure <- subset(all_exposure, tot_precip >= 75 & storm_dist.x <= 100) #4040 observations
-rain_only <- subset(all_exposure, tot_precip >= 75 & storm_dist.x > 100) #4200 obs
-distance_only <- subset(all_exposure, tot_precip < 75 & storm_dist.x <= 100) #0 obs
-sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x <= 100, na.rm = TRUE)
-sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x > 100, na.rm = TRUE)
-sum(all_exposure$tot_precip < 75 & all_exposure$storm_dist.x <= 100, na.rm = TRUE) # It appears that in all county-
-                                                                                   # storm pairs, there is ample
-                                                                                   # rain to meet criteria        
-sum(all_exposure$tot_precip < 75, na.rm = TRUE) # This also returns 0
-```
-
-JMF-- I was surprised to see that there were no observations in the data frame that had total precipitation less than 75 mm. I am wondering if I am missing something in the `full_join` code
+rain_exposure$county_storm <- transmute(rain_exposure,
+                                        county_storm = paste0(fips, " ",
+                                                              storm_id))
+distance_exposure$county_storm <- transmute(distance_exposure,
+                                            county_storm = paste0(fips, " ",
+                                                                  storm_id))
+all_exposure <- rain_exposure %>% 
+        dplyr::select(county_storm, tot_precip) %>%
+        dplyr::full_join(distance_exposure, by = "county_storm")  %>%
+        dplyr::mutate(rain_exposure =  !is.na(tot_precip) & 
+                              tot_precip >= 75,
+                      dist_exposure = !is.na(storm_dist) & 
+                              storm_dist < 100,
+                      both_exposures = rain_exposure & dist_exposure,
+                      rain_only = rain_exposure & !dist_exposure,
+                      dist_only = !rain_exposure & dist_exposure,
+                      neither_exposure = !rain_exposure & !dist_exposure) %>%
+        dplyr::summarize(both_exposure = sum(both_exposures),
+                         rain_only = sum(rain_only),
+                         dist_only = sum(dist_only),
+                         neither_exposure = sum(neither_exposure))
+all_exposure
+
+# both_exposure <- subset(all_exposure, tot_precip >= 75 & 
+#                                 storm_dist <= 100) #4040 observations
+# rain_only <- subset(all_exposure, tot_precip >= 75 & 
+#                             storm_dist.x > 100) #4200 obs
+# distance_only <- subset(all_exposure, tot_precip < 75 & 
+#                                 storm_dist.x <= 100) #0 obs
+# sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x <= 100,
+#     na.rm = TRUE)
+# sum(all_exposure$tot_precip >= 75 & all_exposure$storm_dist.x > 100,
+#     na.rm = TRUE)
+# sum(all_exposure$tot_precip < 75 & all_exposure$storm_dist.x <= 100,
+#     na.rm = TRUE) # It appears that in all county-
+#                                                                                    # storm pairs, there is ample
+#                                                                                    # rain to meet criteria        
+# sum(all_exposure$tot_precip < 75, na.rm = TRUE) # This also returns 0
+```
+
+JMF-- I was surprised to see that there were no observations in the data frame that had total precipitation less than 75 mm.
+I am wondering if I am missing something in the `full_join` code.
+[GBA-- I've re-done this code and it now looks like it's getting the same values as your code below.]
 

### 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
tornado exposure

@@
 
+```{r}
+tornado_exposure <- county_events(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             event_type = "tornado")
+tornado_2000 <- tornado_exposure %>%
+        slice(1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count))
+tornado_2000$value <- cut(tornado_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
+                   include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+map_tornado_2000 <- CountyChoropleth$new(tornado_2000)
+map_tornado_2000$set_zoom(eastern_states)
+map_tornado_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Tornado Events"
+b <- map_tornado_2000$render()
+```
+
 Here are a map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?

### 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
wind and tornado event plots

--- 2016-07-21 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+```{r}
+#Top 2000 storm-county pairs for tornado events during tropical cyclones
+tornado_exposure <- county_events(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             event_type = "tornado")
+tornado_2000 <- tornado_exposure %>%
+        slice(1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count))
+tornado_2000$value <- cut(tornado_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
+                   include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+map_tornado_2000 <- CountyChoropleth$new(tornado_2000)
+map_tornado_2000$set_zoom(eastern_states)
+map_tornado_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Tornado Events"
+a <- map_tornado_2000$render()
+```
+
+```{r}
+#Top 2000 storm-county pairs for wind events during tropical cyclones
+wind_exposure <- county_events(counties = my_fips,
+                              start_year = 1988, end_year = 2011,
+                              event_type = "wind")
+wind_2000 <- wind_exposure %>%
+        slice(1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count))
+wind_2000$value <- cut(wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
+                   include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+map_wind_2000 <- CountyChoropleth$new(wind_2000)
+map_wind_2000$set_zoom(eastern_states)
+map_wind_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Wind Events"
+b <- map_wind_2000$render()
+```
+
+```{r}
+grid.arrange(a, b, ncol = 1)
+```
 Here are a map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?

### 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add flood plot

@@
 
+```{r}
+#Top 2000 storm-county pairs for flooding events during tropical cyclones (since 1996)
+flood_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "flood")
+flood_2000 <- flood_exposure %>%
+        slice(1:2000) %>%
+        group_by(fips) %>%
+        summarise(count = n()) %>%
+        rename(region = fips) %>%
+        mutate(region = as.numeric(region), count = as.numeric(count))
+flood_2000$value <- cut(flood_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
+                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
+                   include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+map_flood_2000 <- CountyChoropleth$new(flood_2000)
+map_flood_2000$set_zoom(eastern_states)
+map_flood_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Flood Events"
+c <- map_flood_2000$render()
+```
+
+```{r}
+grid.arrange(b, c, ncol = 1)
+```
 Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:
@@
 ```
+
 Here are a map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?

### 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
update count plots

@@
 ```
+
 Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:
@@
         mutate(region = as.numeric(region), count = as.numeric(count))
+#Or
+#wind_2000 <- wind_exposure %>%
+        #group_by(fips) %>%
+        #summarise(count = n()) %>%
+        #arrange(desc(count))
+        #slice(1:2000)
+        #rename(region = fips) %>%
+        #mutate(region = as.numeric(region), count = as.numeric(count))
 wind_2000$value <- cut(wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
@@
 distance_count$value <- cut(distance_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24),
+                                                    18, 21, 24, 26),
                            labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24"),
+                                      "16-18", "19-21", "22-24", "25-26"),
                            include.lowest = TRUE)
-exposure_palette <- brewer.pal(7, name = "YlGnBu")
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
 
@@
 rain_count$value <- cut(rain_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24),
+                                                    18, 21, 24, 26),
                            labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24"),
+                                      "16-18", "19-21", "22-24", "25-26"),
                            include.lowest = TRUE)
-exposure_palette <- brewer.pal(7, name = "YlGnBu")
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
 
@@
 wind_count$value <- cut(wind_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24),
+                                                    18, 21, 24, 26),
                            labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24"),
+                                      "16-18", "19-21", "22-24", "25-26"),
                            include.lowest = TRUE)
-exposure_palette <- brewer.pal(7, name = "YlGnBu")
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
 
@@
 grid.arrange(a, b, c, ncol = 1)
+```
+
+```{r}
+tornado_count <- tornado_exposure %>%
+              group_by(fips) %>%
+              summarise(count = n()) %>%
+              rename(region = fips) %>%
+              mutate(region = as.numeric(region), count = as.numeric(count))
+tornado_count$value <- cut(tornado_count$count, c(0, 5, 8, 12, 15,
+                                                    18, 21, 24, 26),
+                           labels = c("0-5", "6-8", "9-12", "13-15",
+                                      "16-18", "19-21", "22-24", "25-26"),
+                           include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+
+map_tornado_count <- CountyChoropleth$new(tornado_count)
+map_tornado_count$set_zoom(eastern_states)
+map_tornado_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_tornado_count$title <- "Total Hits for Storm-County Pairs\nfor Tornado Events"
+d <- map_tornado_count$render()
+```
+
+```{r}
+wind_count <- wind_exposure %>%
+              group_by(fips) %>%
+              summarise(count = n()) %>%
+              rename(region = fips) %>%
+              mutate(region = as.numeric(region), count = as.numeric(count))
+wind_count$value <- cut(wind_count$count, c(0, 5, 8, 12, 15,
+                                                    18, 21, 24, 26),
+                           labels = c("0-5", "6-8", "9-12", "13-15",
+                                      "16-18", "19-21", "22-24", "25-26"),
+                           include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+
+map_wind_count <- CountyChoropleth$new(wind_count)
+map_wind_count$set_zoom(eastern_states)
+map_wind_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_wind_count$title <- "Total Hits for Storm-County Pairs\nfor Wind Events"
+e <- map_wind_count$render()
+```
+
+```{r}
+flood_count <- flood_exposure %>%
+              group_by(fips) %>%
+              summarise(count = n()) %>%
+              rename(region = fips) %>%
+              mutate(region = as.numeric(region), count = as.numeric(count))
+flood_count$value <- cut(flood_count$count, c(0, 5, 8, 12, 15,
+                                                    18, 21, 24, 26),
+                           labels = c("0-5", "6-8", "9-12", "13-15",
+                                      "16-18", "19-21", "22-24", "25-26"),
+                           include.lowest = TRUE)
+exposure_palette <- brewer.pal(8, name = "YlGnBu")
+
+map_flood_count <- CountyChoropleth$new(flood_count)
+map_flood_count$set_zoom(eastern_states)
+map_flood_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+                                                values = exposure_palette,
+                                                na.value = "white")
+map_wind_count$title <- "Total Hits for Storm-County Pairs\nfor Flood Events"
+f <- map_flood_count$render()
 ```
@@
 
-```{r message = FALSE, warning = FALSE}
+```{r message = FALSE, warning = FALSE, echo = FALSE}
 #Test

### 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
event count plots
fix typo

@@
                                                 na.value = "white")
-map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Flood Events"
+map_flood_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Flood Events"
 c <- map_flood_2000$render()
@@
 
+```{r}
+#Event counts
+grid.arrange(d, e, f, ncol = 1)
+```
+
 ```{r echo = FALSE, warning = FALSE, message = FALSE}

### 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
set chunk options for count plots

@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 #Top 2000 storm-county pairs for flooding events during tropical cyclones (since 1996)
@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 #Top 2000 storm-county pairs for tornado events during tropical cyclones
@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 #Top 2000 storm-county pairs for wind events during tropical cyclones
@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 tornado_count <- tornado_exposure %>%
@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 wind_count <- wind_exposure %>%
@@
 
-```{r}
+```{r warning = FALSE, message = FALSE, echo = FALSE}
 flood_count <- flood_exposure %>%

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add ike data

@@
 
+```{r}
+#library(data.table)
+#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
+#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
+load("ike_gauge.Rdata")
+ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
+                             start_year = 2008, end_year = 2008,
+                             rain_limit = 0, dist_limit = 1500,
+                             days_included = c(-2, -1, 0, 1))
+ike_nldas <- ike_nldas %>%
+        filter(storm_id == "Ike-2008")
+ike_gauge$fips <- as.character(ike_gauge$fips)
+compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
+#alex_nldas <- as.data.table(alex_nldas)
+#alex_gauge <- as.data.table(alex_gauge)
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
+b
+```
+
 Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
ike precip validation

--- 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 library(RColorBrewer)
-library(grid)
 ```
@@
 
+```{r}
+library(weathermetrics)
+#library(data.table)
+#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
+#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
+load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
+ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
+                                    "inches", "mm")
+ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
+                             start_year = 2008, end_year = 2008,
+                             rain_limit = 0, dist_limit = 1500,
+                             days_included = c(-2, -1, 0, 1))
+ike_nldas <- ike_nldas %>%
+        filter(storm_id == "Ike-2008")
+ike_gauge$fips <- as.character(ike_gauge$fips)
+compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
+#alex_nldas <- as.data.table(alex_nldas)
+#alex_gauge <- as.data.table(alex_gauge)
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
+b
+```
+
+```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
+compare_rain %>% group_by(fips) %>%
+        summarize(ave_rain_mm = mean(rain_mm),
+                  ave_tot_precip = mean(tot_precip),
+                  n = n(),
+                  within_500_km = first(storm_dist) <= 500) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
+                   size = n, color = within_500_km)) + 
+        geom_hline(yintercept = 75, color = "gray") + 
+        geom_vline(xintercept = 75, color = "gray") + 
+        geom_point(alpha = 0.5) + 
+        theme_bw() +
+        xlim(c(10, 225)) + ylim(c(10, 225)) + 
+        geom_abline(intercept = 0, slope = 1) + 
+        xlab("Rainfall based on NLDAS-2 data (mm)") + 
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        scale_color_manual(values = c("blue", "black"), guide = "none") 
+```
+
 Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
ike precip validation

--- 2016-07-22 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+```{r echo = FALSE, fig.height = 4, fig.width = 6}
+library(weathermetrics)
+#library(data.table)
+#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
+#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
+load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
+ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
+                                    "inches", "mm")
+ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
+                             start_year = 2008, end_year = 2008,
+                             rain_limit = 0, dist_limit = 1500,
+                             days_included = c(-2, -1, 0, 1))
+ike_nldas <- ike_nldas %>%
+        filter(storm_id == "Ike-2008")
+ike_gauge$fips <- as.character(ike_gauge$fips)
+compare_rain1 <- right_join(ike_nldas, ike_gauge, by = "fips")
+#ike_nldas <- as.data.table(alex_nldas)
+#ike_gauge <- as.data.table(alex_gauge)
+#compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
+b
+```
+
+```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
+compare_rain1 %>% group_by(fips) %>%
+        summarize(ave_rain_mm = mean(rain_mm),
+                  ave_tot_precip = mean(tot_precip),
+                  n = n(),
+                  within_500_km = first(storm_dist) <= 500) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
+                   size = n, color = within_500_km)) + 
+        geom_hline(yintercept = 75, color = "gray") + 
+        geom_vline(xintercept = 75, color = "gray") + 
+        geom_point(alpha = 0.5) + 
+        theme_bw() +
+        xlim(c(10, 225)) + ylim(c(10, 225)) + 
+        geom_abline(intercept = 0, slope = 1) + 
+        xlab("Rainfall based on NLDAS-2 data (mm)") + 
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        scale_color_manual(values = c("blue", "black"), guide = "none") 
+```
+
+[JMF-- The two comparison plots look identical. I will keep looking at the code to see why]
+
 Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
note on ike data
add ike wind data

@@
 
+[JMF-- Missing values are those from Arkansas (missing the leading 0 in fips code)]
+
 ```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
@@
 [JMF-- The two comparison plots look identical. I will keep looking at the code to see why]
+
+```{r}
+#Ike-2008 wind
+#ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
+#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
+load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
+```
 

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
ike wind dataset code

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 [JMF-- The two comparison plots look identical. I will keep looking at the code to see why]
+
+```{r}
+#Ike-2008 wind
+#ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
+#save(ike_wind, file = "writing/DraftExposurePaper/ike_wind.RData")
+load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
+```
 

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
begin comparison with countyweather data
correct some code for alex and ike

@@
 #save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
-load("alex_gauge.Rdata")
+load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
 alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
@@
 
-[JMF-- The two comparison plots look identical. I will keep looking at the code to see why]
-
 ```{r}
@@
 load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
+```
+
+####How does NLDAS-2 data compare with rnoaa data from the `countyweather` package in R?
+
+```{r}
+devtools::install_github("ropenscilabs/rnoaa")
+library(rnoaa)
+devtools::install_github("leighseverson/countyweather")
+library(countyweather)
+options("noaakey" = Sys.getenv("noaakey"))
+alex_weather <- weather_fips(fips = alex_nldas$fips,
+                             date_min = "2010-06-25",
+                             date_max = "2010-07-02",
+                             var = "PRCP")
 ```

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
begin coding countyweather data

@@
 ```{r}
-devtools::install_github("ropenscilabs/rnoaa")
+#devtools::install_github("ropenscilabs/rnoaa")
 library(rnoaa)
-devtools::install_github("leighseverson/countyweather")
+#devtools::install_github("leighseverson/countyweather")
 library(countyweather)
 options("noaakey" = Sys.getenv("noaakey"))
-alex_weather <- weather_fips(fips = alex_nldas$fips,
+alex_fips <- alex_nldas$fips
+alex_fips <- unique(alex_fips)
+alex_weather <- weather_fips(fips = alex_fips,
                              date_min = "2010-06-25",

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
edit code for ike precip validation plot

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 #compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+b <- ggplot(compare_rain1, aes(x = tot_precip, y = rain_mm,
                               color = fips)) +
@@
 ```{r}
-devtools::install_github("ropenscilabs/rnoaa")
+#devtools::install_github("ropenscilabs/rnoaa")
 library(rnoaa)
-devtools::install_github("leighseverson/countyweather")
+#devtools::install_github("leighseverson/countyweather")
 library(countyweather)
 options("noaakey" = Sys.getenv("noaakey"))
-alex_weather <- weather_fips(fips = alex_nldas$fips,
-                             date_min = "2010-06-25",
-                             date_max = "2010-07-02",
-                             var = "PRCP")
+alex_fips <- alex_nldas$fips
+alex_fips <- unique(alex_fips)
+#alex_weather <- weather_fips(fips = alex_fips,
+                             #date_min = "2010-06-25",
+                             #date_max = "2010-07-02",
+                             #var = "PRCP") # Returns warning: "sorry, no data found"
 ```

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
more code for countyweather data

@@
 options("noaakey" = Sys.getenv("noaakey"))
+#Alex-2010
 alex_fips <- alex_nldas$fips
@@
                              #var = "PRCP") # Returns warning: "sorry, no data found"
+                                            # and problemin station_ids$meta . . .
+#Ike-2008
+ike_fips <- ike_nldas$fips
+ike_fips <- unique(ike_fips)
+#ike_weather <- weather_fips(fips = ike_fips,
+                             #date_min = "2008-09-01",
+                             #date_max = "2010-09-14",
+                             #var = "PRCP") # Same thing here
 ```

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add Allison MWR data example for rainfall.

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 #save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
+load("alex_gauge.Rdata")
 alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)")  + 
-        scale_color_manual(values = c("blue", "black"), guide = "none") 
+        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
+                           guide = "none") 
 ```
@@
 ```{r echo = FALSE, fig.height = 4, fig.width = 6}
-library(weathermetrics)
-#library(data.table)
-#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
-#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
-ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
-                                    "inches", "mm")
-ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
-                             start_year = 2008, end_year = 2008,
+allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv")
+allison_nldas <- county_rain(counties = unique(allison_gauge$fips),
+                             start_year = 2001, end_year = 2001,
                              rain_limit = 0, dist_limit = 1500,
                              days_included = c(-2, -1, 0, 1))
-ike_nldas <- ike_nldas %>%
-        filter(storm_id == "Ike-2008")
-ike_gauge$fips <- as.character(ike_gauge$fips)
-compare_rain1 <- right_join(ike_nldas, ike_gauge, by = "fips")
-#ike_nldas <- as.data.table(alex_nldas)
-#ike_gauge <- as.data.table(alex_gauge)
-#compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+allison_nldas <- allison_nldas %>%
+        filter(storm_id == "Allison-2001")
+allison_gauge$fips <- as.character(allison_gauge$fips)
+compare_rain <- right_join(allison_nldas, allison_gauge, by = "fips")
+#alex_nldas <- as.data.table(alex_nldas)
+#alex_gauge <- as.data.table(alex_gauge)
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
                               color = fips)) +
@@
         theme_bw()
-b
-```
-
-[JMF-- Missing values are those from Arkansas (missing the leading 0 in fips code)]
+a
+```
 
 ```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain1 %>% group_by(fips) %>%
+compare_rain %>% group_by(fips) %>%
         summarize(ave_rain_mm = mean(rain_mm),
@@
         theme_bw() +
-        xlim(c(10, 225)) + ylim(c(10, 225)) + 
+        xlim(c(25, 200)) + ylim(c(25, 800)) + 
         geom_abline(intercept = 0, slope = 1) + 
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
-        scale_color_manual(values = c("blue", "black"), guide = "none") 
-```
-
-```{r}
-#Ike-2008 wind
-#ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
-#save(ike_wind, file = "writing/DraftExposurePaper/ike_wind.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
-```
-
-####How does NLDAS-2 data compare with rnoaa data from the `countyweather` package in R?
-
-```{r}
-#devtools::install_github("ropenscilabs/rnoaa")
-library(rnoaa)
-#devtools::install_github("leighseverson/countyweather")
-library(countyweather)
-options("noaakey" = Sys.getenv("noaakey"))
-alex_fips <- alex_nldas$fips
-alex_fips <- unique(alex_fips)
-alex_weather <- weather_fips(fips = alex_fips,
-                             date_min = "2010-06-25",
-                             date_max = "2010-07-02",
-                             var = "PRCP")
+        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
+                           guide = "none") 
 ```

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)")  + 
-        scale_color_manual(values = c("blue", "black"), guide = "none") 
+        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
+                           guide = "none") 
+```
+
+```{r echo = FALSE, fig.height = 4, fig.width = 6}
+allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv")
+allison_nldas <- county_rain(counties = unique(allison_gauge$fips),
+                             start_year = 2001, end_year = 2001,
+                             rain_limit = 0, dist_limit = 1500,
+                             days_included = c(-2, -1, 0, 1))
+allison_nldas <- allison_nldas %>%
+        filter(storm_id == "Allison-2001")
+allison_gauge$fips <- as.character(allison_gauge$fips)
+compare_rain <- right_join(allison_nldas, allison_gauge, by = "fips")
+#alex_nldas <- as.data.table(alex_nldas)
+#alex_gauge <- as.data.table(alex_gauge)
+#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
+a
+```
+
+```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
+compare_rain %>% group_by(fips) %>%
+        summarize(ave_rain_mm = mean(rain_mm),
+                  ave_tot_precip = mean(tot_precip),
+                  n = n(),
+                  within_500_km = first(storm_dist) <= 500) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
+                   size = n, color = within_500_km)) + 
+        geom_hline(yintercept = 75, color = "gray") + 
+        geom_vline(xintercept = 75, color = "gray") + 
+        geom_point(alpha = 0.5) + 
+        theme_bw() +
+        xlim(c(25, 200)) + ylim(c(25, 800)) + 
+        geom_abline(intercept = 0, slope = 1) + 
+        xlab("Rainfall based on NLDAS-2 data (mm)") + 
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
+                           guide = "none") 
 ```

### 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Fix example for Ike county weather

@@
 #save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
+#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
+load("ike_gauge.RData")
 ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
@@
 ike_fips <- unique(ike_fips)
-#ike_weather <- weather_fips(fips = ike_fips,
-                             #date_min = "2008-09-01",
-                             #date_max = "2010-09-14",
-                             #var = "PRCP") # Same thing here
+
+ike_weather <- weather_fips(fips = ike_fips[1],
+                             date_min = "2008-09-01",
+                             date_max = "2010-09-14",
+                             var = "PRCP") # Same thing here
 ```

### 2016-07-26 writing/DraftExposurePaper/DraftExposurePaper.Rmd
code with county_timeseries function

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
-        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
-                           guide = "none") 
-```
-
-```{r echo = FALSE, fig.height = 4, fig.width = 6}
-allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv")
-allison_nldas <- county_rain(counties = unique(allison_gauge$fips),
-                             start_year = 2001, end_year = 2001,
-                             rain_limit = 0, dist_limit = 1500,
-                             days_included = c(-2, -1, 0, 1))
-allison_nldas <- allison_nldas %>%
-        filter(storm_id == "Allison-2001")
-allison_gauge$fips <- as.character(allison_gauge$fips)
-compare_rain <- right_join(allison_nldas, allison_gauge, by = "fips")
-#alex_nldas <- as.data.table(alex_nldas)
-#alex_gauge <- as.data.table(alex_gauge)
-#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
-a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw()
-a
-```
-
-```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain %>% group_by(fips) %>%
-        summarize(ave_rain_mm = mean(rain_mm),
-                  ave_tot_precip = mean(tot_precip),
-                  n = n(),
-                  within_500_km = first(storm_dist) <= 500) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
-                   size = n, color = within_500_km)) + 
-        geom_hline(yintercept = 75, color = "gray") + 
-        geom_vline(xintercept = 75, color = "gray") + 
-        geom_point(alpha = 0.5) + 
-        theme_bw() +
-        xlim(c(25, 200)) + ylim(c(25, 800)) + 
-        geom_abline(intercept = 0, slope = 1) + 
-        xlab("Rainfall based on NLDAS-2 data (mm)") + 
-        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
-        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
-                           guide = "none") 
+        scale_color_manual(values = c("blue", "black"), guide = "none") 
 ```
@@
 ike_fips <- unique(ike_fips)
-#ike_weather <- weather_fips(fips = ike_fips,
+#ike_weather <- county_timeseries(fips = ike_fips,
                                  #date_min = "2008-09-01",
-                             #date_max = "2010-09-14",
-                             #var = "PRCP") # Same thing here
+                                 #date_max = "2008-09-14",
+                                 #out_directory = paste0("/Users/joshuaferreri/",
+                                                        #"Documents/ike_timeseries"),
+                                 #out_type = ".csv")
 ```

### 2016-07-26 writing/DraftExposurePaper/DraftExposurePaper.Rmd

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ike_fips <- unique(ike_fips)
-
+#ike_weather <- county_timeseries(fips = ike_fips,
+                                 #date_min = "2008-09-01",
+                                 #date_max = "2008-09-14",
+                                 #out_directory = paste0("/Users/joshuaferreri",
+                                                # "/Documents/ike_timeseries"),
+                                 #out_type = ".csv")
 ike_weather <- weather_fips(fips = ike_fips[1],

### 2016-07-26 writing/DraftExposurePaper/DraftExposurePaper.Rmd
attempts with countyweather dataset

--- 2016-07-25 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ike_fips <- unique(ike_fips)
-
+#ike_weather <- county_timeseries(fips = ike_fips,
+                                 #date_min = "2008-09-01",
+                                 #date_max = "2008-09-14",
+                                 #out_directory = paste0("/Users/joshuaferreri",
+                                                # "/Documents/ike_timeseries"),
+                                 #out_type = ".csv")
 ike_weather <- weather_fips(fips = ike_fips[1],
                              date_min = "2008-09-01",
-                             date_max = "2010-09-14",
+                             date_max = "2008-09-14",
+                             #date_max = "2010-09-14",
                              var = "PRCP") # Same thing here

### 2016-07-26 writing/DraftExposurePaper/DraftExposurePaper.Rmd
id fips with most stations

--- 2016-07-26 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
                              date_min = "2008-09-01",
-                             date_max = "2010-09-14",
-                             var = "PRCP") # Same thing here
+                             date_max = "2008-09-14",
+                             var = "PRCP") 
+#Find fips with greatest number of observations
+test <- data.frame(fips = ike_fips[1:2], station = NA)
+#for(fips in ike_fips[1:2]){ 
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        test["fips", ] <- fips
+        test[ , "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #Still working on it
+ike_table <- data.frame(fips = ike_fips, station = NA)
+#for(fips in ike_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        ike_table[ , "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #still working on it
 ```

### 2016-07-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
work on loop to ID station number

--- 2016-07-26 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
                              date_min = "2008-09-01",
-                             date_max = "2010-09-14",
-                             var = "PRCP") # Same thing here
+                             date_max = "2008-09-14",
+                             var = "PRCP") 
+```
+
+Trying to create a loop to identify the FIPS codes with the greatest number of stations reporting for each time period.
+
+```{r}
+test <- data.frame(fips = ike_fips[1:2], station = NA)
+#for(fips in ike_fips[1:2]){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        test[fips, "stations"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #Still working on it
+ike_table <- data.frame(fips = ike_fips, station = NA)
+#for(fips in ike_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        ike_table[ , "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #still working on it
 ```

### 2016-07-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
work on loop

@@
 
+Comparing MWR precipitation data to NLDAS-2 data for Frances-2004:
+
+```{r}
+
+```
+
+Ike wind data:
+
 ```{r}
@@
 ```{r}
-test <- data.frame(fips = ike_fips[1:2], station = NA)
-#for(fips in ike_fips[1:2]){
+my_fips <- ike_fips[1:3]
+my_stations <- NA
+test <- expand.grid(my_fips, my_stations)
+colnames(test) <- c("fips", "station")
+for(fips in my_fips){
         ike_weather <- weather_fips(fips = fips,
@@
                                     var = "PRCP")
-        test[fips, "stations"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #Still working on it
-ike_table <- data.frame(fips = ike_fips, station = NA)
-#for(fips in ike_fips){
+        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+}
+ike_table <- expand.grid(ike_fips, my_stations)
+colnames(ike_table) <- c("fips", "station")
+for(fips in ike_fips){
         ike_weather <- weather_fips(fips = fips,
@@
                                     var = "PRCP")
-        ike_table[ , "station"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #still working on it
+        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} 
 ```

### 2016-07-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
loop work

--- 2016-07-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
                                     var = "PRCP")
-        test[fips, "stations"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #Still working on it
-ike_table <- data.frame(fips = ike_fips, station = NA)
+        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+} #This works fine (small subset of code below)
+ike_table <- expand.grid(ike_fips, my_stations)
+colnames(ike_table) <- c("fips", "station")
 #for(fips in ike_fips){
@@
                                     var = "PRCP")
-        ike_table[ , "station"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #still working on it
+        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #getting error
 ```

### 2016-07-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Frances precip comparison

@@
 ```{r}
-
+#frances_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_gauge.csv", header = TRUE)
+#save(frances_gauge, file = "writing/DraftExposurePaper/frances_gauge.RData")
+#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_gauge.RData")
+frances_gauge$precip_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
+frances_nldas <- county_rain(counties = unique(frances_gauge$fips),
+                             start_year = 2004, end_year = 2004,
+                             rain_limit = 0, dist_limit = 1500,
+                             days_included = c(-2, -1, 0, 1))
+frances_nldas <- frances_nldas %>%
+        filter(storm_id == "Frances-2004")
+frances_gauge$fips <- as.character(frances_gauge$fips)
+compare_rain <- right_join(frances_nldas, frances_gauge, by = "fips")
+b <- ggplot(compare_rain, aes(x = tot_precip, y = precip_mm,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw()
+b
+```
+
+```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
+compare_rain %>% group_by(fips) %>%
+        summarize(ave_rain_mm = mean(precip_mm),
+                  ave_tot_precip = mean(tot_precip),
+                  n = n(),
+                  within_500_km = first(storm_dist) <= 500) %>%
+        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
+                   size = n, color = within_500_km)) + 
+        geom_hline(yintercept = 75, color = "gray") + 
+        geom_vline(xintercept = 75, color = "gray") + 
+        geom_point(alpha = 0.5) + 
+        theme_bw() +
+        xlim(c(10, 225)) + ylim(c(10, 225)) + 
+        geom_abline(intercept = 0, slope = 1) + 
+        xlab("Rainfall based on NLDAS-2 data (mm)") + 
+        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        scale_color_manual(values = c("blue", "black"), guide = "none") 
 ```

### 2016-07-27 writing/DraftExposurePaper/DraftExposurePaper.Rmd
edited frances wind dataset

@@
 #ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
-#save(ike_wind, file = "writing/DraftExposurePaper/ike_wind.RData")
+#save(ike_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
 load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
+```
+
+Frances wind data:
+
+```{r}
+frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", na.strings = 999,
+                         header = TRUE)
+save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
+load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_wind.RData")
 ```

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
work on code to pull data for all ike_fips available

@@
 ike_gauge$fips <- as.character(ike_gauge$fips)
-compare_rain1 <- right_join(ike_nldas, ike_gauge, by = "fips")
+compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
 #ike_nldas <- as.data.table(alex_nldas)
@@
 ```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain1 %>% group_by(fips) %>%
+compare_rain %>% group_by(fips) %>%
         summarize(ave_rain_mm = mean(rain_mm),
@@
 ```{r}
-my_fips <- ike_fips[1:3]
+my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
 my_stations <- NA
-test <- expand.grid(my_fips, my_stations)
-colnames(test) <- c("fips", "station")
+ike_table <- expand.grid(my_fips, my_stations)
+colnames(ike_table) <- c("fips", "station")
 for(fips in my_fips){
@@
                                     var = "PRCP")
-        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-} #This works fine (small subset of code below)
+        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+} #This works fine (small subset of the code below)
+my_fips <- ike_fips[44:45]
+my_stations <- NA
+test_2 <- expand.grid(my_fips, my_stations)
+colnames(test_2) <- c("fips", "station")
+for(fips in my_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        test_2[test_2$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+}
+#Do not work: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
+ike_table <- ike_table %>%
+        arrange(desc(station)) %>%
+        slice(1:15)
 ike_table <- expand.grid(ike_fips, my_stations)

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
ike reporting stations code

@@
 ```{r}
+#Find the top 15 FIPS by number of reporting stations during Ike-2008
 my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
@@
         ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-} #This works fine (small subset of the code below)
-my_fips <- ike_fips[44:45]
-my_stations <- NA
-test_2 <- expand.grid(my_fips, my_stations)
-colnames(test_2) <- c("fips", "station")
-for(fips in my_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        test_2[test_2$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
 } 
@@
         slice(1:15)
-ike_table <- expand.grid(ike_fips, my_stations)
-colnames(ike_table) <- c("fips", "station")
-#for(fips in ike_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #getting error
 ```

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
find top fips for alex

@@
 ```{r}
-#Find the top 15 FIPS by number of reporting stations during Ike-2008
+#Find the top 15 FIPS by number of reporting stations during Ike-2008 (Warning: Takes a long time)
 my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
@@
 } 
-#Do not work: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
+#Ike FIPS without data: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
 ike_table <- ike_table %>%
@@
         slice(1:15)
+ike_top <- ike_table$fips
+#Find the top 15 FIPS by number of reporting stations for Alex-2010:
+my_fips <- alex_fips
+my_stations <- NA
+alex_table <- expand.grid(my_fips, my_stations)
+colnames(alex_table) <- c("fips", "station")
+for(fips in my_fips){
+        alex_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
+} 
+alex_table <- alex_table %>%
+        arrange(desc(station)) %>%
+        slice(1:15)
+alex_top <- alex_table$fips
 ```

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Alex FIPS for top counties

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
         slice(1:15)
+ike_top <- ike_table$fips
+#Find the top 15 FIPS by number of reporting stations for Alex-2010:
+my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
+my_fips <- alex_fips[18]
+my_stations <- NA
+alex_table <- expand.grid(my_fips, my_stations)
+colnames(alex_table) <- c("fips", "station")
+for(fips in my_fips){
+        alex_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
+} 
+#Alex FIPS without data: 3, 7, 9, 11, 15, 19   
+alex_table <- alex_table %>%
+        arrange(desc(station)) %>%
+        slice(1:15)
+alex_top <- alex_table$fips
 ```

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
begin ike timeseries

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 alex_fips <- unique(alex_fips)
-#alex_weather <- weather_fips(fips = alex_fips,
-                             #date_min = "2010-06-25",
-                             #date_max = "2010-07-02",
-                             #var = "PRCP") # Returns warning: "sorry, no data found"
-                                            # and problemin station_ids$meta . . .
 #Ike-2008
@@
 ike_fips <- unique(ike_fips)
-#ike_weather <- county_timeseries(fips = ike_fips,
-                                 #date_min = "2008-09-01",
-                                 #date_max = "2008-09-14",
-                                 #out_directory = paste0("/Users/joshuaferreri",
-                                                # "/Documents/ike_timeseries"),
-                                 #out_type = ".csv")
-ike_weather <- weather_fips(fips = ike_fips[1],
-                             date_min = "2008-09-01",
-                             date_max = "2008-09-14",
-                             var = "PRCP") 
 ```
@@
         slice(1:15)
+ike_top <- ike_table$fips
+#Find the top 15 FIPS by number of reporting stations for Alex-2010:
+my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
+my_stations <- NA
+alex_table <- expand.grid(my_fips, my_stations)
+colnames(alex_table) <- c("fips", "station")
+for(fips in my_fips){
+        alex_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
+} 
+#Alex FIPS without data: 3, 7, 9, 11, 15, 19   
+alex_table <- alex_table %>%
+        arrange(desc(station)) %>%
+        slice(1:15)
+alex_top <- alex_table$fips
+fips_table <- rbind(alex_table, ike_table)
+fips_table <- fips_table %>%
+        arrange(desc(station)) %>%
+        slice(1:15)
+all_fips <- fips_table$fips
+all_fips <- unique(all_fips)
+```
+
+As of right now, I am not going to do Frances (there are 184 unique fips).
+
+Create a timeseries for the top 15 FIPS by reporting stations for Alex and Ike:
+
+```{r}
+ike_weather <- county_timeseries(fips = ike_top[1:2], 
+                                 percent_coverage = 0, 
+                                 date_min = "2008-09-01", 
+                                 date_max = "2010-09-14", var = "PRCP", 
+                                 out_directory = "~/tmp/ike_ex")
+
 ```

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
comment out weather_fips code

@@
 #Find the top 15 FIPS by number of reporting stations during Ike-2008 (Warning: Takes a long time)
-my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
-my_stations <- NA
-ike_table <- expand.grid(my_fips, my_stations)
-colnames(ike_table) <- c("fips", "station")
-for(fips in my_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-} 
+#my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
+#my_stations <- NA
+#ike_table <- expand.grid(my_fips, my_stations)
+#colnames(ike_table) <- c("fips", "station")
+#for(fips in my_fips){
+        #ike_weather <- weather_fips(fips = fips,
+                                    #date_min = "2008-09-01",
+                                    #date_max = "2008-09-14",
+                                    #var = "PRCP")
+        #ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+#} 
 #Ike FIPS without data: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
-ike_table <- ike_table %>%
-        arrange(desc(station)) %>%
-        slice(1:15)
-ike_top <- ike_table$fips
+#ike_table <- ike_table %>%
+        #arrange(desc(station)) %>%
+        #slice(1:15)
+#ike_top <- ike_table$fips
 #Find the top 15 FIPS by number of reporting stations for Alex-2010:
-my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
-my_stations <- NA
-alex_table <- expand.grid(my_fips, my_stations)
-colnames(alex_table) <- c("fips", "station")
-for(fips in my_fips){
-        alex_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
-} 
+#my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
+#my_stations <- NA
+#alex_table <- expand.grid(my_fips, my_stations)
+#colnames(alex_table) <- c("fips", "station")
+#for(fips in my_fips){
+#        alex_weather <- weather_fips(fips = fips,
+#                                    date_min = "2008-09-01",
+#                                    date_max = "2008-09-14",
+#                                    var = "PRCP")
+#        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
+#} 
 #Alex FIPS without data: 3, 7, 9, 11, 15, 19   
-alex_table <- alex_table %>%
-        arrange(desc(station)) %>%
-        slice(1:15)
-alex_top <- alex_table$fips
-fips_table <- rbind(alex_table, ike_table)
-fips_table <- fips_table %>%
-        arrange(desc(station)) %>%
-        slice(1:15)
-all_fips <- fips_table$fips
-all_fips <- unique(all_fips)
+#alex_table <- alex_table %>%
+#        arrange(desc(station)) %>%
+#        slice(1:15)
+#alex_top <- alex_table$fips
+#fips_table <- rbind(alex_table, ike_table)
+#fips_table <- fips_table %>%
+#        arrange(desc(station)) %>%
+#        slice(1:15)
+#all_fips <- fips_table$fips
+#all_fips <- unique(all_fips)
 ```
@@
 ```{r}
-ike_weather <- county_timeseries(fips = ike_top[1:2], 
+county_timeseries(fips = ike_top[1:2], 
                   percent_coverage = 0, 

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
begin writing generic code to compare nldas to county weather data

@@
                   out_directory = "~/tmp/ike_ex")
-
+ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
+ike_1 <- ike_1 %>%
+        select(date, prcp) %>%
+        filter(date %in% seq(from = as.Date("2008-09-01"), to = as.Date("2008-09-14"), by = 1))
+data(rain)
+ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
+data("closest_dist")
+keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
+keys_cd <- keys_cd$closest_date
+ike_1 <- ike_1 %>%
+        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), 1))
+#assuming the prcp is in inches
+ike_1$prcp_mm <- convert_precip(ike_1$prcp, "inches", "mm")
+ike_sum <- sum(ike_1$prcp_mm)
+ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
+ike_nldas_sum <- ike_nldas_sum$tot_precip
 ```

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
work on code for ike countyweather

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
                   out_directory = "~/tmp/ike_ex")
-
-```
+ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
+ike_1 <- ike_1 %>%
+        select(date, prcp) %>%
+        filter(date %in% seq(from = as.Date("2008-09-01"), to = as.Date("2008-09-14"), by = 1))
+data(rain)
+ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
+data("closest_dist")
+keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
+keys_cd <- keys_cd$closest_date
+ike_1 <- ike_1 %>%
+        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), 1))
+ike_sum <- sum(ike_1$prcp)
+ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
+ike_nldas_sum <- ike_nldas_sum$tot_precip
+```
+
+[JMF-- It looks like precipitation measurements from `countyweather` are already in mm]
 

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
clean up code

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r}
+data("closest_dist")
+data(rain)
 county_timeseries(fips = ike_top[1:2], 
@@
                   out_directory = "~/tmp/ike_ex")
-
-```
+ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
+keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
+keys_cd <- keys_cd$closest_date
+ike_1 <- ike_1 %>%
+        select(date, prcp) %>%
+        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), by = 1))
+ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
+ike_sum <- sum(ike_1$prcp)
+ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
+ike_nldas_sum <- ike_nldas_sum$tot_precip
+```
+
+[JMF-- It looks like precipitation measurements from `countyweather` are already in mm]
 

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add some to text

@@
 
-The most often used hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used?
-We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure] and whether a county's perimeter intercepts a swath describing a buffer a certain distance from the storm's central track.
-These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016].
+Good Characteristics:
+- Flooding-- Most associated with death; property damage
+- Precipitation
+- Wind (potentially weighted)
+
+Bad Characteristics:
+- Distance alone
+- Wind alone
+
+Hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used?
+We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath describing a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale).
+These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016; @Fronstin1994].
 FEMA declaration is largely subjective, requiring state governors and/or tribal chiefs to submit a proposal for government aid on the basis that local resources are overwhelmed [@FEMA].

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Tweak some file paths

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 #save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
+# load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
+load("alex_gauge.Rdata")
 alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
@@
 ike_gauge$fips <- as.character(ike_gauge$fips)
-compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
+compare_rain1 <- right_join(ike_nldas, ike_gauge, by = "fips")
 #ike_nldas <- as.data.table(alex_nldas)
@@
 ```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain %>% group_by(fips) %>%
+compare_rain1 %>% group_by(fips) %>%
         summarize(ave_rain_mm = mean(rain_mm),
@@
 #load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_gauge.RData")
+load("frances_gauge.RData")
 frances_gauge$precip_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
@@
 #save(ike_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
+# load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
+load("ike_wind.RData")
 ```
@@
 
-```{r}
-frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", na.strings = 999,
+```{r eval = FALSE}
+frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
+                         na.strings = 999,
                          header = TRUE)
 save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
-load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_wind.RData")
+```
+
+```{r}
+load("frances_wind.RData")
 ```
@@
 
-```{r}
+```{r eval = FALSE}
 #devtools::install_github("ropenscilabs/rnoaa")
@@
 alex_fips <- unique(alex_fips)
+#alex_weather <- weather_fips(fips = alex_fips,
+                             #date_min = "2010-06-25",
+                             #date_max = "2010-07-02",
+                             #var = "PRCP") # Returns warning: "sorry, no data found"
+                                            # and problemin station_ids$meta . . .
 #Ike-2008
@@
 ike_fips <- unique(ike_fips)
-```
-
-Trying to create a loop to identify the FIPS codes with the greatest number of stations reporting for each time period.
-
-```{r}
-#Find the top 15 FIPS by number of reporting stations during Ike-2008 (Warning: Takes a long time)
-#my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
-#my_stations <- NA
-#ike_table <- expand.grid(my_fips, my_stations)
-#colnames(ike_table) <- c("fips", "station")
-#for(fips in my_fips){
-        #ike_weather <- weather_fips(fips = fips,
+#ike_weather <- county_timeseries(fips = ike_fips,
                                  #date_min = "2008-09-01",
                                  #date_max = "2008-09-14",
-                                    #var = "PRCP")
-        #ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-#} 
-#Ike FIPS without data: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
-#ike_table <- ike_table %>%
-        #arrange(desc(station)) %>%
-        #slice(1:15)
-#ike_top <- ike_table$fips
-#Find the top 15 FIPS by number of reporting stations for Alex-2010:
-#my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
-#my_stations <- NA
-#alex_table <- expand.grid(my_fips, my_stations)
-#colnames(alex_table) <- c("fips", "station")
-#for(fips in my_fips){
-#        alex_weather <- weather_fips(fips = fips,
-#                                    date_min = "2008-09-01",
-#                                    date_max = "2008-09-14",
-#                                    var = "PRCP")
-#        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
-#} 
-#Alex FIPS without data: 3, 7, 9, 11, 15, 19   
-#alex_table <- alex_table %>%
-#        arrange(desc(station)) %>%
-#        slice(1:15)
-#alex_top <- alex_table$fips
-#fips_table <- rbind(alex_table, ike_table)
-#fips_table <- fips_table %>%
-#        arrange(desc(station)) %>%
-#        slice(1:15)
-#all_fips <- fips_table$fips
-#all_fips <- unique(all_fips)
-```
-
-As of right now, I am not going to do Frances (there are 184 unique fips).
-
-Create a timeseries for the top 15 FIPS by reporting stations for Alex and Ike:
-
-```{r}
-data("closest_dist")
-data(rain)
-county_timeseries(fips = ike_top[1:2], 
-                  percent_coverage = 0, 
+                                 #out_directory = paste0("/Users/joshuaferreri",
+                                                # "/Documents/ike_timeseries"),
+                                 #out_type = ".csv")
+ike_weather <- weather_fips(fips = ike_fips[1],
                              date_min = "2008-09-01",
-                  date_max = "2010-09-14", var = "PRCP", 
-                  out_directory = "~/tmp/ike_ex")
-ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
-keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
-keys_cd <- keys_cd$closest_date
-ike_1 <- ike_1 %>%
-        select(date, prcp) %>%
-        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), by = 1))
-ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
-ike_sum <- sum(ike_1$prcp)
-ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
-ike_nldas_sum <- ike_nldas_sum$tot_precip
-```
-
-[JMF-- It looks like precipitation measurements from `countyweather` are already in mm]
+                             date_max = "2008-09-14",
+                             var = "PRCP") 
+```
+
+Trying to create a loop to identify the FIPS codes with the greatest number of stations reporting for each time period.
+
+```{r eval = FALSE}
+my_fips <- ike_fips[1:3]
+my_stations <- NA
+test <- expand.grid(my_fips, my_stations)
+colnames(test) <- c("fips", "station")
+for(fips in my_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+} #This works fine (small subset of code below)
+ike_table <- expand.grid(ike_fips, my_stations)
+colnames(ike_table) <- c("fips", "station")
+#for(fips in ike_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #getting error
+```
 

### 2016-07-28 writing/DraftExposurePaper/example_ike_code.R
Add example file for Ike and countyweather

@@ -1,0 +1,53 @@
+county_timeseries(ike_fips, percent_coverage = 0,
+                  date_min = "1988-01-01", date_max = "2011-12-31",
+                  var = "PRCP")
+
+ike_dir <- "~/tmp/ike_ex/"
+ike_counties <- gsub(".rds", "", list.files(ike_dir))
+
+ike_ave <- vector("list", length = length(ike_counties))
+names(ike_ave) <- ike_counties
+
+for(x in ike_counties){
+        county_weather <- readRDS(paste0(ike_dir, x, ".rds"))$averaged
+        county_weather <- county_weather %>%
+                filter(year(date) == 2008 & month(date) == 9)
+        county_weather$fips <- x
+        ike_ave[[x]] <- county_weather
+}
+ike_ave <- do.call("rbind", ike_ave)
+
+ike_rain <- county_rain(counties = ike_counties, start_year = 2008,
+                        end_year = 2008, rain_limit = 0, dist_limit = 1000) %>%
+        filter(storm_id == "Ike-2008")
+
+library(ggthemes)
+ike_ave %>%
+        ggplot(aes(x = date, y = prcp, color = prcp_reporting)) +
+        geom_line() + geom_point() +
+        geom_segment(data = ike_rain,
+                     aes(x = ymd(closest_date) - ddays(2),
+                         xend = ymd(closest_date) + ddays(1),
+                         y = tot_precip, yend = tot_precip), color = "red") +
+        facet_wrap(~ fips, ncol = 2) +
+        theme_few()
+# Note: different scales on this one-- cumulative rain over four days for
+# `ike_rain`, `ike_ave` is one day at a time
+
+ike_ave %>%
+        left_join(ike_rain, by = "fips") %>%
+        group_by(fips) %>%
+        filter(ymd(closest_date) - ddays(2) <= date &
+                       date <= ymd(closest_date) + ddays(1)) %>%
+        summarize(monitor_rain = sum(prcp),
+                  tot_precip = first(tot_precip),
+                  prcp_reporting = mean(prcp_reporting)) %>%
+        ggplot(aes(x = monitor_rain, y = tot_precip)) +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        geom_point(aes(size = prcp_reporting), alpha = 0.5) +
+        geom_text(aes(label = fips)) +
+        theme_few() +
[…]

### 2016-07-28 writing/DraftExposurePaper/example_ike_code.R
Make some tweaks to the Ike example file

@@ -1,8 +1,15 @@
+options("noaakey" = Sys.getenv("noaakey"))
+
+ike_fips <- c("12021", "12071")
 county_timeseries(ike_fips, percent_coverage = 0,
                   date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP")
+                  var = "PRCP", out_directory = "~/ike_ex/")
 
-ike_dir <- "~/tmp/ike_ex/"
+# Check that it worked...
+list.files("~/ike_ex")
+
+# Do some plots
+ike_dir <- "~/ike_ex/"
 ike_counties <- gsub(".rds", "", list.files(ike_dir))
 
 ike_ave <- vector("list", length = length(ike_counties))

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
some on wind and property damage

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
-While the risks to life from some tropical storm threats (e.g., storm surge) have been greatly reduced over the past century, other risks (e.g., inland flooding) have remained more steady [@Rappaport2000], and costs related to property damage have increased [citation].
+While the risks to life from some tropical storm threats (e.g., storm surge) have been greatly reduced over the past century, other risks (e.g., inland flooding) have remained more steady [@Rappaport2000], and costs related to property damage have increased [citation; Fronstin1994 suggested that there was more property damage for newer homes than older homes during hurricane Andrew due to changes in building codes].
 Determining the metrics of storm exposure that are most associated with loss of life and property damage may help identify and emphasize the important threats that remain from tropical storms in the United States, which could allow similar future success in reducing these threats through community planning, warning systems, etc.
@@
 
-Storm-motion, as well as the direction from which a storm approaches, can play significant roles in determining the extent of precipitation produced by a storm [@Rees2001; @Chang2013].
-(Examples include the south-to-north storm circulation for Hurricane Fran (1996) with an east-to-west movement, causing an amplification of rainband convection, which played a key role in runoff distribution [@Rees2001] and increased precipitation intensity for storms approaching Taiwan from the east [@Chang2013]).
+Storm-motion, as well as the direction from which a storm approaches, can play significant roles in determining the extent of precipitation produced by a storm [@Rees2001; @Chang2012].
+(Examples include the south-to-north storm circulation for Hurricane Fran (1996) with an east-to-west movement, causing an amplification of rainband convection, which played a key role in runoff distribution [@Rees2001] and increased precipitation intensity for storms approaching Taiwan from the east [@Chang2012]).
 
 The topography of a location (e.g., elevation change in a county) could also affect how extreme rainfall is from a particular storm, as well as whether rains from a tropical storm cause a flood, particularly a flash flood [@Rees2001 on orographic enhancement; @Czajkowski2011 mentions this, but is not a primary source for this point].
-For example, topographical characteristics in Taiwan play a key role in regulating rainfall intensity and its location [@Chang2013].
-@Chang2013 found that the most intense rainfall was found in areas with the greatest wind-terrain interaction.
-Storms embodying such characteristics are increasing in both frequency and duration in Taiwan [@Chang2013].[JMF--I am not sure how much of this detail is needed beyond discussing topographical impact, or if mentioning Taiwan distracts from our main focus].
+For example, topographical characteristics in Taiwan play a key role in regulating rainfall intensity and its location [@Chang2012].
+@Chang2012 found that the most intense rainfall was found in areas with the greatest wind-terrain interaction.
+Storms embodying such characteristics are increasing in both frequency and duration in Taiwan [@Chang2012].[JMF--I am not sure how much of this detail is needed beyond discussing topographical impact, or if mentioning Taiwan distracts from our main focus].
 More specifically, the Appalachian mountains provide topography that both enhance precipitation during tropical storms as well as provide ample hydrological conditions for severe flooding [@Rees2001].
@@
 Another study used a criterion of wind speed at or above 74 miles per hour [@Grabich2015].
+
+Some of the most severe regions of property damage during Hurricane Andrew in 1992 were located in regions other than those that experienced maximum sustained winds [@Fronstin1994].
 

### 2016-07-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd

--- 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```
+
+```{r}
+#Find the top 15 FIPS by number of reporting stations during Ike-2008 (Warning: Takes a long time)
+#my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
+#my_stations <- NA
+#ike_table <- expand.grid(my_fips, my_stations)
+#colnames(ike_table) <- c("fips", "station")
+#for(fips in my_fips){
+        #ike_weather <- weather_fips(fips = fips,
+                                    #date_min = "2008-09-01",
+                                    #date_max = "2008-09-14",
+                                    #var = "PRCP")
+        #ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+#} 
+#Ike FIPS without data: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
+#ike_table <- ike_table %>%
+        #arrange(desc(station)) %>%
+        #slice(1:15)
+#ike_top <- ike_table$fips
+#Find the top 15 FIPS by number of reporting stations for Alex-2010:
+#my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
+#my_stations <- NA
+#alex_table <- expand.grid(my_fips, my_stations)
+#colnames(alex_table) <- c("fips", "station")
+#for(fips in my_fips){
+#        alex_weather <- weather_fips(fips = fips,
+#                                    date_min = "2008-09-01",
+#                                    date_max = "2008-09-14",
+#                                    var = "PRCP")
+#        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
+#} 
+#Alex FIPS without data: 3, 7, 9, 11, 15, 19   
+#alex_table <- alex_table %>%
+#        arrange(desc(station)) %>%
+#        slice(1:15)
+#alex_top <- alex_table$fips
+#fips_table <- rbind(alex_table, ike_table)
+#fips_table <- fips_table %>%
+#        arrange(desc(station)) %>%
+#        slice(1:15)
+#all_fips <- fips_table$fips
+#all_fips <- unique(all_fips)
+```
+
+As of right now, I am not going to do Frances (there are 184 unique fips).
+
+Create a timeseries for the top 15 FIPS by reporting stations for Alex and Ike:
+
+```{r}
+data("closest_dist")
+data(rain)
+county_timeseries(fips = ike_top[1:2], 
+                  percent_coverage = 0, 
+                  date_min = "2008-09-01", 
+                  date_max = "2010-09-14", var = "PRCP", 
+                  out_directory = "~/tmp/ike_ex")
+ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
+keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
+keys_cd <- keys_cd$closest_date
+ike_1 <- ike_1 %>%
+        select(date, prcp) %>%
+        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), by = 1))
+ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
+ike_sum <- sum(ike_1$prcp)
+ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
+ike_nldas_sum <- ike_nldas_sum$tot_precip
+```
+
+[JMF-- It looks like precipitation measurements from `countyweather` are already in mm]
 

### 2016-07-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Some tweaks to get Rmd to compile

@@
 #compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
-b <- ggplot(compare_rain1, aes(x = tot_precip, y = rain_mm,
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
                               color = fips)) +
@@
 alex_fips <- unique(alex_fips)
+```
+```{r}
 #Ike-2008
@@
 
-```{r}
-data("closest_dist")
-data(rain)
-county_timeseries(fips = ike_top[1:2], 
+```{r eval = FALSE}
+
+library(countyweather)
+county_timeseries(fips = ike_fips[1:2], 
                   percent_coverage = 0, 
@@
                   date_max = "2010-09-14", var = "PRCP", 
-                  out_directory = "~/tmp/ike_ex")
-ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
+                  out_directory = "ike_ex")
+data("closest_dist")
+data(rain)
+ike_1 <- readRDS("ike_ex/12021.rds")
 keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")

### 2016-07-29 writing/DraftExposurePaper/example_ike_code.R
add .rds for ike fips and run ike_ex

@@ -9,12 +9,16 @@ library(lubridate)
 ike_fips <- c("12021", "12071", "12087", "22051", "22053")
 county_timeseries(ike_fips, percent_coverage = 0,
                   date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP", out_directory = "~/ike_ex/")
+                  var = "PRCP", out_directory = "~/ike_ex/"
+                  #, out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_ex/"
+                  )
 
 # Check that it worked...
 list.files("~/ike_ex")
+#list.files("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_ex/")
 
 # Do some plots
+#ike_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_ex/"
 ike_dir <- "~/ike_ex/"
 ike_counties <- gsub(".rds", "", list.files(ike_dir))
 
@@ -22,7 +26,7 @@ ike_ave <- vector("list", length = length(ike_counties))
 names(ike_ave) <- ike_counties
 
 for(x in ike_counties){
-        county_weather <- readRDS(paste0(ike_dir, x, ".rds"))$averaged
+        county_weather <- readRDS(paste0(ike_dir, x, ".rds"))#$averaged
         county_weather <- county_weather %>%
                 filter(year(date) == 2008 & month(date) == 9)
         county_weather$fips <- x

### 2016-07-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add some code

@@
 ```{r echo = FALSE, fig.height = 4, fig.width = 6}
+#allison_gauge <- read.csv("~/Documents/hurricaneexposure/writing/DraftExposurePaper/MWR_gauge_rain_data/allison_mwr.csv")
 allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv")

### 2016-07-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
unique fips for frances and allison

@@
 ```{r eval = FALSE}
-frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
-                         na.strings = 999,
-                         header = TRUE)
-save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
+#frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
+ #                        na.strings = 999,
+  #                       header = TRUE)
+#save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
 ```
@@
 ike_fips <- unique(ike_fips)
+```
+```{r}
+#Allison-2010
+allison_fips <- allison_nldas$fips
+allison_fips <- unique(allison_fips)
+```
+```{r}
+#Frances-2004
+frances_fips <- frances_nldas$fips
+frances_fips <- unique(frances_fips)
 ```

### 2016-07-29 writing/DraftExposurePaper/dade_storm_plot.R
pull dade timeseries

@@ -1,0 +1,5 @@
+fips <- "12086"
+county_timeseries(fips, percent_coverage = 0,
+                  date_min = "1988-01-01", date_max = "2011-12-31",
+                  var = "PRCP", #out_directory = "~/frances_ex/"
+                  out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")

### 2016-07-29 writing/DraftExposurePaper/dade_storm_plot.R
begin frances code

@@ -3,3 +3,6 @@ county_timeseries(fips, percent_coverage = 0,
                   date_min = "1988-01-01", date_max = "2011-12-31",
                   var = "PRCP", #out_directory = "~/frances_ex/"
                   out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
+frances_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
+county_weather <- readRDS(paste0(frances_dir, fips, ".rds"))
+county_weather <-

### 2016-08-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Edit so bullet points will format
move lables so an see dade plot more clealry

@@
 Good Characteristics:
+- Measureable at the county/sub-county level
+- Internally consistant (either within a study or for a particular storm)
++ Takes into account the unique characteristics of a particular storm and county's location in assessing exposure
 - Flooding-- Most associated with death; property damage
@@
 - Wind alone
+- Metrics that cannot be widely measured at the county or sub-county level (i.e., stress)
 

### 2016-08-01 writing/DraftExposurePaper/dade_storm_plot.R

@@ -98,7 +98,7 @@ county_weather %>%
         ggplot(aes(x = monitor_rain, y = tot_precip)) +
         geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
         geom_point(aes(size = prcp_reporting), alpha = 0.2) +
-        geom_text(aes(label = storm_id)) +
+        geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
         theme_few() +
         scale_size_continuous(guide = "none") +
         xlab("Rainfall (mm) based on \naveraged county monitors") +

### 2016-08-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
work on good/bad list

--- 2016-07-29 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 Good Characteristics:
+- Measureable at the county/sub-county level
+- Internally consistant (either within a study or for a particular storm)
++ Takes into account the unique characteristics of a particular storm and county's location in assessing exposure
 - Flooding-- Most associated with death; property damage
 - Precipitation
-- Wind (potentially weighted)
++ Gauge data if available
+- Wind (potentially weighted based on relation to the storm track), though wind alone has proven not to be sufficient in identifying the most affected region in some cases (e.g., Hurricane Andrew in 1992)[@Fronstin1994].
+Therefore wind alone is not sufficient in all cases (as is commonly done using the Staffir-Simpson scale).
 

### 2016-08-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add some to text and references

@@
 Therefore wind alone is not sufficient in all cases (as is commonly done using the Staffir-Simpson scale).
-
+- Have a metric with finer resolution than the community level (i.d., zip codes or neighborehoods)
++ @Fronstin1994 stated that neighborehoods just 8 blocks away from some of the most impacted areas during Hurricane Andrew (1992) recieved far less damage 
++ Not necessarily possible when guage data may not be available at such a resolution
 Bad Characteristics:
@@
 Hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used?
-We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath describing a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale).
+We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath based on a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale).
 These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016; @Fronstin1994].
@@
 As a result, using FEMA disaster declarations as exposure classification to hurricanes restricts exposure to 1) those states whose resources are overwhelmed and 2) those states who apply for FEMA aid.
-As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane; GBA-- let's think about why that might have had "false positives".
+(FEMA is likely less objective given societal effects (politics) [@Kunkel1999]) As a result, other severely impacted counties and states that may not fit FEMA criteria are excluded and may therefore be incorrectly identified as unexposed, preventing researchers from assessing the full impact of a given hurricane [JMF-- the Grabich 2016 study actually found that FEMA assigned the highest exposure, and may overestimate the health effects of a hurricane; GBA-- let's think about why that might have had "false positives".
 I think that would be an interesting point to include here.].
@@
 For example, property damage, crop damage, injuries, ...
-These are all given in NOAA's Storm Events database, but let's also see if we can find studies that have analyzed these outcomes in relation to tropical storms, if we can.]
+These are all given in NOAA's Storm Events database, but let's also see if we can find studies that have analyzed these outcomes in relation to tropical storms, if we can.] Review has shown the hurricanes are associated with major societal impacts.
+
+Though the loss of life due to hurricanes may have decreased since the mid-1900s, trends show a large increase in property damage [@Kunkel1999] in spite of a statistically significant decrease in both the frequency and intensity of tropical storms since then [@Landsea1996].
+Such a trend is likely attributable to increased population growth in coastal communities [@Pielke1997].
 
@@
 One study of US hurricanes between 1925 and 1995 found that property damage associated with a storm was strongly associated with the intensity of the storm, with most damage occuring during category 3, 4, or 5 storms, even though storms of these categories made up under a quarter of all hurricanes in the study [@Pielke1998].
-Wind-based hazards are also closely associated with power outages [citation], which can introduce a number of threats to human and ecological health and property, including water quality risks if the outage affects wastewater treatment plants [@Mallin2006].
+Wind-based hazards are also closely associated with power outages (both specific to hurricane related hazards; do not specify wind-based or otherwise) [@Liu2005; @Han2009], which can introduce a number of threats to human and ecological health and property, including water quality risks if the outage affects wastewater treatment plants [@Mallin2006].
 Direct fatalities from hurricanes tend to increase with increasing wind exposure from the storm [@Czajkowski2011].

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Edits to draft

--- 2016-08-01 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 Good Characteristics:
-- Measureable at the county/sub-county level
-- Internally consistant (either within a study or for a particular storm)
-+ Takes into account the unique characteristics of a particular storm and county's location in assessing exposure
+
 - Flooding-- Most associated with death; property damage
@@
 Bad Characteristics:
+
 - Distance alone
 - Wind alone
-- Metrics that cannot be widely measured at the county or sub-county level (i.e., stress)
+- Not consistent across storms.
+(For example, if a metric classified a county as "exposed" to a storm late in the dataset, if the same storm had happened earlier in the dataset, it also would have been assessed as "exposed".
+This might be the case for, for example, NOAA Storm Events.
+It might be more common to report an event well enough that it makes it into the database in later years compared to earlier years of the study.)
+- Not consistent across locations for a single storm.
+(For example, some of the metrics depend on subjective judgement of different people in different locations-- I think FEMA disaster declarations and NOAA Storm Events might both be examples in this.
+In this case, you couldn't rely on the assumption that, if two counties have the same level of exposure to a storm, they would have the same status on that criterion.)
 
@@
 
-Our analysis included [x] Atlantic basin tropical storms over the years 1988 to 2011.
-This dataset includes several noteworthy storms, including Hugo (1989), Andrew (1992), Floyd (1999), Katrina (2005), ...
-.
-
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-# Map of tracks of all storms in study.
+# Vector of all storms in study.
 data("closest_dist")
@@
 storms <- storms[which(gsub("*.+-", "", storms) <= 2011)]
+```
+
+Our analysis included `r length(storms)` Atlantic basin tropical storms over the years 1988 to 2011.
+This dataset includes several noteworthy storms, including Hugo (1989), Andrew (1992), Floyd (1999), Katrina (2005), [maybe add Ike, Ivan, others?] ...
+.
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+# Map of tracks of all storms in study.
 map_tracks(storms, plot_points = FALSE, alpha = 0.3)
@@
 
-Justification of NLDAS data: comparison of NLDAS and gauge data for hurricane Alex (2010).
+Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.
+Although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm.
+
+```{r echo = FALSE, fig.width = 6, fig.height = 4}
+ggplot(rain_lag, aes(x = lag, y = precip, group = lag)) + 
+        geom_boxplot() + theme_tufte() + 
+        xlab("Days from storm's closest approach to county") + 
+        ylab("Rainfall in county (mm)") 
+```
+
+We compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season.
+Here are some of the plots comparing county-level estimates of rain exposure from the NLDAS-2 data to all *MWR*-reported rainfall estimates within the county:
 
 ```{r echo = FALSE, fig.height = 4, fig.width = 6}
-#library(data.table)
-#alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
-#save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
-# load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
 load("alex_gauge.Rdata")
@@
 compare_rain <- right_join(alex_nldas, alex_gauge, by = "fips")
-#alex_nldas <- as.data.table(alex_nldas)
-#alex_gauge <- as.data.table(alex_gauge)
-#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
 a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
@@
         geom_point() + 
-        theme_bw()
+        theme_bw() + 
+        ggtitle("Alex, 2010")
 a
+
+county_range <- compare_rain %>% 
+        group_by(fips) %>%
+        summarize(lowest_rain = min(rain_mm),
+                  highest_rain = max(rain_mm),
+                  nldas_rain = mean(tot_precip))
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
+        geom_segment(data = county_range, 
+                     aes(x = nldas_rain, xend = nldas_rain,
+                         y = lowest_rain, yend = highest_rain,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
+        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
+        ggtitle("Alex, 2010") + 
+        xlab("NLDAS-2 rainfall (mm)") + 
+        ylab("MWR rainfall reports (mm)")
+b
 ```
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)")  + 
+        ggtitle("Alex, 2010") + 
         scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
@@
 ```{r echo = FALSE, fig.height = 4, fig.width = 6}
-#allison_gauge <- read.csv("~/Documents/hurricaneexposure/writing/DraftExposurePaper/MWR_gauge_rain_data/allison_mwr.csv")
 allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv")
@@
 compare_rain <- right_join(allison_nldas, allison_gauge, by = "fips")
-#alex_nldas <- as.data.table(alex_nldas)
-#alex_gauge <- as.data.table(alex_gauge)
-#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
+
 a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
@@
         geom_point() + 
-        theme_bw()
+        theme_bw() + 
+        ggtitle("Allison, 2011")
 a
+county_range <- compare_rain %>% 
+        group_by(fips) %>%
+        summarize(lowest_rain = min(rain_mm),
+                  highest_rain = max(rain_mm),
+                  nldas_rain = mean(tot_precip))
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
+        geom_segment(data = county_range, 
+                     aes(x = nldas_rain, xend = nldas_rain,
+                         y = lowest_rain, yend = highest_rain,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
+        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
+        ggtitle("Allison, 2001") + 
+        xlab("NLDAS-2 rainfall (mm)") + 
+        ylab("MWR rainfall reports (mm)")
+b
 ```
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        ggtitle("Allison, 2001") + 
         scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
@@
 library(weathermetrics)
-#library(data.table)
-#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
-#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
-#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
 load("ike_gauge.RData")
@@
                                     "inches", "mm")
+ike_gauge$fips <- sprintf("%05s", ike_gauge$fips)
 ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
@@
 compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
-#ike_nldas <- as.data.table(alex_nldas)
-#ike_gauge <- as.data.table(alex_gauge)
-#compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
+a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
                               color = fips)) +
         geom_point() + 
-        theme_bw()
+        theme_bw() + 
+        ggtitle("Ike, 2008")
+a
+
+county_range <- compare_rain %>% 
+        group_by(fips) %>%
+        summarize(lowest_rain = min(rain_mm),
+                  highest_rain = max(rain_mm),
+                  nldas_rain = mean(tot_precip))
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
+        geom_segment(data = county_range, 
+                     aes(x = nldas_rain, xend = nldas_rain,
+                         y = lowest_rain, yend = highest_rain,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
+        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
+        ggtitle("Ike, 2008") + 
+        xlab("NLDAS-2 rainfall (mm)") + 
+        ylab("MWR rainfall reports (mm)")
 b
@@
 
-[JMF-- Missing values are those from Arkansas (missing the leading 0 in fips code)]
+[JMF-- Missing values are those from Arkansas (missing the leading 0 in fips code); GBA: I fixed this.
+This is ultimately something I need to fix in the `hurricaneexposure` package, but I've got this example fixed to run without losing the FIPS with leading zeros.]
 
@@
         ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
+        ggtitle("Allison, 2001") + 
         scale_color_manual(values = c("blue", "black"), guide = "none") 
@@
 ```{r}
-#frances_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_gauge.csv", header = TRUE)
-#save(frances_gauge, file = "writing/DraftExposurePaper/frances_gauge.RData")
-#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_gauge.RData")
 load("frances_gauge.RData")
-frances_gauge$precip_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
+frances_gauge$rain_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
+frances_gauge$fips <- sprintf("%05s", frances_gauge$fips)
 frances_nldas <- county_rain(counties = unique(frances_gauge$fips),
@@
 compare_rain <- right_join(frances_nldas, frances_gauge, by = "fips")
-b <- ggplot(compare_rain, aes(x = tot_precip, y = precip_mm,
+a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
                               color = fips)) +
         geom_point() + 
-        theme_bw()
+        theme_bw() + 
+        ggtitle("Frances, 2004")
+a
+
+county_range <- compare_rain %>% 
+        group_by(fips) %>%
+        summarize(lowest_rain = min(rain_mm),
+                  highest_rain = max(rain_mm),
+                  nldas_rain = mean(tot_precip))
+b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
+        geom_segment(data = county_range, 
+                     aes(x = nldas_rain, xend = nldas_rain,
+                         y = lowest_rain, yend = highest_rain,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
+        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
+        ggtitle("Frances, 2004") + 
+        xlab("NLDAS-2 rainfall (mm)") + 
+        ylab("MWR rainfall reports (mm)")
 b
@@
 #Ike-2008 wind
-#ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
-#save(ike_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
-# load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
 load("ike_wind.RData")
+ike_wind <- ike_wind %>%
+        filter(!is.na(max_sust_kt)) %>%
+        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
+                                                 old_metric = "knots",
+                                                 new_metric = "mps"),
+               fips = sprintf("%05s", fips))
+
+ike_wind_model <- county_wind(counties = unique(ike_wind$fips),
+                              start_year = 2008, end_year = 2008,
+                              wind_limit = 0) %>%
+        filter(storm_id == "Ike-2008")
+
+compare_wind <- right_join(ike_wind_model, ike_wind, by = "fips")
+a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw() + 
+        ggtitle("Ike, 2008")
+a
+
+county_range <- compare_wind %>% 
+        group_by(fips) %>%
+        summarize(lowest_wind = min(max_sust_mps),
+                  highest_wind = max(max_sust_mps),
+                  model_wind = mean(max_sust))
+b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
+        geom_segment(data = county_range, 
+                     aes(x = model_wind, xend = model_wind,
+                         y = lowest_wind, yend = highest_wind,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
+        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
+        ggtitle("Ike, 2008") + 
+        xlab("Modeled wind speed (m / s)") + 
+        ylab("MWR wind reports (m / s)")
+b
 ```
@@
 ```{r eval = FALSE}
-#frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
- #                        na.strings = 999,
-  #                       header = TRUE)
-#save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
+frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
+                         na.strings = 999,
+                         header = TRUE)
+save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
 ```
@@
 load("frances_wind.RData")
+frances_wind <- frances_wind %>%
+        filter(!is.na(max_sust_kt)) %>%
+        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
+                                                 old_metric = "knots",
+                                                 new_metric = "mps"),
+               fips = sprintf("%05s", fips))
+
+frances_wind_model <- county_wind(counties = unique(frances_wind$fips),
+                              start_year = 2004, end_year = 2004,
+                              wind_limit = 0) %>%
+        filter(storm_id == "Frances-2004")
+
+compare_wind <- right_join(frances_wind_model, frances_wind, by = "fips")
+a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw() + 
+        ggtitle("Frances, 2004")
+a
+
+county_range <- compare_wind %>% 
+        group_by(fips) %>%
+        summarize(lowest_wind = min(max_sust_mps),
+                  highest_wind = max(max_sust_mps),
+                  model_wind = mean(max_sust))
+b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
+        geom_segment(data = county_range, 
+                     aes(x = model_wind, xend = model_wind,
+                         y = lowest_wind, yend = highest_wind,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
+        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
+        ggtitle("Frances, 2004") + 
+        xlab("Modeled wind speed (m / s)") + 
+        ylab("MWR wind reports (m / s)")
+b
 ```
@@
 ike_fips <- unique(ike_fips)
-```
-```{r}
-#Allison-2010
-allison_fips <- allison_nldas$fips
-allison_fips <- unique(allison_fips)
-```
-```{r}
-#Frances-2004
-frances_fips <- frances_nldas$fips
-frances_fips <- unique(frances_fips)
 ```
@@
 [JMF-- It looks like precipitation measurements from `countyweather` are already in mm]
-
-Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.
-Although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm.
-
-```{r echo = FALSE, fig.width = 6, fig.height = 4}
-ggplot(rain_lag, aes(x = lag, y = precip, group = lag)) + 
-        geom_boxplot() + theme_tufte() + 
-        xlab("Days from storm's closest approach to county") + 
-        ylab("Rainfall in county (mm)") 
-```
 

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd

@@
 
+- Measureable at the county/sub-county level
+- Internally consistant (either within a study or for a particular storm)
++ Takes into account the unique characteristics of a particular storm and county's location in assessing exposure
 - Flooding-- Most associated with death; property damage
 - Precipitation
-- Wind (potentially weighted)
++ Gauge data if available
+- Wind (potentially weighted based on relation to the storm track), though wind alone has proven not to be sufficient in identifying the most affected region in some cases (e.g., Hurricane Andrew in 1992)[@Fronstin1994].
+Therefore wind alone is not sufficient in all cases (as is commonly done using the Staffir-Simpson scale).
+- Have a metric with finer resolution than the community level (i.d., zip codes or neighborehoods)
++ @Fronstin1994 stated that neighborehoods just 8 blocks away from some of the most impacted areas during Hurricane Andrew (1992) recieved far less damage 
++ Not necessarily possible when guage data may not be available at such a resolution
 
@@
 In this case, you couldn't rely on the assumption that, if two counties have the same level of exposure to a storm, they would have the same status on that criterion.)
+- Metrics that cannot be widely measured at the county or sub-county level (i.e., stress)
 
 Hurricane exposure criteria at the county-level include FEMA disaster declarations [GBA-- are these that commonly used?
-We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath describing a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale).
+We should certainly put them in as one way to measure exposure, but I don't know if we want to include them, with distance, as "most often used" unless we can find quite a few examples of them being used to determine exposure], whether a county's perimeter intercepts a swath based on a buffer a certain distance from the storm's central track, or exposure metrics based primarily on sustained wind speeds (i.e., the Staffir-Simpson scale).
 These ways of assessing exposure can be nonspecific and potentially lead to misclassification of exposure at the county-level [@Grabich2016; @Fronstin1994].

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Get merged version running.

@@
 compare_rain %>% group_by(fips) %>%
-        summarize(ave_rain_mm = mean(precip_mm),
+        summarize(ave_rain_mm = mean(rain_mm),
                   ave_tot_precip = mean(tot_precip),
@@
 frances_wind <- frances_wind %>%
-        filter(!is.na(max_sust_kt)) %>%
-        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
+        filter(!is.na(wind_sust)) %>%
+        mutate(max_sust_mps = convert_wind_speed(wind_sust, 
                                                  old_metric = "knots",

### 2016-08-02 writing/DraftExposurePaper/dade_storm_plot.R
Make some tweaks to the Miami-Dade plot

@@ -37,8 +37,8 @@ all_fips <- c(all_fips, as.integer(12025))
 check_dates[check_dates$fips == 12086, "fips"] <- 12025
 
 dade_rain <- data.table::fread(
-        "/Users/joshuaferreri/Documents/nasa_precip_export_2.txt",
-        #"data-raw/nasa_precip_export_2.txt",
+        #"/Users/joshuaferreri/Documents/nasa_precip_export_2.txt",
+        "~/Documents/CSU2016/hurricaneproject/hurricaneexposuredata/data-raw/nasa_precip_export_2.txt",
         header = TRUE,
         select = c("county", "year_month_day", "precip", "precip_max")) %>%
         dplyr::filter(county %in% all_fips,
@@ -66,11 +66,13 @@ dade_rain <- dade_rain %>%
 fips <- "12086"
 county_timeseries(fips, percent_coverage = 0,
                   date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP", #out_directory = "~/frances_ex/"
-                  out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
+                  var = "PRCP",
+                  out_directory = "~/dade_data/")
+                  #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
 
-frances_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
-county_weather <- readRDS(paste0(frances_dir, fips, ".rds"))
+# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
+dade_dir <- "dade_data/"
+county_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
 
 county_weather <- county_weather %>%
         filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
@@ -96,9 +98,13 @@ county_weather %>%
                   tot_precip = sum(precip),
                   prcp_reporting = mean(prcp_reporting)) %>%
         ggplot(aes(x = monitor_rain, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
         geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
[…]

### 2016-08-02 writing/DraftExposurePaper/dade_storm_plot.R
code to merge new data with closest_dist

@@ -110,6 +110,57 @@ county_weather %>%
         xlab("Rainfall (mm) based on \naveraged county monitors") +
         ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
         ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade county by Storm")
+
+lag_sum <- function(counties = NULL, county_weather, df2, year_range = NULL,
+                              distance_limit = NULL, rain_limit = NULL,
+                              include_rain = FALSE, days_included = NULL,
+                              output_vars = c("fips")){
+
+        county_weather <- as.data.frame(county_weather)
+        df <- inner_join(county_weather, df2, by = "date") %>%
+                arrange(storm_id)
+        df <- rename(df, fips = fips.x)
+
+        closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
+
+        if(!is.null(counties)){
+                closest_dist <- closest_dist[get("fips") %in% counties]
+        }
+
+        if(!is.null(year_range)){
+                closest_dist <- closest_dist[ , .(storm_id,
+                                                  fips,
+                                                  closest_date, storm_dist,
+                                                  local_time,
+                                                  closest_time_utc,
+                                                  year = substring(get("closest_date"),
+                                                                   1, 4)), ][
+                                                                           get("year") %in%
+                                                                                   year_range[1]:year_range[2]
+                                                                           ]
+        }
+
+        if(!is.null(distance_limit)){
+                closest_dist <- closest_dist[get("storm_dist") <=
+                                                     distance_limit]
+        }
+
+        if(include_rain){
+                df <- df[get("lag") %in% days_included]
+                df <- df[ , .(tot_precip = sum(get("prcp"))),
+                              by = .(fips, storm_id)]
+                closest_dist <- merge(closest_dist, df, all.x = TRUE,
+                                      by = c("storm_id", "fips"))
+                if(!is.null(rain_limit)){
+                        closest_dist <- closest_dist[get("tot_precip") >=
+                                                             rain_limit]
+                }
+        }
+
+        closest_dist <- closest_dist[ , .SD, , .SDcols = output_vars]
+        return(closest_dist)
+}
+
 check <- county_weather %>%
         inner_join(dade_rain, "date") %>%
         filter(lag == -2 | lag == -1 | lag == 0 | lag == 1) %>%

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Move all library loading to first code chunk

@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
+library(hurricaneexposuredata)
 library(hurricaneexposure)
-# install hurricane exposure data package if you don't already have it
-# library(devtools)
-# install_github("geanders/hurricaneexposuredata")
-library(hurricaneexposuredata)
 library(knitr)
@@
 library(RColorBrewer)
+library(weathermetrics)
+
+#devtools::install_github("ropenscilabs/rnoaa")
+library(rnoaa)
+#devtools::install_github("leighseverson/countyweather")
+library(countyweather)
 ```
@@
 ```{r echo = FALSE, fig.height = 4, fig.width = 6}
-library(weathermetrics)
+
 load("ike_gauge.RData")
@@
 ```{r eval = FALSE}
-#devtools::install_github("ropenscilabs/rnoaa")
-library(rnoaa)
-#devtools::install_github("leighseverson/countyweather")
-library(countyweather)
 options("noaakey" = Sys.getenv("noaakey"))
@@
 ```{r eval = FALSE}
-
-library(countyweather)
 county_timeseries(fips = ike_fips[1:2], 

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Load data from packages in first code chunk

@@
 library(countyweather)
+
+data(closest_dist)
+data(rain)
 ```
@@
 # Vector of all storms in study.
-data("closest_dist")
+
 storms <- unique(closest_dist$storm_id)
@@
 ```{r warning = FALSE, message = FALSE, echo = FALSE}
-data(closest_dist)
-
 # Generate a list of unique study fips (all in the eastern US)
@@
         mutate(storm_county = paste0(storm_id, "_", fips))
-data(rain)
+
 rain_lag <- rain %>%
@@
                   out_directory = "ike_ex")
-data("closest_dist")
-data(rain)
 ike_1 <- readRDS("ike_ex/12021.rds")
@@
 ```{r echo = FALSE}
-data(closest_dist)
-
 # Generate a list of unique study fips (all in the eastern US)
@@
 ```{r echo = FALSE, warning=FALSE, message=FALSE}
-data(closest_dist)
-
 # Generate a list of unique study fips (all in the eastern US)
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-data("closest_dist")
 all_counties <- unique(closest_dist$fips)

### 2016-08-02 writing/DraftExposurePaper/dade_storm_plot.R
work on fxns to merge countyweather data

@@ -114,7 +114,7 @@ county_weather %>%
 lag_sum <- function(counties = NULL, county_weather, df2, year_range = NULL,
                               distance_limit = NULL, rain_limit = NULL,
                               include_rain = FALSE, days_included = NULL,
-                              output_vars = c("fips")){
+                              output_vars = c("fips", "cw_precip")){
 
         county_weather <- as.data.frame(county_weather)
         df <- inner_join(county_weather, df2, by = "date") %>%
@@ -146,19 +146,41 @@ lag_sum <- function(counties = NULL, county_weather, df2, year_range = NULL,
         }
 
         if(include_rain){
+                df <- data.table::data.table(df)
                 df <- df[get("lag") %in% days_included]
-                df <- df[ , .(tot_precip = sum(get("prcp"))),
+                df <- df[ , .(cw_precip = sum(get("prcp"))),
                               by = .(fips, storm_id)]
                 closest_dist <- merge(closest_dist, df, all.x = TRUE,
                                       by = c("storm_id", "fips"))
                 if(!is.null(rain_limit)){
-                        closest_dist <- closest_dist[get("tot_precip") >=
+                        closest_dist <- closest_dist[get("cw_precip") >=
                                                              rain_limit]
                 }
         }
 
         closest_dist <- closest_dist[ , .SD, , .SDcols = output_vars]
         return(closest_dist)
+}
+
+countyweather_rain <- function(counties, county_weather, df2, start_year, end_year,
+                        rain_limit, dist_limit,
+                        days_included = c(-2, -1, 0, 1)){
+
+        rain_storm_df <- lag_sum(counties = counties,
+                                           county_weather = county_weather,
+                                           df2 = df2,
+                                           year_range = c(start_year, end_year),
+                                           distance_limit = dist_limit,
+                                           rain_limit = rain_limit,
+                                           include_rain = TRUE,
+                                           days_included = days_included,
+                                           output_vars = c("storm_id", "fips",
+                                                           "closest_date",
+                                                           "storm_dist",
+                                                           "cw_precip",
+                                                           "local_time",
+                                                           "closest_time_utc"))
+        return(rain_storm_df)
 }
 
 check <- county_weather %>%

### 2016-08-02 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Move creation of all storms and all FIPS to set-up

@@
 data(rain)
+
+# Generate a list of unique study fips (all in the eastern US)
+my_fips <- unique(closest_dist$fips)
+
+# Vector of all storms in study.
+storms <- unique(closest_dist$storm_id)
+storms <- storms[gsub("*.+-", "", storms) <= 2011 & 
+                        gsub("*.+-", "", storms) >= 1988]
 ```
@@
 
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-# Vector of all storms in study.
-
-storms <- unique(closest_dist$storm_id)
-storms <- storms[which(gsub("*.+-", "", storms) <= 2011)]
-```
-
 Our analysis included `r length(storms)` Atlantic basin tropical storms over the years 1988 to 2011.
-This dataset includes several noteworthy storms, including Hugo (1989), Andrew (1992), Floyd (1999), Katrina (2005), [maybe add Ike, Ivan, others?] ...
-.
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
+This dataset covers several noteworthy storms, including Hugo (1989), Bob (1991), Andrew (1992), Opal (1995), Fran (1996), Georges (1998), Floyd (1999), Allison (2001), Isabel (2003), Charley (2004), Frances (2004), Ivan (2004), Jeanne (2004), Dennis (2005), Katrina (2005), Rita (2005), Gustav (2008), Ike (2008), and Irene (2011), all of which were severe enough their names were retired.
+
+```{r echo = FALSE, fig.align = 'center', fig.cap = "Tracks of tropical cyclones included in this study. Darker tracks indicate storms whose names were retired."}
 # Map of tracks of all storms in study.
-map_tracks(storms, plot_points = FALSE, alpha = 0.3)
+a <- map_tracks(storms, plot_points = FALSE, alpha = 0.3, color = "darkcyan")
+# Add storms whose names were retired
+# Source: http://www.nhc.noaa.gov/aboutnames_history.shtml
+map_tracks(c("Hugo-1989", "Bob-1991", "Andrew-1992", 
+             "Opal-1995", "Fran-1996", "Georges-1998", "Mitch-1998",
+             "Floyd-1999", "Allison-2001", "Michelle-2001",
+             "Isidore-2002", "Lili-2002", "Isabel-2003",
+             "Charley-2004", "Frances-2004",
+             "Ivan-2004", "Jeanne-2004", 
+             "Dennis-2005", "Katrina-2005", "Rita-2005",
+             "Wilma-2005", "Noel-2007",
+             "Gustav-2008", "Ike-2008", "Paloma-2008",
+             "Irene-2011"),
+           plot_object = a, plot_points = FALSE, color = "darkcyan") + 
+        theme(panel.grid.major = element_blank(),
+           panel.grid.minor = element_blank(),
+           panel.background = element_rect(colour = "black", size = 1))
 ```
@@
 
-[Spline vs. linear interpolation? Is 15 minutes adequate? Should we use shapefiles instead? IBTrACS probably has this (`rnoaa`).]
+[Is 15 minutes adequate? We could test if we get different results using a finer time resolution.]
 
@@
 
-To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").
+```{r warning = FALSE, message = FALSE, echo = FALSE}
+storm_lag <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 150, dist_limit = 500,
+                             days_include = -5:3)
+storm_lag <- storm_lag %>%
+        mutate(storm_county = paste0(storm_id, "_", fips))
+
+rain_lag <- rain %>%
+        mutate(storm_county = paste0(storm_id, "_", fips)) %>%
+        filter(storm_county %in% storm_lag$storm_county) %>%
+        group_by(storm_county) %>%
+        mutate(cumulative_rain = cumsum(precip),
+               difference_rain = lead(cumulative_rain) - cumulative_rain,
+               next_lag = lead(lag),
+               next_cumulative_rain = lead(cumulative_rain),
+               most_rain = difference_rain == max(difference_rain,
+                                                  na.rm = TRUE),
+               over_50 = lead(precip) > 50)
+to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1), y = c(0, 500, 500, 0))
+
+rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = cumulative_rain)) +
+        theme_tufte() + 
+        geom_segment(aes(xend = next_lag, yend = next_cumulative_rain,
+                         color = over_50),
+                     alpha = 0.1, size = 0.3) +
+        geom_polygon(data = to_highlight, aes(x = x, y = y),
+                     alpha = 0.2, color = "red", fill = NA) +
+        xlab("Days from storm's closest approach to county") + 
+        ylab("Cumulative rainfall in \ncounty from storm (mm)") + 
+        scale_color_manual(values = c("darkgray", "blue"),
+                           guide = FALSE) 
+rain_lag_plot
+```
+
+Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.
+Although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm.
+
+```{r echo = FALSE, fig.width = 6, fig.height = 4}
+ggplot(rain_lag, aes(x = lag, y = precip, group = lag)) + 
+        geom_boxplot() + theme_tufte() + 
+        xlab("Days from storm's closest approach to county") + 
+        ylab("Rainfall in county (mm)") 
+```
+
+To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we also investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").
 
@@
 
-```{r warning = FALSE, message = FALSE, echo = FALSE}
-# Generate a list of unique study fips (all in the eastern US)
-my_fips <- closest_dist$fips
-my_fips <- unique(my_fips)
-storm_lag <- county_rain(counties = my_fips,
-                             start_year = 1988, end_year = 2011,
-                             rain_limit = 150, dist_limit = 500,
-                             days_include = -5:3)
-storm_lag <- storm_lag %>%
-        mutate(storm_county = paste0(storm_id, "_", fips))
-
-rain_lag <- rain %>%
-        mutate(storm_county = paste0(storm_id, "_", fips)) %>%
-        filter(storm_county %in% storm_lag$storm_county) %>%
-        group_by(storm_county) %>%
-        mutate(cumulative_rain = cumsum(precip),
-               difference_rain = lead(cumulative_rain) - cumulative_rain,
-               next_lag = lead(lag),
-               next_cumulative_rain = lead(cumulative_rain),
-               most_rain = difference_rain == max(difference_rain,
-                                                  na.rm = TRUE),
-               over_50 = lead(precip) > 50)
-to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1), y = c(0, 500, 500, 0))
-
-rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = cumulative_rain)) +
-        theme_tufte() + 
-        geom_segment(aes(xend = next_lag, yend = next_cumulative_rain,
-                         color = over_50),
-                     alpha = 0.1, size = 0.3) +
-        geom_polygon(data = to_highlight, aes(x = x, y = y),
-                     alpha = 0.2, color = "red", fill = NA) +
-        xlab("Days from storm's closest approach to county") + 
-        ylab("Cumulative rainfall in \ncounty from storm (mm)") + 
-        scale_color_manual(values = c("darkgray", "blue"),
-                           guide = FALSE) 
-rain_lag_plot
-```
-
-Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.
-Although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm.
-
-```{r echo = FALSE, fig.width = 6, fig.height = 4}
-ggplot(rain_lag, aes(x = lag, y = precip, group = lag)) + 
-        geom_boxplot() + theme_tufte() + 
-        xlab("Days from storm's closest approach to county") + 
-        ylab("Rainfall in county (mm)") 
-```
-
 We compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season.
@@
 ```{r echo = FALSE}
-# Generate a list of unique study fips (all in the eastern US)
-my_fips <- closest_dist$fips
-my_fips <- unique(my_fips)
-
 all_storm_counties <- closest_dist %>%
@@
 ```{r echo = FALSE, warning=FALSE, message=FALSE}
-# Generate a list of unique study fips (all in the eastern US)
-my_fips <- closest_dist$fips
-my_fips <- unique(my_fips)
-
 # Get distance-based exposure for all study counties, all years, using a 
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_counties <- unique(closest_dist$fips)
-
-all_distance <- county_distance(counties = all_counties, start_year = 1988, 
+all_distance <- county_distance(counties = my_fips, start_year = 1988, 
                         end_year = 2011, dist_limit =100)
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_counties <- unique(closest_dist$fips)
-
-all_rain <- county_rain(counties = all_counties, start_year = 1988, 
+all_rain <- county_rain(counties = my_fips, start_year = 1988, 
                         end_year = 2011, rain_limit = 75, dist_limit = 500,
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_counties <- unique(closest_dist$fips)
-
-all_wind <- county_wind(counties = all_counties, start_year = 1988, 
+all_wind <- county_wind(counties = my_fips, start_year = 1988, 
                         end_year = 2011, wind_limit = 18)
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_counties <- unique(closest_dist$fips)
-
-all_wind <- county_wind(counties = all_counties, start_year = 1988, 
+all_wind <- county_wind(counties = my_fips, start_year = 1988, 
                         end_year = 2011, wind_limit = 33)
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_counties <- unique(closest_dist$fips)
-
-all_wind <- county_wind(counties = all_counties, start_year = 1988, 
+all_wind <- county_wind(counties = my_fips, start_year = 1988, 
                         end_year = 2011, wind_limit = 50)

### 2016-08-02 writing/DraftExposurePaper/dade_storm_plot.R
create functions to create new dataset with countyweather and nldas totals

@@ -166,7 +166,7 @@ countyweather_rain <- function(counties, county_weather, df2, start_year, end_ye
                         rain_limit, dist_limit,
                         days_included = c(-2, -1, 0, 1)){
 
-        rain_storm_df <- lag_sum(counties = counties,
+        county_rain_storm_df <- lag_sum(counties = counties,
                                            county_weather = county_weather,
                                            df2 = df2,
                                            year_range = c(start_year, end_year),
@@ -180,12 +180,22 @@ countyweather_rain <- function(counties, county_weather, df2, start_year, end_ye
                                                            "cw_precip",
                                                            "local_time",
                                                            "closest_time_utc"))
-        return(rain_storm_df)
+        rain_storm_df <- filter_storm_data(counties = counties,
+                                           year_range = c(start_year, end_year),
+                                           distance_limit = dist_limit,
+                                           rain_limit = rain_limit,
+                                           include_rain = TRUE,
+                                           days_included = days_included,
+                                           output_vars = c("storm_id", "fips",
+                                                           "closest_date",
+                                                           "storm_dist",
+                                                           "tot_precip",
+                                                           "local_time",
+                                                           "closest_time_utc"))
+        rain_df <- merge(county_rain_storm_df, rain_storm_df, all.x = TRUE,
+                         by = c("fips", "storm_id", "closest_date", "storm_dist",
+                                "local_time", "closest_time_utc"))
+        return(rain_df)
 }
 
-check <- county_weather %>%
-        inner_join(dade_rain, "date") %>%
-        filter(lag == -2 | lag == -1 | lag == 0 | lag == 1) %>%
-        arrange(storm_id) %>%
-        #mutate(monitor_rain = sum(prcp))
-        summarize(monitor_rain = sum(prcp))
+

### 2016-08-02 writing/DraftExposurePaper/dade_storm_plot.R
create miami comparison plot using new functions

@@ -198,4 +198,34 @@ countyweather_rain <- function(counties, county_weather, df2, start_year, end_ye
         return(rain_df)
 }
 
+miami_rain <- countyweather_rain(counties = "12086", county_weather = county_weather,
+                                 df2 = dade_rain, start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 1000)
+miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade county by Storm")
 
+
+        ggplot(aes(x = monitor_rain, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade county by Storm")

### 2016-08-02 writing/DraftExposurePaper/dade_storm_plot.R
edit miami comparison plot

@@ -72,12 +72,12 @@ county_timeseries(fips, percent_coverage = 0,
 
 # miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
 dade_dir <- "dade_data/"
-county_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
+dade_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
 
-county_weather <- county_weather %>%
+dade_weather <- dade_weather %>%
         filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
 
-county_weather$fips <- fips
+dade_weather$fips <- fips
 #dade_rain <- county_rain(counties = fips, start_year = 1988, end_year = 2011,
  #                        rain_limit = 0, dist_limit = 1000) %>%
   #      filter(fips == "12086")
@@ -87,7 +87,7 @@ county_weather$fips <- fips
 
 #Join `dade_rain` and `county_weather` datasets by date and create a plot comparing rain by storm (not fips)
 
-county_weather %>%
+dade_weather %>%
         inner_join(dade_rain, "date") %>%
         group_by(storm_id) %>%
         filter(lag == -3 | lag == -2 | lag == -1 |lag == 0 | lag == 1) %>%
@@ -198,7 +198,7 @@ countyweather_rain <- function(counties, county_weather, df2, start_year, end_ye
         return(rain_df)
 }
 
-miami_rain <- countyweather_rain(counties = "12086", county_weather = county_weather,
+miami_rain <- countyweather_rain(counties = "12086", county_weather = dade_weather,
                                  df2 = dade_rain, start_year = 1988, end_year = 2011,
                                  rain_limit = 0, dist_limit = 1000)
 miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work more on draft

@@
 
-Our analysis included `r length(storms)` Atlantic basin tropical storms over the years 1988 to 2011.
+Our analysis included `r length(storms)` Atlantic basin tropical storms over the years 1988 to 2011 (Figure \ref{fig:all_storms}).
 This dataset covers several noteworthy storms, including Hugo (1989), Bob (1991), Andrew (1992), Opal (1995), Fran (1996), Georges (1998), Floyd (1999), Allison (2001), Isabel (2003), Charley (2004), Frances (2004), Ivan (2004), Jeanne (2004), Dennis (2005), Katrina (2005), Rita (2005), Gustav (2008), Ike (2008), and Irene (2011), all of which were severe enough their names were retired.
 
-```{r echo = FALSE, fig.align = 'center', fig.cap = "Tracks of tropical cyclones included in this study. Darker tracks indicate storms whose names were retired."}
+```{r all_storms, echo = FALSE, fig.align = 'center', fig.cap = "Tracks of tropical cyclones included in this study. Darker tracks indicate storms whose names were retired.", message = FALSE}
 # Map of tracks of all storms in study.
@@
 
-For rainfall, we are using the cumulative rainfall from two days before to one day after the storm is closest to the county.
-This is based on analysis of the precipitation data for storms in our study period.
+For rainfall, we calculate total precipitation within a county during a storm by calculating the cumulative rainfall from two days before to one day after the storm's closest approach to the county.
 If we go larger, there is a danger of picking up rainfall that isn't related to the storm.
-However, some storms move slowly, so if we limit to just the day the storm is closest to the county, we would miss some of the associated rainfall.
-
-```{r warning = FALSE, message = FALSE, echo = FALSE}
+However, some storms move slowly or are large, so if we limit to just the day the storm is closest to the county, we are likely to miss some of storm-associated precipitation.
+
+We explored whether this window was adequate for capturing storm-related precipitation for most storms in most counties.
+We looked at all storm-county observations for which a county received 150 millimeters or more of precipitation within the window of five days before to three days after the storm's closest approach to a county.
+
+```{r cumulative_rain, echo = FALSE, fig.align = "center", fig.cap = "Cumulative rainfall within a county for a storm for which the county had 150 mm or more of rain over the period from five days before to three days after the storm's closest approach to the county.
+The line for any day for which a county received 50 mm or more of precipitation are shown in blue, while the line for any day for which a county received less than 50 mm of precipitation are shown in gray.
+The red box highlights the days used throughout the rest of the paper to measure total storm precipitation for each county.
+Numbers at the top of the graph give the total number of county-storm observations on that lag for which total precipitation was 50 mm or more (i.e., number of blue lines for that lag)."}
+# Determine all storm-county pairs with 150 mm or more of rainfall over 
+# lag -5 to 3
 storm_lag <- county_rain(counties = my_fips,
@@
                              rain_limit = 150, dist_limit = 500,
-                             days_include = -5:3)
-storm_lag <- storm_lag %>%
+                             days_include = -5:3) %>%
         mutate(storm_county = paste0(storm_id, "_", fips))
 
-rain_lag <- rain %>%
+lag_6 <- storm_lag %>%
+        select(fips, storm_id) %>%
+        mutate(lag = -6, precip = 0, precip_max = 0)
+
+# Pull in rain data and limit to storm-county pairs with 150 mm or more 
+# of rainfall over lag -5 to 3. Figure out the cumulative values at each 
+# day within a county-storm for each lag day.
+rain_lag <- rbind(rain, lag_6) %>%
+        arrange(storm_id, fips, lag) %>%
         mutate(storm_county = paste0(storm_id, "_", fips)) %>%
@@
         mutate(cumulative_rain = cumsum(precip),
-               difference_rain = lead(cumulative_rain) - cumulative_rain,
                next_lag = lead(lag),
                next_cumulative_rain = lead(cumulative_rain),
-               most_rain = difference_rain == max(difference_rain,
-                                                  na.rm = TRUE),
-               over_50 = lead(precip) > 50)
-to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1), y = c(0, 500, 500, 0))
-
-rain_lag_plot <- ggplot(rain_lag, aes(x = lag, y = cumulative_rain)) +
-        theme_tufte() + 
-        geom_segment(aes(xend = next_lag, yend = next_cumulative_rain,
+               over_50 = lead(precip) >= 50) %>%
+        filter(lag != 3)
+high_rain <- rain_lag %>%
+        ungroup() %>%
+        group_by(lag) %>%
+        summarize(n_high = sum(over_50)) %>%
+        ungroup() %>%
+        mutate(lag = lag + 1)
+to_highlight <- data.frame(x = c(-3.1, -3.1, 1.1, 1.1) + 0.5, 
+                           y = c(-15, 500, 500, -15))
+
+# Adding + 0.5 so labels show up centered for the day for which we're showing
+# cumulative rainfall.
+ggplot(rain_lag, aes(x = lag + 0.5, y = cumulative_rain)) +
+        theme_few() + 
+        geom_segment(aes(xend = next_lag + 0.5,
+                         yend = next_cumulative_rain,
                          color = over_50),
@@
                      alpha = 0.2, color = "red", fill = NA) +
-        xlab("Days from storm's closest approach to county") + 
+        geom_text(data = high_rain, 
+                  aes(x = lag, y = 475,
+                      label = prettyNum(n_high, big.mark = ","))) + 
+        xlab("Lag period from storm's closest approach to county") + 
         ylab("Cumulative rainfall in \ncounty from storm (mm)") + 
@@
                            guide = FALSE) 
-rain_lag_plot
-```
-
-Here is a histogram of the amount of rain per day during a storm for the county-storm pairs with 150 mm or more rain associated with a storm.
-Although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm.
-
-```{r echo = FALSE, fig.width = 6, fig.height = 4}
-ggplot(rain_lag, aes(x = lag, y = precip, group = lag)) + 
-        geom_boxplot() + theme_tufte() + 
-        xlab("Days from storm's closest approach to county") + 
-        ylab("Rainfall in county (mm)") 
-```
-
-To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we also investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").
+```
+
+By far, the majority of days in which precipitation in a county exceeded 50 millimeters were within two days before and one day after the storm's closest approach to the county (Figure \ref{fig:cumulative_rain}); in fact, `r round(100 * high_rain$n_high[high_rain$lag == 0] / sum(high_rain$n_high))`% of these days were the day of the storm's closest approach to the county.
+By using a period from two days before to one day after the storm's closest approach, we are capturing `r round(100 * sum(high_rain$n_high[high_rain$lag %in% -2:1]) / sum(high_rain$n_high))`% of days with precipitation of 50 millimeters or more within county-storm observations where the county received 150 millimeters or more rain between five days before and three days after the storm's closest approach.
+
+Similarly, a histogram of the amount of rain within a county per day during a storm for the county-storm pairs with 150 mm or more storm-associated rain shows that, although there are outliers with high rainfalls on all days, for most county-storm pairs, the only days with notable rain are one and two days before the storm, the day of the storm, and one day after the storm (Figure \ref{fig:daily_rain}).
+
+```{r daily_rain, echo = FALSE, fig.width = 6, fig.height = 4}
+rain_lag %>%
+        filter(lag != -6) %>%
+        ggplot(aes(x = lag, y = precip, group = lag)) + 
+        geom_boxplot(outlier.size = 0.5) + theme_few() + 
+        xlab("Lag period from storm's closest approach to county") + 
+        ylab("Daily rainfall in county (mm)") 
+```
+
+We did a more detailed investigation of some of the storms that were notable for being slow-moving or having a looped pattern, which could result in our window underestimating storm-associated precipitation.
+For example, [storm example] moved very slowly and was very large, and so if there is significant rainfall outside of our window because of storm size and speed, it should show up for this storm.
+As another example, Tropical Storm Allison (2001) followed a looped pattern, and so passed near some counties twice, with around [x] days between these two approaches in some cases.
+Our method of estimating rainfall may not include all rain from both passes of the storm.
+
+[To explore if this window (two days before to one day after the storm's closest approach) was adequately large, we also investigated patterns in rainfall by day (from four days before to two days after the storm's closest approach) for a few storms that had the potential to not fit this pattern, Floyd (1999), Irene (2011), Lee (2011), Ike (2008), and Allison (2001) (results shown in "validate_precip.pdf").]
 
@@
 
-For the most part, using a four-day window from two days before to one day after a tropical cyclone in assessing precipitation exposure seemed to capture the effects of the storm while excluding potential false positives from precipitation from other weather systems.
-However, for storms that have tracks that make a loop, such as Allison (2001), this four-day window may not be adequate, and lead to an underestimation of total rainfall associated with the storm relative to estimates for faster-moving storms.
+For the most part, using our window (two days before to one day after a storm's closest approach) captured most of the storm-associated precipitation that would have been counted using a wider window from five days before to two days after a storm's closest approach, even for a large, slow storm.
+However, for a storm that loops and so passes some counties twice, such as Allison (2001), this four-day window may not be adequate, and lead to an underestimation of total rainfall associated with the storm relative to estimates for faster-moving storms.
+Other storms that looped over or near land and so passed close to US counties twice include Allison (1989), Alberto (1994), Gordon (1994), Dennis (1999), Edouard (2002), Fay (2002), Alex (2004), Ivan (2004), Dennis (2005), and Ophelia (2005).
 

### 2016-08-03 writing/DraftExposurePaper/dade_storm_plot.R
Trying out Miami-Dade example

@@ -37,8 +37,8 @@ all_fips <- c(all_fips, as.integer(12025))
 check_dates[check_dates$fips == 12086, "fips"] <- 12025
 
 dade_rain <- data.table::fread(
-        "/Users/joshuaferreri/Documents/nasa_precip_export_2.txt",
-        #"~/Documents/CSU2016/hurricaneproject/hurricaneexposuredata/data-raw/nasa_precip_export_2.txt",
+        #"/Users/joshuaferreri/Documents/nasa_precip_export_2.txt",
+        "~/Documents/CSU2016/hurricaneproject/hurricaneexposuredata/data-raw/nasa_precip_export_2.txt",
         header = TRUE,
         select = c("county", "year_month_day", "precip", "precip_max")) %>%
         dplyr::filter(county %in% all_fips,
@@ -64,17 +64,19 @@ dade_rain <- dade_rain %>%
 #create timeseries for Dade precipitation data from the `countyweather` package
 
 fips <- "12086"
-county_timeseries(fips, percent_coverage = 0,
-                  date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP",
-                  out_directory = "dade_data/")
-                  #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "dade_data/")
+#                   #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
 
 # miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
 dade_dir <- "dade_data/"
 dade_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
 
-dade_weather <- dade_weather %>%
+dade_weather <- dade_weather$averaged %>%
         filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
 
 dade_weather$fips <- fips
@@ -213,5 +215,5 @@ miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +
         scale_size_continuous(guide = "none") +
[…]

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on draft

@@
 # day within a county-storm for each lag day.
-rain_lag <- rbind(rain, lag_6) %>%
+rain_lag_1 <- rbind(rain, lag_6) %>%
         arrange(storm_id, fips, lag) %>%
@@
                next_cumulative_rain = lead(cumulative_rain),
-               over_50 = lead(precip) >= 50) %>%
-        filter(lag != 3)
+               over_50 = lead(precip) >= 50)
+rain_lag <- rain_lag_1 %>% filter(lag != 3)
 high_rain <- rain_lag %>%
@@
 
-```{r daily_rain, echo = FALSE, fig.width = 6, fig.height = 4}
-rain_lag %>%
+```{r daily_rain, echo = FALSE, fig.width = 6, fig.height = 4, fig.align = "center", fig.cap = "Boxplots of daily rainfall for days at a lag of five days before (-5) to three days after (3) compared to the date of the storm's closest approach to a county.
+These values are only for county-storm observations for which a total of at least 150 millimeters of rain fell over the full period (lags -5 to 3)."}
+rain_lag_1 %>%
         filter(lag != -6) %>%
@@
 
-We compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season.
+The method we use to assess storm-related rain exposure is spatially and temporally complete, so it allows us to assess storm exposure for every eastern US county for every storm from 1988 to 2011.
+However, we wanted to compare this method of assessing storm-related rain with other ways of measuring storm-related rain.
+We compared our method with two other assessments.
+First, we compared our results to rain totals from ground-based monitors, using county-level values averaged across all available ground-based precipitation monitors in the county.
+For this assessment, we picked several counties and compared storm-related rain estimates using our data and ground-based monitor estimates for all storms over our study period.
+We also picked several storms and compared our method to values from ground-based monitors for several counties for that storm.
+
+Second, researchers from the [NWS?
+Hurricane Center?] publish yearly tropical cyclone season summaries in the journal *Monthly Weather Review*.
+For some storms, this report includes multiple ground-based measurements of total precipitation and wind speeds (both sustained and gusts).
+We picked a few storms with multiple ground-based measurements included in the *Monthly Weather Review* report, and compared assessments using our method to these reported storm-related measurements.
+
+Here are results for comparing our method to using ground-based monitors.
+First, we looked at several counties, comparing estimates of storm-associated rain using NLDAS-2 versus ground-based monitors (Figure [x]).
+
+[Add in figures by county here for county averages of ground-based monitors versus our method.]
+
+Next, we looked at several different counties within a few notable storms.
+These storms were picked based on storms that had a large number of ground-based precipitation observations in either the main text or supplemental material of the *Monthly Weather Review* report.
+
+[Move figures here for NLDAS vs. county average values.]
+
+Finally, we compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season (Figure \ref{fig:mwr_comparison}).
 Here are some of the plots comparing county-level estimates of rain exposure from the NLDAS-2 data to all *MWR*-reported rainfall estimates within the county:
 
-```{r echo = FALSE, fig.height = 4, fig.width = 6}
-load("alex_gauge.Rdata")
-alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
-                             start_year = 2010, end_year = 2010,
-                             rain_limit = 0, dist_limit = 1500,
-                             days_included = c(-2, -1, 0, 1))
-alex_nldas <- alex_nldas %>%
-        filter(storm_id == "Alex-2010")
-alex_gauge$fips <- as.character(alex_gauge$fips)
-compare_rain <- right_join(alex_nldas, alex_gauge, by = "fips")
-a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Alex, 2010")
-a
+```{r mwr_comparison, fig.width = 10, fig.height = 10, echo = FALSE, fig.align = "center", fig.cap = "Comparison of estimated storm-associated rainfall from our method (using NLDAS-2 data for a window from two days before to one day after the storm's closest approach to the county) with ground-based observations reported in the Monthly Weather Review report.
+Observations within the same county are linked by a line segment.
+Blue points / lines indicate a county for which the storm came within 500 kilometers, while gray points / lines indicate that the storm did not come within 500 kilometers of the county (and so would be excluded as an 'exposed' county based on distance).
+The diagonal gray line shows where points would fall if assessments were identical.
+Horizontal and vertical gray lines show the 75 millimeter threshold for storm-associated rainfall we use for rain-related storm exposure classification."}
+# Function to plot comparison between two datasets
+plot_mwr_compare <- function(data){
+        storm_year <- as.numeric(gsub(".+-", "", data$storm_id[1]))
+        data$fips <- sprintf("%05s", data$fips)
+        storm_title <- paste(unlist(strsplit(as.character(data$storm_id[1]), 
+                                       split = "-")), collapse = ", ")
+
+        compare_rain <- county_rain(counties = unique(data$fips),
+                             start_year = storm_year, 
+                             end_year = storm_year,
+                             rain_limit = 0, dist_limit = 2000) %>%
+                dplyr::filter(storm_id == data$storm_id[1]) %>%
+                dplyr::right_join(data, by = "fips") %>%
+                dplyr::mutate(within_500_km = storm_dist <= 500)
 
@@
                   highest_rain = max(rain_mm),
-                  nldas_rain = mean(tot_precip))
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
+                  nldas_rain = mean(tot_precip),
+                  within_500_km = first(storm_dist) <= 500)
+
+        plot_mwr <- ggplot(compare_rain, 
+                   aes(x = tot_precip, y = rain_mm,
+                       color = within_500_km)) + 
+        geom_hline(yintercept = 75, color = "lightgray") + 
+        geom_vline(xintercept = 75, color = "lightgray") + 
+                geom_abline(aes(intercept = 0, slope = 1), color = "gray") + 
         geom_segment(data = county_range, 
@@
                          y = lowest_rain, yend = highest_rain,
-                         group = fips), color = "gray") + 
+                         group = fips, color = within_500_km)) + 
         geom_point(aes(group = NA), alpha = 0.5) + 
         ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
         xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
         ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
-        ggtitle("Alex, 2010") + 
+        ggtitle(storm_title) + 
         xlab("NLDAS-2 rainfall (mm)") + 
-        ylab("MWR rainfall reports (mm)")
-b
-```
-
-```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain %>% group_by(fips) %>%
-        summarize(ave_rain_mm = mean(rain_mm),
-                  ave_tot_precip = mean(tot_precip),
-                  n = n(),
-                  within_500_km = first(storm_dist) <= 500) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
-                   size = n, color = within_500_km)) + 
-        geom_hline(yintercept = 75, color = "gray") + 
-        geom_vline(xintercept = 75, color = "gray") + 
-        geom_point(alpha = 0.5) + 
-        theme_bw() +
-        xlim(c(10, 225)) + ylim(c(10, 225)) + 
-        geom_abline(intercept = 0, slope = 1) + 
-        xlab("Rainfall based on NLDAS-2 data (mm)") + 
-        ylab("Rainfall reported in Monthly \nWeather Report (mm)")  + 
-        ggtitle("Alex, 2010") + 
+        ylab("Monthly Weather Report\n rainfall reports (mm)") + 
         scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
                            guide = "none") 
-```
-
-```{r echo = FALSE, fig.height = 4, fig.width = 6}
-#allison_gauge <- read.csv("~/Documents/hurricaneexposure/writing/DraftExposurePaper/MWR_gauge_rain_data/allison_mwr.csv")
+        return(plot_mwr)
+}
+
 allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
-allison_nldas <- county_rain(counties = unique(allison_gauge$fips),
-                             start_year = 2001, end_year = 2001,
-                             rain_limit = 0, dist_limit = 1500,
-                             days_included = c(-2, -1, 0, 1))
-allison_nldas <- allison_nldas %>%
-        filter(storm_id == "Allison-2001")
-allison_gauge$fips <- as.character(allison_gauge$fips)
-compare_rain <- right_join(allison_nldas, allison_gauge, by = "fips")
-
-a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Allison, 2011")
-a
-county_range <- compare_rain %>% 
-        group_by(fips) %>%
-        summarize(lowest_rain = min(rain_mm),
-                  highest_rain = max(rain_mm),
-                  nldas_rain = mean(tot_precip))
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
-        geom_segment(data = county_range, 
-                     aes(x = nldas_rain, xend = nldas_rain,
-                         y = lowest_rain, yend = highest_rain,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
-        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
-        ggtitle("Allison, 2001") + 
-        xlab("NLDAS-2 rainfall (mm)") + 
-        ylab("MWR rainfall reports (mm)")
-b
-```
-
-```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain %>% group_by(fips) %>%
-        summarize(ave_rain_mm = mean(rain_mm),
-                  ave_tot_precip = mean(tot_precip),
-                  n = n(),
-                  within_500_km = first(storm_dist) <= 500) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
-                   size = n, color = within_500_km)) + 
-        geom_hline(yintercept = 75, color = "gray") + 
-        geom_vline(xintercept = 75, color = "gray") + 
-        geom_point(alpha = 0.5) + 
-        theme_bw() +
-        xlim(c(25, 200)) + ylim(c(25, 800)) + 
-        geom_abline(intercept = 0, slope = 1) + 
-        xlab("Rainfall based on NLDAS-2 data (mm)") + 
-        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
-        ggtitle("Allison, 2001") + 
-        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
-                           guide = "none") 
-```
-
-```{r echo = FALSE, fig.height = 4, fig.width = 6}
+allison_mwr <- plot_mwr_compare(allison_gauge)
+
+load("frances_gauge.RData")
+frances_gauge$rain_mm <- convert_precip(frances_gauge$precip_in,
+                                        "inches", "mm")
+frances_mwr <- plot_mwr_compare(frances_gauge)
 
@@
                                     "inches", "mm")
-ike_gauge$fips <- sprintf("%05s", ike_gauge$fips)
-ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
-                             start_year = 2008, end_year = 2008,
-                             rain_limit = 0, dist_limit = 1500,
-                             days_included = c(-2, -1, 0, 1))
-ike_nldas <- ike_nldas %>%
-        filter(storm_id == "Ike-2008")
-ike_gauge$fips <- as.character(ike_gauge$fips)
-compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
-a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Ike, 2008")
-a
-
-county_range <- compare_rain %>% 
-        group_by(fips) %>%
-        summarize(lowest_rain = min(rain_mm),
-                  highest_rain = max(rain_mm),
-                  nldas_rain = mean(tot_precip))
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
-        geom_segment(data = county_range, 
-                     aes(x = nldas_rain, xend = nldas_rain,
-                         y = lowest_rain, yend = highest_rain,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
-        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
-        ggtitle("Ike, 2008") + 
-        xlab("NLDAS-2 rainfall (mm)") + 
-        ylab("MWR rainfall reports (mm)")
-b
-```
-
-[JMF-- Missing values are those from Arkansas (missing the leading 0 in fips code); GBA: I fixed this.
-This is ultimately something I need to fix in the `hurricaneexposure` package, but I've got this example fixed to run without losing the FIPS with leading zeros.]
-
-```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain %>% group_by(fips) %>%
-        summarize(ave_rain_mm = mean(rain_mm),
-                  ave_tot_precip = mean(tot_precip),
-                  n = n(),
-                  within_500_km = first(storm_dist) <= 500) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
-                   size = n, color = within_500_km)) + 
-        geom_hline(yintercept = 75, color = "gray") + 
-        geom_vline(xintercept = 75, color = "gray") + 
-        geom_point(alpha = 0.5) + 
-        theme_bw() +
-        xlim(c(10, 225)) + ylim(c(10, 225)) + 
-        geom_abline(intercept = 0, slope = 1) + 
-        xlab("Rainfall based on NLDAS-2 data (mm)") + 
-        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
-        ggtitle("Allison, 2001") + 
-        scale_color_manual(values = c("blue", "black"), guide = "none") 
-```
-
-Comparing MWR precipitation data to NLDAS-2 data for Frances-2004:
-
-```{r}
-load("frances_gauge.RData")
-frances_gauge$rain_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
-frances_gauge$fips <- sprintf("%05s", frances_gauge$fips)
-frances_nldas <- county_rain(counties = unique(frances_gauge$fips),
-                             start_year = 2004, end_year = 2004,
-                             rain_limit = 0, dist_limit = 1500,
-                             days_included = c(-2, -1, 0, 1))
-frances_nldas <- frances_nldas %>%
-        filter(storm_id == "Frances-2004")
-frances_gauge$fips <- as.character(frances_gauge$fips)
-compare_rain <- right_join(frances_nldas, frances_gauge, by = "fips")
-a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Frances, 2004")
-a
-
-county_range <- compare_rain %>% 
-        group_by(fips) %>%
-        summarize(lowest_rain = min(rain_mm),
-                  highest_rain = max(rain_mm),
-                  nldas_rain = mean(tot_precip))
-b <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm)) + 
-        geom_segment(data = county_range, 
-                     aes(x = nldas_rain, xend = nldas_rain,
-                         y = lowest_rain, yend = highest_rain,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_rain$tot_precip, compare_rain$rain_mm)) +
-        ylim(range(compare_rain$tot_precip, compare_rain$rain_mm)) + 
-        ggtitle("Frances, 2004") + 
-        xlab("NLDAS-2 rainfall (mm)") + 
-        ylab("MWR rainfall reports (mm)")
-b
-```
-
-```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
-compare_rain %>% group_by(fips) %>%
-        summarize(ave_rain_mm = mean(rain_mm),
-                  ave_tot_precip = mean(tot_precip),
-                  n = n(),
-                  within_500_km = first(storm_dist) <= 500) %>%
-        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
-                   size = n, color = within_500_km)) + 
-        geom_hline(yintercept = 75, color = "gray") + 
-        geom_vline(xintercept = 75, color = "gray") + 
-        geom_point(alpha = 0.5) + 
-        theme_bw() +
-        xlim(c(10, 225)) + ylim(c(10, 225)) + 
-        geom_abline(intercept = 0, slope = 1) + 
-        xlab("Rainfall based on NLDAS-2 data (mm)") + 
-        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
-        scale_color_manual(values = c("blue", "black"), guide = "none") 
-```
-
-Ike wind data:
-
-```{r}
-#Ike-2008 wind
-load("ike_wind.RData")
-ike_wind <- ike_wind %>%
-        filter(!is.na(max_sust_kt)) %>%
-        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
-                                                 old_metric = "knots",
-                                                 new_metric = "mps"),
-               fips = sprintf("%05s", fips))
-
-ike_wind_model <- county_wind(counties = unique(ike_wind$fips),
-                              start_year = 2008, end_year = 2008,
-                              wind_limit = 0) %>%
-        filter(storm_id == "Ike-2008")
-
-compare_wind <- right_join(ike_wind_model, ike_wind, by = "fips")
-a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Ike, 2008")
-a
-
-county_range <- compare_wind %>% 
-        group_by(fips) %>%
-        summarize(lowest_wind = min(max_sust_mps),
-                  highest_wind = max(max_sust_mps),
-                  model_wind = mean(max_sust))
-b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
-        geom_segment(data = county_range, 
-                     aes(x = model_wind, xend = model_wind,
-                         y = lowest_wind, yend = highest_wind,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
-        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
-        ggtitle("Ike, 2008") + 
-        xlab("Modeled wind speed (m / s)") + 
-        ylab("MWR wind reports (m / s)")
-b
-```
-
-Frances wind data:
-
-```{r eval = FALSE}
-#frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
- #                        na.strings = 999,
-  #                       header = TRUE)
-#save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
-```
-
-```{r}
-load("frances_wind.RData")
-frances_wind <- frances_wind %>%
-        filter(!is.na(wind_sust)) %>%
-        mutate(max_sust_mps = convert_wind_speed(wind_sust, 
-                                                 old_metric = "knots",
-                                                 new_metric = "mps"),
-               fips = sprintf("%05s", fips))
-
-frances_wind_model <- county_wind(counties = unique(frances_wind$fips),
-                              start_year = 2004, end_year = 2004,
-                              wind_limit = 0) %>%
-        filter(storm_id == "Frances-2004")
-
-compare_wind <- right_join(frances_wind_model, frances_wind, by = "fips")
-a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Frances, 2004")
-a
-
-county_range <- compare_wind %>% 
-        group_by(fips) %>%
-        summarize(lowest_wind = min(max_sust_mps),
-                  highest_wind = max(max_sust_mps),
-                  model_wind = mean(max_sust))
-b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
-        geom_segment(data = county_range, 
-                     aes(x = model_wind, xend = model_wind,
-                         y = lowest_wind, yend = highest_wind,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
-        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
-        ggtitle("Frances, 2004") + 
-        xlab("Modeled wind speed (m / s)") + 
-        ylab("MWR wind reports (m / s)")
-b
-```
-
-####How does NLDAS-2 data compare with rnoaa data from the `countyweather` package in R?
-
-```{r eval = FALSE}
-options("noaakey" = Sys.getenv("noaakey"))
-#Alex-2010
-alex_fips <- alex_nldas$fips
-alex_fips <- unique(alex_fips)
-```
-```{r}
-#Ike-2008
-ike_fips <- ike_nldas$fips
-ike_fips <- unique(ike_fips)
-```
-```{r}
-#Allison-2010
-allison_fips <- allison_nldas$fips
-allison_fips <- unique(allison_fips)
-```
-```{r}
-#Frances-2004
-frances_fips <- frances_nldas$fips
-frances_fips <- unique(frances_fips)
-```
-
-Trying to create a loop to identify the FIPS codes with the greatest number of stations reporting for each time period.
-
-```{r eval = FALSE}
-my_fips <- ike_fips[1:3]
-my_stations <- NA
-test <- expand.grid(my_fips, my_stations)
-colnames(test) <- c("fips", "station")
-for(fips in my_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-} #This works fine (small subset of code below)
-ike_table <- expand.grid(ike_fips, my_stations)
-colnames(ike_table) <- c("fips", "station")
-#for(fips in ike_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #getting error
+ike_mwr <- plot_mwr_compare(ike_gauge)
+
+load("alex_gauge.Rdata")
+alex_mwr <- plot_mwr_compare(alex_gauge)
+
+grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
 ```
@@
 
-As of right now, I am not going to do Frances (there are 184 unique fips).
-
-Create a timeseries for the top 15 FIPS by reporting stations for Alex and Ike:
-
 ```{r eval = FALSE}
@@
 
-[JMF-- It looks like precipitation measurements from `countyweather` are already in mm]
+```{r eval = FALSE}
+options("noaakey" = Sys.getenv("noaakey"))
+#Alex-2010
+alex_fips <- alex_nldas$fips
+alex_fips <- unique(alex_fips)
+
+#Ike-2008
+ike_fips <- ike_nldas$fips
+ike_fips <- unique(ike_fips)
+
+#Allison-2010
+allison_fips <- allison_nldas$fips
+allison_fips <- unique(allison_fips)
+
+#Frances-2004
+frances_fips <- frances_nldas$fips
+frances_fips <- unique(frances_fips)
+```
+
+Trying to create a loop to identify the FIPS codes with the greatest number of stations reporting for each time period.
+
+```{r eval = FALSE}
+my_fips <- ike_fips[1:3]
+my_stations <- NA
+test <- expand.grid(my_fips, my_stations)
+colnames(test) <- c("fips", "station")
+for(fips in my_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
+} #This works fine (small subset of code below)
+ike_table <- expand.grid(ike_fips, my_stations)
+colnames(ike_table) <- c("fips", "station")
+#for(fips in ike_fips){
+        ike_weather <- weather_fips(fips = fips,
+                                    date_min = "2008-09-01",
+                                    date_max = "2008-09-14",
+                                    var = "PRCP")
+        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
+} #getting error
+```
 
@@
 [Validate windspeedmodel functions using Extended Best Tracks Data]
+
+Ike wind data:
+
+```{r}
+#Ike-2008 wind
+load("ike_wind.RData")
+ike_wind <- ike_wind %>%
+        filter(!is.na(max_sust_kt)) %>%
+        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
+                                                 old_metric = "knots",
+                                                 new_metric = "mps"),
+               fips = sprintf("%05s", fips))
+
+ike_wind_model <- county_wind(counties = unique(ike_wind$fips),
+                              start_year = 2008, end_year = 2008,
+                              wind_limit = 0) %>%
+        filter(storm_id == "Ike-2008")
+
+compare_wind <- right_join(ike_wind_model, ike_wind, by = "fips")
+a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw() + 
+        ggtitle("Ike, 2008")
+a
+
+county_range <- compare_wind %>% 
+        group_by(fips) %>%
+        summarize(lowest_wind = min(max_sust_mps),
+                  highest_wind = max(max_sust_mps),
+                  model_wind = mean(max_sust))
+b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
+        geom_segment(data = county_range, 
+                     aes(x = model_wind, xend = model_wind,
+                         y = lowest_wind, yend = highest_wind,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
+        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
+        ggtitle("Ike, 2008") + 
+        xlab("Modeled wind speed (m / s)") + 
+        ylab("MWR wind reports (m / s)")
+b
+```
+
+Frances wind data:
+
+```{r eval = FALSE}
+#frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
+ #                        na.strings = 999,
+  #                       header = TRUE)
+#save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
+```
+
+```{r}
+load("frances_wind.RData")
+frances_wind <- frances_wind %>%
+        filter(!is.na(wind_sust)) %>%
+        mutate(max_sust_mps = convert_wind_speed(wind_sust, 
+                                                 old_metric = "knots",
+                                                 new_metric = "mps"),
+               fips = sprintf("%05s", fips))
+
+frances_wind_model <- county_wind(counties = unique(frances_wind$fips),
+                              start_year = 2004, end_year = 2004,
+                              wind_limit = 0) %>%
+        filter(storm_id == "Frances-2004")
+
+compare_wind <- right_join(frances_wind_model, frances_wind, by = "fips")
+a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
+                              color = fips)) +
+        geom_point() + 
+        theme_bw() + 
+        ggtitle("Frances, 2004")
+a
+
+county_range <- compare_wind %>% 
+        group_by(fips) %>%
+        summarize(lowest_wind = min(max_sust_mps),
+                  highest_wind = max(max_sust_mps),
+                  model_wind = mean(max_sust))
+b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
+        geom_segment(data = county_range, 
+                     aes(x = model_wind, xend = model_wind,
+                         y = lowest_wind, yend = highest_wind,
+                         group = fips), color = "gray") + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        geom_abline(aes(intercept = 0, slope = 1)) + 
+        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
+        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
+        ggtitle("Frances, 2004") + 
+        xlab("Modeled wind speed (m / s)") + 
+        ylab("MWR wind reports (m / s)")
+b
+```
 

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on draft

@@
 library(weathermetrics)
+library(lubridate)
 
@@
 #devtools::install_github("leighseverson/countyweather")
+options("noaakey" = Sys.getenv("noaakey"))
 library(countyweather)
@@
 
-Next, we looked at several different counties within a few notable storms.
+Next, we looked at several different counties within a few notable storms, to compare county-averaged ground-based monitor values with our estimates (Figure \ref{fig:county_ave_comparison}).
 These storms were picked based on storms that had a large number of ground-based precipitation observations in either the main text or supplemental material of the *Monthly Weather Review* report.
 
-[Move figures here for NLDAS vs. county average values.]
+```{r echo = FALSE, eval = FALSE}
+options("noaakey" = Sys.getenv("noaakey"))
+
+allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
+allison_fips <- sprintf("%05s", as.character(unique(allison_gauge$fips)))
+allison_fips <- sample(allison_fips, 10)
+date_range <- closest_dist %>% 
+        filter(storm_id == "Allison-2001") %>%
+        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
+                  latest_date = max(ymd(closest_date)) + ddays(2))
+county_timeseries(allison_fips, percent_coverage = 0,
+                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
+                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
+                  var = "PRCP", out_directory = "county_ave/allison/")
+
+load("MWR_gauge_rain_data/frances_gauge.RData")
+frances_fips <- sprintf("%05s", as.character(unique(frances_gauge$fips)))
+frances_fips <- sample(frances_fips, 10)
+date_range <- closest_dist %>% 
+        filter(storm_id == "Frances-2004") %>%
+        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
+                  latest_date = max(ymd(closest_date)) + ddays(2))
+county_timeseries(frances_fips, percent_coverage = 0,
+                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
+                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
+                  var = "PRCP", out_directory = "county_ave/frances/")
+
+load("MWR_gauge_rain_data/ike_gauge.RData")
+ike_fips <- sprintf("%05s", as.character(unique(ike_gauge$fips)))
+ike_fips <- sample(ike_fips, 10)
+date_range <- closest_dist %>% 
+        filter(storm_id == "Ike-2008") %>%
+        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
+                  latest_date = max(ymd(closest_date)) + ddays(2))
+county_timeseries(ike_fips, percent_coverage = 0,
+                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
+                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
+                  var = "PRCP", out_directory = "county_ave/ike/")
+
+load("MWR_gauge_rain_data/alex_gauge.RData")
+alex_fips <- sprintf("%05s", as.character(unique(alex_gauge$fips)))
+alex_fips <- sample(alex_fips, 10)
+date_range <- closest_dist %>% 
+        filter(storm_id == "Alex-2010") %>%
+        summarize(earliest_date = min(ymd(closest_date)) - ddays(2),
+                  latest_date = max(ymd(closest_date)) + ddays(2))
+county_timeseries(alex_fips, percent_coverage = 0,
+                  date_min = format(date_range[1, 1], "%Y-%m-%d"),
+                  date_max = format(date_range[1, 2], "%Y-%m-%d"),
+                  var = "PRCP", out_directory = "county_ave/alex/")
+```
+
+```{r county_ave_comparison, echo = FALSE, fig.width = 10, fig.height = 10, fig.align = "center", fig.cap = "Comparison of county-level average cumulative rainfall for the period from two days before to one day after the storm's closest approach to a county, using NLDAS-2 versus ground-based monitors, for counties with both estimates available for four example storms.
+The size of each point indicates the average number of monitors reporting from that county during the window (larger size indicates more monitors).
+The text over each point gives the FIPS county identifier."}
+plot_county_ave_compare <- function(county_dir, storm_id){
+        storm_title <- paste0(paste(unlist(strsplit(
+                as.character(storm_id), "-")),
+                             collapse = " ("), ")")
+        storm_year <- gsub(".+-", "", storm_id)
+        counties <- gsub(".rds", "",
+                         list.files(paste0("county_ave/", county_dir)))
+
+        storm_ave <- vector("list", length = length(counties))
+        names(storm_ave) <- counties
+
+        for(x in counties){
+                county_weather <- readRDS(paste0("county_ave/", 
+                                                 county_dir, "/", x,
+                                                 ".rds"))$averaged %>%
+                        dplyr::ungroup() %>%
+                        dplyr::mutate(date = ymd(date)) %>%
+                        dplyr::mutate(fips = x)
+                storm_ave[[x]] <- county_weather
+                }
+        storm_ave <- do.call("rbind", storm_ave)
+
+        storm_rain <- county_rain(counties = counties, 
+                                  start_year = storm_year,
+                                  end_year = storm_year,
+                                  rain_limit = 0, dist_limit = 2000) %>%
+                dplyr::filter(storm_id == storm_id)
+
+        county_ave <- storm_ave %>%
+                dplyr::left_join(storm_rain, by = "fips") %>%
+                dplyr::group_by(fips) %>%
+                dplyr::filter(ymd(closest_date) - ddays(2) <= date &
+                       date <= ymd(closest_date) + ddays(1)) %>%
+                dplyr::summarize(monitor_rain = sum(prcp),
+                  tot_precip = first(tot_precip),
+                  prcp_reporting = mean(prcp_reporting))
+        plot_range <- range(county_ave[ , c("monitor_rain", "tot_precip")]) + 
+                c(-15, 15)
+
+        county_ave_plot <- ggplot(county_ave, 
+                                  aes(x = tot_precip, 
+                                      y = monitor_rain)) +
+                geom_abline(aes(intercept = 0, slope = 1), color = "gray",
+                    alpha = 0.5) +
+                geom_point(aes(size = prcp_reporting), alpha = 0.3) +
+                geom_text(aes(label = fips)) +
+                theme_few() +
+                scale_size_continuous(guide = "none") +
+                ylab("Rainfall (mm) based on \naveraged county monitors") +
+                xlab("Rainfall (mm) based on NLDAS-2 county data") +
+                xlim(plot_range) +
+                ylim(plot_range) + 
+                ggtitle(storm_title)
+        return(county_ave_plot)
+}
+
+allison_county_ave <- plot_county_ave_compare("allison", 
+                                           storm_id = "Allison-2001")
+
+frances_county_ave <- plot_county_ave_compare("frances", 
+                                           storm_id = "Frances-2004")
+
+ike_county_ave <- plot_county_ave_compare("ike", 
+                                           storm_id = "Ike-2008")
+
+alex_county_ave <- plot_county_ave_compare("alex", 
+                                           storm_id = "Alex-2010")
+
+grid.arrange(allison_county_ave, frances_county_ave,
+             ike_county_ave, alex_county_ave, ncol = 2)
+```
 
 Finally, we compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season (Figure \ref{fig:mwr_comparison}).
-Here are some of the plots comparing county-level estimates of rain exposure from the NLDAS-2 data to all *MWR*-reported rainfall estimates within the county:
 
@@
 
-load("frances_gauge.RData")
+load("MWR_gauge_rain_data/frances_gauge.RData")
 frances_gauge$rain_mm <- convert_precip(frances_gauge$precip_in,
@@
 
-load("ike_gauge.RData")
+load("MWR_gauge_rain_data/ike_gauge.RData")
 ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
@@
 
-load("alex_gauge.Rdata")
+load("MWR_gauge_rain_data/alex_gauge.Rdata")
 alex_mwr <- plot_mwr_compare(alex_gauge)
@@
 grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
-```
-
-```{r}
-#Find the top 15 FIPS by number of reporting stations during Ike-2008 (Warning: Takes a long time)
-#my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
-#my_stations <- NA
-#ike_table <- expand.grid(my_fips, my_stations)
-#colnames(ike_table) <- c("fips", "station")
-#for(fips in my_fips){
-        #ike_weather <- weather_fips(fips = fips,
-                                    #date_min = "2008-09-01",
-                                    #date_max = "2008-09-14",
-                                    #var = "PRCP")
-        #ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-#} 
-#Ike FIPS without data: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
-#ike_table <- ike_table %>%
-        #arrange(desc(station)) %>%
-        #slice(1:15)
-#ike_top <- ike_table$fips
-#Find the top 15 FIPS by number of reporting stations for Alex-2010:
-#my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
-#my_stations <- NA
-#alex_table <- expand.grid(my_fips, my_stations)
-#colnames(alex_table) <- c("fips", "station")
-#for(fips in my_fips){
-#        alex_weather <- weather_fips(fips = fips,
-#                                    date_min = "2008-09-01",
-#                                    date_max = "2008-09-14",
-#                                    var = "PRCP")
-#        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
-#} 
-#Alex FIPS without data: 3, 7, 9, 11, 15, 19   
-#alex_table <- alex_table %>%
-#        arrange(desc(station)) %>%
-#        slice(1:15)
-#alex_top <- alex_table$fips
-#fips_table <- rbind(alex_table, ike_table)
-#fips_table <- fips_table %>%
-#        arrange(desc(station)) %>%
-#        slice(1:15)
-#all_fips <- fips_table$fips
-#all_fips <- unique(all_fips)
-```
-
-```{r eval = FALSE}
-county_timeseries(fips = ike_fips[1:2], 
-                  percent_coverage = 0, 
-                  date_min = "2008-09-01", 
-                  date_max = "2010-09-14", var = "PRCP", 
-                  out_directory = "ike_ex")
-ike_1 <- readRDS("ike_ex/12021.rds")
-keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
-keys_cd <- keys_cd$closest_date
-ike_1 <- ike_1 %>%
-        select(date, prcp) %>%
-        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), by = 1))
-ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
-ike_sum <- sum(ike_1$prcp)
-ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
-ike_nldas_sum <- ike_nldas_sum$tot_precip
-```
-
-```{r eval = FALSE}
-options("noaakey" = Sys.getenv("noaakey"))
-#Alex-2010
-alex_fips <- alex_nldas$fips
-alex_fips <- unique(alex_fips)
-
-#Ike-2008
-ike_fips <- ike_nldas$fips
-ike_fips <- unique(ike_fips)
-
-#Allison-2010
-allison_fips <- allison_nldas$fips
-allison_fips <- unique(allison_fips)
-
-#Frances-2004
-frances_fips <- frances_nldas$fips
-frances_fips <- unique(frances_fips)
-```
-
-Trying to create a loop to identify the FIPS codes with the greatest number of stations reporting for each time period.
-
-```{r eval = FALSE}
-my_fips <- ike_fips[1:3]
-my_stations <- NA
-test <- expand.grid(my_fips, my_stations)
-colnames(test) <- c("fips", "station")
-for(fips in my_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
-} #This works fine (small subset of code below)
-ike_table <- expand.grid(ike_fips, my_stations)
-colnames(ike_table) <- c("fips", "station")
-#for(fips in ike_fips){
-        ike_weather <- weather_fips(fips = fips,
-                                    date_min = "2008-09-01",
-                                    date_max = "2008-09-14",
-                                    var = "PRCP")
-        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
-} #getting error
 ```
@@
 For example, Lee in 2011 was only tracked through the best tracks until it was near the Georgia-Tennessee border.
-However, remnants of the system brought extreme rain further north, into the Mid-Atlantic and New England [I believe that the Monthly Weather Review for the 2011 season includes a description of these rain impacts.
-We likely want to look over that and incorporate it here]:
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
+However, remnants of the system brought extreme rain further north, into the Mid-Atlantic and New England (Figure \ref{fig:lee_rain_example}).
+[I believe that the Monthly Weather Review for the 2011 season includes a description of these rain impacts.
+We likely want to look over that and incorporate it here].
+Other possible examples of this during our study period include Nicole (2010) [check storm history in MWR to confirm that Mid-Atlantic and New England rain during this time was related to storm], Ida (2009), Humberto (2007), Katrina (2005), Grace (2003) [confirm rain was from storm remnants], Frances (1998), and Marco (1990).
+
+```{r lee_rain_example, echo = FALSE, fig.width = 6, fig.height = 8, fig.align = "center", fig.cap = "Example of a tropical storm for which the 'best tracks' track ends early enough to miss significant later effects.
+These two maps show rainfall during the days surrounding Lee (2011) (top) as well as the counties classified as 'exposed' (bottom) based on criterion of (1) storm approaches within 500 kilometers of the county's population mean center and (2) county gets 75 millimeters or more of precipitation over the period between two days before and one day after the storm's closest approach to the county.
+The 'best track' is shown in red in both plots and illustrates that tracking ceased too early to capture more northern storm-associated rainfall."}
 storm <- "Lee-2011"
@@
 
-Other possible examples of this during our study period include Nicole (2010) [check storm history in MWR to confirm that Mid-Atlantic and New England rain during this time was related to storm], Ida (2009), Humberto (2007), Katrina (2005), Grace (2003) [confirm rain was from storm remnants], Frances (1998), and Marco (1990).
-
 For rain-based exposure, we use a distance constraint of 500 km.
@@
 There are only a few very large storms where this might not be the case.
-One example is Ike (2008), in which case the 500 km constraint excludes some of the counties in northern Texas that got heavy rain from the system:
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
+One example is Ike (2008), in which case the 500 km constraint excludes some of the counties in northern Texas that got heavy rain from the system (Figure \ref{fig:ike_rain_example}).
+Other large systems where this might be a concern include Wilma (2005), Hermine (2004) [confirm rain in SC, NC, and VA during this time was from this storm], Arlene (1993) [confirm rain through to LA is from this storm].
+
+```{r ike_rain_example, echo = FALSE, fig.width = 6, fig.height = 8, fig.align = "center", fig.cap = "Example of a storm large enough that some storm-associated rainfall was likely missed when using a 500 kilometer distance criterion with the rain-based exposure metric.
+These two maps show rainfall during the days surrounding Ike (2008) (top) as well as the counties classified as 'exposed' (bottom) based on criterion of (1) storm approaches within 500 kilometers of the county's population mean center and (2) county gets 75 millimeters or more of precipitation over the period between two days before and one day after the storm's closest approach to the county.
+The storm's 'best track' is plotted on each map in red.
+This example storm shows that some counties in northern Texas were incorrectly classified as 'unexposed' based on storm-associated rain because they fell outside the distance of 500 kilometers."}
 storm <- "Ike-2008"
@@
 
-Other large systems where this might be a concern include Wilma (2005), Hermine (2004) [confirm rain in SC, NC, and VA during this time was from this storm], Arlene (1993) [confirm rain through to LA is from this storm].
-
-For storms in 1996 or later, when flooding reports are included in NOAA's Storm Events data, exposure patterns based on floods reports generally pick up similar regions as rain-based exposure, and in some cases (e.g., Floyd, 1999), counties classified as exposed using rain-based exposure are very similar to the counties in which flood events were recorded.
-For example, here are comparisons for some of the worst storms in terms of rain-based exposures:
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Frances-2004"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Frances, 2004")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Frances, 2004")
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Ivan-2004"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Ivan, 2004")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Ivan, 2004")
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Isidore-2002"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Isidore, 2002")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Isidore, 2002")
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Floyd-1999"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Floyd, 1999")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Floyd, 1999")
-grid.arrange(a, b, ncol = 1)
-```
-
-For all storms, we divided counties into four categories:
-
-1. "Exposed" based on both rainfall and reported flood event;
-2. "Exposed" based on rainfall, "unexposed" based on flood event; 
-3. "Unexposed" based on rainfall, "exposed" based on flood event; and
-4. "Unexposed" based on both rainfall and reported flood event.
-
-All counties in the first and fourth categories are consistently classified based on using rainfall and flood events as exposure metrics, while those in categories 2 and 3 are inconsistent, in that one metric marks the county as exposed to the storm while the other metric classifies it as unexposed.
-
-```{r echo = FALSE}
-all_storm_counties <- closest_dist %>%
-        dplyr::select(storm_id, fips) %>%
-        dplyr::mutate(storm_county = paste(storm_id, fips, sep = "-")) %>%
-        dplyr::filter(as.numeric(gsub(".*-", "", storm_id)) >= 1996 & 
-                              as.numeric(gsub(".*-", "", storm_id)) <= 2011)
-
-rain_exposure <- county_rain(counties = my_fips, start_year = 1996,
-                             end_year = 2011, rain_limit = 75, 
-                             dist_limit = 500) %>%
-        dplyr::select(storm_id, fips) %>%
-        dplyr::mutate(rain_exposed = 1,
-                      storm_county = paste(storm_id, fips, sep = "-")) %>%
-        dplyr::select(storm_county, rain_exposed)
-flood_exposure <- county_events(counties = my_fips, start_year = 1996,
-                             end_year = 2011, event_type = "flood") %>%
-        dplyr::mutate(flood_exposed = 1,
-                      storm_county = paste(storm_id, fips, sep = "-")) %>%
-        dplyr::select(storm_county, flood_exposed)
-
-all_exposed <- rain_exposure %>%
-        dplyr::full_join(flood_exposure, by = "storm_county") %>%
-        dplyr::full_join(all_storm_counties, by = "storm_county") %>%
-        dplyr::mutate(rain_exposed = !is.na(rain_exposed),
-                      flood_exposed = !is.na(flood_exposed))
-
-storm_exposed <- all_exposed %>%
-        dplyr::group_by(storm_id) %>%
-        dplyr::summarize(both_exposed = sum(rain_exposed & flood_exposed),
-                         rain_only = sum(rain_exposed & !flood_exposed),
-                         flood_only = sum(!rain_exposed & flood_exposed),
-                         neither_exposed = sum(!rain_exposed & 
-                                                       !flood_exposed)) 
-storm_exposed %>% 
-        dplyr::mutate(percent_agreement = round(100 * (both_exposed + 
-                                                         neither_exposed) / 
-                              (both_exposed + rain_only + flood_only +
-                              neither_exposed))) %>%
-        dplyr::filter(both_exposed + rain_only + flood_only > 0) %>%
-        dplyr::arrange(neither_exposed, desc(both_exposed), 
-                       desc(rain_only)) %>%
-        kable()
-```
-
-Among the storms with a large number of counties exposed under one or both metrics, Irene (2011), Floyd (1999), and Fran (1996) are notable for very high agreement between the two metrics (98%, 96%, and 95%, respectively).
-Agreement is comparatively lower (although in no cases is the agreement below [x]%, because so many counties are labeled as unexposed under both metrics for most counties) for Frances (2004; 82% agreement), Ivan (2004, 86% agreement), Lee (2011, 88% agreement), Katrina (2005, 89% agreement), and Isidore (2002; 89% agreement).
-In all these cases, most of the discordant classifications were for counties that were classified as exposed based on rain but did not have a reported flood event.
-
-Here are histograms showing the number of counties classified as exposed for both rain and flood metrics, for flood but not rain, and for rain but not flood.
-This is limited to storms between 1996 and 2011 and only includes storm-county pairs where the county was classified as exposed to the storm under at least one of these two metrics (rain and flood).
-It was more common for a county to be classified as exposed based on rain but unexposed by flood then to be assessed as exposed by both metrics or by flood but not rain.
-In a handful of storms, more than 150 counties were assessed as exposed to the storm based on both rain and flood exposure metrics.
-In some storms, quite a number of counties (in some cases, well over 200) were classified as exposed based on the rain metric but unexposed based on the flood metric.
-In almost every storm, however, fewer than 100 counties (out of 2,396 considered for each storm) were classified as exposed based on the flood metric without also being classified as exposed based on the rain metric.
-This suggests that the rain metric might be less stringent than the flood metric and, in some cases, the two metrics identify different risks (i.e., flood risk is not completely determined by rainfall, as pre-existing local conditions can make a flood more or less likely).
-
-```{r echo = FALSE, fig.height = 6, fig.width = 5}
-storm_exposed %>%
-        dplyr::select(-neither_exposed) %>%
-        dplyr::filter(both_exposed > 0 & rain_only > 0 & flood_only > 0) %>%
-        tidyr::gather(classification, n_counties, -storm_id) %>%
-        ggplot(aes(x = n_counties)) + 
-        geom_histogram(bins = 30) + 
-        facet_wrap(~ classification, ncol = 1) + 
-        theme_tufte() + 
-        xlab("Number of counties classified as exposed") + 
-        ylab("Number of storms")
-```
-
-Here is a table showing, for all storms where 200 or more counties were classified as exposed based on at least one of the two metrics, what percent were classified in a way that agreed (i.e., classified as exposed by both metrics):
-
-```{r echo = FALSE}
-storm_exposed %>%
-        dplyr::mutate(any_exposed = both_exposed + rain_only + flood_only,
-                      both_of_any = round(100 * both_exposed /
-                                                  any_exposed)) %>%
-        dplyr::filter(any_exposed >= 200) %>%
-        dplyr::select(-neither_exposed, -any_exposed) %>%
-        dplyr::arrange(desc(both_of_any)) %>%
-        kable()
-```
-
-Only two storms (Irene in 2011 and Floyd and 1999) have over 50% agreement among counties classified as exposed by at least one of the two metrics.
-Throughout, in almost every storm, among the discordant counties (classified as exposed by one but not both metrics), there were many more counties classified as exposed based on the rain metric but where no flood events were reported in the NOAA Storm Events database.
-The only exceptions, where more discordant counties were classified as exposed based on flood than rain, were Allison (2001) and Cindy (2005).
-
 #### Are our methods of assessing wind speed reasonable
@@
 #Ike-2008 wind
-load("ike_wind.RData")
+load("MWR_gauge_wind_data/ike_wind.RData")
 ike_wind <- ike_wind %>%
@@
 
-```{r eval = FALSE}
-#frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
- #                        na.strings = 999,
-  #                       header = TRUE)
-#save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
-```
-
 ```{r}
-load("frances_wind.RData")
+load("MWR_gauge_wind_data/frances_wind.RData")
 frances_wind <- frances_wind %>%
@@
 
-Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less [GBA-- These are **awesome**.
-I think we want to tweak the color scale and the legend a bit so that they're consistent between the three plots showing the top 2,000 storm-county pairs for the three metrics (distance, rain, and wind).
-That will make it easier to compare across the three maps, which is what we want readers to be able to do here.]:
+Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
 
@@
 
+For storms in 1996 or later, when flooding reports are included in NOAA's Storm Events data, exposure patterns based on floods reports generally pick up similar regions as rain-based exposure, and in some cases (e.g., Floyd, 1999), counties classified as exposed using rain-based exposure are very similar to the counties in which flood events were recorded.
+For example, here are comparisons for some of the worst storms in terms of rain-based exposures:
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Frances-2004"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Frances, 2004")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Frances, 2004")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Ivan-2004"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Ivan, 2004")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Ivan, 2004")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Isidore-2002"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Isidore, 2002")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Isidore, 2002")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r echo = FALSE, fig.width = 6, fig.height = 8}
+storm <- "Floyd-1999"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Floyd, 1999")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Floyd, 1999")
+grid.arrange(a, b, ncol = 1)
+```
+
+For all storms, we divided counties into four categories:
+
+1. "Exposed" based on both rainfall and reported flood event;
+2. "Exposed" based on rainfall, "unexposed" based on flood event; 
+3. "Unexposed" based on rainfall, "exposed" based on flood event; and
+4. "Unexposed" based on both rainfall and reported flood event.
+
+All counties in the first and fourth categories are consistently classified based on using rainfall and flood events as exposure metrics, while those in categories 2 and 3 are inconsistent, in that one metric marks the county as exposed to the storm while the other metric classifies it as unexposed.
+
+```{r echo = FALSE}
+all_storm_counties <- closest_dist %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::filter(as.numeric(gsub(".*-", "", storm_id)) >= 1996 & 
+                              as.numeric(gsub(".*-", "", storm_id)) <= 2011)
+
+rain_exposure <- county_rain(counties = my_fips, start_year = 1996,
+                             end_year = 2011, rain_limit = 75, 
+                             dist_limit = 500) %>%
+        dplyr::select(storm_id, fips) %>%
+        dplyr::mutate(rain_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, rain_exposed)
+flood_exposure <- county_events(counties = my_fips, start_year = 1996,
+                             end_year = 2011, event_type = "flood") %>%
+        dplyr::mutate(flood_exposed = 1,
+                      storm_county = paste(storm_id, fips, sep = "-")) %>%
+        dplyr::select(storm_county, flood_exposed)
+
+all_exposed <- rain_exposure %>%
+        dplyr::full_join(flood_exposure, by = "storm_county") %>%
+        dplyr::full_join(all_storm_counties, by = "storm_county") %>%
+        dplyr::mutate(rain_exposed = !is.na(rain_exposed),
+                      flood_exposed = !is.na(flood_exposed))
+
+storm_exposed <- all_exposed %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(both_exposed = sum(rain_exposed & flood_exposed),
+                         rain_only = sum(rain_exposed & !flood_exposed),
+                         flood_only = sum(!rain_exposed & flood_exposed),
+                         neither_exposed = sum(!rain_exposed & 
+                                                       !flood_exposed)) 
+storm_exposed %>% 
+        dplyr::mutate(percent_agreement = round(100 * (both_exposed + 
+                                                         neither_exposed) / 
+                              (both_exposed + rain_only + flood_only +
+                              neither_exposed))) %>%
+        dplyr::filter(both_exposed + rain_only + flood_only > 0) %>%
+        dplyr::arrange(neither_exposed, desc(both_exposed), 
+                       desc(rain_only)) %>%
+        kable()
+```
+
+Among the storms with a large number of counties exposed under one or both metrics, Irene (2011), Floyd (1999), and Fran (1996) are notable for very high agreement between the two metrics (98%, 96%, and 95%, respectively).
+Agreement is comparatively lower (although in no cases is the agreement below [x]%, because so many counties are labeled as unexposed under both metrics for most counties) for Frances (2004; 82% agreement), Ivan (2004, 86% agreement), Lee (2011, 88% agreement), Katrina (2005, 89% agreement), and Isidore (2002; 89% agreement).
+In all these cases, most of the discordant classifications were for counties that were classified as exposed based on rain but did not have a reported flood event.
+
+Here are histograms showing the number of counties classified as exposed for both rain and flood metrics, for flood but not rain, and for rain but not flood.
+This is limited to storms between 1996 and 2011 and only includes storm-county pairs where the county was classified as exposed to the storm under at least one of these two metrics (rain and flood).
+It was more common for a county to be classified as exposed based on rain but unexposed by flood then to be assessed as exposed by both metrics or by flood but not rain.
+In a handful of storms, more than 150 counties were assessed as exposed to the storm based on both rain and flood exposure metrics.
+In some storms, quite a number of counties (in some cases, well over 200) were classified as exposed based on the rain metric but unexposed based on the flood metric.
+In almost every storm, however, fewer than 100 counties (out of 2,396 considered for each storm) were classified as exposed based on the flood metric without also being classified as exposed based on the rain metric.
+This suggests that the rain metric might be less stringent than the flood metric and, in some cases, the two metrics identify different risks (i.e., flood risk is not completely determined by rainfall, as pre-existing local conditions can make a flood more or less likely).
+
+```{r echo = FALSE, fig.height = 6, fig.width = 5}
+storm_exposed %>%
+        dplyr::select(-neither_exposed) %>%
+        dplyr::filter(both_exposed > 0 & rain_only > 0 & flood_only > 0) %>%
+        tidyr::gather(classification, n_counties, -storm_id) %>%
+        ggplot(aes(x = n_counties)) + 
+        geom_histogram(bins = 30) + 
+        facet_wrap(~ classification, ncol = 1) + 
+        theme_tufte() + 
+        xlab("Number of counties classified as exposed") + 
+        ylab("Number of storms")
+```
+
+Here is a table showing, for all storms where 200 or more counties were classified as exposed based on at least one of the two metrics, what percent were classified in a way that agreed (i.e., classified as exposed by both metrics):
+
+```{r echo = FALSE}
+storm_exposed %>%
+        dplyr::mutate(any_exposed = both_exposed + rain_only + flood_only,
+                      both_of_any = round(100 * both_exposed /
+                                                  any_exposed)) %>%
+        dplyr::filter(any_exposed >= 200) %>%
+        dplyr::select(-neither_exposed, -any_exposed) %>%
+        dplyr::arrange(desc(both_of_any)) %>%
+        kable()
+```
+
+Only two storms (Irene in 2011 and Floyd and 1999) have over 50% agreement among counties classified as exposed by at least one of the two metrics.
+Throughout, in almost every storm, among the discordant counties (classified as exposed by one but not both metrics), there were many more counties classified as exposed based on the rain metric but where no flood events were reported in the NOAA Storm Events database.
+The only exceptions, where more discordant counties were classified as exposed based on flood than rain, were Allison (2001) and Cindy (2005).
+
 Here is a comparison of pair-wise metric agreement for all combinations of metrics.

### 2016-08-03 writing/DraftExposurePaper/example_alex_code.R

@@ -10,7 +10,6 @@ alex_fips <- c("48355", "48061", "48215", "48057", "12087")
 county_timeseries(alex_fips, percent_coverage = 0,
                   date_min = "1988-01-01", date_max = "2011-12-31",
                   var = "PRCP", out_directory = "~/alex_ex/"
-                  #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_ex/"
 )
 
 # Check that it worked...

### 2016-08-03 writing/DraftExposurePaper/example_alex_code.R
Remove `example_*_code.R` files (all in main text now)

@@ -1,71 +1,0 @@
-options("noaakey" = Sys.getenv("noaakey"))
-library(countyweather)
-library(hurricaneexposure)
-library(dplyr)
-library(ggplot2)
-library(ggthemes)
-library(lubridate)
-
-alex_fips <- c("48355", "48061", "48215", "48057", "12087")
-county_timeseries(alex_fips, percent_coverage = 0,
-                  date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP", out_directory = "~/alex_ex/"
-)
-
-# Check that it worked...
-list.files("~/alex_ex")
-#list.files("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_ex/")
-
-# Do some plots
-#alex_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_ex/"
-alex_dir <- "~/alex_ex/"
-alex_counties <- gsub(".rds", "", list.files(alex_dir))
-
-alex_ave <- vector("list", length = length(alex_counties))
-names(alex_ave) <- alex_counties
-
-for(x in alex_counties){
-        county_weather <- readRDS(paste0(alex_dir, x, ".rds"))#$averaged
-        county_weather <- county_weather %>%
-                filter(date %in% seq(from = as.Date("2010-06-01"), to = as.Date("2010-07-01"), by = 1))
-                       #year(date) == 2010 & month(date) == 6 | month(date) == 7)
-        county_weather$fips <- x
-        alex_ave[[x]] <- county_weather
-}
-alex_ave <- do.call("rbind", alex_ave)
-
-alex_rain <- county_rain(counties = alex_counties, start_year = 2010,
-                        end_year = 2010, rain_limit = 0, dist_limit = 1000) %>%
-        filter(storm_id == "Alex-2010")
-
-library(ggthemes)
-alex_ave %>%
-        ggplot(aes(x = date, y = prcp, color = prcp_reporting)) +
-        geom_line() + geom_point() +
-        geom_segment(data = alex_rain,
-                     aes(x = ymd(closest_date) - ddays(2),
-                         xend = ymd(closest_date) + ddays(1),
-                         y = tot_precip, yend = tot_precip), color = "red") +
-        facet_wrap(~ fips, ncol = 2) +
-        theme_few()
-# Note: different scales on this one-- cumulative rain over four days for
-# `alex_rain`, `alex_ave` is one day at a time
-
-alex_ave %>%
[…]

### 2016-08-03 writing/DraftExposurePaper/example_frances_code.R

@@ -1,72 +1,0 @@
-options("noaakey" = Sys.getenv("noaakey"))
-library(countyweather)
-library(hurricaneexposure)
-library(dplyr)
-library(ggplot2)
-library(ggthemes)
-library(lubridate)
-
-frances_fips <- c("12099", "12086", "13311", "37011", "12121")
-county_timeseries(frances_fips, percent_coverage = 0,
-                  date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP", #out_directory = "~/frances_ex/"
-                  out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_ex/"
-)
-
-# Check that it worked...
-list.files("~/frances_ex")
-#list.files("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_ex/")
-
-# Do some plots
-#frances_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_ex/"
-frances_dir <- "~/frances_ex/"
-frances_counties <- gsub(".rds", "", list.files(frances_dir))
-
-frances_ave <- vector("list", length = length(frances_counties))
-names(frances_ave) <- frances_counties
-
-for(x in frances_counties){
-        county_weather <- readRDS(paste0(frances_dir, x, ".rds"))#$averaged
-        county_weather <- county_weather %>%
-                filter(date %in% seq(from = as.Date("2004-08-20"), to = as.Date("2004-09-11"), by = 1))
-        #year(date) == 2010 & month(date) == 6 | month(date) == 7)
-        county_weather$fips <- x
-        frances_ave[[x]] <- county_weather
-}
-frances_ave <- do.call("rbind", frances_ave)
-
-frances_rain <- county_rain(counties = frances_counties, start_year = 2004,
-                         end_year = 2004, rain_limit = 0, dist_limit = 1000) %>%
-        filter(storm_id == "Frances-2004")
-
-library(ggthemes)
-frances_ave %>%
-        ggplot(aes(x = date, y = prcp, color = prcp_reporting)) +
-        geom_line() + geom_point() +
-        geom_segment(data = frances_rain,
-                     aes(x = ymd(closest_date) - ddays(2),
-                         xend = ymd(closest_date) + ddays(1),
-                         y = tot_precip, yend = tot_precip), color = "red") +
-        facet_wrap(~ fips, ncol = 2) +
-        theme_few()
[…]

### 2016-08-03 writing/DraftExposurePaper/example_ike_code.R

@@ -1,70 +1,0 @@
-options("noaakey" = Sys.getenv("noaakey"))
-library(countyweather)
-library(hurricaneexposure)
-library(dplyr)
-library(ggplot2)
-library(ggthemes)
-library(lubridate)
-
-ike_fips <- c("12021", "12071", "12087", "22051", "22053")
-county_timeseries(ike_fips, percent_coverage = 0,
-                  date_min = "1988-01-01", date_max = "2011-12-31",
-                  var = "PRCP", out_directory = "~/ike_ex/"
-                  #, out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_ex/"
-                  )
-
-# Check that it worked...
-list.files("~/ike_ex")
-#list.files("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_ex/")
-
-# Do some plots
-#ike_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_ex/"
-ike_dir <- "~/ike_ex/"
-ike_counties <- gsub(".rds", "", list.files(ike_dir))
-
-ike_ave <- vector("list", length = length(ike_counties))
-names(ike_ave) <- ike_counties
-
-for(x in ike_counties){
-        county_weather <- readRDS(paste0(ike_dir, x, ".rds"))#$averaged
-        county_weather <- county_weather %>%
-                filter(year(date) == 2008 & month(date) == 9)
-        county_weather$fips <- x
-        ike_ave[[x]] <- county_weather
-}
-ike_ave <- do.call("rbind", ike_ave)
-
-ike_rain <- county_rain(counties = ike_counties, start_year = 2008,
-                        end_year = 2008, rain_limit = 0, dist_limit = 1000) %>%
-        filter(storm_id == "Ike-2008")
-
-library(ggthemes)
-ike_ave %>%
-        ggplot(aes(x = date, y = prcp, color = prcp_reporting)) +
-        geom_line() + geom_point() +
-        geom_segment(data = ike_rain,
-                     aes(x = ymd(closest_date) - ddays(2),
-                         xend = ymd(closest_date) + ddays(1),
-                         y = tot_precip, yend = tot_precip), color = "red") +
-        facet_wrap(~ fips, ncol = 2) +
-        theme_few()
-# Note: different scales on this one-- cumulative rain over four days for
-# `ike_rain`, `ike_ave` is one day at a time
-
-ike_ave %>%
[…]

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Pull all available FIPS for precip for four storms

@@
 allison_fips <- sprintf("%05s", as.character(unique(allison_gauge$fips)))
-allison_fips <- sample(allison_fips, 10)
+# allison_fips <- sample(allison_fips, 10)
 date_range <- closest_dist %>% 
@@
 frances_fips <- sprintf("%05s", as.character(unique(frances_gauge$fips)))
-frances_fips <- sample(frances_fips, 10)
+# frances_fips <- sample(frances_fips, 10)
 date_range <- closest_dist %>% 
@@
 ike_fips <- sprintf("%05s", as.character(unique(ike_gauge$fips)))
-ike_fips <- sample(ike_fips, 10)
+# ike_fips <- sample(ike_fips, 10)
 date_range <- closest_dist %>% 
@@
 alex_fips <- sprintf("%05s", as.character(unique(alex_gauge$fips)))
-alex_fips <- sample(alex_fips, 10)
+# alex_fips <- sample(alex_fips, 10)
 date_range <- closest_dist %>% 
@@
                 geom_point(aes(size = prcp_reporting), alpha = 0.3) +
-                geom_text(aes(label = fips)) +
+                # geom_text(aes(label = fips)) +
                 theme_few() +

### 2016-08-03 writing/DraftExposurePaper/dade_storm_plot.R
revise code so does not require nasa data

@@ -20,46 +20,46 @@ check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
                       day_0 = day_0 + days(0),
                       day_b1 = day_0 - days(1),
                       day_b2 = day_0 - days(2),
-                      day_b3 = day_0 - days(3),
-                      day_b4 = day_0 - days(4),
-                      day_b5 = day_0 - days(5),
                       day_a1 = day_0 + days(1),
                       day_a2 = day_0 + days(2),
                       day_a3 = day_0 + days(3)) %>%
-        dplyr::select(storm_id, fips, day_b5, day_b4, day_b3, day_b2, day_b1,
+        dplyr::select(storm_id, fips, day_b2, day_b1,
                       day_0, day_a1, day_a2, day_a3) %>%
         tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
-        dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "12086") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
 
-all_dates <- unique(check_dates$day)
+all_dates <- unique(check_dates$date)
 all_fips <- unique(check_dates$fips) # has Miami as "12086", and is still in check_dates here
-all_fips <- c(all_fips, as.integer(12025))
-check_dates[check_dates$fips == 12086, "fips"] <- 12025
+#all_fips <- c(all_fips, as.integer(12025))
+#check_dates[check_dates$fips == 12086, "fips"] <- 12025
 
-dade_rain <- data.table::fread(
-        "/Users/joshuaferreri/Documents/nasa_precip_export_2.txt",
-        #"~/Documents/CSU2016/hurricaneproject/hurricaneexposuredata/data-raw/nasa_precip_export_2.txt",
-        header = TRUE,
-        select = c("county", "year_month_day", "precip", "precip_max")) %>%
-        dplyr::filter(county %in% all_fips,
-                      year_month_day %in% all_dates) %>%
-        dplyr::rename(fips = county, day = year_month_day) %>%
-        dplyr::right_join(data.table(check_dates),
-                          by = c("fips" = "fips", "day" = "day")) %>%
-        dplyr::filter(!is.na(precip) & !is.na(precip_max)) %>%
[…]

### 2016-08-03 writing/DraftExposurePaper/dade_storm_plot.R
Updated code for merging county_weather and nldas precip data for comparison plots

@@ -126,15 +126,14 @@ dade_weather <- dade_weather %>%
 #        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
 #        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade county by Storm")
 
-lag_sum <- function(counties = NULL, county_weather, df2, year_range = NULL,
+lag_sum <- function(counties = NULL, county_weather, year_range = NULL,
                               distance_limit = NULL, rain_limit = NULL,
                               include_rain = FALSE, days_included = NULL,
                               output_vars = c("fips", "cw_precip")){
 
-        county_weather <- as.data.frame(county_weather)
-        df <- inner_join(county_weather, df2, by = "date") %>%
-                arrange(storm_id)
-        df <- rename(df, fips = fips.x)
+#        df <- inner_join(county_weather, df2, by = "date") %>%
+#                arrange(storm_id)
+#        df <- rename(df, fips = fips.x)
 
         closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
 
@@ -161,7 +160,7 @@ lag_sum <- function(counties = NULL, county_weather, df2, year_range = NULL,
         }
 
         if(include_rain){
-                df <- data.table::data.table(df)
+                df <- data.table::data.table(county_weather)
                 df <- df[get("lag") %in% days_included]
                 df <- df[ , .(cw_precip = sum(get("prcp"))),
                               by = .(fips, storm_id)]
@@ -177,13 +176,12 @@ lag_sum <- function(counties = NULL, county_weather, df2, year_range = NULL,
         return(closest_dist)
 }
 
-countyweather_rain <- function(counties, county_weather, df2, start_year, end_year,
+countyweather_rain <- function(counties, county_weather, start_year, end_year,
                         rain_limit, dist_limit,
                         days_included = c(-2, -1, 0, 1)){
 
         county_rain_storm_df <- lag_sum(counties = counties,
                                            county_weather = county_weather,
-                                           df2 = df2,
                                            year_range = c(start_year, end_year),
                                            distance_limit = dist_limit,
                                            rain_limit = rain_limit,
@@ -214,7 +212,7 @@ countyweather_rain <- function(counties, county_weather, df2, start_year, end_ye
 }
 
 miami_rain <- countyweather_rain(counties = "12086", county_weather = dade_weather,
[…]

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work some on "total hits" maps

@@
                         gsub("*.+-", "", storms) >= 1988]
+
+# All eastern states (for mapping)
+eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
+                    "district of columbia", "florida", "georgia", "illinois",
+                    "indiana", "iowa", "kansas", "kentucky", "louisiana",
+                    "maine", "maryland", "massachusetts", "michigan",
+                    "mississippi", "missouri", "new hampshire", "new jersey",
+                    "new york", "north carolina", "ohio", "oklahoma",
+                    "pennsylvania", "rhode island", "south carolina",
+                    "tennessee", "texas", "vermont", "virginia",
+                    "west virginia", "wisconsin")
 ```
@@
 
-[Add in figures by county here for county averages of ground-based monitors versus our method.]
+[Add in figures by county here for county averages of ground-based monitors versus our method. (for example, Miami-Dade figure)]
 
@@
 
-Finally, we compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season (Figure \ref{fig:mwr_comparison}).
-
-```{r mwr_comparison, fig.width = 10, fig.height = 10, echo = FALSE, fig.align = "center", fig.cap = "Comparison of estimated storm-associated rainfall from our method (using NLDAS-2 data for a window from two days before to one day after the storm's closest approach to the county) with ground-based observations reported in the Monthly Weather Review report.
+Finally, we compared exposure results based on the NLDAS-2 precipitation data with values from monitors reported for certain storms in the *Monthly Weather Reviews* yearly summary of the Atlantic basin tropical storm season (Figure \ref{fig:mwr_rain_comparison}).
+
+```{r mwr_rain_comparison, fig.width = 10, fig.height = 10, echo = FALSE, fig.align = "center", fig.cap = "Comparison of estimated storm-associated rainfall from our method (using NLDAS-2 data for a window from two days before to one day after the storm's closest approach to the county) with ground-based observations reported in the Monthly Weather Review report.
 Observations within the same county are linked by a line segment.
@@
 
-Ike wind data:
-
-```{r}
-#Ike-2008 wind
+For four example storms, we also assessed how estimated wind speeds from our model compared to observed wind speeds reported in the *Monthy Weather Review* tropical cyclone reports (Figure \ref{fig:mwr_wind_comparison}).
+
+```{r mwr_wind_comparison, fig.width = 10, fig.height = 10, echo = FALSE, fig.align = "center", fig.cap = "Comparison of estimated storm sustained winds from our method (using a model based on Willoughby's model) with ground-based observations reported in the Monthly Weather Review report.
+Observations within the same county are linked by a line segment.
+The diagonal gray line shows where points would fall if assessments were identical.
+Horizontal and vertical gray lines show the 15 m / s threshold we use for wind-related storm exposure classification."}
+# Function to plot comparison between two datasets
+plot_mwr_compare_wind <- function(data){
+        storm_year <- as.numeric(gsub(".+-", "", data$storm_id[1]))
+        data$fips <- sprintf("%05s", data$fips)
+        storm_title <- paste(unlist(strsplit(as.character(data$storm_id[1]), 
+                                       split = "-")), collapse = ", ")
+
+        compare_wind <- county_wind(counties = unique(data$fips),
+                             start_year = storm_year, 
+                             end_year = storm_year,
+                             wind_limit = 0,
+                             wind_var = "max_sust") %>%
+                dplyr::filter(storm_id == data$storm_id[1]) %>%
+                dplyr::right_join(data, by = "fips")
+
+        county_range <- compare_wind %>% 
+                group_by(fips) %>%
+                summarize(lowest_wind = min(wind_sust),
+                  highest_wind = max(wind_sust),
+                  model_wind = mean(max_sust))
+
+        plot_mwr <- ggplot(compare_wind, 
+                   aes(x = max_sust, y = wind_sust)) + 
+        geom_hline(yintercept = 15, color = "lightgray") + 
+        geom_vline(xintercept = 15, color = "lightgray") + 
+                geom_abline(aes(intercept = 0, slope = 1), color = "gray") + 
+        geom_segment(data = county_range, 
+                     aes(x = model_wind, xend = model_wind,
+                         y = lowest_wind, yend = highest_wind,
+                         group = fips)) + 
+        geom_point(aes(group = NA), alpha = 0.5) + 
+        ggthemes::theme_few() + 
+        xlim(range(compare_wind$max_sust, compare_wind$wind_sust)) +
+        ylim(range(compare_wind$max_sust, compare_wind$wind_sust)) + 
+        ggtitle(storm_title) + 
+        xlab("Modeled sustained wind (m / s)") + 
+        ylab("Monthly Weather Report\n sustained wind reports (m / s)") + 
+        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
+                           guide = "none") 
+        return(plot_mwr)
+}
+
+#allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
+#allison_mwr <- plot_mwr_compare(allison_gauge)
+
+load("MWR_gauge_wind_data/frances_wind.RData")
+frances_wind <- frances_wind %>%
+        filter(!is.na(wind_sust)) %>%
+        mutate(wind_sust = convert_wind_speed(wind_sust, "knots", "mps"))
+frances_wind_mwr <- plot_mwr_compare_wind(frances_wind)
+
 load("MWR_gauge_wind_data/ike_wind.RData")
@@
         filter(!is.na(max_sust_kt)) %>%
-        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
-                                                 old_metric = "knots",
-                                                 new_metric = "mps"),
-               fips = sprintf("%05s", fips))
-
-ike_wind_model <- county_wind(counties = unique(ike_wind$fips),
-                              start_year = 2008, end_year = 2008,
-                              wind_limit = 0) %>%
-        filter(storm_id == "Ike-2008")
-
-compare_wind <- right_join(ike_wind_model, ike_wind, by = "fips")
-a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Ike, 2008")
-a
-
-county_range <- compare_wind %>% 
-        group_by(fips) %>%
-        summarize(lowest_wind = min(max_sust_mps),
-                  highest_wind = max(max_sust_mps),
-                  model_wind = mean(max_sust))
-b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
-        geom_segment(data = county_range, 
-                     aes(x = model_wind, xend = model_wind,
-                         y = lowest_wind, yend = highest_wind,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
-        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
-        ggtitle("Ike, 2008") + 
-        xlab("Modeled wind speed (m / s)") + 
-        ylab("MWR wind reports (m / s)")
-b
-```
-
-Frances wind data:
-
-```{r}
-load("MWR_gauge_wind_data/frances_wind.RData")
-frances_wind <- frances_wind %>%
-        filter(!is.na(wind_sust)) %>%
-        mutate(max_sust_mps = convert_wind_speed(wind_sust, 
-                                                 old_metric = "knots",
-                                                 new_metric = "mps"),
-               fips = sprintf("%05s", fips))
-
-frances_wind_model <- county_wind(counties = unique(frances_wind$fips),
-                              start_year = 2004, end_year = 2004,
-                              wind_limit = 0) %>%
-        filter(storm_id == "Frances-2004")
-
-compare_wind <- right_join(frances_wind_model, frances_wind, by = "fips")
-a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Frances, 2004")
-a
-
-county_range <- compare_wind %>% 
-        group_by(fips) %>%
-        summarize(lowest_wind = min(max_sust_mps),
-                  highest_wind = max(max_sust_mps),
-                  model_wind = mean(max_sust))
-b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
-        geom_segment(data = county_range, 
-                     aes(x = model_wind, xend = model_wind,
-                         y = lowest_wind, yend = highest_wind,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
-        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
-        ggtitle("Frances, 2004") + 
-        xlab("Modeled wind speed (m / s)") + 
-        ylab("MWR wind reports (m / s)")
-b
-```
-
-#### Are exposure patterns different when using distance-, rain-, and wind-based metrics?
+        mutate(wind_sust = convert_wind_speed(max_sust_kt, "knots", "mps"))
+ike_wind_mwr <- plot_mwr_compare_wind(ike_wind)
+
+# load("MWR_gauge_rain_data/alex_gauge.Rdata")
+# alex_mwr <- plot_mwr_compare(alex_gauge)
+
+# grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
+grid.arrange(frances_wind_mwr, ike_wind_mwr,
+             frances_wind_mwr, ike_wind_mwr, ncol = 2)
+```
+
+#### Are exposure patterns different when using different exposure metrics?
 In particular, (1) do some exposure metrics result in more "exposed" classifications than others and (2) are geographic patterns of exposure different among the different metrics?
 
-Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
-
-```{r echo = FALSE, warning=FALSE, message=FALSE}
-# Get distance-based exposure for all study counties, all years, using a 
-# distance limit that will ensure at least 1000 storm-county pairs
+We determined the average number of hits per year for each eastern US county for the exposure criteria based on distance (Figure \ref{fig:distance_hits}), rain (Figure \ref{fig:rain_hits}), and wind (Figure \ref{fig:wind_hits}), as well as the total number of hits based on storm-associated events listed in the NOAA Storm Events database for flood events (Figure \ref{fig:flood_hits}), wind events (Figure \ref{fig:wind_event_hits}), and tornado events (Figure \ref{fig:tornado_hits}) (all of these maps based on NOAA Storm Events are restricted to storms between 1996 and 2011).
+
+```{r echo = FALSE}
+# Map number of hits per county based on exposure data
+map_total_hits <- function(exposure_data, my_fips = my_fips, 
+                           eastern_states = eastern_states,
+                           title, start_year = 1988, end_year = 2011){
+        missing_fips <- my_fips[!(my_fips %in% exposure_data$fips)]
+        missing_counties <- data.frame(region = as.numeric(missing_fips),
+                                       count = 0)
+
+        exposure_count <- exposure_data %>%
+                dplyr::group_by(fips) %>%
+                dplyr::summarize(count = n() /(end_year - start_year + 1)) %>%
+                dplyr::rename(region = fips) %>%
+                dplyr::mutate(region = as.numeric(region),
+                              count = as.numeric(count)) %>%
+                rbind(missing_counties)
+
+        exposure_count$value <- cut(exposure_count$count, c(0, 
+                                                            0.11,
+                                                            0.22,
+                                                            0.33,
+                                                            0.44,
+                                                            0.56,
+                                                            0.67,
+                                                            0.78,
+                                                            0.89,
+                                                            10),
+                           labels = c("0-.10", ".11-.21",
+                                      ".22-.32", ".33-.43",
+                                      ".44-.55", ".56-.66",
+                                      ".67-.77", ".78-.88", ".89+"),
+                           include.lowest = TRUE)
+        exposure_palette <- c("#FFFFFF", brewer.pal(8, name = "YlGnBu"))
+
+        map_dist_count <- CountyChoropleth$new(exposure_count)
+        map_dist_count$set_zoom(eastern_states)
+        map_dist_count$ggplot_scale <- scale_fill_manual(
+                name = "Hits per\ncounty per\nyear",
+                values = exposure_palette,
+                na.value = "white")
+        map_dist_count$title <- title
+        a <- map_dist_count$render()
+
+        storms_to_plot <- exposure_data %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(n = n()) %>%
+                dplyr::arrange(desc(n)) %>%
+                slice(1:5)
+        # print(storms_to_plot)
+        out <- map_tracks(storms_to_plot$storm_id, plot_object = a,
+                          alpha = 0.75, plot_points = FALSE)
+        return(out)
+}
+```
+
+```{r distance_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on a criterion of the county being within 100 kilometers of the storm track, for storms from 1988 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Beryl, 1994: 346 counties; Frances, 2004: 321 counties; Cindy, 2005: 317 counties; Jeanne, 2004: 296 counties; and Dennis, 2005: 290 counties)."}
 distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
                                      end_year = 2011, dist_limit = 100)
-
-# distance_1000 <- distance_exposure %>%
-#         arrange(desc(storm_dist)) %>%
-#         group_by(fips) %>%
-#         summarise(count = n()) %>%
-#         rename(region = fips) %>%
-#         mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-#         arrange(desc(count)) %>%
-#         slice(1:1000)
-
+map_total_hits(distance_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on distance criterion")
+```
+
+```{r rain_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on a criteria of the county getting 75 millimeters or more of storm-associated rain in the window from two days before to one day after the storm's closest approach to the county and the county being within 500 kilometers of the storm track, for storms from 1988 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criteria over the same period (Frances, 2004: 464 counties; Ivan, 2004, 411 counties; Opal, 1995: 407 counties; Ike, 2008: 338 counties, Lee, 2011: 355 counties)."}
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500)
+map_total_hits(rain_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on rain criterion")
+```
+
+```{r wind_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on a criteria of the county experiencing storm-related sustained gusts of 15 m / s or more, for storms from 1988 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Bertha, 1996: 199 counties; Irene, 2011: 139 counties; Floyd, 1999: 135 counties; Isabel, 2003: 128 counties; Bob, 1991: 118 counties)."}
+wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2011, wind_limit = 15)
+map_total_hits(wind_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on wind criterion")
+```
+
+```{r flood_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated flood event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Floyd, 1999: 218 counties; Allison, 2001: 187 counties; Irene, 2011: 187 counties; Ike, 2008: 167 counties; Ivan, 2004: 153 counties)."}
+flood_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "flood")
+map_total_hits(flood_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on flood events",
+               start_year = 1996, end_year = 2011)
+```
+
+```{r wind_event_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated wind event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
+Light red lines show the storm tracks over the same period.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Allison, 2001: 435 counties; Arlene, 2005: 360 counties; Danny, 1997: 310 counties; Cindy, 2005: 189 counties; Cristobal, 2008: 183 counties)."}
+wind_event_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "wind")
+map_total_hits(wind_event_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on wind events")
+```
+
+```{r tornado_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated tornado event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
+Light red lines show the storm tracks over the same period.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Ivan, 2004: 95 counties; Frances, 2004: 72 counties; Rita, 2005: 52 counties; Katrina, 2005: 45 counties; Fay, 2008: 43 counties)."}
+tornado_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "tornado")
+map_total_hits(tornado_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on tornado events",
+               start_year = 1996, end_year = 2011)
+```
+
+Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
+
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 distance_2000 <- distance_exposure %>%
-        arrange(storm_dist) %>% # For distance, I don't think we want `desc`
-        slice(1:2000) %>% # I think we need to slice out the top 1000 before 
-                          # counting up by county. We want the top 1000
-                          # county-storm pairs, not to count up the numbers for
-                          # all storms within 100 km and then take the top 1000
-                          # counties from that (which is what the old code did)
+        arrange(storm_dist) %>% 
+        slice(1:2000) %>% 
         group_by(fips) %>%
@@
 exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
-                    "district of columbia", "florida", "georgia", "illinois",
-                    "indiana", "iowa", "kansas", "kentucky", "louisiana",
-                    "maine", "maryland", "massachusetts", "michigan",
-                    "mississippi", "missouri", "new hampshire", "new jersey",
-                    "new york", "north carolina", "ohio", "oklahoma",
-                    "pennsylvania", "rhode island", "south carolina",
-                    "tennessee", "texas", "vermont", "virginia",
-                    "west virginia", "wisconsin")
 
@@
 
-# rain_1000 <- rain_exposure %>%
-#         arrange(desc(tot_precip)) %>%
-#         group_by(fips) %>%
-#         summarise(count = n()) %>%
-#         rename(region = fips) %>%
-#         mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-#         arrange(desc(count)) %>%
-#         slice(1:1000)
 rain_2000 <- rain_exposure %>%
@@
 
-```{r warning = FALSE, message = FALSE, echo = FALSE}
-#Top 2000 storm-county pairs for flooding events during tropical cyclones (since 1996)
-flood_exposure <- county_events(counties = my_fips,
-                              start_year = 1996, end_year = 2011,
-                              event_type = "flood")
-flood_2000 <- flood_exposure %>%
-        slice(1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count))
-flood_2000$value <- cut(flood_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-map_flood_2000 <- CountyChoropleth$new(flood_2000)
-map_flood_2000$set_zoom(eastern_states)
-map_flood_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_flood_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Flood Events"
-c <- map_flood_2000$render()
-```
-
-```{r}
-grid.arrange(b, c, ncol = 1)
-```
-
-Map of county exposures to the top 1000 county-storm pairs by wind-based exposure:
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
+Map of county exposures to the top 2,000 county-storm pairs by wind-based exposure:
+
+```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
 #TS winds and above
@@
                              wind_var = "max_sust")
-#HU exposures (32.1 mps and above)
-HU_wind_exposure <- county_wind(counties = my_fips,
-                                start_year = 1988,
-                                end_year = 2011,
-                                wind_limit = 32.1,
-                                wind_var = "max_sust")
+
 TS_wind_2000 <- TS_wind_exposure %>% 
@@
 TS_wind_2000 <- subset(TS_wind_2000, !is.na(TS_wind_2000$region)) # Remove NA observation
-HU_wind_2000 <- HU_wind_exposure %>% 
-        arrange(desc(max_sust)) %>%
-        slice (1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) # Max of 1 again for hits for top Storm-
                                                                        # County pairs
@@
 map_TS_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor TS Wind Metric"
-a <- map_TS_2000$render()
-HU_wind_2000$value <- cut(HU_wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-map_HU_2000 <- CountyChoropleth$new(HU_wind_2000)
-map_HU_2000$set_zoom(eastern_states)
-map_HU_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_HU_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor HU Wind Metric"
-b <- map_HU_2000$render()
-```
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-#Category 3 and above 
-cat3_exposure <- county_wind(counties = my_fips,
-                                start_year = 1988,
-                                end_year = 2011,
-                                wind_limit = 50,
-                                wind_var = "max_sust")
-cat3_2000 <- cat3_exposure %>% 
-        arrange(desc(max_sust)) %>%
-        slice (1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count))
-cat3_2000$value <- cut(cat3_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-map_cat3_2000 <- CountyChoropleth$new(cat3_2000)
-map_cat3_2000$set_zoom(eastern_states)
-map_cat3_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_cat3_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Category 3 and Above Wind Metric"
-c <- map_cat3_2000$render()
-```
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE}
-grid.arrange(a, b, c, ncol = 1)
-```
-
-```{r warning = FALSE, message = FALSE, echo = FALSE}
-#Top 2000 storm-county pairs for tornado events during tropical cyclones
-tornado_exposure <- county_events(counties = my_fips,
-                             start_year = 1988, end_year = 2011,
-                             event_type = "tornado")
-tornado_2000 <- tornado_exposure %>%
-        slice(1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count))
-tornado_2000$value <- cut(tornado_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-map_tornado_2000 <- CountyChoropleth$new(tornado_2000)
-map_tornado_2000$set_zoom(eastern_states)
-map_tornado_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Tornado Events"
-a <- map_tornado_2000$render()
-```
-
-```{r warning = FALSE, message = FALSE, echo = FALSE}
-#Top 2000 storm-county pairs for wind events during tropical cyclones
-wind_exposure <- county_events(counties = my_fips,
-                              start_year = 1988, end_year = 2011,
-                              event_type = "wind")
-wind_2000 <- wind_exposure %>%
-        slice(1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count))
-#Or
-#wind_2000 <- wind_exposure %>%
-        #group_by(fips) %>%
-        #summarise(count = n()) %>%
-        #arrange(desc(count))
-        #slice(1:2000)
-        #rename(region = fips) %>%
-        #mutate(region = as.numeric(region), count = as.numeric(count))
-wind_2000$value <- cut(wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-map_wind_2000 <- CountyChoropleth$new(wind_2000)
-map_wind_2000$set_zoom(eastern_states)
-map_wind_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_tornado_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Wind Events"
-b <- map_wind_2000$render()
-```
-
-```{r}
-grid.arrange(a, b, ncol = 1)
-```
-
-Here are a map of total count of "exposures" by county, based on "exposed" = distance of 100 [or do we want a different distance limit for this?
-What seems reasonable?; JMF-- Based on what I have read so far, I think this seems reasonable, unless we want to push it out to 120 km to loosely reflect the Czajkowski study which seems to have found a great number of fatalities in tertiary counties] kilometers or less and a map of total count of "exposures" by county, based on "exposed" = rainfall over a three-day window of 75 millimeters or more and distance of 500 kilometers or less:
-
-```{r echo = FALSE, warning=FALSE, message=FALSE}
+map_TS_2000$render()
+```
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
 distance_count <- distance_exposure %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) 
-distance_count$value <- cut(distance_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24, 26),
-                           labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24", "25-26"),
-                           include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_dist_count <- CountyChoropleth$new(distance_count)
-map_dist_count$set_zoom(eastern_states)
-map_dist_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_dist_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance <= 100 km"
-a <- map_dist_count$render()
-```
-
-```{r echo = FALSE, warning=FALSE, message=FALSE}
+                dplyr::group_by(fips) %>%
+                dplyr::summarize(count = n()) %>%
+                dplyr::rename(region = fips) %>%
+                dplyr::mutate(region = as.numeric(region),
+                              count = as.numeric(count))
 rain_count <- rain_exposure %>%
-              group_by(fips) %>%
-              summarise(count = n()) %>%
-              rename(region = fips) %>%
-              mutate(region = as.numeric(region), count = as.numeric(count))
-rain_count$value <- cut(rain_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24, 26),
-                           labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24", "25-26"),
-                           include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_rain_count <- CountyChoropleth$new(rain_count)
-map_rain_count$set_zoom(eastern_states)
-map_rain_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_rain_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain > 75 mm and Distance <= 500 km"
-b <- map_rain_count$render()
-```
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-wind_count <- TS_wind_exposure %>%
-              group_by(fips) %>%
-              summarise(count = n()) %>%
-              rename(region = fips) %>%
-              mutate(region = as.numeric(region), count = as.numeric(count)) #Same size df as before (only 474 obs)
-wind_count$value <- cut(wind_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24, 26),
-                           labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24", "25-26"),
-                           include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_wind_count <- CountyChoropleth$new(wind_count)
-map_wind_count$set_zoom(eastern_states)
-map_wind_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_wind_count$title <- "Hits for top 2,000 Storm-County Pairs\nfor Wind > 18 m/s"
-c <- map_wind_count$render()
-```
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE}
-grid.arrange(a, b, c, ncol = 1)
-```
-
-```{r warning = FALSE, message = FALSE, echo = FALSE}
-tornado_count <- tornado_exposure %>%
-              group_by(fips) %>%
-              summarise(count = n()) %>%
-              rename(region = fips) %>%
-              mutate(region = as.numeric(region), count = as.numeric(count))
-tornado_count$value <- cut(tornado_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24, 26),
-                           labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24", "25-26"),
-                           include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_tornado_count <- CountyChoropleth$new(tornado_count)
-map_tornado_count$set_zoom(eastern_states)
-map_tornado_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_tornado_count$title <- "Total Hits for Storm-County Pairs\nfor Tornado Events"
-d <- map_tornado_count$render()
-```
-
-```{r warning = FALSE, message = FALSE, echo = FALSE}
+                dplyr::group_by(fips) %>%
+                dplyr::summarize(count = n()) %>%
+                dplyr::rename(region = fips) %>%
+                dplyr::mutate(region = as.numeric(region),
+                              count = as.numeric(count))
+
 wind_count <- wind_exposure %>%
-              group_by(fips) %>%
-              summarise(count = n()) %>%
-              rename(region = fips) %>%
-              mutate(region = as.numeric(region), count = as.numeric(count))
-wind_count$value <- cut(wind_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24, 26),
-                           labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24", "25-26"),
-                           include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_wind_count <- CountyChoropleth$new(wind_count)
-map_wind_count$set_zoom(eastern_states)
-map_wind_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_wind_count$title <- "Total Hits for Storm-County Pairs\nfor Wind Events"
-e <- map_wind_count$render()
-```
-
-```{r warning = FALSE, message = FALSE, echo = FALSE}
-flood_count <- flood_exposure %>%
-              group_by(fips) %>%
-              summarise(count = n()) %>%
-              rename(region = fips) %>%
-              mutate(region = as.numeric(region), count = as.numeric(count))
-flood_count$value <- cut(flood_count$count, c(0, 5, 8, 12, 15,
-                                                    18, 21, 24, 26),
-                           labels = c("0-5", "6-8", "9-12", "13-15",
-                                      "16-18", "19-21", "22-24", "25-26"),
-                           include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_flood_count <- CountyChoropleth$new(flood_count)
-map_flood_count$set_zoom(eastern_states)
-map_flood_count$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_wind_count$title <- "Total Hits for Storm-County Pairs\nfor Flood Events"
-f <- map_flood_count$render()
-```
-
-```{r}
-#Event counts
-grid.arrange(d, e, f, ncol = 1)
-```
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
+                dplyr::group_by(fips) %>%
+                dplyr::summarize(count = n()) %>%
+                dplyr::rename(region = fips) %>%
+                dplyr::mutate(region = as.numeric(region),
+                              count = as.numeric(count))
+
 exposure_name <- c("Distance_Exposure", "Rain_Exposure", "Wind_Exposure")
@@
 ```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Ivan-2004"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Ivan, 2004")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Ivan, 2004")
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Isidore-2002"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Isidore, 2002")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Isidore, 2002")
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
 storm <- "Floyd-1999"
@@
 
-```{r echo = FALSE}
+```{r echo = FALSE, eval = FALSE}
 all_storm_counties <- closest_dist %>%
@@
 
-```{r echo = FALSE, fig.width = 12, fig.height = 12}
+```{r echo = FALSE, fig.width = 12, fig.height = 12, eval = FALSE}
 a <- agreement_hist(all_exposed, "distance", "rain")

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
add miami plot to draft

--- 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 
+Compare data from `countyweather` and NLDAS-2 data for Miami-Dade county by storm:
+
+```{r}
+data("closest_dist")
+
+# Create a new dataset for precip with lag -5 to 3 for fips 12086 only
+
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "12086") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips) # has Miami as "12086", and is still in check_dates here
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "12086"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "dade_data/")
+#                   #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
+
+# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
+dade_dir <- "dade_data/"
+dade_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
+
+dade_weather <- dade_weather$averaged %>%
+        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+dade_weather$fips <- fips
+dade_weather <- as.data.frame(dade_weather)
+
+dade_weather <- dade_weather %>%
+        dplyr::filter(date %in% all_dates) %>%
+        dplyr::right_join(data.table(check_dates),
+                          by = c("fips" = "fips", "date" = "date")) %>%
+        dplyr::filter(!is.na(prcp)) %>%
+        dplyr::arrange(storm_id, fips) %>%
+        dplyr::mutate(fips = as.integer(fips)) %>%
+        dplyr::mutate(fips = sprintf("%05d", fips),
+                      lag = gsub("day_", "", lag),
+                      lag = gsub("b", "-", lag),
+                      lag = gsub("a", "", lag),
+                      lag = as.numeric(lag))
+
+lag_sum <- function(counties = NULL, county_weather, year_range = NULL,
+                              distance_limit = NULL, rain_limit = NULL,
+                              include_rain = FALSE, days_included = NULL,
+                              output_vars = c("fips", "cw_precip")){
+
+        closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
+
+        if(!is.null(counties)){
+                closest_dist <- closest_dist[get("fips") %in% counties]
+        }
+
+        if(!is.null(year_range)){
+                closest_dist <- closest_dist[ , .(storm_id,
+                                                  fips,
+                                                  closest_date, storm_dist,
+                                                  local_time,
+                                                  closest_time_utc,
+                                                  year = substring(get("closest_date"),
+                                                                   1, 4)), ][
+                                                                           get("year") %in%
+                                                                                   year_range[1]:year_range[2]
+                                                                           ]
+        }
+
+        if(!is.null(distance_limit)){
+                closest_dist <- closest_dist[get("storm_dist") <=
+                                                     distance_limit]
+        }
+
+        if(include_rain){
+                df <- data.table::data.table(county_weather)
+                df <- df[get("lag") %in% days_included]
+                df <- df[ , .(cw_precip = sum(get("prcp"))),
+                              by = .(fips, storm_id)]
+                closest_dist <- merge(closest_dist, df, all.x = TRUE,
+                                      by = c("storm_id", "fips"))
+                if(!is.null(rain_limit)){
+                        closest_dist <- closest_dist[get("cw_precip") >=
+                                                             rain_limit]
+                }
+        }
+
+        closest_dist <- closest_dist[ , .SD, , .SDcols = output_vars]
+        return(closest_dist)
+}
+
+countyweather_rain <- function(counties, county_weather, start_year, end_year,
+                        rain_limit, dist_limit,
+                        days_included = c(-2, -1, 0, 1)){
+
+        county_rain_storm_df <- lag_sum(counties = counties,
+                                           county_weather = county_weather,
+                                           year_range = c(start_year, end_year),
+                                           distance_limit = dist_limit,
+                                           rain_limit = rain_limit,
+                                           include_rain = TRUE,
+                                           days_included = days_included,
+                                           output_vars = c("storm_id", "fips",
+                                                           "closest_date",
+                                                           "storm_dist",
+                                                           "cw_precip",
+                                                           "local_time",
+                                                           "closest_time_utc"))
+        rain_storm_df <- filter_storm_data(counties = counties,
+                                           year_range = c(start_year, end_year),
+                                           distance_limit = dist_limit,
+                                           rain_limit = rain_limit,
+                                           include_rain = TRUE,
+                                           days_included = days_included,
+                                           output_vars = c("storm_id", "fips",
+                                                           "closest_date",
+                                                           "storm_dist",
+                                                           "tot_precip",
+                                                           "local_time",
+                                                           "closest_time_utc"))
+        rain_df <- merge(county_rain_storm_df, rain_storm_df, all.x = TRUE,
+                         by = c("fips", "storm_id", "closest_date", "storm_dist",
+                                "local_time", "closest_time_utc"))
+        return(rain_df)
+}
+
+miami_rain <- countyweather_rain(counties = "12086", county_weather = dade_weather,
+                                 start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 1000)
+miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade County, FL, by storm")
+
+miami_plot
+```
+
 #### Are exposure patterns different when using distance-, rain-, and wind-based metrics?

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
edit text

@@
 
-# Create a new dataset for precip with lag -5 to 3 for fips 12086 only
+# Create a new dataset for precip with lag -2 to 3 for fips 12086 only
 
@@
 all_dates <- unique(check_dates$date)
-all_fips <- unique(check_dates$fips) # has Miami as "12086", and is still in check_dates here
+all_fips <- unique(check_dates$fips)
 

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
miami dade code

@@
                                  start_year = 1988, end_year = 2011,
-                                 rain_limit = 0, dist_limit = 1000)
+                                 rain_limit = 0, dist_limit = 2000)
 miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +

### 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
resolve merge conflict

@@
 
-Ike wind data:
-
-```{r}
-#Ike-2008 wind
-load("MWR_gauge_wind_data/ike_wind.RData")
-ike_wind <- ike_wind %>%
-        filter(!is.na(max_sust_kt)) %>%
-        mutate(max_sust_mps = convert_wind_speed(max_sust_kt, 
-                                                 old_metric = "knots",
-                                                 new_metric = "mps"),
-               fips = sprintf("%05s", fips))
-
-ike_wind_model <- county_wind(counties = unique(ike_wind$fips),
-                              start_year = 2008, end_year = 2008,
-                              wind_limit = 0) %>%
-        filter(storm_id == "Ike-2008")
-
-compare_wind <- right_join(ike_wind_model, ike_wind, by = "fips")
-a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Ike, 2008")
-a
+For four example storms, we also assessed how estimated wind speeds from our model compared to observed wind speeds reported in the *Monthy Weather Review* tropical cyclone reports (Figure \ref{fig:mwr_wind_comparison}).
+
+```{r mwr_wind_comparison, fig.width = 10, fig.height = 10, echo = FALSE, fig.align = "center", fig.cap = "Comparison of estimated storm sustained winds from our method (using a model based on Willoughby's model) with ground-based observations reported in the Monthly Weather Review report.
+Observations within the same county are linked by a line segment.
+The diagonal gray line shows where points would fall if assessments were identical.
+Horizontal and vertical gray lines show the 15 m / s threshold we use for wind-related storm exposure classification."}
+# Function to plot comparison between two datasets
+plot_mwr_compare_wind <- function(data){
+        storm_year <- as.numeric(gsub(".+-", "", data$storm_id[1]))
+        data$fips <- sprintf("%05s", data$fips)
+        storm_title <- paste(unlist(strsplit(as.character(data$storm_id[1]), 
+                                       split = "-")), collapse = ", ")
+
+        compare_wind <- county_wind(counties = unique(data$fips),
+                             start_year = storm_year, 
+                             end_year = storm_year,
+                             wind_limit = 0,
+                             wind_var = "max_sust") %>%
+                dplyr::filter(storm_id == data$storm_id[1]) %>%
+                dplyr::right_join(data, by = "fips")
 
@@
                 group_by(fips) %>%
-        summarize(lowest_wind = min(max_sust_mps),
-                  highest_wind = max(max_sust_mps),
+                summarize(lowest_wind = min(wind_sust),
+                  highest_wind = max(wind_sust),
                   model_wind = mean(max_sust))
-b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
+
+        plot_mwr <- ggplot(compare_wind, 
+                   aes(x = max_sust, y = wind_sust)) + 
+        geom_hline(yintercept = 15, color = "lightgray") + 
+        geom_vline(xintercept = 15, color = "lightgray") + 
+                geom_abline(aes(intercept = 0, slope = 1), color = "gray") + 
         geom_segment(data = county_range, 
@@
                          y = lowest_wind, yend = highest_wind,
-                         group = fips), color = "gray") + 
+                         group = fips)) + 
         geom_point(aes(group = NA), alpha = 0.5) + 
         ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
-        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
-        ggtitle("Ike, 2008") + 
-        xlab("Modeled wind speed (m / s)") + 
-        ylab("MWR wind reports (m / s)")
-b
-```
-
-Frances wind data:
-
-```{r}
+        xlim(range(compare_wind$max_sust, compare_wind$wind_sust)) +
+        ylim(range(compare_wind$max_sust, compare_wind$wind_sust)) + 
+        ggtitle(storm_title) + 
+        xlab("Modeled sustained wind (m / s)") + 
+        ylab("Monthly Weather Report\n sustained wind reports (m / s)") + 
+        scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "black"), 
+                           guide = "none") 
+        return(plot_mwr)
+}
+
+#allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
+#allison_mwr <- plot_mwr_compare(allison_gauge)
+
 load("MWR_gauge_wind_data/frances_wind.RData")
@@
         filter(!is.na(wind_sust)) %>%
-        mutate(max_sust_mps = convert_wind_speed(wind_sust, 
-                                                 old_metric = "knots",
-                                                 new_metric = "mps"),
-               fips = sprintf("%05s", fips))
-
-frances_wind_model <- county_wind(counties = unique(frances_wind$fips),
-                              start_year = 2004, end_year = 2004,
-                              wind_limit = 0) %>%
-        filter(storm_id == "Frances-2004")
-
-compare_wind <- right_join(frances_wind_model, frances_wind, by = "fips")
-a <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps,
-                              color = fips)) +
-        geom_point() + 
-        theme_bw() + 
-        ggtitle("Frances, 2004")
-a
-
-county_range <- compare_wind %>% 
-        group_by(fips) %>%
-        summarize(lowest_wind = min(max_sust_mps),
-                  highest_wind = max(max_sust_mps),
-                  model_wind = mean(max_sust))
-b <- ggplot(compare_wind, aes(x = max_sust, y = max_sust_mps)) + 
-        geom_segment(data = county_range, 
-                     aes(x = model_wind, xend = model_wind,
-                         y = lowest_wind, yend = highest_wind,
-                         group = fips), color = "gray") + 
-        geom_point(aes(group = NA), alpha = 0.5) + 
-        ggthemes::theme_few() + 
-        geom_abline(aes(intercept = 0, slope = 1)) + 
-        xlim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) +
-        ylim(range(compare_wind$max_sust, compare_wind$max_sust_mps)) + 
-        ggtitle("Frances, 2004") + 
-        xlab("Modeled wind speed (m / s)") + 
-        ylab("MWR wind reports (m / s)")
-b
+        mutate(wind_sust = convert_wind_speed(wind_sust, "knots", "mps"))
+frances_wind_mwr <- plot_mwr_compare_wind(frances_wind)
 ```
@@
 
-#### Are exposure patterns different when using distance-, rain-, and wind-based metrics?
+```{r}
+load("MWR_gauge_wind_data/ike_wind.RData")
+ike_wind <- ike_wind %>%
+        filter(!is.na(max_sust_kt)) %>%
+        mutate(wind_sust = convert_wind_speed(max_sust_kt, "knots", "mps"))
+ike_wind_mwr <- plot_mwr_compare_wind(ike_wind)
+
+# load("MWR_gauge_rain_data/alex_gauge.Rdata")
+# alex_mwr <- plot_mwr_compare(alex_gauge)
+
+# grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
+grid.arrange(frances_wind_mwr, ike_wind_mwr,
+             frances_wind_mwr, ike_wind_mwr, ncol = 2)
+```
+
+#### Are exposure patterns different when using different exposure metrics?
 In particular, (1) do some exposure metrics result in more "exposed" classifications than others and (2) are geographic patterns of exposure different among the different metrics?
 
-Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
-
-```{r echo = FALSE, warning=FALSE, message=FALSE}
-# Get distance-based exposure for all study counties, all years, using a 
-# distance limit that will ensure at least 1000 storm-county pairs
+We determined the average number of hits per year for each eastern US county for the exposure criteria based on distance (Figure \ref{fig:distance_hits}), rain (Figure \ref{fig:rain_hits}), and wind (Figure \ref{fig:wind_hits}), as well as the total number of hits based on storm-associated events listed in the NOAA Storm Events database for flood events (Figure \ref{fig:flood_hits}), wind events (Figure \ref{fig:wind_event_hits}), and tornado events (Figure \ref{fig:tornado_hits}) (all of these maps based on NOAA Storm Events are restricted to storms between 1996 and 2011).
+
+```{r echo = FALSE}
+# Map number of hits per county based on exposure data
+map_total_hits <- function(exposure_data, my_fips = my_fips, 
+                           eastern_states = eastern_states,
+                           title, start_year = 1988, end_year = 2011){
+        missing_fips <- my_fips[!(my_fips %in% exposure_data$fips)]
+        missing_counties <- data.frame(region = as.numeric(missing_fips),
+                                       count = 0)
+
+        exposure_count <- exposure_data %>%
+                dplyr::group_by(fips) %>%
+                dplyr::summarize(count = n() /(end_year - start_year + 1)) %>%
+                dplyr::rename(region = fips) %>%
+                dplyr::mutate(region = as.numeric(region),
+                              count = as.numeric(count)) %>%
+                rbind(missing_counties)
+
+        exposure_count$value <- cut(exposure_count$count, c(0, 
+                                                            0.11,
+                                                            0.22,
+                                                            0.33,
+                                                            0.44,
+                                                            0.56,
+                                                            0.67,
+                                                            0.78,
+                                                            0.89,
+                                                            10),
+                           labels = c("0-.10", ".11-.21",
+                                      ".22-.32", ".33-.43",
+                                      ".44-.55", ".56-.66",
+                                      ".67-.77", ".78-.88", ".89+"),
+                           include.lowest = TRUE)
+        exposure_palette <- c("#FFFFFF", brewer.pal(8, name = "YlGnBu"))
+
+        map_dist_count <- CountyChoropleth$new(exposure_count)
+        map_dist_count$set_zoom(eastern_states)
+        map_dist_count$ggplot_scale <- scale_fill_manual(
+                name = "Hits per\ncounty per\nyear",
+                values = exposure_palette,
+                na.value = "white")
+        map_dist_count$title <- title
+        a <- map_dist_count$render()
+
+        storms_to_plot <- exposure_data %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(n = n()) %>%
+                dplyr::arrange(desc(n)) %>%
+                slice(1:5)
+        # print(storms_to_plot)
+        out <- map_tracks(storms_to_plot$storm_id, plot_object = a,
+                          alpha = 0.75, plot_points = FALSE)
+        return(out)
+}
+```
+
+```{r distance_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on a criterion of the county being within 100 kilometers of the storm track, for storms from 1988 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Beryl, 1994: 346 counties; Frances, 2004: 321 counties; Cindy, 2005: 317 counties; Jeanne, 2004: 296 counties; and Dennis, 2005: 290 counties)."}
 distance_exposure <- county_distance(counties = my_fips, start_year = 1988,
                                      end_year = 2011, dist_limit = 100)
-
-# distance_1000 <- distance_exposure %>%
-#         arrange(desc(storm_dist)) %>%
-#         group_by(fips) %>%
-#         summarise(count = n()) %>%
-#         rename(region = fips) %>%
-#         mutate(region = as.numeric(region), count = as.numeric(count)) %>%
-#         arrange(desc(count)) %>%
-#         slice(1:1000)
-
+map_total_hits(distance_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on distance criterion")
+```
+
+```{r rain_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on a criteria of the county getting 75 millimeters or more of storm-associated rain in the window from two days before to one day after the storm's closest approach to the county and the county being within 500 kilometers of the storm track, for storms from 1988 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criteria over the same period (Frances, 2004: 464 counties; Ivan, 2004, 411 counties; Opal, 1995: 407 counties; Ike, 2008: 338 counties, Lee, 2011: 355 counties)."}
+rain_exposure <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 75, dist_limit = 500)
+map_total_hits(rain_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on rain criterion")
+```
+
+```{r wind_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on a criteria of the county experiencing storm-related sustained gusts of 15 m / s or more, for storms from 1988 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Bertha, 1996: 199 counties; Irene, 2011: 139 counties; Floyd, 1999: 135 counties; Isabel, 2003: 128 counties; Bob, 1991: 118 counties)."}
+wind_exposure <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2011, wind_limit = 15)
+map_total_hits(wind_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on wind criterion")
+```
+
+```{r flood_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated flood event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Floyd, 1999: 218 counties; Allison, 2001: 187 counties; Irene, 2011: 187 counties; Ike, 2008: 167 counties; Ivan, 2004: 153 counties)."}
+flood_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "flood")
+map_total_hits(flood_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on flood events",
+               start_year = 1996, end_year = 2011)
+```
+
+```{r wind_event_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated wind event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
+Light red lines show the storm tracks over the same period.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Allison, 2001: 435 counties; Arlene, 2005: 360 counties; Danny, 1997: 310 counties; Cindy, 2005: 189 counties; Cristobal, 2008: 183 counties)."}
+wind_event_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "wind")
+map_total_hits(wind_event_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on wind events")
+```
+
+```{r tornado_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated tornado event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
+Light red lines show the storm tracks over the same period.
+Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Ivan, 2004: 95 counties; Frances, 2004: 72 counties; Rita, 2005: 52 counties; Katrina, 2005: 45 counties; Fay, 2008: 43 counties)."}
+tornado_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "tornado")
+map_total_hits(tornado_exposure, my_fips = my_fips, 
+               eastern_states = eastern_states,
+               title = "Hits per county based on tornado events",
+               start_year = 1996, end_year = 2011)
+```
+
+Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
+
+```{r echo = FALSE, warning=FALSE, message=FALSE}
 distance_2000 <- distance_exposure %>%
-        arrange(storm_dist) %>% # For distance, I don't think we want `desc`
-        slice(1:2000) %>% # I think we need to slice out the top 1000 before 
-                          # counting up by county. We want the top 1000
-                          # county-storm pairs, not to count up the numbers for
-                          # all storms within 100 km and then take the top 1000
-                          # counties from that (which is what the old code did)
+        arrange(storm_dist) %>% 
+        slice(1:2000) %>% 
         group_by(fips) %>%

### 2016-08-04 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add comparison for orleans, fulton, floyd, and charleston counties

@@
 
+Compare data from `countyweather` and NLDAS-2 data for Orleans perish by storm:
+
+```{r}
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "22071") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips)
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "22071"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "orleans_data/")
+
+# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
+orleans_dir <- "orleans_data/"
+orleans_weather <- readRDS(paste0(orleans_dir, fips, ".rds"))
+
+#orleans_weather <- orleans_weather$averaged %>%
+#        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+orleans_weather <- orleans_weather %>%
+        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+orleans_weather$fips <- fips
+orleans_weather <- as.data.frame(orleans_weather)
+
+orleans_weather <- orleans_weather %>%
+        dplyr::filter(date %in% all_dates) %>%
+        dplyr::right_join(data.table(check_dates),
+                          by = c("fips" = "fips", "date" = "date")) %>%
+        dplyr::filter(!is.na(prcp)) %>%
+        dplyr::arrange(storm_id, fips) %>%
+        dplyr::mutate(fips = as.integer(fips)) %>%
+        dplyr::mutate(fips = sprintf("%05d", fips),
+                      lag = gsub("day_", "", lag),
+                      lag = gsub("b", "-", lag),
+                      lag = gsub("a", "", lag),
+                      lag = as.numeric(lag))
+
+orleans_rain <- countyweather_rain(counties = fips, county_weather = orleans_weather,
+                                 start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 2000)
+orleans_plot <- ggplot(orleans_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Orleans perish, LA, by storm")
+
+orleans_plot
+```
+
+Compare data from `countyweather` and NLDAS-2 data for Charleston county by storm:
+
+```{r}
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "45019") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips)
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "45019"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "charleston_data/")
+
+# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
+charleston_dir <- "charleston_data/"
+charleston_weather <- readRDS(paste0(charleston_dir, fips, ".rds"))
+
+#orleans_weather <- orleans_weather$averaged %>%
+#        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+charleston_weather <- charleston_weather %>%
+        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+charleston_weather$fips <- fips
+charleston_weather <- as.data.frame(charleston_weather)
+
+charleston_weather <- charleston_weather %>%
+        dplyr::filter(date %in% all_dates) %>%
+        dplyr::right_join(data.table(check_dates),
+                          by = c("fips" = "fips", "date" = "date")) %>%
+        dplyr::filter(!is.na(prcp)) %>%
+        dplyr::arrange(storm_id, fips) %>%
+        dplyr::mutate(fips = as.integer(fips)) %>%
+        dplyr::mutate(fips = sprintf("%05d", fips),
+                      lag = gsub("day_", "", lag),
+                      lag = gsub("b", "-", lag),
+                      lag = gsub("a", "", lag),
+                      lag = as.numeric(lag))
+
+charleston_rain <- countyweather_rain(counties = fips, county_weather = charleston_weather,
+                                 start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 2000)
+charleston_plot <- ggplot(charleston_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Charleston county, SC, by storm")
+
+charleston_plot
+```
+
+Compare data from `countyweather` and NLDAS-2 data for Floyd county by storm:
+
+```{r}
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "51063") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips)
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "51063"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "floyd_data/")
+
+floyd_dir <- "floyd_data/"
+floyd_weather <- readRDS(paste0(floyd_dir, fips, ".rds"))
+
+#orleans_weather <- orleans_weather$averaged %>%
+#        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+floyd_weather <- floyd_weather %>%
+        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+floyd_weather$fips <- fips
+floyd_weather <- as.data.frame(floyd_weather)
+
+floyd_weather <- floyd_weather %>%
+        dplyr::filter(date %in% all_dates) %>%
+        dplyr::right_join(data.table(check_dates),
+                          by = c("fips" = "fips", "date" = "date")) %>%
+        dplyr::filter(!is.na(prcp)) %>%
+        dplyr::arrange(storm_id, fips) %>%
+        dplyr::mutate(fips = as.integer(fips)) %>%
+        dplyr::mutate(fips = sprintf("%05d", fips),
+                      lag = gsub("day_", "", lag),
+                      lag = gsub("b", "-", lag),
+                      lag = gsub("a", "", lag),
+                      lag = as.numeric(lag))
+
+floyd_rain <- countyweather_rain(counties = fips, county_weather = floyd_weather,
+                                 start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 2000)
+floyd_plot <- ggplot(floyd_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Floyd county, VA, by storm")
+
+floyd_plot
+```
+
+Compare data from `countyweather` and NLDAS-2 data for Fulton county by storm:
+
+```{r}
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "13121") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips)
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "13121"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "fulton_data/")
+
+fulton_dir <- "fulton_data/"
+fulton_weather <- readRDS(paste0(fulton_dir, fips, ".rds"))
+
+#orleans_weather <- orleans_weather$averaged %>%
+#        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+fulton_weather <- fulton_weather %>%
+        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+fulton_weather$fips <- fips
+fulton_weather <- as.data.frame(fulton_weather)
+
+fulton_weather <- fulton_weather %>%
+        dplyr::filter(date %in% all_dates) %>%
+        dplyr::right_join(data.table(check_dates),
+                          by = c("fips" = "fips", "date" = "date")) %>%
+        dplyr::filter(!is.na(prcp)) %>%
+        dplyr::arrange(storm_id, fips) %>%
+        dplyr::mutate(fips = as.integer(fips)) %>%
+        dplyr::mutate(fips = sprintf("%05d", fips),
+                      lag = gsub("day_", "", lag),
+                      lag = gsub("b", "-", lag),
+                      lag = gsub("a", "", lag),
+                      lag = as.numeric(lag))
+
+fulton_rain <- countyweather_rain(counties = fips, county_weather = fulton_weather,
+                                 start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 2000)
+fulton_plot <- ggplot(fulton_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Fulton county, GA, by storm")
+
+fulton_plot
+```
+
 ```{r}

### 2016-08-04 writing/DraftExposurePaper/DraftExposurePaper.Rmd
rearrange code sections for wind plots
add plot for baltimore county comparison

@@
 
+```{r}
+load("MWR_gauge_wind_data/ike_wind.RData")
+ike_wind <- ike_wind %>%
+        filter(!is.na(max_sust_kt)) %>%
+        mutate(wind_sust = convert_wind_speed(max_sust_kt, "knots", "mps"))
+ike_wind_mwr <- plot_mwr_compare_wind(ike_wind)
+
+# load("MWR_gauge_rain_data/alex_gauge.Rdata")
+# alex_mwr <- plot_mwr_compare(alex_gauge)
+
+# grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
+grid.arrange(frances_wind_mwr, ike_wind_mwr,
+             frances_wind_mwr, ike_wind_mwr, ncol = 2)
+```
+
 Compare data from `countyweather` and NLDAS-2 data for Miami-Dade county by storm:
@@
 
+Compare data from `countyweather` and NLDAS-2 data for Baltimore county by storm:
+
 ```{r}
-load("MWR_gauge_wind_data/ike_wind.RData")
-ike_wind <- ike_wind %>%
-        filter(!is.na(max_sust_kt)) %>%
-        mutate(wind_sust = convert_wind_speed(max_sust_kt, "knots", "mps"))
-ike_wind_mwr <- plot_mwr_compare_wind(ike_wind)
-
-# load("MWR_gauge_rain_data/alex_gauge.Rdata")
-# alex_mwr <- plot_mwr_compare(alex_gauge)
-
-# grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
-grid.arrange(frances_wind_mwr, ike_wind_mwr,
-             frances_wind_mwr, ike_wind_mwr, ncol = 2)
+check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
+        dplyr::mutate(closest_date = ymd(closest_date)) %>%
+        dplyr::rename(day_0 = closest_date) %>%
+        dplyr::mutate(fips = as.integer(fips),
+                      day_0 = day_0 + days(0),
+                      day_b1 = day_0 - days(1),
+                      day_b2 = day_0 - days(2),
+                      day_a1 = day_0 + days(1),
+                      day_a2 = day_0 + days(2),
+                      day_a3 = day_0 + days(3)) %>%
+        dplyr::select(storm_id, fips, day_b2, day_b1,
+                      day_0, day_a1, day_a2, day_a3) %>%
+        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
+        dplyr::rename(date = day) %>%
+        dplyr::filter(fips == "24005") %>%
+        dplyr::mutate(fips = as.character(fips))
+        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
+
+all_dates <- unique(check_dates$date)
+all_fips <- unique(check_dates$fips)
+
+#create timeseries for Dade precipitation data from the `countyweather` package
+
+fips <- "24005"
+# I've commented this out because we're saving the data, so we only need
+# to run it once.
+# county_timeseries(fips, percent_coverage = 0,
+#                   date_min = "1988-01-01", date_max = "2011-12-31",
+#                   var = "PRCP",
+#                   out_directory = "baltimore_data/")
+
+baltimore_dir <- "baltimore_data/"
+baltimore_weather <- readRDS(paste0(baltimore_dir, fips, ".rds"))
+
+#orleans_weather <- orleans_weather$averaged %>%
+#        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+baltimore_weather <- baltimore_weather %>%
+        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
+
+baltimore_weather$fips <- fips
+baltimore_weather <- as.data.frame(baltimore_weather)
+
+baltimore_weather <- baltimore_weather %>%
+        dplyr::filter(date %in% all_dates) %>%
+        dplyr::right_join(data.table(check_dates),
+                          by = c("fips" = "fips", "date" = "date")) %>%
+        dplyr::filter(!is.na(prcp)) %>%
+        dplyr::arrange(storm_id, fips) %>%
+        dplyr::mutate(fips = as.integer(fips)) %>%
+        dplyr::mutate(fips = sprintf("%05d", fips),
+                      lag = gsub("day_", "", lag),
+                      lag = gsub("b", "-", lag),
+                      lag = gsub("a", "", lag),
+                      lag = as.numeric(lag))
+
+baltimore_rain <- countyweather_rain(counties = fips, county_weather = baltimore_weather,
+                                 start_year = 1988, end_year = 2011,
+                                 rain_limit = 0, dist_limit = 2000)
+baltimore_plot <- ggplot(baltimore_rain, aes(x = cw_precip, y = tot_precip)) +
+        geom_hline(aes(yintercept = 75), color = "lightgray") +
+        geom_vline(aes(xintercept = 75), color = "lightgray") +
+        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
+        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
+        geom_point(alpha = 0.5) +
+        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
+        xlim(c(0, 275)) + ylim(c(0, 275)) +
+        theme_few() +
+        scale_size_continuous(guide = "none") +
+        xlab("Rainfall (mm) based on \naveraged county monitors") +
+        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
+        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Baltimore county, MD, by storm")
+
+baltimore_plot
 ```

### 2016-08-04 writing/DraftExposurePaper/DraftExposurePaper.Rmd
More work on draft

--- 2016-08-03 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 library(lubridate)
+library(xtable)
 
@@
 frances_wind_mwr <- plot_mwr_compare_wind(frances_wind)
-```
-
-Compare data from `countyweather` and NLDAS-2 data for Miami-Dade county by storm:
-
-```{r}
-data("closest_dist")
-
-# Create a new dataset for precip with lag -2 to 3 for fips 12086 only
-
-check_dates <- dplyr::select(closest_dist, -storm_dist) %>%
-        dplyr::mutate(closest_date = ymd(closest_date)) %>%
-        dplyr::rename(day_0 = closest_date) %>%
-        dplyr::mutate(fips = as.integer(fips),
-                      day_0 = day_0 + days(0),
-                      day_b1 = day_0 - days(1),
-                      day_b2 = day_0 - days(2),
-                      day_a1 = day_0 + days(1),
-                      day_a2 = day_0 + days(2),
-                      day_a3 = day_0 + days(3)) %>%
-        dplyr::select(storm_id, fips, day_b2, day_b1,
-                      day_0, day_a1, day_a2, day_a3) %>%
-        tidyr::gather(key = lag, value = day, -storm_id, -fips) %>%
-        dplyr::rename(date = day) %>%
-        dplyr::filter(fips == "12086") %>%
-        dplyr::mutate(fips = as.character(fips))
-        #dplyr::mutate(day = as.numeric(format(day, "%Y%m%d")))
-
-all_dates <- unique(check_dates$date)
-all_fips <- unique(check_dates$fips)
-
-#create timeseries for Dade precipitation data from the `countyweather` package
-
-fips <- "12086"
-# I've commented this out because we're saving the data, so we only need
-# to run it once.
-# county_timeseries(fips, percent_coverage = 0,
-#                   date_min = "1988-01-01", date_max = "2011-12-31",
-#                   var = "PRCP",
-#                   out_directory = "dade_data/")
-#                   #out_directory = "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/")
-
-# miami_dade_dir <- "~/Documents/hurricaneexposure/writing/DraftExposurePaper/dade_data/"
-dade_dir <- "dade_data/"
-dade_weather <- readRDS(paste0(dade_dir, fips, ".rds"))
-
-dade_weather <- dade_weather$averaged %>%
-        filter(date %in% seq(from = as.Date("1988-01-01"), to = as.Date("2011-12-31"), by = 1))
-
-dade_weather$fips <- fips
-dade_weather <- as.data.frame(dade_weather)
-
-dade_weather <- dade_weather %>%
-        dplyr::filter(date %in% all_dates) %>%
-        dplyr::right_join(data.table(check_dates),
-                          by = c("fips" = "fips", "date" = "date")) %>%
-        dplyr::filter(!is.na(prcp)) %>%
-        dplyr::arrange(storm_id, fips) %>%
-        dplyr::mutate(fips = as.integer(fips)) %>%
-        dplyr::mutate(fips = sprintf("%05d", fips),
-                      lag = gsub("day_", "", lag),
-                      lag = gsub("b", "-", lag),
-                      lag = gsub("a", "", lag),
-                      lag = as.numeric(lag))
-
-lag_sum <- function(counties = NULL, county_weather, year_range = NULL,
-                              distance_limit = NULL, rain_limit = NULL,
-                              include_rain = FALSE, days_included = NULL,
-                              output_vars = c("fips", "cw_precip")){
-
-        closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
-
-        if(!is.null(counties)){
-                closest_dist <- closest_dist[get("fips") %in% counties]
-        }
-
-        if(!is.null(year_range)){
-                closest_dist <- closest_dist[ , .(storm_id,
-                                                  fips,
-                                                  closest_date, storm_dist,
-                                                  local_time,
-                                                  closest_time_utc,
-                                                  year = substring(get("closest_date"),
-                                                                   1, 4)), ][
-                                                                           get("year") %in%
-                                                                                   year_range[1]:year_range[2]
-                                                                           ]
-        }
-
-        if(!is.null(distance_limit)){
-                closest_dist <- closest_dist[get("storm_dist") <=
-                                                     distance_limit]
-        }
-
-        if(include_rain){
-                df <- data.table::data.table(county_weather)
-                df <- df[get("lag") %in% days_included]
-                df <- df[ , .(cw_precip = sum(get("prcp"))),
-                              by = .(fips, storm_id)]
-                closest_dist <- merge(closest_dist, df, all.x = TRUE,
-                                      by = c("storm_id", "fips"))
-                if(!is.null(rain_limit)){
-                        closest_dist <- closest_dist[get("cw_precip") >=
-                                                             rain_limit]
-                }
-        }
-
-        closest_dist <- closest_dist[ , .SD, , .SDcols = output_vars]
-        return(closest_dist)
-}
-
-countyweather_rain <- function(counties, county_weather, start_year, end_year,
-                        rain_limit, dist_limit,
-                        days_included = c(-2, -1, 0, 1)){
-
-        county_rain_storm_df <- lag_sum(counties = counties,
-                                           county_weather = county_weather,
-                                           year_range = c(start_year, end_year),
-                                           distance_limit = dist_limit,
-                                           rain_limit = rain_limit,
-                                           include_rain = TRUE,
-                                           days_included = days_included,
-                                           output_vars = c("storm_id", "fips",
-                                                           "closest_date",
-                                                           "storm_dist",
-                                                           "cw_precip",
-                                                           "local_time",
-                                                           "closest_time_utc"))
-        rain_storm_df <- filter_storm_data(counties = counties,
-                                           year_range = c(start_year, end_year),
-                                           distance_limit = dist_limit,
-                                           rain_limit = rain_limit,
-                                           include_rain = TRUE,
-                                           days_included = days_included,
-                                           output_vars = c("storm_id", "fips",
-                                                           "closest_date",
-                                                           "storm_dist",
-                                                           "tot_precip",
-                                                           "local_time",
-                                                           "closest_time_utc"))
-        rain_df <- merge(county_rain_storm_df, rain_storm_df, all.x = TRUE,
-                         by = c("fips", "storm_id", "closest_date", "storm_dist",
-                                "local_time", "closest_time_utc"))
-        return(rain_df)
-}
-
-miami_rain <- countyweather_rain(counties = "12086", county_weather = dade_weather,
-                                 start_year = 1988, end_year = 2011,
-                                 rain_limit = 0, dist_limit = 2000)
-miami_plot <- ggplot(miami_rain, aes(x = cw_precip, y = tot_precip)) +
-        geom_hline(aes(yintercept = 75), color = "lightgray") +
-        geom_vline(aes(xintercept = 75), color = "lightgray") +
-        geom_abline(aes(intercept = 0, slope = 1), color = "gray", alpha = 0.5) +
-        # geom_point(aes(size = prcp_reporting), alpha = 0.2) +
-        geom_point(alpha = 0.5) +
-        # geom_text(aes(x = monitor_rain + 100, label = storm_id)) +
-        xlim(c(0, 275)) + ylim(c(0, 275)) +
-        theme_few() +
-        scale_size_continuous(guide = "none") +
-        xlab("Rainfall (mm) based on \naveraged county monitors") +
-        ylab("Rainfall (mm) based on \nNLDAS-2 county data") +
-        ggtitle("Monitor versus NLDAS rainfall estimates \nfor Miami-Dade County, FL, by storm")
-
-miami_plot
-```
-
-```{r}
+
 load("MWR_gauge_wind_data/ike_wind.RData")
@@
 
-Map of the count of county exposures to the top 2,000 county-storm pairs by distance-based exposure, based on "exposed" = distance of 100 kilometers or less:
+We first investigated whether different exposure metrics resulted in similar numbers of counties classified as "exposed" to a storm (Figure \ref{fig:county_number_cor}, Table \ref{tab:county_number}).
+
+```{r county_number_cor, echo = FALSE, fig.width = 8, fig.height = 8, fig.align = "center", fig.cap = "Association between the number of counties classified as `exposed` for a storm using different exposure metrics.
+Each small plot shows this association, with each point representing a storm, with positions on the x- and y-axes showing the number of counties classified as `exposed` for that storm based on one of the exposure metrics considered in this study.
+Study storms from 1996 to 2011 are included in each plot."}
+dist_exp <- distance_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(dist_exposed = 1)
+rain_exp <- rain_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed = 1)
+wind_exp <- wind_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed = 1)
+flood_exp <- flood_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(flood_exposed = 1,
+               storm_id = as.character(storm_id))
+tornado_exp <- tornado_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(tornado_exposed = 1,
+               storm_id = as.character(storm_id))
+
+all_exp <- dist_exp %>%
+        full_join(rain_exp, by = c("storm_id", "fips")) %>%
+        full_join(wind_exp, by = c("storm_id", "fips")) %>%
+        filter(as.numeric(gsub(".+-", "", storm_id)) >= 1996) %>%
+        full_join(flood_exp, by = c("storm_id", "fips")) %>%
+        full_join(tornado_exp, by = c("storm_id", "fips")) %>%
+        mutate(dist_exposed = !is.na(dist_exposed),
+               rain_exposed = !is.na(rain_exposed),
+               wind_exposed = !is.na(wind_exposed),
+               flood_exposed = !is.na(flood_exposed),
+               tornado_exposed = !is.na(tornado_exposed)) %>%
+        group_by(storm_id) %>%
+        summarize(any_exposed = n(),
+                  Distance = sum(dist_exposed),
+                  Rain = sum(rain_exposed),
+                  Wind = sum(wind_exposed),
+                  Flood = sum(flood_exposed),
+                  Tornado = sum(tornado_exposed)) %>%
+        arrange(desc(any_exposed))
+
+GGally::ggpairs(select(all_exp, -storm_id, -any_exposed),
+                aes(alpha = 0.5)) + theme_few()
+```
+
+```{r county_number, echo = FALSE, results = "asis"}
+a <- all_exp %>% 
+        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")")) %>%
+        slice(1:20) %>%
+        rename(Storm = storm_id, 
+               "At least one metric" = any_exposed) %>%
+        xtable(caption = "Top 10 storms in terms of numbers of counties assessed as `exposed` based on at least one of the exposure metrics, for storms between 1996 and 2011.",
+                       label = "tab:county_number")
+print(a, include.rownames = FALSE, floating = TRUE,
+      table.placement = "p", caption.placement = "top", 
+      booktabs = TRUE, comment = FALSE, timestamp = NULL)
+```
+
+We also mapped geographic patterns in the number of hits per county for distance, rain, and wind exposure when we incorporated the intensity of exposure for each county-storm pair (Figures \ref{fip:top_dist_map}, \ref{fip:top_rain_map}, \ref{fip:top_wind_map}); Tables \ref{tab:top_distance}, \ref{tab:top_rain}, \ref{tab:top_wind}).
+For these maps, we only plotted "hits" for the top 2,000 county-storm pairs (i.e., the 2,000 storm-county pairs with smallest distance to storm's closest approach, in the case of distance-based exposure, or the 2,000 storm-county pairs with highest precipitation, in the case of rain-based exposure).
+
+```{r top_distance, echo = FALSE, results = "asis"}
+a <- distance_exposure %>%
+        arrange(storm_dist) %>% 
+        slice(1:20) %>%
+        left_join(county_centers, by = "fips") %>%
+        select(storm_id, name, storm_dist) %>%
+        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")")) %>%
+        rename(Storm = storm_id,
+               County = name,
+               "Distance (km)" = storm_dist) %>%
+        xtable(caption = "Top storm-county pairs in terms of closest distance between county's population mean center and storm's closest approach.", 
+               label = "tab:top_distance")
+print(a, include.rownames = FALSE, floating = TRUE,
+      table.placement = "p", caption.placement = "top", 
+      booktabs = TRUE, comment = FALSE, timestamp = NULL)
+```
+
+```{r top_rain, echo = FALSE, results = "asis"}
+a <- rain_exposure %>%
+        arrange(desc(tot_precip)) %>% 
+        slice(1:20) %>%
+        left_join(county_centers, by = "fips") %>%
+        select(storm_id, name, tot_precip) %>%
+        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")")) %>%
+        rename(Storm = storm_id,
+               County = name,
+               "Rainfall (mm)" = tot_precip) %>%
+        xtable(caption = "Top storm-county pairs in terms of most cumulative precipitation in the window from two days before to one day after a storm's closest approach to the county.", 
+               label = "tab:top_rain")
+print(a, include.rownames = FALSE, floating = TRUE,
+      table.placement = "p", caption.placement = "top", 
+      booktabs = TRUE, comment = FALSE, timestamp = NULL)
+```
+
+```{r top_wind, echo = FALSE, results = "asis"}
+a <- wind_exposure %>%
+        arrange(desc(max_sust)) %>% 
+        slice(1:20) %>%
+        left_join(county_centers, by = "fips") %>%
+        select(storm_id, name, max_sust) %>%
+        mutate(storm_id = paste0(gsub("-", " (", storm_id), ")"),
+               max_sust = round(max_sust)) %>%
+        rename(Storm = storm_id,
+               County = name,
+               "Maximum sustained wind (m / s)" = max_sust) %>%
+        xtable(caption = "Top storm-county pairs in terms of most intense sustained winds.", 
+               label = "tab:top_wind")
+print(a, include.rownames = FALSE, floating = TRUE,
+      table.placement = "p", caption.placement = "top", 
+      booktabs = TRUE, comment = FALSE, timestamp = NULL)
+```
 
 ```{r echo = FALSE, warning=FALSE, message=FALSE}
-distance_2000 <- distance_exposure %>%
-        arrange(storm_dist) %>% 
-        slice(1:2000) %>% 
+plot_top_x <- function(exposure_data, x = 2000, desc = TRUE,
+                       exposure_var, eastern_states = eastern_states,
+                       my_fips = my_fips, title){
+        missing_fips <- my_fips[!(my_fips %in% exposure_data$fips)]
+        missing_counties <- data.frame(fips = as.numeric(missing_fips),
+                                       count = 0)
+
+        exposure_data <- as.data.frame(exposure_data)
+        exposure_data$exp_var <- exposure_data[ , exposure_var]
+        if(!desc){ exposure_data$exp_var <- -1 * exposure_data$exp_var}
+
+        exposure_x <- exposure_data %>%
+                arrange(desc(exp_var)) %>%
+                slice(1:x) %>% 
                 group_by(fips) %>%
                 summarise(count = n()) %>%
+                rbind(missing_counties) %>%
                 rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count))
-
-distance_2000$value <- cut(distance_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
+                mutate(region = as.numeric(region),
+                       count = as.numeric(count))
+
+        exposure_x$value <- cut(exposure_x$count,
+                                c(0, 1, 3, 5, 7, 9, 11, 13, 100), 
+                   labels = c("0", "1-2", "3-4", "5-6", "7-8",
+                              "9-10", "11-12", "13+"), 
                    include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_dist_2000 <- CountyChoropleth$new(distance_2000)
-map_dist_2000$set_zoom(eastern_states)
-map_dist_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
+        exposure_palette <- c("#FFFFFF", brewer.pal(8, name = "YlGnBu"))
+
+        map_top_x <- CountyChoropleth$new(exposure_x)
+        map_top_x$set_zoom(eastern_states)
+        map_top_x$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
                                                 values = exposure_palette,
                                                 na.value = "white")
-map_dist_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Distance Metric"
-a <- map_dist_2000$render()
-```
-
-```{r echo = FALSE, warning=FALSE, message=FALSE}
-# Get the rain-based exposure for all counties over the study period, 
-# using a rain limit of 75 mm and a distance limit of 500 km
-rain_exposure <- county_rain(counties = my_fips,
-                             start_year = 1988, end_year = 2011,
-                             rain_limit = 75, dist_limit = 500,
-                             days_include = c(-1, 0, 1))
-
-rain_2000 <- rain_exposure %>%
-        arrange(desc(tot_precip)) %>%
-        slice(1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count))
-
-rain_2000$value <- cut(rain_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-
-map_rain_2000 <- CountyChoropleth$new(rain_2000)
-map_rain_2000$set_zoom(eastern_states)
-map_rain_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_rain_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor Rain Metric"
-b <- map_rain_2000$render()
-```
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
-grid.arrange(a, b, ncol = 1)
-```
-
-Map of county exposures to the top 2,000 county-storm pairs by wind-based exposure:
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE, message = FALSE, warning = FALSE}
-#TS winds and above
-TS_wind_exposure <- county_wind(counties = my_fips,
-                             start_year = 1988,
-                             end_year = 2011,
-                             wind_limit = 18,
-                             wind_var = "max_sust")
-
-TS_wind_2000 <- TS_wind_exposure %>% 
-        arrange(desc(max_sust)) %>%
-        slice (1:2000) %>%
-        group_by(fips) %>%
-        summarise(count = n()) %>%
-        rename(region = fips) %>%
-        mutate(region = as.numeric(region), count = as.numeric(count)) # Max count of 178 is a NA value
-TS_wind_2000 <- subset(TS_wind_2000, !is.na(TS_wind_2000$region)) # Remove NA observation
-                                                                       # County pairs
-TS_wind_2000$value <- cut(TS_wind_2000$count, c(0, 1, 3, 5, 7, 9, 11, 13, 15), 
-                   labels = c("0-1", "2-3", "4-5", "6-7", "8-9", "10-11", "12-13", "14-15"), 
-                   include.lowest = TRUE)
-exposure_palette <- brewer.pal(8, name = "YlGnBu")
-map_TS_2000 <- CountyChoropleth$new(TS_wind_2000)
-map_TS_2000$set_zoom(eastern_states)
-map_TS_2000$ggplot_scale <- scale_fill_manual(name = "Hits per\nCounty",
-                                                values = exposure_palette,
-                                                na.value = "white")
-map_TS_2000$title <- "Hits for top 2,000 Storm-County Pairs\nfor TS Wind Metric"
-map_TS_2000$render()
+        map_top_x$title <- title
+        out <- map_top_x$render()
+        return(out)
+}
+```
+
+```{r top_dist_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance."}
+plot_top_x(distance_exposure, desc = FALSE, exposure_var = "storm_dist",
+           eastern_states = eastern_states, my_fips = my_fips,
+           title = "Hits for top 2,000 Storm-County Pairs\nfor Distance Metric")
+```
+
+```{r top_rain_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for rain."}
+plot_top_x(rain_exposure, exposure_var = "tot_precip",
+           eastern_states = eastern_states, my_fips = my_fips,
+           title = "Hits for top 2,000 Storm-County Pairs\nfor Rain Metric")
+```
+
+```{r top_wind_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance"}
+plot_top_x(wind_exposure, exposure_var = "max_sust",
+           eastern_states = eastern_states, my_fips = my_fips,
+           title = "Hits for top 2,000 Storm-County Pairs\nfor Wind Metric")
 ```

### 2016-08-05 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work some on draft

--- 2016-08-04 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 Here are results for comparing our method to using ground-based monitors.
-First, we looked at several counties, comparing estimates of storm-associated rain using NLDAS-2 versus ground-based monitors (Figure [x]).
-
-[Add in figures by county here for county averages of ground-based monitors versus our method. (for example, Miami-Dade figure)]
+First, we looked at several counties, comparing estimates of storm-associated rain using NLDAS-2 versus ground-based monitors (Figure \ref{fig:compare_rain_ex_counties}).
+
+```{r compare_rain_ex_counties, echo = FALSE, fig.width = 10, fig.height = 10, efig.align = "center", fig.cap = "Comparison of rain estimates for example counties."}
+plot_county_rain_compare <- function(ex_fips, ex_dir, ex_title, 
+                                     get_data = FALSE){
+        if(get_data){
+                county_timeseries(ex_fips, percent_coverage = 0,
+                                  date_min = "1988-01-01",
+                                  date_max = "2011-12-31",
+                                  var = "PRCP",
+                                  out_directory = ex_dir)
+        }
+
+        check_dates <- closest_dist %>%
+                dplyr::filter(fips == ex_fips) %>%
+                dplyr::select(-storm_dist) %>%
+                dplyr::mutate(closest_date = ymd(closest_date)) %>%
+                dplyr::rename(day_0 = closest_date) %>%
+                dplyr::mutate(fips = as.integer(fips),
+                              day_0 = day_0 + days(0),
+                              day_b1 = day_0 - days(1),
+                              day_b2 = day_0 - days(2),
+                              day_a1 = day_0 + days(1)) %>%
+                dplyr::select(storm_id, day_b2, day_b1, day_0, day_a1) %>%
+                tidyr::gather(key = lag, value = day, -storm_id) %>%
+                dplyr::rename(date = day)
+        all_dates <- unique(check_dates$date)
+
+        ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
+        ex_weather <- as.data.frame(ex_weather$averaged)
+
+        ex_weather <- ex_weather %>%
+                dplyr::filter(date %in% all_dates) %>%
+                dplyr::right_join(check_dates, by = "date") %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(prcp = sum(prcp),
+                                ave_n = mean(prcp_reporting))
+
+        ex_rain <- county_rain(counties = ex_fips,
+                               start_year = 1988, end_year = 2011,
+                               rain_limit = 0, dist_limit = 2000) %>%
+                full_join(ex_weather, by = "storm_id") %>%
+                filter(!is.na(prcp) & !is.na(tot_precip))
+
+        plot_range <- range(ex_rain[ , c("prcp", "tot_precip")],
+                            na.rm = TRUE)
+
+        ex_plot <- ggplot(ex_rain, aes(x = tot_precip, y = prcp)) +
+                geom_hline(aes(yintercept = 75), color = "lightgray") +
+                geom_vline(aes(xintercept = 75), color = "lightgray") +
+                geom_abline(aes(intercept = 0, slope = 1), color = "gray",
+                    alpha = 0.5) +
+                geom_point(alpha = 0.5) +
+                theme_few() +
+                scale_size_continuous(guide = "none") +
+                xlab("Rainfall (mm)  from\nNLDAS-2 data") +
+                ylab("Rainfall (mm)\nfrom monitors") +
+                xlim(plot_range) + ylim(plot_range) + 
+                ggtitle(ex_title)
+        return(ex_plot)
+}
+
+a <- plot_county_rain_compare(ex_fips = "12086", ex_dir = "dade_data/",
+                         ex_title = "Miami-Dade, FL")
+
+b <- plot_county_rain_compare(ex_fips = "22071", ex_dir = "orleans_data/",
+                         ex_title = "Orleans Parish, LA")
+
+c <- plot_county_rain_compare(ex_fips = "45019", ex_dir = "charleston_data/",
+                         ex_title = "Charleston County, SC")
+
+d <- plot_county_rain_compare(ex_fips = "37183", ex_dir = "wake_data/",
+                         ex_title = "Wake County, NC")
+
+e <- plot_county_rain_compare(ex_fips = "13121", ex_dir = "fulton_data/",
+                         ex_title = "Fulton County, GA")
+
+f <- plot_county_rain_compare(ex_fips = "24005", ex_dir = "baltimore_data/",
+                         ex_title = "Baltimore County, MD")
+
+g <- plot_county_rain_compare(ex_fips = "42101",
+                              ex_dir = "philadelphia_data/",
+                         ex_title = "Philadelphia County, PA")
+
+h <- plot_county_rain_compare(ex_fips = "48201", ex_dir = "harris_data/",
+                         ex_title = "Harris County, TX")
+
+i <- plot_county_rain_compare(ex_fips = "01097", ex_dir = "mobile_data/",
+                         ex_title = "Mobile County, AL", get_data = TRUE)
+
+grid.arrange(a, b, c, d, e, f, g, h, i, ncol = 3)
+```
 
@@
 
-# load("MWR_gauge_rain_data/alex_gauge.Rdata")
-# alex_mwr <- plot_mwr_compare(alex_gauge)
-
-# grid.arrange(allison_mwr, frances_mwr, ike_mwr, alex_mwr, ncol = 2)
+load("MWR_gauge_wind_data/alex_wind.Rdata")
+alex_wind <- alex_wind %>%
+        filter(!is.na(max_sust)) %>%
+        mutate(wind_sust = convert_wind_speed(max_sust, "knots", "mps")) %>%
+        select(storm_id, fips, wind_sust)
+alex_wind_mwr <- plot_mwr_compare_wind(alex_wind)
+
 grid.arrange(frances_wind_mwr, ike_wind_mwr,
-             frances_wind_mwr, ike_wind_mwr, ncol = 2)
+             frances_wind_mwr, alex_wind_mwr, ncol = 2)
 ```
@@
 
-We determined the average number of hits per year for each eastern US county for the exposure criteria based on distance (Figure \ref{fig:distance_hits}), rain (Figure \ref{fig:rain_hits}), and wind (Figure \ref{fig:wind_hits}), as well as the total number of hits based on storm-associated events listed in the NOAA Storm Events database for flood events (Figure \ref{fig:flood_hits}), wind events (Figure \ref{fig:wind_event_hits}), and tornado events (Figure \ref{fig:tornado_hits}) (all of these maps based on NOAA Storm Events are restricted to storms between 1996 and 2011).
+We determined the average number of hits per year for each eastern US county for the exposure criteria based on distance (Figure \ref{fig:distance_hits}), rain (Figure \ref{fig:rain_hits}), and wind (Figure \ref{fig:wind_hits}), as well as the total number of hits based on storm-associated events listed in the NOAA Storm Events database for flood events (Figure \ref{fig:flood_hits}) and tornado events (Figure \ref{fig:tornado_hits}) (all of these maps based on NOAA Storm Events are restricted to storms between 1996 and 2011).
 
@@
                start_year = 1996, end_year = 2011)
-```
-
-```{r wind_event_hits, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Average number of hits per year for eastern US counties based on the county having a storm-associated wind event listed in the NOAA Storm Events database, for storms from 1996 to 2011.
-Light red lines show the storm tracks over the same period.
-Red lines show the storm tracks for the top 5 storms in terms of number of counties classified as 'exposed' to the storm based on this criterion over the same period (Allison, 2001: 435 counties; Arlene, 2005: 360 counties; Danny, 1997: 310 counties; Cindy, 2005: 189 counties; Cristobal, 2008: 183 counties)."}
-wind_event_exposure <- county_events(counties = my_fips,
-                              start_year = 1996, end_year = 2011,
-                              event_type = "wind")
-map_total_hits(wind_event_exposure, my_fips = my_fips, 
-               eastern_states = eastern_states,
-               title = "Hits per county based on wind events")
 ```
@@
 
-We also mapped geographic patterns in the number of hits per county for distance, rain, and wind exposure when we incorporated the intensity of exposure for each county-storm pair (Figures \ref{fip:top_dist_map}, \ref{fip:top_rain_map}, \ref{fip:top_wind_map}); Tables \ref{tab:top_distance}, \ref{tab:top_rain}, \ref{tab:top_wind}).
+Some storms can very substantially in their exposure patterns when different metrics are used.
+For example, Tropical Storm Allison (2001) was notable for its rainfall rather than wind intensity; as a result, many more counties are classified as "exposed" to that storm when the exposure classification is based on rainfall rather than wind (Figure \ref{fig:allison_wind_rain_exp}).
+Conversely, Hurricane Hugo (1989) was more notable for wind than rainfall, which is reflected in more counties assessed as "exposed" based on wind than rain (Figure \ref{fig:hugo_wind_rain_exp}).
+
+[One interesting contrast is between storms that were noteworthy for direct wind-related deaths (Anderew in 1992, Hugo in 1989) versus storms noteworthy for direct freshwater flooding-related deaths (Floyd in 1999, Alberto in 1994, Charley in 1998) for the 1970--1999 period [@Rappaport2000].
+(Another study notes the following storms were notable for inland impacts-- Hugo, 1989 (wind and rain); Floyd, 1999 (flooding); Allison, 2001 (flooding); Fran, 1996 (wind) [@Kruk2010].) Here are two examples, one of each of these two types:]
+
+```{r allison_wind_rain_exp, fig.width = 6, fig.height = 8, echo = FALSE, fig.align = "center", fig.cap = "Counties that were classified as 'exposed' to Tropical Storm Allison (2001) based on rain (top) and wind (bottom) exposure criteria.
+Rain-based exposure was assessed based on storm-associated rainfall of 75 millimeters or more in the county and a distance of 500 kilometers or less between the county's population mean center and the storm's closest approach.
+Wind-based exposure was based on maximum sustained winds of 15 meters per second or more in the county during the storm."}
+storm <- "Allison-2001"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                          rain_limit = 75,
+                                          dist_limit = 500)) + 
+        ggtitle(paste0(storm, ", Rain-based exposure"))
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_wind_exposure(storm = storm, 
+                                           wind_limit = 15)) + 
+        ggtitle(paste0(storm, ", Wind-based exposure"))
+
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r hugo_wind_rain_exp, fig.width = 6, fig.height = 8, echo = FALSE, fig.align = "center", fig.cap = "Counties that were classified as 'exposed' to Hurricane Hugo (1989) based on rain (top) and wind (bottom) exposure criteria.
+Rain-based exposure was assessed based on storm-associated rainfall of 75 millimeters or more in the county and a distance of 500 kilometers or less between the county's population mean center and the storm's closest approach.
+Wind-based exposure was based on maximum sustained winds of 15 meters per second or more in the county during the storm."}
+storm <- "Hugo-1989"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                          rain_limit = 75,
+                                          dist_limit = 500)) + 
+        ggtitle(paste0(storm, ", Rain-based exposure"))
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_wind_exposure(storm = storm, 
+                                           wind_limit = 15)) + 
+        ggtitle(paste0(storm, ", Wind-based exposure"))
+
+grid.arrange(a, b, ncol = 1)
+```
+
+As another example, exposure classifications based on flood events in some cases align well with rain-based classifications (e.g., for Hurricane Floyd in 1999; Figure \ref{fig:floyd_rain_flood_exp}) but for other storms only a small percent of the counties assessed as "exposed" based on rain criteria are also assessed as "exposed" based on flood exposure (e.g., Hurricane Frances, 2004; Figure \ref{fig:frances_rain_flood_exp}).
+
+[For storms in 1996 or later, when flooding reports are included in NOAA's Storm Events data, exposure patterns based on floods reports generally pick up similar regions as rain-based exposure, and in some cases (e.g., Floyd, 1999), counties classified as exposed using rain-based exposure are very similar to the counties in which flood events were recorded.
+For example, here are comparisons for some of the worst storms in terms of rain-based exposures:]
+
+```{r frances_rain_flood_exp, echo = FALSE, fig.width = 6, fig.height = 8, fig.align = "center", fig.cap = "Counties that were classified as 'exposed' to Hurricane Frances (2004) based on rain (top) and flood event (bottom) exposure criteria.
+Rain-based exposure was assessed based on storm-associated rainfall of 75 millimeters or more in the county and a distance of 500 kilometers or less between the county's population mean center and the storm's closest approach.
+Flood event-based exposure was based on the county have a flood event listing in the NOAA Storm Events database and being within [x] kilometers of the storm's track."}
+storm <- "Frances-2004"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Frances, 2004")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Frances, 2004")
+grid.arrange(a, b, ncol = 1)
+```
+
+```{r floyd_rain_flood_exp, echo = FALSE, fig.width = 6, fig.height = 8, fig.align = "center", fig.cap = "Counties that were classified as 'exposed' to Hurricane Floyd (1999) based on rain (top) and flood event (bottom) exposure criteria.
+Rain-based exposure was assessed based on storm-associated rainfall of 75 millimeters or more in the county and a distance of 500 kilometers or less between the county's population mean center and the storm's closest approach.
+Flood event-based exposure was based on the county have a flood event listing in the NOAA Storm Events database and being within [x] kilometers of the storm's track."}
+storm <- "Floyd-1999"
+a <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_rain_exposure(storm = storm, 
+                                           dist_limit = 500,
+                                           rain_limit = 75)) + 
+        ggtitle("Rain-based exposure during Floyd, 1999")
+b <- map_tracks(storms = storm, plot_points = FALSE, 
+                plot_object = map_event_exposure(storm = storm, 
+                                           event_type = "flood")) + 
+        ggtitle("Flood events during Floyd, 1999")
+grid.arrange(a, b, ncol = 1)
+```
+
+We mapped geographic patterns in the number of hits per county for distance, rain, and wind exposure when we incorporated the intensity of exposure for each county-storm pair (Figures \ref{fig:top_dist_map}, \ref{fig:top_rain_map}, \ref{fig:top_wind_map}); Tables \ref{tab:top_distance}, \ref{tab:top_rain}, \ref{tab:top_wind}).
 For these maps, we only plotted "hits" for the top 2,000 county-storm pairs (i.e., the 2,000 storm-county pairs with smallest distance to storm's closest approach, in the case of distance-based exposure, or the 2,000 storm-county pairs with highest precipitation, in the case of rain-based exposure).
@@
                        my_fips = my_fips, title){
-        missing_fips <- my_fips[!(my_fips %in% exposure_data$fips)]
-        missing_counties <- data.frame(fips = as.numeric(missing_fips),
-                                       count = 0)
 
@@
                 group_by(fips) %>%
-                summarise(count = n()) %>%
-                rbind(missing_counties) %>%
+                summarise(count = n())
+
+        zero_fips <- my_fips[!(my_fips %in% exposure_x$fips)]
+        zero_counties <- data.frame(fips = as.numeric(zero_fips),
+                                       count = 0)
+
+        exposure_x <- exposure_x %>%
+                rbind(zero_counties) %>%
                 rename(region = fips) %>%
@@
 
-```{r}
+```{r echo = FALSE}
 rain_exposure$county_storm <- transmute(rain_exposure,
@@
 
-One interesting contrast is between storms that were noteworthy for direct wind-related deaths (Anderew in 1992, Hugo in 1989) versus storms noteworthy for direct freshwater flooding-related deaths (Floyd in 1999, Alberto in 1994, Charley in 1998) for the 1970--1999 period [@Rappaport2000].
-(Another study notes the following storms were notable for inland impacts-- Hugo, 1989 (wind and rain); Floyd, 1999 (flooding); Allison, 2001 (flooding); Fran, 1996 (wind) [@Kruk2010].) Here are two examples, one of each of these two types:
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE}
-storm <- "Hugo-1989"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                          rain_limit = 75,
-                                          dist_limit = 500)) + 
-        ggtitle(paste0(storm, ", Rain-based exposure"))
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_wind_exposure(storm = storm, 
-                                           wind_limit = 15)) + 
-        ggtitle(paste0(storm, ", Wind-based exposure"))
-
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r fig.width = 6, fig.height = 8, echo = FALSE}
-storm <- "Allison-2001"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                          rain_limit = 75,
-                                          dist_limit = 500)) + 
-        ggtitle(paste0(storm, ", Rain-based exposure"))
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_wind_exposure(storm = storm, 
-                                           wind_limit = 15)) + 
-        ggtitle(paste0(storm, ", Wind-based exposure"))
-
-grid.arrange(a, b, ncol = 1)
-```
-
 #### Which storms exposed the most counties when using distance-, rain-, and wind-based metrics? What were the top county-storm exposures for rain- and wind-based metrics?
@@
 
-Here is a map of the rain totals for Frances:
-
-```{r echo = FALSE}
-map_tracks(storms = "Frances-2004", plot_points = FALSE, 
-                plot_object = map_counties(storm = "Frances-2004", 
-                                           metric = "rainfall")) + 
-        ggtitle(paste("Frances-2004", "rainfall"))
-```
-
-Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 18 m/s or greater (e.g., tropical storm strength or greater):
+Here is a table of the top 10 storms by number of counties exposed, based on maximum sustained winds of 15 m / s or greater (e.g., tropical storm strength or greater):
 
@@
 all_wind <- county_wind(counties = my_fips, start_year = 1988, 
-                        end_year = 2011, wind_limit = 18)
+                        end_year = 2011, wind_limit = 15)
 
@@
 
-Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 33 m/s or greater (e.g., Hurricane strength, category 1 or greater):
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_wind <- county_wind(counties = my_fips, start_year = 1988, 
-                        end_year = 2011, wind_limit = 33)
-
-hu_exposed <- all_wind %>%
-        group_by(storm_id) %>%
-        summarize(number_of_counties_exposed = n()) %>%
-        arrange(desc(number_of_counties_exposed)) %>%
-        top_n(10)
-
-hu_exposed %>%
-        mutate(storm_id = gsub("-", " (", storm_id),
-               storm_id = paste0(storm_id, ")")) %>%
-        kable(col.names = c("Storm", "# of exposed counties"),
-              caption = "Top 10 storms based on the number of counties exposed, using a metric of wind.")
-```
-
-Here is a table of the top 10 storms by number of counties exposed, based on a wind metric of 50 m/s or greater (e.g., Intense Hurricane strength, category 3 or greater):
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_wind <- county_wind(counties = my_fips, start_year = 1988, 
-                        end_year = 2011, wind_limit = 50)
-
-cat3_exposed <- all_wind %>%
-        group_by(storm_id) %>%
-        summarize(number_of_counties_exposed = n()) %>%
-        arrange(desc(number_of_counties_exposed)) %>%
-        top_n(10)
-
-cat3_exposed %>%
-        mutate(storm_id = gsub("-", " (", storm_id),
-               storm_id = paste0(storm_id, ")")) %>%
-        kable(col.names = c("Storm", "# of exposed counties"),
-              caption = "Top 10 storms based on the number of counties exposed, using a metric of wind.")
-```
-
-For all of the above exposure categories, Lili (1996), Karl (2004), and Ophelia (2005) remained the top 3 storms by number of counties exposed based on the various wind exposure metrics, each with 2396 counties exposed [JMF-- I think something weird is going on here].
-
 #### Comparing across all exposure metrics
-
-For storms in 1996 or later, when flooding reports are included in NOAA's Storm Events data, exposure patterns based on floods reports generally pick up similar regions as rain-based exposure, and in some cases (e.g., Floyd, 1999), counties classified as exposed using rain-based exposure are very similar to the counties in which flood events were recorded.
-For example, here are comparisons for some of the worst storms in terms of rain-based exposures:
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Frances-2004"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Frances, 2004")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Frances, 2004")
-grid.arrange(a, b, ncol = 1)
-```
-
-```{r echo = FALSE, fig.width = 6, fig.height = 8}
-storm <- "Floyd-1999"
-a <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_rain_exposure(storm = storm, 
-                                           dist_limit = 500,
-                                           rain_limit = 75)) + 
-        ggtitle("Rain-based exposure during Floyd, 1999")
-b <- map_tracks(storms = storm, plot_points = FALSE, 
-                plot_object = map_event_exposure(storm = storm, 
-                                           event_type = "flood")) + 
-        ggtitle("Flood events during Floyd, 1999")
-grid.arrange(a, b, ncol = 1)
-```
 

### 2016-08-05 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work more on draft.

@@
 
-Here is a table of the top 10 storms by number of counties exposed, based on a distance metric of within 100 kilometers or less of the storm track:
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_distance <- county_distance(counties = my_fips, start_year = 1988, 
-                        end_year = 2011, dist_limit =100)
-
-distance_exposed <- all_distance %>%
-        group_by(storm_id) %>%
-        summarize(number_of_counties_exposed = n()) %>%
-        arrange(desc(number_of_counties_exposed)) %>%
-        top_n(10)
-
-distance_exposed %>%
-        mutate(storm_id = gsub("-", " (", storm_id),
-               storm_id = paste0(storm_id, ")")) %>%
-        kable(col.names = c("Storm", "# of exposed counties"),
-              caption = "Top 10 storms based on the number of counties exposed, using a metric of distance.")
-```
-
-Here are plots of the top four storms based on number of counties exposed by distance:
-
-```{r echo = FALSE, fig.width = 10, fig.height = 8}
+We created plots of the top four storms, in terms of the number of counties classified as exposed, for exposure metrics of distance (Figure \ref{fig:top_four_distance}), rain (Figure \ref{fig:top_four_rain}), wind (Figure \ref{fig:top_four_wind}), flood events (Figure \ref{fig:top_four_flood}), and tornado events (Figure \ref{fig:top_four_tornado}).
+
+```{r top_four_distance, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Counties exposed for the top four storms over the study period, in terms of distance-based exposure."}
 top_distance_exposed <- function(storm_id){
@@
 
+all_distance <- county_distance(counties = my_fips, start_year = 1988, 
+                        end_year = 2011, dist_limit =100)
+
+distance_exposed <- all_distance %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
 n <- 4
@@
 
-Here is a table of the top 10 storms by number of counties exposed, based on a rain metric of 75 millimeters or more and within 500 kilometers or less of the storm track:
-
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_rain <- county_rain(counties = my_fips, start_year = 1988, 
-                        end_year = 2011, rain_limit = 75, dist_limit = 500,
-                        days_included = -1:1)
-
-rain_exposed <- all_rain %>%
-        group_by(storm_id) %>%
-        summarize(number_of_counties_exposed = n()) %>%
-        arrange(desc(number_of_counties_exposed)) %>%
-        top_n(10)
-
-rain_exposed %>%
-        mutate(storm_id = gsub("-", " (", storm_id),
-               storm_id = paste0(storm_id, ")")) %>%
-        kable(col.names = c("Storm", "# of exposed counties"),
-              caption = "Top 10 storms based on the number of counties exposed, using a metric of rainfall.")
-```
-
-Here are plots of the exposed counties for the top 4 storms based on that rain metric:
-
-```{r echo = FALSE, fig.width = 10, fig.height = 8}
+```{r top_four_rain, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Counties exposed for the top four storms over the study period, in terms of distance-based exposure."}
 top_rain_exposed <- function(storm_id){
@@
 
+all_rain <- county_rain(counties = my_fips, start_year = 1988, 
+                        end_year = 2011, rain_limit = 75, dist_limit = 500,
+                        days_included = -1:1)
+
+rain_exposed <- all_rain %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
 n <- 4
@@
 grid.arrange(rain_plots$a, rain_plots$b, rain_plots$c, rain_plots$d, ncol = 2)
+```
+
+```{r top_four_wind, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Counties exposed for the top four storms over the study period, in terms of distance-based exposure."}
+top_wind_exposed <- function(storm_id){
+        my_map <- map_wind_exposure(storm_id, wind_limit = 15)
+        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
+                             plot_points = FALSE) + 
+                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
+}
+
+all_wind <- county_wind(counties = my_fips, start_year = 1988, 
+                        end_year = 2011, wind_limit = 15)
+
+wind_exposed <- all_wind %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10)
+
+n <- 4
+wind_plots <- lapply(wind_exposed$storm_id[1:n], top_wind_exposed)
+names(wind_plots) <- letters[1:n]
+grid.arrange(wind_plots$a, wind_plots$b, wind_plots$c,
+             wind_plots$d, ncol = 2)
+```
+
+```{r top_four_flood, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Counties exposed for the top four storms over the study period, in terms of flood event-based exposure."}
+top_flood_exposed <- function(storm_id){
+        my_map <- map_event_exposure(storm_id, event_type = "flood")
+        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
+                             plot_points = FALSE) + 
+                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
+}
+
+all_flood <- county_events(counties = my_fips, start_year = 1996, 
+                        end_year = 2011, event_type = "flood")
+
+flood_exposed <- all_flood %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10) %>%
+        ungroup() %>%
+        mutate(storm_id = as.character(storm_id))
+
+n <- 4
+flood_plots <- lapply(flood_exposed$storm_id[1:n], top_flood_exposed)
+names(flood_plots) <- letters[1:n]
+grid.arrange(flood_plots$a, flood_plots$b, flood_plots$c,
+             flood_plots$d, ncol = 2)
+```
+
+```{r top_four_tornado, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Counties exposed for the top four storms over the study period, in terms of tornado event-based exposure."}
+top_tornado_exposed <- function(storm_id){
+        my_map <- map_event_exposure(storm_id, event_type = "tornado")
+        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
+                             plot_points = FALSE) + 
+                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
+}
+
+all_tornado <- county_events(counties = my_fips, start_year = 1996, 
+                        end_year = 2011, event_type = "tornado")
+
+tornado_exposed <- all_tornado %>%
+        group_by(storm_id) %>%
+        summarize(number_of_counties_exposed = n()) %>%
+        arrange(desc(number_of_counties_exposed)) %>%
+        top_n(10) %>%
+        ungroup() %>%
+        mutate(storm_id = as.character(storm_id))
+
+n <- 4
+tornado_plots <- lapply(tornado_exposed$storm_id[1:n], top_tornado_exposed)
+names(tornado_plots) <- letters[1:n]
+grid.arrange(tornado_plots$a, tornado_plots$b, tornado_plots$c,
+             tornado_plots$d, ncol = 2)
+```
+
+Here is a table of the top 10 storms by number of counties exposed, based on a distance metric of within 100 kilometers or less of the storm track:
+
+```{r echo = FALSE}
+distance_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of distance.")
+```
+
+Here is a table of the top 10 storms by number of counties exposed, based on a rain metric of 75 millimeters or more and within 500 kilometers or less of the storm track:
+
+```{r echo = FALSE}
+rain_exposed %>%
+        mutate(storm_id = gsub("-", " (", storm_id),
+               storm_id = paste0(storm_id, ")")) %>%
+        kable(col.names = c("Storm", "# of exposed counties"),
+              caption = "Top 10 storms based on the number of counties exposed, using a metric of rainfall.")
 ```
@@
 
-```{r echo = FALSE, message = FALSE, warning = FALSE}
-all_wind <- county_wind(counties = my_fips, start_year = 1988, 
-                        end_year = 2011, wind_limit = 15)
-
-wind_exposed <- all_wind %>%
-        group_by(storm_id) %>%
-        summarize(number_of_counties_exposed = n()) %>%
-        arrange(desc(number_of_counties_exposed)) %>%
-        top_n(10)
-
+```{r echo = FALSE}
 wind_exposed %>%
@@
               caption = "Top 10 storms based on the number of counties exposed, using a metric of wind.")
-```
-
-Here are plots of the top four storms based on number of counties exposed by wind:
-
-```{r echo = FALSE, fig.width = 10, fig.height = 8}
-top_wind_exposed <- function(storm_id){
-        my_map <- map_wind_exposure(storm_id, wind_limit = 18)
-        my_map <- map_tracks(storm = storm_id, plot_object = my_map, 
-                             plot_points = FALSE) + 
-                ggtitle(paste0(gsub("-", " (", storm_id), ")"))
-}
-
-n <- 4
-wind_plots <- lapply(wind_exposed$storm_id[1:n], top_wind_exposed)
-names(wind_plots) <- letters[1:n]
-grid.arrange(wind_plots$a, wind_plots$b, wind_plots$c,
-             wind_plots$d, ncol = 2)
 ```

### 2016-08-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work on draft

@@
                 dplyr::group_by(fips) %>%
-                dplyr::summarize(count = n() /(end_year - start_year + 1)) %>%
+                dplyr::summarize(count = 10 * n() /(end_year - start_year + 1)) %>%
                 dplyr::rename(region = fips) %>%
@@
         exposure_count$value <- cut(exposure_count$count, c(0, 
-                                                            0.11,
-                                                            0.22,
-                                                            0.33,
-                                                            0.44,
-                                                            0.56,
-                                                            0.67,
-                                                            0.78,
-                                                            0.89,
-                                                            10),
-                           labels = c("0-.10", ".11-.21",
-                                      ".22-.32", ".33-.43",
-                                      ".44-.55", ".56-.66",
-                                      ".67-.77", ".78-.88", ".89+"),
+                                                            1,
+                                                            2,
+                                                            3,
+                                                            4,
+                                                            5,
+                                                            6,
+                                                            7,
+                                                            8,
+                                                            20),
+                           labels = c("0-1", "1-2",
+                                      "2-3", "3-4",
+                                      "4-5", "5-6",
+                                      "6-7", "7-8", "8+"),
                            include.lowest = TRUE)
@@
                start_year = 1996, end_year = 2011)
+```
+
+We assessed some statistics on the number of counties exposed to storms under different exposure metrics (Table \ref{tab:num_counties_exposed}).
+
+```{r echo = FALSE, results = "asis"}
+distance_exposure_50 <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2011, dist_limit = 50)
+distance_exposure_75 <- county_distance(counties = my_fips, start_year = 1988,
+                                     end_year = 2011, dist_limit = 75)
+rain_exposure_125 <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 125, dist_limit = 500)
+rain_exposure_175 <- county_rain(counties = my_fips,
+                             start_year = 1988, end_year = 2011,
+                             rain_limit = 175, dist_limit = 500)
+wind_exposure_20 <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2011, wind_limit = 20)
+wind_exposure_25 <- county_wind(counties = my_fips, start_year = 1988,
+                             end_year = 2011, wind_limit = 25)
+
+dist_exp <- distance_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(dist_exposed = 1)
+dist_exp_50 <- distance_exposure_50 %>%
+        select(storm_id, fips) %>%
+        mutate(dist_exposed_50 = 1)
+dist_exp_75 <- distance_exposure_75 %>%
+        select(storm_id, fips) %>%
+        mutate(dist_exposed_75 = 1)
+rain_exp <- rain_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed = 1)
+rain_exp <- rain_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed = 1)
+rain_exp_125 <- rain_exposure_125 %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed_125 = 1)
+rain_exp_175 <- rain_exposure_175 %>%
+        select(storm_id, fips) %>%
+        mutate(rain_exposed_175 = 1)
+wind_exp <- wind_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed = 1)
+wind_exp_20 <- wind_exposure_20 %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed_20 = 1)
+wind_exp_25 <- wind_exposure_25 %>%
+        select(storm_id, fips) %>%
+        mutate(wind_exposed_25 = 1)
+
+all_exp <- dist_exp %>%
+        full_join(dist_exp_75, by = c("storm_id", "fips")) %>%
+        full_join(dist_exp_50, by = c("storm_id", "fips")) %>%
+        full_join(rain_exp, by = c("storm_id", "fips")) %>%
+        full_join(rain_exp_125, by = c("storm_id", "fips")) %>%
+        full_join(rain_exp_175, by = c("storm_id", "fips")) %>%
+        full_join(wind_exp, by = c("storm_id", "fips")) %>%
+        full_join(wind_exp_20, by = c("storm_id", "fips")) %>%
+        full_join(wind_exp_25, by = c("storm_id", "fips")) %>%
+        gather(metric, exposure, -storm_id, -fips) %>%
+        mutate(exposure = !is.na(exposure))
+
+flood_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "flood")
+tornado_exposure <- county_events(counties = my_fips,
+                              start_year = 1996, end_year = 2011,
+                              event_type = "tornado")
+
+flood_exp <- flood_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(flood_exposed = 1)
+tornado_exp <- tornado_exposure %>%
+        select(storm_id, fips) %>%
+        mutate(tornado_exposed = 1)
+
+all_exp_2 <- flood_exp %>%
+        full_join(tornado_exp, by = c("storm_id", "fips")) %>%
+        gather(metric, exposure, -storm_id, -fips) %>%
+        mutate(exposure = !is.na(exposure),
+               storm_id = as.character(storm_id))  %>%
+        dplyr::group_by(storm_id, metric) %>%
+        dplyr::summarize(n_exposed = sum(exposure)) %>%
+        dplyr::ungroup() %>%
+        dplyr::group_by(metric) %>%
+        dplyr::summarize(median_exposure = median(n_exposed),
+                         iqr_exposure = paste(quantile(n_exposed, c(0.25, 0.75)),
+                                              collapse = ", "), 
+                         median_exposure = paste0(median_exposure, " (",
+                                                  iqr_exposure, ")"),
+                         max_exposure = max(n_exposed),
+                         which_max_exposure = storm_id[which.max(n_exposed)],
+                         which_max_exposure = gsub("-", ", ", which_max_exposure),
+                         max_exposure = paste0(max_exposure, " (",
+                                               which_max_exposure, ")"), 
+                         n_100_plus = sum(n_exposed >= 100) / 10) %>%
+        dplyr::select(-iqr_exposure, -which_max_exposure)
+
+all_exp %>%
+        dplyr::group_by(storm_id, metric) %>%
+        dplyr::summarize(n_exposed = sum(exposure)) %>%
+        dplyr::ungroup() %>%
+        dplyr::group_by(metric) %>%
+        dplyr::summarize(median_exposure = median(n_exposed),
+                         iqr_exposure = paste(quantile(n_exposed, c(0.25, 0.75)),
+                                              collapse = ", "), 
+                         median_exposure = paste0(median_exposure, " (",
+                                                  iqr_exposure, ")"),
+                         max_exposure = max(n_exposed),
+                         which_max_exposure = storm_id[which.max(n_exposed)],
+                         which_max_exposure = gsub("-", ", ", which_max_exposure),
+                         max_exposure = paste0(max_exposure, " (",
+                                               which_max_exposure, ")"), 
+                         n_100_plus = sum(n_exposed >= 100) / 10) %>%
+        dplyr::select(-iqr_exposure, -which_max_exposure) %>%
+        rbind(all_exp_2) %>%
+        dplyr::mutate(metric = c("Distance (100 km or less)",
+                                 "Distance (75 km or less)",
+                                 "Distance (50 km or less)",
+                                 "Rainfall (75 mm or more)",
+                                 "Rainfall (125 mm or more)",
+                                 "Rainfall (175 mm or more)",
+                                 "Wind (15 m/s or more)",
+                                 "Wind (20 m/s or more)",
+                                 "Wind (25 m/s or more)",
+                                 "Flood event",
+                                 "Tornado event")) %>%
+        xtable(caption = "County-level exposure statistics for different exposure metrics",
+               label = "tab:num_counties_exposed",
+               digits = c(0, 0, 0, 0, 1)) %>%
+        print(include.rownames = FALSE, floating = TRUE,
+              table.placement = "p", caption.placement = "top", 
+              booktabs = TRUE, comment = FALSE, timestamp = NULL)
 ```

### 2016-08-07 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Work some on agreement between metrics

@@
 
+Here is some analysis of agreement and disagreement among different metrics:
+
+```{r echo = FALSE, results = "asis"}
+all_exp <- dist_exp %>%
+        full_join(rain_exp, by = c("storm_id", "fips")) %>%
+        full_join(wind_exp, by = c("storm_id", "fips")) %>%
+        gather(metric, exposure, -storm_id, -fips) %>%
+        mutate(exposure = !is.na(exposure)) %>%
+        spread(metric, exposure)
+
+check_agreement <- function(data, x1, x2, n_counties = 2396){
+        data$x1 <- data[ , x1]
+        data$x2 <- data[ , x2]
+        agreement <- data %>%
+                dplyr::select(storm_id, x1, x2) %>%
+                dplyr::mutate(agree_exp = x1 & x2,
+                              x1_exp_x2_unexp = x1 & !x2,
+                              x1_unexp_x2_exp = !x1 & x2,
+                              any_exp = x1 | x2) %>%
+                dplyr::group_by(storm_id) %>%
+                dplyr::summarize(disagree = sum(!agree_exp),
+                                 agree = (n_counties - disagree),
+                                 agree_exp = sum(agree_exp),
+                                 agree_unexp = (n_counties - disagree - agree_exp),
+                                 disagree_x1_exp = sum(x1_exp_x2_unexp),
+                                 disagree_x2_exp = sum(x1_unexp_x2_exp),
+                                 agree_any = round(100 * agree_exp / sum(any_exp), 1),
+                                 disagree_any_x1_exp = round(100 * disagree_x1_exp /
+                                                                     sum(any_exp), 1),
+                                 disagree_any_x2_exp = round(100 * disagree_x2_exp /
+                                                                     sum(any_exp), 1))
+        agreement[ , 2:7] <- round(100 * agreement[ , 2:7] / n_counties, 1)
+        return(agreement)
+}
+
+summarize_agreement <- function(agreement){
+        med_val <- apply(agreement[ , -1], MARGIN = 2, FUN = median, na.rm = TRUE)
+        min_val <- apply(agreement[ , -1], MARGIN = 2, FUN = min, na.rm = TRUE)
+        max_val <- apply(agreement[ , -1], MARGIN = 2, FUN = max, na.rm = TRUE)
+        min_storm <- apply(agreement[ , -1], MARGIN = 2,
+                           FUN = function(x){
+                                   i <- which.min(x)
+                                   out <- agreement$storm_id[i]
+                                   return(out)
+                                   })
+        max_storm <- apply(agreement[ , -1], MARGIN = 2,
+                           FUN = function(x){
+                                   i <- which.max(x)
+                                   out <- agreement$storm_id[i]
+                                   return(out)
+                           })
+        out <- rbind(t(as.data.frame(med_val)), t(as.data.frame(min_val)),
+                     t(as.data.frame(max_val)), t(as.data.frame(min_storm)),
+                     t(as.data.frame(max_storm)))
+        out <- as.data.frame(out)
+        out$metric <- row.names(out)
+        return(out)
+}
+
+exp_metrics <- c("dist_exposed", "rain_exposed", "wind_exposed")
+
+agg_exp_matrix <- matrix(NA, ncol = length(exp_metrics), nrow = length(exp_metrics))
+colnames(agg_exp_matrix) <- rownames(agg_exp_matrix) <- exp_metrics
+agg_unexp_matrix <- x1_exp_matrix <- x2_exp_matrix <- agg_exp_matrix
+for(i in 1:length(exp_metrics)){
+        for(j in 1:length(exp_metrics)){
+                x1 <- exp_metrics[i]
+                x2 <- exp_metrics[j]
+                # if(x1 == x2) next
+
+                agreement <- check_agreement(all_exp, x1 = x1, x2 = x2)
+                agg_exp_matrix[i, j] <- median(agreement$agree_exp, na.rm = TRUE)
+                agg_unexp_matrix[i, j] <- median(agreement$agree_unexp, na.rm = TRUE)
+                x1_exp_matrix[i, j] <- median(agreement$disagree_x1_exp, na.rm = TRUE)
+                x2_exp_matrix[i, j] <- median(agreement$disagree_x2_exp, na.rm = TRUE)
+        }
+}
+
+out <- rbind(cbind(agg_exp_matrix, x1_exp_matrix),
+             cbind(x2_exp_matrix, agg_unexp_matrix)) 
+row.names(out) <- NULL
+colnames(out) <- NULL
+out <- xtable(as.data.frame(out), caption = "Agreement between metrics.",
+              align = rep("c", 7), digits = 1)
+print(out, include.rownames = FALSE, floating = TRUE,
+      table.placement = "p", caption.placement = "top", 
+      booktabs = TRUE, comment = FALSE, timestamp = NULL)
+```
+
 We first investigated whether different exposure metrics resulted in similar numbers of counties classified as "exposed" to a storm (Figure \ref{fig:county_number_cor}, Table \ref{tab:county_number}).

### 2016-08-09 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add in temporal trend plots for exposures

@@
 
+Here are time trends for some different metrics:
+
+```{r echo = FALSE}
+tornado_events <- county_events(counties = my_fips, event_type = "tornado",
+                                start_year = 1988, end_year = 2011) %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(n_counties = n()) %>%
+        dplyr::mutate(year = as.numeric(gsub(".+-", "", storm_id)))
+
+tornado_year_mean <- tornado_events %>%
+        dplyr::group_by(year) %>%
+        dplyr::summarize(year_mean = mean(n_counties),
+                         year_tot = sum(n_counties))
+
+a <- ggplot(tornado_events, aes(x = year, y = n_counties)) + 
+        geom_point(alpha = 0.5) + 
+        geom_line(data = tornado_year_mean, aes(y = year_tot), 
+                   color = "blue", alpha = 0.2, size = 1) + 
+        geom_point(data = tornado_year_mean, aes(y = year_tot), 
+                   color = "blue", alpha = 0.2, size = 3) + 
+        theme_few() + 
+        xlab("Year") + ylab("Number of exposed counties")+ 
+        ggtitle("Patterns in county exposure to tornado events")
+
+flood_events <- county_events(counties = my_fips, event_type = "flood",
+                                start_year = 1988, end_year = 2011) %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(n_counties = n()) %>%
+        dplyr::mutate(year = as.numeric(gsub(".+-", "", storm_id)))
+
+flood_year_mean <- flood_events %>%
+        dplyr::group_by(year) %>%
+        dplyr::summarize(year_mean = mean(n_counties),
+                         year_tot = sum(n_counties))
+
+b <- ggplot(flood_events, aes(x = year, y = n_counties)) + 
+        geom_point(alpha = 0.5) + 
+        geom_line(data = flood_year_mean, aes(y = year_tot), 
+                  color = "blue", alpha = 0.2, size = 1) + 
+        geom_point(data = flood_year_mean, aes(y = year_tot), 
+                   color = "blue", alpha = 0.2, size = 3) + 
+        theme_few() + 
+        xlab("Year") + ylab("Number of exposed counties") + 
+        ggtitle("Patterns in county exposure to flood events")
+
+rain_events <- county_rain(counties = my_fips, rain_limit = 75, dist_limit = 500,
+                              start_year = 1988, end_year = 2011) %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(n_counties = n()) %>%
+        dplyr::mutate(year = as.numeric(gsub(".+-", "", storm_id)))
+
+rain_year_mean <- rain_events %>%
+        dplyr::group_by(year) %>%
+        dplyr::summarize(year_mean = mean(n_counties),
+                         year_tot = sum(n_counties))
+
+c <- ggplot(rain_events, aes(x = year, y = n_counties)) + 
+        geom_point(alpha = 0.5) + 
+        geom_line(data = rain_year_mean, aes(y = year_tot), 
+                  color = "blue", alpha = 0.2, size = 1) + 
+        geom_point(data = rain_year_mean, aes(y = year_tot), 
+                   color = "blue", alpha = 0.2, size = 3) + 
+        theme_few() + 
+        xlab("Year") + ylab("Number of exposed counties") + 
+        ggtitle("Patterns in county exposure to rain")
+
+wind_events <- county_wind(counties = my_fips, wind_limit = 15, 
+                           start_year = 1988, end_year = 2011) %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(n_counties = n()) %>%
+        dplyr::mutate(year = as.numeric(gsub(".+-", "", storm_id)))
+
+wind_year_mean <- wind_events %>%
+        dplyr::group_by(year) %>%
+        dplyr::summarize(year_mean = mean(n_counties),
+                         year_tot = sum(n_counties))
+
+d <- ggplot(wind_events, aes(x = year, y = n_counties)) + 
+        geom_point(alpha = 0.5) + 
+        geom_line(data = wind_year_mean, aes(y = year_tot), 
+                  color = "blue", alpha = 0.2, size = 1) + 
+        geom_point(data = wind_year_mean, aes(y = year_tot), 
+                   color = "blue", alpha = 0.2, size = 3) + 
+        theme_few() + 
+        xlab("Year") + ylab("Number of exposed counties") + 
+        ggtitle("Patterns in county exposure to wind")
+
+distance_events <- county_distance(counties = my_fips, dist_limit = 100, 
+                           start_year = 1988, end_year = 2011) %>%
+        dplyr::group_by(storm_id) %>%
+        dplyr::summarize(n_counties = n()) %>%
+        dplyr::mutate(year = as.numeric(gsub(".+-", "", storm_id)))
+
+distance_year_mean <- distance_events %>%
+        dplyr::group_by(year) %>%
+        dplyr::summarize(year_mean = mean(n_counties),
+                         year_tot = sum(n_counties))
+
+e <- ggplot(distance_events, aes(x = year, y = n_counties)) + 
+        geom_point(alpha = 0.5) + 
+        geom_line(data = distance_year_mean, aes(y = year_tot), 
+                  color = "blue", alpha = 0.2, size = 1) + 
+        geom_point(data = distance_year_mean, aes(y = year_tot), 
+                   color = "blue", alpha = 0.2, size = 3) + 
+        theme_few() + 
+        xlab("Year") + ylab("Number of exposed counties") + 
+        ggtitle("Patterns in county exposure to distance")
+
+grid.arrange(a, c, d, e, ncol = 1)
+```
+
 We assessed some statistics on the number of counties exposed to storms under different exposure metrics (Table \ref{tab:num_counties_exposed}).

### 2016-08-10 R/map_exposure.R
Change mapping code to map spline-interpolated tracks

@@ -68,8 +68,9 @@ default_map <- function(){
 #' @examples
 #' map_tracks(storms = "Sandy-2012")
 #' map_tracks(storms = "Floyd-1999", plot_points = TRUE)
-#' a <- map_tracks(storms = "Sandy-2012")
-#' b <- map_tracks(storms = "Floyd-1999", plot_object = a, plot_points = FALSE)
+#' map_tracks(storms = c("Sandy-2012", "Floyd-1999"))
+#' a <- map_tracks(storms = "Sandy-2012", color = "blue", alpha = 0.3)
+#' b <- map_tracks(storms = "Floyd-1999", plot_object = a)
 #' b
 #'
 #' @importFrom dplyr %>%
@@ -87,14 +88,21 @@ map_tracks <- function(storms, plot_object = NULL,
                          MARGIN = 2,
                          function(x) range(x) + c(-1, 1) * padding)
         tracks <- hurricaneexposuredata::hurr_tracks %>%
-                dplyr::select_(~ latitude, ~ longitude, ~ storm_id) %>%
+                dplyr::select_(~ latitude, ~ longitude, ~ storm_id,
+                               ~ date) %>%
                 dplyr::filter_(~ as.character(storm_id) %in% storms &
                               longitude > map_dim[1, 1] &
                               longitude < map_dim[2, 1] &
                               latitude > map_dim[1, 2] &
-                              latitude < map_dim[2, 2])
+                              latitude < map_dim[2, 2]) %>%
+                dplyr::mutate_(date = ~ lubridate::ymd_hm(date))
+
+        splt_tracks <- split(tracks, tracks$storm_id)
+        full_tracks <- lapply(splt_tracks, interp_track)
+        full_tracks <- do.call("rbind", full_tracks)
+
         out <- plot_object +
-                        ggplot2::geom_path(data = tracks,
+                        ggplot2::geom_path(data = full_tracks,
                                            ggplot2::aes_(x = ~ longitude,
                                                         y = ~ latitude,
                                                         group = ~ storm_id),
@@ -109,6 +117,43 @@ map_tracks <- function(storms, plot_object = NULL,
                                                     alpha = alpha)
         }
         return(out)
+}
+
+#' Interpolate a storm track
+#'
+#' This function takes a wider-spaced storm track (e.g., every 6 hours) and
+#' interpolates to a finer interval (e.g., every 15 minutes). To do this, it
[…]

### 2016-08-10 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add method details for spline interpolation of tracks.

@@
 These positions are given to within 0.1 degrees latitude / longitude at these synoptic times [@Landsea2013].
-We interpolated these location values to every 15 minutes during the period when the storm was active, using a linear interpolation [should we somehow be using great circle distance to do this?] between each measured point.
+
+We interpolated these location values to every 15 minutes during the period when the storm was active, using spline interpolation.
+We fit linear models to predict latitude, longitude, and wind intensity based on a linear spline of date-time using the best tracks data, and then predicted to 15 minute-intervals using these models to generate interpolated tracks.
+We determined the smoothness of these interpolation splines by using degrees of freedom equal to half the number of observations in the original best tracks.
 
@@
 From these distance calculations, we identified the date-time and distance between the storm track and the county center at the 15-minute interval when the storm was closest to the county center.
+
 We identified the date when the storm was closest to each county and converted this date-time  to the county's local time zone [citation-- `countytimezones` R package].

### 2016-09-08 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Update now that `rnoaa` is on CRAN

@@
 library(xtable)
-
-#devtools::install_github("ropenscilabs/rnoaa")
 library(rnoaa)
+
 #devtools::install_github("leighseverson/countyweather")
@@
         ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
-        ex_weather <- as.data.frame(ex_weather$averaged)
+        ex_weather <- as.data.frame(ex_weather$daily_weather)
 
@@
                                start_year = 1988, end_year = 2011,
-                               rain_limit = 0, dist_limit = 2000) %>%
+                               rain_limit = 0, dist_limit = 500) %>%
                 full_join(ex_weather, by = "storm_id") %>%
                 filter(!is.na(prcp) & !is.na(tot_precip))
+
+        ave_n <- round(mean(ex_rain$ave_n, na.rm = TRUE))
+        n_storms <- nrow(ex_rain)
+        ex_title <- paste0(ex_title, "\n(", n_storms, "; ", ave_n, ")")
 
@@
 i <- plot_county_rain_compare(ex_fips = "01097", ex_dir = "mobile_data/",
-                         ex_title = "Mobile County, AL", get_data = TRUE)
+                         ex_title = "Mobile County, AL")
 

### 2016-09-08 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Update rain data pulling for newest version of countyweather

@@
         if(get_data){
-                county_timeseries(ex_fips, percent_coverage = 0,
+                daily_timeseries(ex_fips, coverage = 0,
                                   date_min = "1988-01-01",
@@
         ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
-        ex_weather <- as.data.frame(ex_weather$daily_weather)
+        ex_weather <- as.data.frame(ex_weather$daily_data)
 
@@
                 full_join(ex_weather, by = "storm_id") %>%
-                filter(!is.na(prcp) & !is.na(tot_precip))
+                filter(!is.na(prcp) & !is.na(tot_precip)) %>%
+                mutate(prcp = prcp / 10) ## Units for countyweather are now 10ths
+                                         ## of millimeters for precipitation
 
@@
         n_storms <- nrow(ex_rain)
-        ex_title <- paste0(ex_title, "\n(", n_storms, "; ", ave_n, ")")
+        ex_title <- paste0(ex_title, "\n(storms: ", n_storms,
+                           "; monitors:", ave_n, ")")
 

### 2016-11-12 R/filter_storm_data.R
update max_sust to vmax_sust, etc., based on swm updates

@@ -98,7 +98,7 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 #' @param wind_limit A numeric vector of length one giving the minimum
 #'    wind speed (in meters per second) to use in the filter
 #' @param wind_var A character string giving the wind variable to use. Choices
-#'    are "max_sust" (maximum sustained winds; default) or "max_gust" (maximum
+#'    are "vmax_sust" (maximum sustained winds; default) or "vmax_gust" (maximum
 #'    gust winds).
 #'
 #' @return A dataframe with storms filtered based on the input criteria to the
@@ -108,17 +108,14 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 #' @examples
 #' filter_wind_data(counties = c("22071", "51700"), year_range = c(1988, 2011),
 #'                  wind_limit = 20,
-#'                  output_vars = c("fips", "storm_id", "max_sust"))
-#' filter_wind_data(storm = "Floyd-1999", include_rain = TRUE,
-#'                  days_included = c(-1, 0, 1),
-#'                  output_vars = c("fips", "tot_precip"))
+#'                  output_vars = c("fips", "storm_id", "vmax_sust"))
 #'
 #' @import data.table
 #'
 #' @export
 filter_wind_data <- function(counties = NULL, storm = NULL, year_range = NULL,
                              wind_limit = NULL, output_vars = "fips",
-                             wind_var = "max_sust"){
+                             wind_var = "vmax_sust"){
         storm_winds <- data.table::data.table(hurricaneexposuredata::storm_winds)
 
         if(!is.null(counties)){
@@ -132,7 +129,7 @@ filter_wind_data <- function(counties = NULL, storm = NULL, year_range = NULL,
         if(!is.null(year_range)){
                 storm_winds <- storm_winds[ , .(storm_id,
                                                 fips,
-                                                max_gust, max_sust,
+                                                vmax_gust, vmax_sust,
                                                 year = gsub("*.+-", "", get("storm_id"))), ][
                                                         get("year") %in%
                                                                 year_range[1]:year_range[2]
@@ -140,11 +137,11 @@ filter_wind_data <- function(counties = NULL, storm = NULL, year_range = NULL,
         }
 
         if(!is.null(wind_limit)){
-                if(wind_var == "max_sust"){
-                        storm_winds <- storm_winds[get("max_sust") >=
[…]

### 2016-11-12 R/map_exposure.R

@@ -63,8 +63,6 @@ default_map <- function(){
 #'    of the selected storms. This object can be printed directly or added
 #'    on to with other ggplot commands.
 #'
-#' @export
-#'
 #' @examples
 #' map_tracks(storms = "Sandy-2012")
 #' map_tracks(storms = "Floyd-1999", plot_points = TRUE)
@@ -74,6 +72,8 @@ default_map <- function(){
 #' b
 #'
 #' @importFrom dplyr %>%
+#'
+#' @export
 map_tracks <- function(storms, plot_object = NULL,
                       padding = 2,
                       plot_points = FALSE,
@@ -135,24 +135,32 @@ map_tracks <- function(storms, plot_object = NULL,
 #' @return A dataframe with hurricane track data for a single storm,
 #'    interpolated to the interval specified by \code{tint}.
 interp_track <- function(track, tint = 0.25){
-        interp_df <- floor(nrow(track) / 2)
-        interp_date <- seq(from = min(track$date),
-                           to = max(track$date),
-                           by = 900) # interpolate to 15 minutes
-        interp_date <- data.frame(date = interp_date)
 
-        lat_spline <- stats::glm(latitude ~ splines::ns(date, df = interp_df),
-                                 data = track)
-        interp_lat <- stats::predict.glm(lat_spline,
+        if(nrow(track) < 3){
+                return(track)
+        } else {
+                interp_df <- floor(nrow(track) / 2)
+                interp_date <- seq(from = min(track$date),
+                                to = max(track$date),
+                                by = 900) # interpolate to 15 minutes
+                interp_date <- data.frame(date = interp_date)
+
+                lat_spline <- stats::glm(latitude ~ splines::ns(date,
+                                                                df = interp_df),
+                                        data = track)
+                interp_lat <- stats::predict.glm(lat_spline,
                                          newdata = as.data.frame(interp_date))
-        lon_spline <- stats::glm(longitude ~ splines::ns(date, df = interp_df),
-                                 data = track)
-        interp_lon <- stats::predict.glm(lon_spline, newdata = interp_date)
+                lon_spline <- stats::glm(longitude ~ splines::ns(date,
+                                                                 df = interp_df),
+                                        data = track)
+                interp_lon <- stats::predict.glm(lon_spline,
+                                                 newdata = interp_date)
 
-        full_track <- data.frame(storm_id = track$storm_id[1],
-                                 date = interp_date,
-                                 latitude = interp_lat,
[…]

### 2016-11-12 R/wind_exposure.R

@@ -31,20 +31,20 @@
 #' @examples
 #' county_wind(counties = c("22071", "51700"),
 #'             start_year = 1988, end_year = 2005,
-#'             wind_limit = 20, wind_var = "max_sust")
+#'             wind_limit = 20, wind_var = "vmax_sust")
 #'
 #' @export
 #'
 #' @importFrom dplyr %>%
 county_wind <- function(counties, start_year, end_year, wind_limit,
-                        wind_var = "max_sust"){
+                        wind_var = "vmax_sust"){
 
         wind_df <- filter_wind_data(counties = counties,
                                          year_range = c(start_year, end_year),
                                          wind_limit = wind_limit,
                                          output_vars = c("storm_id", "fips",
-                                                         "max_sust",
-                                                         "max_gust")) %>%
+                                                         "vmax_sust",
+                                                         "vmax_gust")) %>%
                 dplyr::left_join(hurricaneexposuredata::closest_dist,
                                  by = c("storm_id", "fips"))
 
@@ -91,14 +91,14 @@ multi_county_wind <- function(communities, start_year, end_year,
                 dplyr::left_join(hurricaneexposuredata::closest_dist,
                                  by = c("storm_id", "fips")) %>%
                 dplyr::group_by_(~ commun, ~ storm_id) %>%
-                dplyr::mutate_(max_wind = ~ max(max_sust),
+                dplyr::mutate_(max_wind = ~ max(vmax_sust),
                                min_dist = ~ min(storm_dist)) %>%
                 dplyr::filter_(~ max_wind >= wind_limit) %>%
                 dplyr::summarize_(closest_date = ~ dplyr::first(closest_date),
                                   local_time = ~ dplyr::first(local_time),
                                   closest_time_utc = ~ dplyr::first(closest_time_utc),
                                   mean_dist = ~ mean(storm_dist),
-                                  mean_wind = ~ mean(max_sust),
+                                  mean_wind = ~ mean(vmax_sust),
                                   min_dist = ~ dplyr::first(min_dist),
                                   max_wind = ~ dplyr::first(max_wind))
         return(wind_df)

### 2016-11-17 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Update README

@@
                                   var = "PRCP",
-                                  out_directory = ex_dir)
+                                  out_directory = ex_dir,
+                                  keep_map = FALSE)
         }
@@
         ex_title <- paste0(ex_title, "\n(storms: ", n_storms,
-                           "; monitors:", ave_n, ")")
+                           "; monitors: ", ave_n, ")")
 
@@
 a <- plot_county_rain_compare(ex_fips = "12086", ex_dir = "dade_data/",
-                         ex_title = "Miami-Dade, FL")
+                         ex_title = "Miami-Dade, FL",
+                         get_data = FALSE)
 
 b <- plot_county_rain_compare(ex_fips = "22071", ex_dir = "orleans_data/",
-                         ex_title = "Orleans Parish, LA")
+                         ex_title = "Orleans Parish, LA",
+                         get_data = FALSE)
 
 c <- plot_county_rain_compare(ex_fips = "45019", ex_dir = "charleston_data/",
-                         ex_title = "Charleston County, SC")
+                         ex_title = "Charleston County, SC",
+                         get_data = FALSE)
 
 d <- plot_county_rain_compare(ex_fips = "37183", ex_dir = "wake_data/",
-                         ex_title = "Wake County, NC")
+                         ex_title = "Wake County, NC",
+                         get_data = FALSE)
 
 e <- plot_county_rain_compare(ex_fips = "13121", ex_dir = "fulton_data/",
-                         ex_title = "Fulton County, GA")
+                         ex_title = "Fulton County, GA",
+                         get_data = FALSE)
 
 f <- plot_county_rain_compare(ex_fips = "24005", ex_dir = "baltimore_data/",
-                         ex_title = "Baltimore County, MD")
+                         ex_title = "Baltimore County, MD",
+                         get_data = FALSE)
 
@@
                               ex_dir = "philadelphia_data/",
-                         ex_title = "Philadelphia County, PA")
+                         ex_title = "Philadelphia County, PA",
+                         get_data = FALSE)
 
 h <- plot_county_rain_compare(ex_fips = "48201", ex_dir = "harris_data/",
-                         ex_title = "Harris County, TX")
+                         ex_title = "Harris County, TX",
+                         get_data = FALSE)
 
 i <- plot_county_rain_compare(ex_fips = "01097", ex_dir = "mobile_data/",
-                         ex_title = "Mobile County, AL")
+                         ex_title = "Mobile County, AL",
+                         get_data = FALSE)
 
@@
 ```{r echo = FALSE, eval = FALSE}
-options("noaakey" = Sys.getenv("noaakey"))
-
 allison_gauge <- read.csv("MWR_gauge_rain_data/allison_mwr.csv") 
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-county_timeseries(allison_fips, percent_coverage = 0,
+daily_timeseries(allison_fips, coverage = 0,
                   date_min = format(date_range[1, 1], "%Y-%m-%d"),
                   date_max = format(date_range[1, 2], "%Y-%m-%d"),
-                  var = "PRCP", out_directory = "county_ave/allison/")
+                  var = "PRCP", out_directory = "county_ave/allison/",
+                 keep_map = FALSE)
 
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-county_timeseries(frances_fips, percent_coverage = 0,
+daily_timeseries(frances_fips, coverage = 0,
                   date_min = format(date_range[1, 1], "%Y-%m-%d"),
                   date_max = format(date_range[1, 2], "%Y-%m-%d"),
-                  var = "PRCP", out_directory = "county_ave/frances/")
+                  var = "PRCP", out_directory = "county_ave/frances/",
+                 keep_map = FALSE)
 
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-county_timeseries(ike_fips, percent_coverage = 0,
+daily_timeseries(ike_fips, coverage = 0,
                   date_min = format(date_range[1, 1], "%Y-%m-%d"),
                   date_max = format(date_range[1, 2], "%Y-%m-%d"),
-                  var = "PRCP", out_directory = "county_ave/ike/")
+                  var = "PRCP", out_directory = "county_ave/ike/",
+                 keep_map = FALSE)
 
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-county_timeseries(alex_fips, percent_coverage = 0,
+daily_timeseries(alex_fips, coverage = 0,
                   date_min = format(date_range[1, 1], "%Y-%m-%d"),
                   date_max = format(date_range[1, 2], "%Y-%m-%d"),
-                  var = "PRCP", out_directory = "county_ave/alex/")
+                  var = "PRCP", out_directory = "county_ave/alex/",
+                 keep_map = FALSE)
 ```
@@
                          list.files(paste0("county_ave/", county_dir)))
+        if(county_dir == "frances"){
+                counties <- counties[counties != "51019"] # Missing all data
+        }
 
@@
                                                  county_dir, "/", x,
-                                                 ".rds"))$averaged %>%
-                        dplyr::ungroup() %>%
-                        dplyr::mutate(date = ymd(date)) %>%
+                                                 ".rds"))$daily_data %>%
+                        dplyr::mutate(date = ymd(date),
+                                      prcp = prcp / 10) %>%
                         dplyr::mutate(fips = x)
@@
                   prcp_reporting = mean(prcp_reporting))
-        plot_range <- range(county_ave[ , c("monitor_rain", "tot_precip")]) + 
+        plot_range <- range(county_ave[ , c("monitor_rain", "tot_precip")], 
+                            na.rm = TRUE) + 
                 c(-15, 15)
@@
 
-grid.arrange(allison_county_ave, frances_county_ave,
+grid.arrange(allison_county_ave,
+             frances_county_ave,
              ike_county_ave, alex_county_ave, ncol = 2)
@@
 
+```{r pareto_1, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Number of total county exposures summed across percentile bins of storms"}
+my_pareto_1 <- function(data, title = "Rain exposure"){
+        data <- data %>%
+                select(storm_id, fips) %>%
+                group_by(storm_id) %>%
+                summarize(n = n()) %>%
+                mutate(cat = ntile(n, 10)) %>%
+                group_by(cat) %>%
+                summarize(n = sum(n)) %>%
+                ungroup() %>%
+                arrange(desc(n)) %>%
+                mutate(cumulative = cumsum(n),
+                       group = 1:n(),
+                       cat_labs = c("90-100th", "80-90th", "70-80th",
+                               "60-70th", "50-60th", "40-50th",
+                               "30-40th", "20-30th", "10-20th",
+                               "0-10th")) %>%
+                mutate(percent = paste0(round(100 * cumulative / sum(n)), "%"))
+        my_pareto_plot <- ggplot(data, aes(x = group, y = n)) + 
+                geom_bar(stat = "identity") + 
+                geom_point(aes(x = group, y = cumulative), alpha = 0.25) + 
+                geom_line(aes(x = group, y = cumulative), alpha = 0.25) + 
+                geom_text(aes(y = cumulative, label = percent), 
+                        hjust = "inward", vjust = "inward") + 
+                theme_few() + 
+                ylab("Number of\nexposed counties") + 
+                ggtitle(title) + 
+                scale_x_continuous(name = "", breaks = 1:10,
+                                labels = data$cat_labs, 
+                                trans = "reverse") +
+                coord_flip()
+        return(my_pareto_plot)
+}
+
+data <- county_distance(my_fips, start_year = 1988, end_year = 2011,
+                               dist_limit = 100) 
+a <- my_pareto_1(data, title = "Distance")
+
+data <- county_rain(my_fips, start_year = 1988, end_year = 2011,
+                               rain_limit = 75, dist_limit = 500) 
+b <- my_pareto_1(data, title = "Rain")
+
+data <- county_wind(my_fips, start_year = 1988, end_year = 2011,
+                               wind_limit = 15) 
+c <- my_pareto_1(data, title = "Wind")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "flood") 
+d <- my_pareto_1(data, title = "Floods")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "tornado") 
+e <- my_pareto_1(data, title = "Tornadoes")
+
+grid.arrange(a, b, c, d, e, nrow = 2)
+```
+
+```{r pareto_2, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Number of total storm hits summed across percentile bins of counties"}
+my_pareto_2 <- function(data, title = "Rain exposure"){
+        data <- data %>%
+                select(storm_id, fips) %>%
+                group_by(fips) %>%
+                summarize(n = n()) %>%
+                mutate(cat = ntile(n, 10)) %>%
+                group_by(cat) %>%
+                summarize(n = sum(n)) %>%
+                ungroup() %>%
+                arrange(desc(n)) %>%
+                mutate(cumulative = cumsum(n),
+                       group = 1:n(),
+                       cat_labs = c("90-100th", "80-90th", "70-80th",
+                               "60-70th", "50-60th", "40-50th",
+                               "30-40th", "20-30th", "10-20th",
+                               "0-10th")) %>%
+                mutate(percent = paste0(round(100 * cumulative / sum(n)), "%"))
+        my_pareto_plot <- ggplot(data, aes(x = group, y = n)) + 
+                geom_bar(stat = "identity") + 
+                geom_point(aes(x = group, y = cumulative), alpha = 0.25) + 
+                geom_line(aes(x = group, y = cumulative), alpha = 0.25) + 
+                geom_text(aes(y = cumulative, label = percent), 
+                        hjust = "inward", vjust = "inward") + 
+                theme_few() + 
+                ylab("Number of storm hits") + 
+                ggtitle(title) + 
+                scale_x_continuous(name = "", breaks = 1:10,
+                                labels = data$cat_labs, 
+                                trans = "reverse") +
+                coord_flip()
+        return(my_pareto_plot)
+}
+
+data <- county_distance(my_fips, start_year = 1988, end_year = 2011,
+                               dist_limit = 100) 
+a <- my_pareto_2(data, title = "Distance")
+
+data <- county_rain(my_fips, start_year = 1988, end_year = 2011,
+                               rain_limit = 75, dist_limit = 500) 
+b <- my_pareto_2(data, title = "Rain")
+
+data <- county_wind(my_fips, start_year = 1988, end_year = 2011,
+                               wind_limit = 15) 
+c <- my_pareto_2(data, title = "Wind")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "flood") 
+d <- my_pareto_2(data, title = "Floods")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "tornado") 
+e <- my_pareto_2(data, title = "Tornadoes")
+
+grid.arrange(a, b, c, d, e, nrow = 2)
+```
+
+```{r power_1, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Distribution of number of exposed counties by rank of storm"}
+my_power_1 <- function(data, title){
+        data <- data %>%
+                select(storm_id, fips) %>%
+                group_by(storm_id) %>%
+                summarize(n = n()) %>%
+                ungroup() %>%
+                arrange(desc(n)) %>%
+                mutate(rank = 1:n())
+        power_plot <- ggplot(data, aes(x = rank, y = n)) + 
+                geom_point(alpha = 0.5) + 
+                scale_y_continuous(trans = "log", breaks = c(1, 10, 100)) + 
+                theme_few() + 
+                ggtitle(title) + 
+                xlab("Rank of storm") + ylab("# of counties exposed")
+        return(power_plot)
+}
+
+data <- county_distance(my_fips, start_year = 1988, end_year = 2011,
+                               dist_limit = 100) 
+a <- my_power_1(data, title = "Distance")
+
+data <- county_rain(my_fips, start_year = 1988, end_year = 2011,
+                               rain_limit = 75, dist_limit = 500) 
+b <- my_power_1(data, title = "Rain")
+
+data <- county_wind(my_fips, start_year = 1988, end_year = 2011,
+                               wind_limit = 15) 
+c <- my_power_1(data, title = "Wind")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "flood") 
+d <- my_power_1(data, title = "Floods")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "tornado") 
+e <- my_power_1(data, title = "Tornadoes")
+
+grid.arrange(a, b, c, d, e, nrow = 2)
+```
+
+```{r power_2, echo = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", fig.cap = "Distribution of number of storm hits by rank of county"}
+my_power_2 <- function(data, title){
+        data <- data %>%
+                select(storm_id, fips) %>%
+                group_by(fips) %>%
+                summarize(n = n()) %>%
+                ungroup() %>%
+                arrange(desc(n)) %>%
+                mutate(rank = 1:n())
+        power_plot <- ggplot(data, aes(x = rank, y = n)) + 
+                geom_point(alpha = 0.5) + 
+                scale_y_continuous(trans = "log", breaks = c(1, 10, 100)) + 
+                theme_few() + 
+                ggtitle(title) + 
+                xlab("Rank of county") + ylab("# of storm hits")
+        return(power_plot)
+}
+
+data <- county_distance(my_fips, start_year = 1988, end_year = 2011,
+                               dist_limit = 100) 
+a <- my_power_2(data, title = "Distance")
+
+data <- county_rain(my_fips, start_year = 1988, end_year = 2011,
+                               rain_limit = 75, dist_limit = 500) 
+b <- my_power_2(data, title = "Rain")
+
+data <- county_wind(my_fips, start_year = 1988, end_year = 2011,
+                               wind_limit = 15) 
+c <- my_power_2(data, title = "Wind")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "flood") 
+d <- my_power_2(data, title = "Floods")
+
+data <- county_events(my_fips, start_year = 1996, end_year = 2011,
+                      event_type = "tornado") 
+e <- my_power_2(data, title = "Tornadoes")
+
+grid.arrange(a, b, c, d, e, nrow = 2)
+```
+
+```{r message = FALSE}
+all_storm_counties <- closest_dist %>%
+        dplyr::select(storm_id, fips)
+metric_a <- county_rain(my_fips, start_year = 1988, end_year = 2011,
+                               rain_limit = 75, dist_limit = 500) %>%
+        select(storm_id, fips) %>%
+        mutate(exposed_a = 1)
+metric_b <- county_wind(my_fips, start_year = 1988, end_year = 2011,
+                               wind_limit = 15) %>%
+        select(storm_id, fips) %>%
+        mutate(exposed_b = 1)
+two_metrics <- metric_a %>%
+        full_join(metric_b, by = c("storm_id", "fips")) %>%
+        full_join(all_storm_counties, by = c("storm_id", "fips")) %>%
+        mutate(exposed_a = ifelse(is.na(exposed_a), 0, 1),
+               exposed_b = ifelse(is.na(exposed_b), 0, 1)) %>%
+        group_by(storm_id) %>%
+        summarize(a = sum(exposed_a & exposed_b),
+                  b = sum(exposed_a & !exposed_b),
+                  c = sum(!exposed_a & exposed_b),
+                  d = sum(!exposed_a & !exposed_b),
+                  f1 = sum(exposed_b),
+                  f2 = sum(!exposed_b),
+                  g1 = sum(exposed_a),
+                  g2 = sum(!exposed_a),
+                  n = n(),
+                  no_exposure = a + b + c == 0,
+                  prev = abs(a - d) / n,
+                  bias = abs(b - c) / n,
+                  prop_agree = (a + d) / n,
+                  exp_agree = (f1 * g1 + f2 * g2) / n^2,
+                  kappa = (prop_agree - exp_agree) / 
+                                         (1 - exp_agree),
+                  kappa = ifelse(is.nan(kappa) | is.infinite(kappa),
+                                 0, kappa))
+
+ggplot(two_metrics, aes(x = kappa, fill = no_exposure)) + 
+        geom_histogram(color = "black")
+```
+
 #### Are storms with heavy rain exposure likely to occur early in the hurricane season?

### 2016-11-17 README.Rmd
Add travis
Turn off Travis badge for the moment
Update README

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
+
+```{r eval = FALSE, echo = FALSE}
+[![Travis-CI Build Status](https://travis-ci.org/geanders/hurricaneexposure.svg?branch=master)](https://travis-ci.org/geanders/hurricaneexposure)
+```
 
@@
 
-Once you've loaded the dataframe, you can load the included data using the `data` function. For example:
+The following datasets are included with the `hurricaneexposuredata` package:
+
+- `closest_dist`: A dataframe that gives the distance and date-time for the closest approach of each tropical storm to the mean population center of each US county in states in the eastern half of the United States.
+- `county_centers`: A dataset containing the latitude and longitude of the center of population of every United State county, based on population as of the 2010 US Census.
+- `hurr_tracks`: A dataset containing the storm tracks for Atlantic basin tropical storms between 1988 and 2012, from the Extended Best Track Dataset for the Atlantic basin.
+- `rain`: A dataframe that gives the total rainfall in US counties for a one-week window centered at the date on which the tropical storm was closest to the county for Atlantic basin storms between 1988 and 2011.
+- `storm_winds`: A dataframe that gives modeled windspeeds in US counties for all eastern US counties for Atlantic basin tropical storms between 1988 and 2011.
+- `storm_events`: A dataframe that gives county-specific severe storm events listed in US counties temporally and spatially close to storm tracks for all eastern US counties for Atlantic basin tropical storms between 1988 and 2011.
+
+Once you've loaded the `hurricaneexposuredata` package, you can load the included data using the `data` function. For example:
 
@@
 
-The following datasets are included with the package:
+For each dataset, you can see the helpfiles for the data for more information. For example:
 
-- `closest_dist`
-- `county_centers`
-- `hurr_tracks`
-- `rain`
-
-For each, you can see the helpfiles for the data for more information about the data included in each.
+```{r eval = FALSE}
+?hurr_tracks
+```
 

### 2016-11-18 README.Rmd
Update README

@@
 ```
+
+## Overview
+
+This package is meant to add users in creating datasets of tropical storm exposure at the county level for US counties.
+It covers counties in the eastern half of the United States and includes all Atlantic basin tropical storms between 1988 and 2015 (although data for some hazards are only available for some of those years).
+Tropical storm exposure datasets can be generated based on distance to storm tracks, rain, wind, or NOAA storm event listings (including tornado and flood events).
+
+For non-R users, we are developing [a web application](https://zhimingguo.shinyapps.io/Hurricane_Map1/) to allow similar functionality.
+The exposure datasets generated by this web application can be downloaded as flat files and read into other software.
 
@@
 
-For each dataset, you can see the helpfiles for the data for more information. For example:
+For each dataset, you can see the helpfiles for the data for more information by using the `?` call with the dataset name.
+For example, to access the documentation for the `hurr_tracks` dataset, you can run:
 
@@
 
-There are several functions in this package that help to take the data on rainfall and storm tracks stored in the package and use them, in conjunction with thresholds for the distance to the storm tracks and the amount of rainfall required for exposure, to create time series of exposure that can be easily integrated into time series datasets of, for example, health data.
+There are several functions in this package that help to take the data on rainfall, wind, storm events, and storm tracks stored in the package and use them, in conjunction with user-specified thresholds, to create county-level time series of exposure that can be easily integrated into time series datasets of, for example, health data.
 
@@
 
-For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over three days was 100 millimeters or more, you could run:
+For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over three days was 100 millimeters or more, you can run:
 
@@
 
-In addition to giving you the names and closest dates of each storm for each county (`closest_date`-- note, this is given using the UTC timezone), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).
+In addition to giving you the names and closest dates of each storm for each county (`closest_date`-- note, this is given based on local time), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).
+The returned dataframe also includes columns with the local time (based on the counties timezone using the [`countytimezones`](https://cran.r-project.org/web/packages/countytimezones/index.html) R package) when the storm was closest to that county and the time in UTC (to aid in pairing with other weather data in UTC).
 
-To get a dataframe listing the relevant storms for multi-county communities, you can use the `multi_county_rain` function in a similar way:
+Some health data is aggregated at the multi-county, rather than county, level (e.g., data is available for New York City as a whole, rather than for each county in the city).
+To get a dataframe listing the relevant storms for multi-county communities, you can use the `multi_county_rain` function in a similar way, passing in a `communities` dataframe that lists all counties associated with each community:
 
@@
                                   "22071", "51700"))
-rain_storm_df <- multi_county_rain(communities = communities,
+multi_county_rain(communities = communities,
                   start_year = 1995, end_year = 2005,
@@
 
-This output includes columns for the average closest distance for any of the counties in the community (`mean_dist`), the average precipitation for all the counties (`mean_precip`), the highest precipitation for any of the counties (`max_rain`), and the smallest distance between the storm track and any of the county population-weighted centers (`min_dist`).
+This output includes columns for the average closest distance, averaged across all counties in the community (`mean_dist`), the average precipitation for all community counties (`mean_precip`), the highest precipitation for any community county (`max_rain`), and the smallest distance between the storm track and any of the county population-weighted centers (`min_dist`).
+
+Similarly, `county_wind` and `multi_county_wind` functions can be used to generate exposure datasets based on wind exposure during tropical storms.
+For example, to determine all storms with sustained winds of 20 m / s or more for Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700) between 1995 and 2005, you can run:
+
+```{r}
+county_wind(counties = c("22071", "51700"),
+            start_year = 1995, end_year = 2005,
+            wind_limit = 20)
+```
+
+This function returns a dataframe that lists all the storms to which each county was exposed.
+The `vmax_sust` column gives the highest sustained wind modeled at the county's population mean center, while the `vmax_gust` gives the highest gust wind speed modeled there.
 

### 2016-11-18 README.Rmd
Update README

@@
 ```{r fig.width = 7, fig.height = 4, message = FALSE}
-map_1 <- map_counties(storm = "Floyd-1999", metric = "rainfall")
-map_1
+map_counties(storm = "Floyd-1999", metric = "rainfall")
 ```
 
-You can also use this function to plot the closest distance between the storm and each county. For this, you use the argument `metric = "distance"`.
+You can also use this function to plot the closest distance between the storm and each county.
+For this, you use the argument `metric`, which you can set equal to "rainfall", "wind", or "distance".
+For example, here is a map of wind-based exposure to Hurricane Katrina:
+
+```{r}
+map_counties(storm = "Katrina-2005", metric = "wind")
+```
+
+Here is a map of distance-based exposure to the same storm:
+
+```{r}
+map_counties(storm = "Katrina-2005", metric = "distance")
+```
+
+You can also create maps that show binary classification of exposure for each county--- whether that county exceeded a threshold used to define tropical storm exposure based on one of the hazards.
+For example, here is a map of all counties within 75 kilometers of the track of Tropical Storm Allison in 2001:
 
 ```{r fig.width = 7, fig.height = 4}
-map_2 <- map_counties(storm = "Sandy-2012", metric = "distance")
-map_2
+map_distance_exposure(storm = "Allison-2001", dist_limit = 75)
 ```
 
-You can map a binary variable of distance-based exposure using `map_distance_exposure`:
+Similarly, here is a map of all counties with sustained winds of 20 m / s or higher during Katrina:
+
+```{r}
+map_wind_exposure(storm = "Katrina-2005", wind_limit = 20)
+```
+
+Here is a map of all counties with 75 mm or more or rainfall over the day of, day before, and day after Hurricane Floyd, constrained to counties within 500 km of Floyd's track:
 
 ```{r fig.width = 7, fig.height = 4}
-allison_map <- map_distance_exposure(storm = "Allison-2001",
-                                     dist_limit = 75)
-plot(allison_map)
-```
-
-You can also map a binary variable of rain exposure for the communities that were exposed, based on a certain rainfall limit and distance limit:
-
-```{r fig.width = 7, fig.height = 4}
-map_3 <- map_rain_exposure(storm = "Floyd-1999", rain_limit = 125,
-                           dist_limit = 500, 
-                           days_included = c(-1, 0, 1))
-plot(map_3)
+map_rain_exposure(storm = "Floyd-1999", rain_limit = 75,
+                  dist_limit = 500, days_included = c(-1, 0, 1))
 ```
@@
 ```{r fig.width = 5.5, fig.height = 4}
-map_4 <- map_tracks(storms = "Floyd-1999")
-map_4
+map_tracks(storms = "Floyd-1999")
 ```
@@
 There are some different options you can use for the tracks' appearance.
-For example, if you wanted to plot the tracks of several storms, not plot each point when the track locations were measured (typically every six hours), and use some transparency so you can see all the lines, you can use:
+For example, if you wanted to plot the tracks of several storms, plot a point each time the track locations were measured (typically every six hours), and use some transparency so you can see all the lines, you can use:
 
 ```{r fig.width = 5.5, fig.height = 4}
-map_5 <- map_tracks(storms = c("Floyd-1999", "Sandy-2012",
+map_tracks(storms = c("Floyd-1999", "Sandy-2012",
                                "Katrina-2005"), 
-                    plot_points = FALSE,
-                    alpha = 0.5)
-map_5
+           plot_points = TRUE, alpha = 0.5)
 ```
@@
 ```{r fig.width = 7, fig.height = 4}
-map_6 <- map_tracks(storms = "Sandy-2012", plot_object = map_2,
-                    plot_points = FALSE)
-map_6
+floyd_map <- map_counties(storm = "Floyd-1999", metric = "rainfall")
+map_tracks(storms = "Floyd-1999", plot_object = floyd_map)
 ```
-
-```{r fig.width = 7, fig.height = 4}
-map_7 <- map_tracks(storms = "Floyd-1999", plot_object = map_3,
-                    plot_points = FALSE)
-map_7
-```

### 2016-11-18 README.Rmd
Update README

@@
 
+Here is a map of all counties with flood events listed in the NOAA Storm Events database near the time and location of Hurricane Floyd:
+
+```{r}
+map_event_exposure(storm_id = "Floyd-1999", event_type = "flood")
+```
+
+Here are all the counties with tornado events listed for the same storm:
+
+```{r}
+map_event_exposure(storm_id = "Floyd-1999", event_type = "tornado")
+```
+
 ### Plotting storm tracks

### 2016-12-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Beginning to re-work end table with exposure metrics

@@
 kable(exposure_parameters, "markdown")
+
+#studies <- c("Davidson 2001", "Atallah 2007", "Kinney 2008", "Zandbergen 2009", "Czajkowski 2010", "Zaharan #2010", "Kruk 2010", "Villarini 2011", "Rezempour 2014", "Currie 2013", "Grabich 2015", "Grabich 2016")
+#distance_km <- c("120", "", "Buffer", "25, 30, 60, 75", "30, 60, 100", 
+                 "2 counties", "Radial Max Wind", "100-500", "")
+#rain_mm <- c("", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
+#wind_kt <- c("", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
+#exposure_parameters <- data.frame(studies, distance_km, rain_mm, wind_kt)
+#kable(exposure_parameters, "markdown")
 ```

### 2016-12-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
continued edit on table

@@
 #studies <- c("Davidson 2001", "Atallah 2007", "Kinney 2008", "Zandbergen 2009", "Czajkowski 2010", "Zaharan #2010", "Kruk 2010", "Villarini 2011", "Rezempour 2014", "Currie 2013", "Grabich 2015", "Grabich 2016")
-#distance_km <- c("120", "", "Buffer", "25, 30, 60, 75", "30, 60, 100", 
-                 "2 counties", "Radial Max Wind", "100-500", "")
-#rain_mm <- c("", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
-#wind_kt <- c("", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
+#distance_km <- c("Hurricane Disaster Risk Index (HDRI)", "", "County Center",)
+#rain_mm <- c("HDRI", "", "", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
+#wind_kt <- c("HDRI", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
 #exposure_parameters <- data.frame(studies, distance_km, rain_mm, wind_kt)

### 2017-01-06 writing/DraftExposurePaper/DraftExposurePaper.Rmd

--- 2016-11-17 writing/DraftExposurePaper/DraftExposurePaper.Rmd
@@
 ```{r echo = FALSE}
-studies <- c("Czajkowski 2010", "Rezempour 2014", "Zandbergen 2009", 
-             "Currie 2013", "Grabich 2016", "Zaharan 2010", "Kruk 2010", 
-             "Villarini 2011", "Atallah 2007")
-distance_km <- c("120", "", "Buffer", "25, 30, 60, 75", "30, 60, 100", 
-                 "2 counties", "Radial Max Wind", "100-500", "")
-rain_mm <- c("", "2/hr, 48/d", "", "", "", "", "", "", "200-500 (Floyd)")
-wind_kt <- c("", "", "SSWS", "", "SSWS", "", "34, 50, 64", "", "")
+studies <- c("Davidson 2001", "Atallah 2007", "Kinney 2008", "Zandbergen 2009", "Czajkowski 2010", 
+             "Zaharan 2010", "Kruk 2010", "Villarini 2011", "Rezempour 2014", "Currie 2013", "Grabich 2015", 
+             "Grabich 2016")
+distance_km <- c("Hurricane Disaster Risk Index (HDRI)", "", "County Center", "30 Buffer and Decay", "120", 
+                 "2 counties", "Radial Max Wind", "100-500", "", "25, 30, 60, 75", "60", "30, 60, 100")
+rain_mm <- c("HDRI", "200-500 (Floyd)", "", "", "", "", "", "", "2/hr, 48/d", "", "", "")
+wind_kt <- c("HDRI", "", "", "SSWS", "", "", "34, 50, 64", "", "", "", "64", "SSWS")
 exposure_parameters <- data.frame(studies, distance_km, rain_mm, wind_kt)

### 2017-01-17 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Make some updates on paper draft

@@
 
-#devtools::install_github("leighseverson/countyweather")
 options("noaakey" = Sys.getenv("noaakey"))
@@
            panel.grid.minor = element_blank(),
-           panel.background = element_rect(colour = "black", size = 1))
+           panel.background = element_rect(colour = "black", size = 1)) + 
+        coord_map()
 ```
@@
 Rain-based exposure was assessed based on storm-associated rainfall of 75 millimeters or more in the county and a distance of 500 kilometers or less between the county's population mean center and the storm's closest approach.
-Flood event-based exposure was based on the county have a flood event listing in the NOAA Storm Events database and being within [x] kilometers of the storm's track."}
+Flood event-based exposure was based on the county have a flood event listing in the NOAA Storm Events database and being within 500 kilometers of the storm's track."}
 storm <- "Frances-2004"
@@
 Rain-based exposure was assessed based on storm-associated rainfall of 75 millimeters or more in the county and a distance of 500 kilometers or less between the county's population mean center and the storm's closest approach.
-Flood event-based exposure was based on the county have a flood event listing in the NOAA Storm Events database and being within [x] kilometers of the storm's track."}
+Flood event-based exposure was based on the county have a flood event listing in the NOAA Storm Events database and being within 500 kilometers of the storm's track."}
 storm <- "Floyd-1999"

### 2017-01-17 R/event_exposure.R
Add closest date, distance to `storm_events` output

@@ -11,6 +11,8 @@
 #'               start_year = 1988, end_year = 1999,
 #'               event_type = "flood")
 #'
+#' @importFrom dplyr %>%
+#'
 #' @export
 county_events <- function(counties, start_year, end_year, event_type){
         events <- hurricaneexposuredata::storm_events
@@ -29,7 +31,10 @@ county_events <- function(counties, start_year, end_year, event_type){
                                       grepl("Tropical Storm", events) |
                                       grepl("Tropical Depression", events),
                               wind = tropical_storm | grepl("Wind", events))
-        events <- events[events[ , event_type], c("fips", "storm_id")]
+        events <- events[events[ , event_type], c("fips", "storm_id")] %>%
+                dplyr::mutate(storm_id = as.character(storm_id)) %>%
+                dplyr::left_join(hurricaneexposuredata::closest_dist,
+                                 by = c("storm_id", "fips"))
         return(events)
 }
 

### 2017-01-19 R/zzz.R
Use .onAttach, not .onLoad
I was getting a check NOTE when using `packageStartupMessage` in `.onLoad`.

@@ -1,6 +1,6 @@
 .pkgglobalenv <- new.env(parent=emptyenv())
 
-.onLoad <- function(libname, pkgname) {
+.onAttach <- function(libname, pkgname) {
   has_data_package <- requireNamespace("hurricaneexposuredata")
   if(!has_data_package){
     packageStartupMessage(paste("To use this package, you must install the",
@@ -9,6 +9,7 @@
                                 "`install.packages('hurricaneexposuredata',",
                                 "repos = 'https://geanders.github.io/drat/',",
                                 "type = 'source')`."))
+    packageStartupMessage("See the `hurricaneexposure` vignette for more details.")
   }
   assign("has_data", has_data_package, envir = .pkgglobalenv)
 }

### 2017-01-19 R/distance_exposure.R
Add check of available data package for functions that use that data

@@ -99,6 +99,12 @@ multi_county_distance <- function(communities, start_year, end_year,
 
         communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
 
+        if(!.pkgglobalenv$has_data){
+          stop(paste("To use this function, you must have the",
+                     "`hurricaneexposuredata` package installed. See the",
+                     "`hurricaneexposure` package vignette for more details."))
+        }
+
         distance_df <- hurricaneexposuredata::closest_dist %>%
                 dplyr::mutate_(closest_date =
                                        ~ lubridate::ymd(closest_date)) %>%

### 2017-01-19 R/event_exposure.R

@@ -20,6 +20,13 @@
 #'
 #' @export
 county_events <- function(counties, start_year, end_year, event_type){
+
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         events <- hurricaneexposuredata::storm_events
 
         event_years <- as.numeric(gsub(".+-", "", names(events)))
@@ -58,6 +65,13 @@ county_events <- function(counties, start_year, end_year, event_type){
 #'
 #' @export
 map_event_exposure <- function(storm_id, event_type){
+
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         storm <- storm_id
         storm_year <- gsub("*.+-", "", storm_id)
         counties <- hurricaneexposuredata::closest_dist %>%

### 2017-01-19 R/filter_storm_data.R

@@ -44,6 +44,12 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
                               include_rain = FALSE, days_included = NULL,
                               output_vars = c("fips")){
 
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         closest_dist <- data.table::data.table(hurricaneexposuredata::closest_dist)
 
         if(!is.null(counties)){
@@ -116,6 +122,13 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 filter_wind_data <- function(counties = NULL, storm = NULL, year_range = NULL,
                              wind_limit = NULL, output_vars = "fips",
                              wind_var = "vmax_sust"){
+
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         storm_winds <- data.table::data.table(hurricaneexposuredata::storm_winds)
 
         if(!is.null(counties)){

### 2017-01-19 R/map_exposure.R

@@ -79,6 +79,13 @@ map_tracks <- function(storms, plot_object = NULL,
                       plot_points = FALSE,
                       alpha = 1,
                       color = "firebrick"){
+
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         if(is.null(plot_object)){
                 plot_object <- default_map()
         }

### 2017-01-19 R/rain_exposure.R

@@ -170,6 +170,12 @@ multi_county_rain <- function(communities, start_year, end_year,
                          rain_limit, dist_limit,
                          days_included = c(-2, -1, 0, 1)){
 
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
 
         dots <- stats::setNames(list(lazyeval::interp(~ lubridate::ymd(x),

### 2017-01-19 R/wind_exposure.R

@@ -39,6 +39,12 @@
 county_wind <- function(counties, start_year, end_year, wind_limit,
                         wind_var = "vmax_sust"){
 
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
+
         wind_df <- filter_wind_data(counties = counties,
                                          year_range = c(start_year, end_year),
                                          wind_limit = wind_limit,
@@ -79,6 +85,12 @@ county_wind <- function(counties, start_year, end_year, wind_limit,
 #' @importFrom dplyr %>%
 multi_county_wind <- function(communities, start_year, end_year,
                               wind_limit){
+
+        if(!.pkgglobalenv$has_data){
+                stop(paste("To use this function, you must have the",
+                           "`hurricaneexposuredata` package installed. See the",
+                           "`hurricaneexposure` package vignette for more details."))
+        }
 
         communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
 

### 2017-01-19 writing/drat_for_data_package/drat_for_data_package.Rmd
Add write-up on coordinating a drat data package with a CRAN package

---
title: "Setting up drat repo for a data package"
author: "Brooke Anderson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Motivation

R data packages can be large enough to exceed CRAN size requirements, but you may still want to use the data as a dependency for a smaller package that you submit to CRAN.

For example, I am working on a package called `hurricaneexposuredata` that includes data for U.S. counties on rain, wind, and some other exposures for all historical tropical storms over a couple of decades. This package includes only data, but I've written a package called `hurricaneexposure` with functions to map and interact with this data. The `hurricaneexposure` package is not very large, and I'd like to post it to CRAN, but it completely depends on being able to access the data on `hurricaneexposuredata`, which I'd like to store in a drat repository given its large size.

This document outlines the steps you can take to use drat to host a larger data R package that you use within a package you plan to submit to CRAN. Broadly, the steps involved are:

1. Create your own drat repository
2. Create the data package and post it in your drat repository
3. Create the package you plan to submit to CRAN and add some infrastructure that will allow it to use data in the data package but still pass CRAN checks

# Posting data package to a drat repository

### Creating a drat repository

First, if you do not already have a drat repository, you will need to create one. See the documentation for the `drat` package (https://github.com/eddelbuettel/drat) for more details on how to create your own drat repository. Consider putting the local copy of your drat directory in the `~/git/` directory of your computer, as this will make some of the following instructions go a bit more smoothly.

### Creating a data package

Next, you will need to create the data package and post that to a drat repository. For full details on creating R packages, including data-only packages, see Hadley Wickham's [*R Packages*](http://r-pkgs.had.co.nz/) book. While you will not have to pass CRAN checks for a package posted on drat, it is wise to resolve as many ERRORS, WARNINGS, and NOTES from CRAN checks as possible, even if you are hosting the package yourself. You can use the `check` function from the `devtools` package to identify these issues. For a large data package, it would be preferable to resolve all issues except the NOTE on the package size being large.

If you would like to see an example data package, all code for the `hurricaneexposuredata` data package described in the "Motivation" section can be explored at https://github.com/geanders/hurricaneexposuredata.

### Building your data package

Next, you need to build the package into a `.tar.gz` file on your computer. To do this, open a shell. You can do this by clicking on the gear icon in the "Git" window and selecting "Shell...":

```{r echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/opening_shell.png")
```

Use `cd` in the shell to navigate to the parent directory of your package's directory. If you opened the Shell from RStudio's Git window, the shell will start in your package's directory, so you can navigate to the parent directory with `cd ..`. Once you are in the correct directory (you can check by running `ls`, and your package's directory should be one of the listings), build the package using `R CMD build` with your package's name. For example, to build the package `hurricaneexposuredata`, from the shell I called:

```{r engine = 'sh', eval = FALSE}
R CMD build hurricaneexposuredata
```

If the build is successful, you will now have a `.tar.gz` file on your computer with your package's source code. Try running `ls` in the shell, and you should now see that file listed. The file's name will include your package's name and its current version number. This file will be in the directory from which you built your package.

Next, move this file into the directory of your home directory. If the location of your drat repository on your local computer is `~/git/drat`, you can move the file from the shell using the following command (this code example is for the `tar.gz` file for version 0.0.1 of `hurricaneexposuredata`-- for your own package, change the package name and version number in the file name appropriately):

```{r engine = 'sh', eval = FALSE}
mv hurricaneexposuredata_0.0.1.tar.gz ~/git/drat
```

This shell command is specifying to move the file "hurricaneexposuredata_0.0.1.tar.gz" from the current working directory to the `~/git/drat` directory.

### Adding the data package to your drat repository

Finally, you can add the data package you created to your drat repository. Now, you can use functions from the `drat` package within R to add this `.tar.gz` source code file to your drat repository. Open the project file for your drat repository directory in RStudio (from RStudio, you can do this using "File" -> "Open Project..." and then navigate to your drat directory, then doubleclick on the .Rproj file in that directory). Your drat directory will have two Git branches, "Master" and "gh-pages". Before you run the commands to add your data package to the repository, make sure you are in the "gh-pages" branch. In the Git pane in RStudio, there is a spot on the top left of the pane that shows your current branch, and if you click on this you can change to other available branches.

```{r echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/changing_branches.png")
```

Once you are in the "gh-pages" branch, you can use the `insertPackage` function from the `drat` package to add the `.tar.gz.` file you built for your package to your drat repository. If you have an older version of your package in the repository, you may want to use the option `action` to either `"archive"` (move the older package version to an "archive" folder) or `"prune"` (completely remove) older versions of that package.

For example, to add version 0.0.1 of `hurricaneexposuredata` to my drat repository, I ran:

```{r eval = FALSE}
drat::insertPackage("hurricaneexposuredata_0.0.1.tar.gz",  action = "prune")
```

# Setting up a CRAN package to use the data package

Finally, you need to add some infrastructure to the package that you would like to submit to CRAN (`hurricaneexposure` in my example), so that it will both pass CRAN checks and also appropriately load and access the data in the data package you posted to your drat repository.

### Add infrastructure to the DESCRIPTION file

First, add three things to the DESCRIPTION file of the package you plan to submit to CRAN:

1. Add the data package to the "Suggests" list of packages in the DESCRIPTION file of the package you plan to submit to CRAN. The following figure shows an example of the part of the DESCRIPTION file for the `hurricaneexposure` package that suggests the `hurricaneexposuredata` package hosted on my drat repository (you can look at the full DESCRIPTION file for `hurricaneexposure` at https://github.com/geanders/hurricaneexposure/blob/master/DESCRIPTION).

```{r echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/desc_example_suggests.png")
```

2. Add the address of the drat repository that stores the data package to the DESCRIPTION file of the package you want to add to CRAN, for "Additional_repositories". If you are hosting the drat repository through GitHub pages, this repository address will look something like "http://yourgithubusername.github.io/drat".

```{r echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/desc_example_add_repos.png")
```

3. Add some text to the "Description" file of the DESCRIPTION file to let users know how to install the data package and how large it is (this is inspired by the `grattan` package: https://cran.r-project.org/web/packages/grattan/index.html). Here is the example in the `hurricaneexposure` DESCRIPTION file:

```{r echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/desc_details.png")
```

### Customize behavior when the package is attached

Next, you should add some infrastructure to the source code that will check if the user has the data package available when the CRAN package is loaded. If it is not, this infrastructure will give the user a startup message to tell him or her how to install the data package. Finally, you will save a boolean variable of whether the data package is available, which you will store in a package global environment you create and use in the code of your package functions to give users an error if they try to use a function that requires the data without having the data package available.

First, create a file called "zzz.R" in the `R` directory of the package you plan to submit. Within this file, first create an environment called `.pkgglobalenv`:

```{r eval = FALSE}
.pkgglobalenv <- new.env(parent=emptyenv())
```

Then, create a function called `.onAttach` that is a function of `libname` and `pkgname`. Within this function, use `requireNamespace` to figure out if the user has the data package available. In my example, I did this with:

```{r eval = FALSE}
has_data_package <- requireNamespace("hurricaneexposuredata")
```

Since `requireNamespace` returns a boolean of whether the user has the package available, you can then use `has_data_package` to print a startup message (use the `packageStartupMessage` function) for users who lack the data package. For example:

```{r eval = FALSE}
if(!has_data_package){
    packageStartupMessage(paste("To use this package, you must install the",
                                "hurricaneexposuredata package."))
    packageStartupMessage(paste("To install that package, run",
                                "`install.packages('hurricaneexposuredata',",
                                "repos = 'https://geanders.github.io/drat/',",
                                "type = 'source')`."))
    packageStartupMessage("See the `hurricaneexposure` vignette for more details.")
  }
```

Finally, store the boolean `has_data_package` in the package global environment you created:

```{r eval = FALSE}
assign("has_data", has_data_package, envir = .pkgglobalenv)
```

This will allow you to check whether the user has the data package available within functions in the package that use the data (see next step).

The full "zzz.R" file for the package you plan to submit to CRAN might look something like this:

```{r eval = FALSE}
.pkgglobalenv <- new.env(parent=emptyenv())

.onAttach <- function(libname, pkgname) {
  has_data_package <- requireNamespace("hurricaneexposuredata")
  if(!has_data_package){
    packageStartupMessage(paste("To use this package, you must install the",
                                "hurricaneexposuredata package."))
    packageStartupMessage(paste("To install that package, run",
                                "`install.packages('hurricaneexposuredata',",
                                "repos = 'https://geanders.github.io/drat/',",
                                "type = 'source')`."))
    packageStartupMessage("See the `hurricaneexposure` vignette for more details.")
  }
  assign("has_data", has_data_package, envir = .pkgglobalenv)
}
```

### Add code to stop functions that require data if user does not have data package

Finally, you need to add code to any functions that use the data in the data package. The function should check to see if the user has the data package available (which you've stored in `.pkgglobalenv$has_data`). If not, you should stop the function and give the user an error message explaining that they need to install the data package before they can use that function.

For the package that you plan to submit in CRAN, look for any functions that use the data from the data package on your drat repository. Before you use data from that package, check the flag of whether they have the data package available. If not, use `stop` to stop the function, and provide an error message that gives some guidance on how the user can install the required package. For example, I use the following piece of code to do this in the `hurricaneexposure` package:

```{r eval = FALSE}
if(!.pkgglobalenv$has_data){ 
        stop(paste("To use this function, you must have the",
                   "`hurricaneexposuredata` package installed. See the",
                   "`hurricaneexposure` package vignette for more details."))
}
```

Here is a screenshot of how this warning gets worked into one of the `hurricaneexposure` package functions (you can see that this function later calls data from the `hurricaneexposuredata` package with `hurricaneexposuredata::closest_dist`):

```{r echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/stopping_without_data.png")
```

### 2017-01-20 writing/r_journal_draft/anderson-eddelbuettel.Rmd
Add directory for R Journal version of write-up

--- 2017-01-19 writing/drat_for_data_package/drat_for_data_package.Rmd
@@
 ---
-title: "Setting up drat repo for a data package"
-author: "Brooke Anderson"
-date: "`r Sys.Date()`"
-output: html_document
+title: Capitalized Title Here
+author:
+  - name: G. Brooke Anderson
+    affiliation: Colorado State University
+    address:
+    - Lake Street
+    - Fort Collins, CO
+    email:  brooke.anderson@colostate.edu
+  - name: Dirk Eddelbuettel
+    affiliation: Affiliation
+    address:
+    - line 1
+    - line 2
+    email:  author2@work
+abstract: >
+  An abstract of less than 150 words.
+preamble: >
+  % Any extra latex you need in the preamble
+output: rticles::rjournal_article
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-# Motivation
+## Motivation
 
@@
 
-# Posting data package to a drat repository
+## Posting data package to a drat repository
 
@@
 
-# Setting up a CRAN package to use the data package
+## Setting up a CRAN package to use the data package
 

### 2017-01-20 R/distance_exposure.R
make examples conditional on data package

@@ -46,7 +46,7 @@
 #' @note Only counties in states in the eastern half of the United States can
 #'    be processed by this function.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #'
 #' county_distance(counties = c("22071", "51700"),
 #'             start_year = 1995, end_year = 2005,
@@ -86,7 +86,7 @@ county_distance <- function(counties, start_year, end_year, dist_limit){
 #'
 #' @export
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
 #'                          fips = c("36005", "36047", "36061",
 #'                                   "36085", "36081", "36119",
@@ -145,7 +145,7 @@ multi_county_distance <- function(communities, start_year, end_year,
 #'    documentation for \code{\link{county_rain}} and
 #'    \code{\link{multi_county_rain}}.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' # By county
 #' distance_exposure(locations = c("22071", "51700"),
 #'               start_year = 1995, end_year = 2005,

### 2017-01-20 R/event_exposure.R

@@ -11,7 +11,7 @@
 #'    event type, you should not use a start year prior to 1996, as events of
 #'    other types were not recorded in the database before 1996.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' county_events(counties = c("37031", "37053", "37055"),
 #'               start_year = 1988, end_year = 1999,
 #'               event_type = "flood")
@@ -57,7 +57,7 @@ county_events <- function(counties, start_year, end_year, event_type){
 #' @inheritParams county_distance
 #' @inheritParams county_events
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' map_event_exposure(storm_id = "Floyd-1999", event_type = "flood")
 #' map_event_exposure(storm_id = "Floyd-1999", event_type = "tornado")
 #' map_event_exposure(storm_id = "Floyd-1999", event_type = "wind")

### 2017-01-20 R/filter_storm_data.R

@@ -26,7 +26,7 @@
 #'    function. Columns in the output will vary depending on the user's
 #'    selections for the \code{output_vars} argument.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' filter_storm_data(counties = c("22071", "51700"), year_range = c(1988, 2011),
 #'                   distance_limit = 250, rain_limit = 150,
 #'                  include_rain = TRUE, days_included = c(-1, 0, 1),
@@ -111,7 +111,7 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 #'    function. Columns in the output will vary depending on the user's
 #'    selections for the \code{output_vars} argument.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' filter_wind_data(counties = c("22071", "51700"), year_range = c(1988, 2011),
 #'                  wind_limit = 20,
 #'                  output_vars = c("fips", "storm_id", "vmax_sust"))

### 2017-01-20 R/map_exposure.R

@@ -63,7 +63,7 @@ default_map <- function(){
 #'    of the selected storms. This object can be printed directly or added
 #'    on to with other ggplot commands.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' map_tracks(storms = "Sandy-2012")
 #' map_tracks(storms = "Floyd-1999", plot_points = TRUE)
 #' map_tracks(storms = c("Sandy-2012", "Floyd-1999"))
@@ -185,7 +185,7 @@ interp_track <- function(track, tint = 0.25){
 #'    part of the United States, showing distance from a storm track or total
 #'    rainfall over a given window of one or more days.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #' floyd_map <- map_counties("Floyd-1999", metric = "rainfall",
 #'                            days_included = c(-2, -1, 0, 1))
 #' floyd_map
@@ -235,7 +235,7 @@ map_counties <-function(storm, metric = "distance",
 #' @inheritParams map_counties
 #'
 #' @examples
-#'\dontrun{
+#'if (requireNamespace(""hurricandataexposuredata")) {
 #' floyd_map <- map_rain_exposure(storm = "Floyd-1999", rain_limit = 50,
 #'                                dist_limit = 100)
 #' floyd_map
@@ -296,7 +296,7 @@ map_rain_exposure <- function(storm, rain_limit, dist_limit,
 #' @return Plots a map showing whether eastern US counties were exposed or
 #'    unexposed to a specific storm based on a distance criterion.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #'
 #' floyd_map <- map_distance_exposure(storm = "Floyd-1999", dist_limit = 75)
 #' floyd_map
@@ -351,7 +351,7 @@ map_distance_exposure <- function(storm, dist_limit){
 #' @return Plots a map showing whether eastern US counties were exposed or
 #'    unexposed to a specific storm based on a wind criterion.
 #'
-#' @examples \dontrun{
+#' @examples if (requireNamespace(""hurricandataexposuredata")) {
 #'
 #' beryl_map <- map_wind_exposure(storm = "Beryl-1988", wind_limit = 15)
 #' beryl_map

### 2017-01-26 R/rain_exposure.R
Work some on vignette / updating documentation

@@ -18,8 +18,8 @@
 #'    \code{c(-1, 0, 1)} would calculate rain for a county as the sum of the
 #'    rainfall for the day before, the day of, and the day after the date when
 #'    the storm center was closest to the county center. Values can range from
-#'    -3 to 3 (i.e., at most, you can calculate the total rainfall for the
-#'    seven-day window centered on the day when the storm is closest to the
+#'    -5 to 3 (i.e., at most, you can calculate the total rainfall from five days
+#'    to three days after the day when the storm is closest to the
 #'    county).
 #' @inheritParams county_distance
 #'

### 2017-01-26 R/map_exposure.R
Change to default to include storm track on maps

@@ -206,6 +206,8 @@ interp_track <- function(track, tint = 0.25){
 #' the county's population mean center and a certain amount of rain
 #' \code{rain_limit} fell during a specified window of days (\code{days_included}).
 #'
+#' @param add_track TRUE / FALSE of whether to add the storm's track to the map. The
+#'    default is TRUE.
 #' @inheritParams county_distance
 #' @inheritParams county_rain
 #' @inheritParams map_counties
@@ -218,15 +220,17 @@ interp_track <- function(track, tint = 0.25){
 #'
 #' map_rain_exposure(storm = "Floyd-1999", rain_limit = 50, dist_limit = 100)
 #'
-#' allison_map <- map_rain_exposure(storm = "Allison-2001", rain_limit = 20,
-#'                                  dist_limit = 100, days_included = 0)
-#' map_tracks("Allison-2001", plot_points = FALSE, plot_object = allison_map)
+#' # Example of customizine track appearance
+#' allison_map <- map_rain_exposure(storm = "Allison-2001", rain_limit = 125,
+#'                                  dist_limit = 100, days_included = -5:3,
+#'                                  add_track = FALSE)
+#' map_tracks("Allison-2001", plot_object = allison_map, plot_points = TRUE)
 #'}
 #' @importFrom dplyr %>%
 #'
 #' @export
 map_rain_exposure <- function(storm, rain_limit, dist_limit,
-                              days_included = c(-2, -1, 0, 1)){
+                              days_included = c(-2, -1, 0, 1), add_track = TRUE){
 
         map_data <- filter_storm_data(storm = storm,
                                            days_included = days_included,
@@ -262,6 +266,11 @@ map_rain_exposure <- function(storm, rain_limit, dist_limit,
                 ggplot2::scale_fill_manual(name = paste("Rain >", rain_limit, "mm"),
                                            values = c("white", "navy"),
                                            labels = c("Unexposed", "Exposed"))
+
+        if(add_track){
+                out <- map_tracks(storm, plot_object = out)
+        }
+
         return(out)
 }
 
@@ -274,6 +283,7 @@ map_rain_exposure <- function(storm, rain_limit, dist_limit,
 #' @inheritParams county_distance
 #' @inheritParams county_rain
 #' @inheritParams map_counties
+#' @inheritParams map_rain_exposure
 #'
 #' @return Plots a map showing whether eastern US counties were exposed or
 #'    unexposed to a specific storm based on a distance criterion.
[…]

### 2017-01-26 README.Rmd
Update README

--- 2016-11-18 README.Rmd
@@
 ```{r echo = FALSE, message = FALSE, warning = FALSE}
-# If you need to install the hurricaneexposuredata package: 
-# library(devtools)
-# install_github("geanders/hurricaneexposuredata")
+# If you need to, install the `hurricaneexposuredata` package from the drat repo:
+# library(drat)
+# addRepo("geanders")
+# install.packages("hurricaneexposuredata")
 library(hurricaneexposuredata)
@@
 
-## Overview
+## Package overview
 
-This package is meant to add users in creating datasets of tropical storm exposure at the county level for US counties.
-It covers counties in the eastern half of the United States and includes all Atlantic basin tropical storms between 1988 and 2015 (although data for some hazards are only available for some of those years).
-Tropical storm exposure datasets can be generated based on distance to storm tracks, rain, wind, or NOAA storm event listings (including tornado and flood events).
+This package allows you to explore and map data on county-level exposures to Atlantic-basin tropical storms between 1988 and 2015 for a number of storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
+This package allows the user to map exposures by county for a selected storm or to identify all storms to which selected counties were exposed, based on user-specified thresholds (for example, the package allows the user to identify all storms that brought wind of 34 knots or higher to Miami-Dade County in Florida).
 
-For non-R users, we are developing [a web application](https://zhimingguo.shinyapps.io/Hurricane_Map1/) to allow similar functionality.
-The exposure datasets generated by this web application can be downloaded as flat files and read into other software.
+If you use this package and the data in the associated data package (`hurricaneexposuredata`) for research, please cite both packages.
+In particular, be sure to include the version of the packages that you used, as this will make your research project more reproducible, and the data will likely be updated as we get further years of data and find improved ways to measure tropical storm exposure.
+The two packages can be cited as:
 
-## Loading the package
+- Anderson B, Yan M, Ferreri J, Crosson W, Al-Hamdan M, Schumacher A and
+ Eddelbuettel D (2017). _hurricaneexposure: Explore and Map County-Level
+ Hurricane Exposure in the United States_. R package version 0.0.1, <URL:
+ http://CRAN.R-project.org/package=hurricaneexposure>.
 
-The package currently exists in [a development version](https://github.com/geanders/hurricaneexposure) on GitHub, and all the associated data is in its own [development version package](https://github.com/geanders/hurricaneexposuredata), also on GitHub.
-You can use the following code to load both:
+- Anderson B, Schumacher A, Crosson W, Al-Hamdan M, Yan M, Ferreri J, Chen Z,
+ Quiring S and Guikema S (2017). _hurricaneexposuredata: Data Characterizing
+ Exposure to Hurricanes in United States Counties_. R package version 0.0.1,
+ <URL: https://github.com/geanders/hurricaneexposuredata>.
+
+To generate BibTex entries for the packages, you can use the `citation` function in R.
+
+## Required set-up to use this package
+
+This package depends on data in a data package (`hurricaneexposuredata`) that is available through a `drat` repository on GitHub.
+To use the `hurricaneexposure` package, you will need to install `hurricaneexposuredata` on your computer.
+You can do that by adding the `drat` repository to your repositories; once you do this, you can install the `hurricaneexposuredata` using the `install.packages` function (and later update it using the `update.packages` function):
 
 ```{r eval = FALSE}
-library(devtools)
-install_github("geanders/hurricaneexposuredata")
-install_github("geanders/hurricaneexposure", 
-               build_vignettes = TRUE)
-library(hurricaneexposuredata)
-library(hurricaneexposure)
+library(drat)
+addRepo("geanders")
+install.packages("hurricaneexposuredata")
 ```
 
-The following datasets are included with the `hurricaneexposuredata` package:
+The `hurricaneexposuredata` data package includes data that characterizes county-level exposure to tropical storms in counties in the eastern half of the United States between 1988 and 2015 (for some hazards, exposure data is only included for a subset of these years).
+Tropical storms that did not pass within at least 250 km of at least one US county were excluded from these datasets.
+The following datasets are included with the `hurricaneexposuredata` data package:
 
-- `closest_dist`: A dataframe that gives the distance and date-time for the closest approach of each tropical storm to the mean population center of each US county in states in the eastern half of the United States.
-- `county_centers`: A dataset containing the latitude and longitude of the center of population of every United State county, based on population as of the 2010 US Census.
-- `hurr_tracks`: A dataset containing the storm tracks for Atlantic basin tropical storms between 1988 and 2012, from the Extended Best Track Dataset for the Atlantic basin.
-- `rain`: A dataframe that gives the total rainfall in US counties for a one-week window centered at the date on which the tropical storm was closest to the county for Atlantic basin storms between 1988 and 2011.
-- `storm_winds`: A dataframe that gives modeled windspeeds in US counties for all eastern US counties for Atlantic basin tropical storms between 1988 and 2011.
-- `storm_events`: A dataframe that gives county-specific severe storm events listed in US counties temporally and spatially close to storm tracks for all eastern US counties for Atlantic basin tropical storms between 1988 and 2011.
+- `county_centers`: Location of United States county centers of population
+- `hurr_tracks`: Storm tracks for Atlantic-basin storms, 1988-2015
+- `closest_dist`: Closest distances between counties and a storm track, for Atlantic-basin storms, 1988-2015
+- `rain`: Rainfall for US counties during Atlantic basin tropical storms, 1988-2011; daily rainfall is given from five days before to three days after the storm's closest approach to the county
+- `storm_winds`: Modeled county wind speeds for Atlantic-basin storms, 1988-2015
+- `storm_events`: Listings from the NOAA Storm Events database that occurred near in time and location to tropical storms, 1988-2015.
+This database changed the types of events it reported in 1996, which should be considered when using the data.
+- `ext_tracks_wind`: Estimated county wind speeds for Atlantic-basin storms, 1988-2015, based on the wind radii listed in the Extended Best Tracks dataset
 
-Once you've loaded the `hurricaneexposuredata` package, you can load the included data using the `data` function. For example:
+Once you've installed and loaded `hurricaneexposuredata`, you can load the included data using the `data` function. For example:
 
 ```{r}
+library(hurricaneexposuredata)
 data("hurr_tracks")
@@
 
-For each dataset, you can see the helpfiles for the data for more information by using the `?` call with the dataset name.
-For example, to access the documentation for the `hurr_tracks` dataset, you can run:
+For each dataset included in `hurricaneexposuredata`, you can see the helpfiles for the data for more information about the data included in each (e.g., `?hurr_tracks` to read documentation on the hurricane tracks data shown above).
+This data documentation includes both explanations of how the dataset was created, definitions of the variables included in each dataset, and the sources of data included.
 
-```{r eval = FALSE}
-?hurr_tracks
+The following table shows the storms covered by the `hurricaneexposuredata` data package, as well as which hazard metrics are available for each year.
+All storms passed within 250 km of at least one U.S.
+county.
+
+```{r echo = FALSE, message = FALSE, warning = FALSE}
+library(dplyr)
+library(tidyr)
+hurr_tracks %>%
+        select(storm_id) %>%
+        distinct() %>%
+        separate(storm_id, into = c("storm_name", "Year"), sep = "-") %>%
+        group_by(Year) %>%
+        summarize(Storms = paste(storm_name, collapse = ", ")) %>%
+        mutate(Distance = "x",
+               Wind = "x",
+               Rain = ifelse(Year <= 2011, "x", ""),
+               Flood = ifelse(Year >= 1996, "x", ""),
+               Tornado = "x") %>%
+        pander::pander(justify = "clccccc")
+```
+
+In functions throughout the `hurricaneexposure` package, storms are identified based on their name and year (e.g., Hurricane Floyd in 1999 is identified as "Floyd-1999").
+Note that a few storms satisfied this criteria but were unnamed (e.g., "Notnamed" in 1991, "Subtrop" in 1997, etc.).
+These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table (e.g., "Notnamed-1991"), although for many applications you may want to consider excluding these unnamed storms from the full analysis.
+
+## Mapping hurricane exposure
+
+The `hurricaneexposure` package has functions to interact with the data in the `hurricaneexposuredata` package.
+First, this package has functions that can be used to create some different maps of hurricane exposures based on distance to the storm track and rainfall.
+
+### Plotting county-level continuous exposure
+
+First, for several of the storm hazard measurements (currently rain, wind, and distance), you can use the `map_counties` function in `hurricaneexposure` to plot a map for a specific storm showing the continuous value of that metric for all eastern U.S.
+counties.
+For example, you can map county-level rain during Hurricane Floyd (1999) using the following call:
+
+```{r fig.width = 7, fig.height = 4}
+map_counties(storm = "Floyd-1999", metric = "rainfall")
+```
+
+By default, this map shows the cumulative rain in each county for two days before to one day after the date that the storm passed closest to the county (see the help file for the `closest_dist` dataset in `hurricaneexposuredata` to find out more about how this date was determined for each county-storm combination).
+The period used to calculate cumulative rainfall for the map can be customized using the `days_included` option in the `map_counties` function.
+The two following examples show the difference in rain maps for Tropical Storm Allison (2001) when using rain only from the day before and the day the storm was closest to each county (`days_included = -1:0`) versus from five days before to three days after the storm's closest approach (`days_included = -5:3`):
+
+```{r fig.width = 7, fig.height = 4}
+map_counties(storm = "Allison-2001", metric = "rainfall", days_included = -1:0) + 
+        ggplot2::ggtitle("Rain during Allison (2001) for day before and day of closest approach")
+```
+
+```{r fig.width = 7, fig.height = 4}
+map_counties(storm = "Allison-2001", metric = "rainfall", days_included = -5:3) + 
+        ggplot2::ggtitle("Rain during Allison (2001) for five days before to three days\nafter the day of closest approach")
+```
+
+You can also use the `map_counties` function to plot the maximum wind during the storm for each county.
+For this, you use the argument `metric = "wind"`.
+For example, you can plot county-specific storm winds during Hurricane Katrina (2005) with the call:
+
+```{r fig.width = 7, fig.height = 4}
+map_counties(storm = "Katrina-2005", metric = "wind")
+```
+
+These county winds are determined based on a model of wind speeds from Best Tracks hurricane data.
+See the documentation for the `storm_winds` dataset (`?storm_winds`) to find out more about this data; further details on the wind model are available in the vignettes for the [`stormwindmodel` package](https://cran.r-project.org/web/packages/stormwindmodel/index.html).
+
+[Add details about how to map this from the wind radii estimates, instead]
+
+Finally, you can also use this function to plot the closest distance between the storm and each county.
+For this, you use the argument `metric = "distance"`.
+For example, the following call plots county-level distances from the track of Hurricane Sandy (2012):
+
+```{r fig.width = 7, fig.height = 4}
+map_counties(storm = "Sandy-2012", metric = "distance")
+```
+
+See the documentation for the `closest_dist` dataset (`?closest_dist`) for more details on how distances from the storm track were calculated for each county.
+
+### Mapping county-level binary exposure
+
+You can also map binary county-level exposures, where each county is classified as "exposed" based on whether it meets a threshold for one of the hazard-based metrics.
+These maps can be created for the metrics shown in the continuous exposure maps above (wind, rain, and distance from the storm track).
+In addition, binary classification maps can also be created based on storm event listings for flood events and tornado events from the NOAA Storm Events database.
+
+For example, you can map a binary variable of distance-based exposure using `map_distance_exposure`, with "exposure" defined as the county being within the number of kilometers given in `dist_limit` of the storm track.
+For example, you can map the counties that were within 75 km of Hurricane Sandy's storm track with the call:
+
+```{r fig.width = 7, fig.height = 4}
+map_distance_exposure(storm = "Sandy-2012", dist_limit = 75)
+```
+
+Similarly, you can map a binary variable county exposure based on rain using `map_rain_exposure`.
+In the case of rain, you must define exposure thresholds for both rainfall and distance.
+This is because the rain data is observed data, so neglecting to limit "exposed" counties to those close to the storm track may pick up counties far from the storm that experienced a lot of rain over the same period.
+For example, you can use the following call to map rain exposure to Tropical Storm Allison (2001), where exposure is defined as a county where cumulative rainfall from fives days before to three days after the storm's closest approach was 175 mm or more and where the county was within 500 km of the storm's track:
+
+```{r fig.width = 7, fig.height = 4}
+map_rain_exposure(storm = "Allison-2001", rain_limit = 175, dist_limit = 500,
+                  days_included = -5:3)
+```
+
+As with the continuous maps of rain exposure, the number of days included to calculate cumulative rain can be adjusted with the `days_included` option in `map_rain_exposure` (see the help file for `map_rain_exposure` for more details).
+
+You can use the `map_wind_exposure` function to map counties exposed to a certain level of wind during a storm.
+For example, to identify all counties exposed to sustained winds of 20 m / s or higher during Hurricane Katrina, you can run:
+
+```{r fig.width = 7, fig.height = 4}
+map_wind_exposure(storm = "Katrina-2005", wind_limit = 20)
+```
+
+If you would like to base a wind threshold on knots rather than m / s, you can use the `convert_wind_speed` function from the `weathermetrics` package to do that.
+For example, to map counties in which winds were 34 knots or higher during the storm, you can run:
+
+```{r fig.width = 7, fig.height = 4}
+library(weathermetrics)
+map_wind_exposure(storm = "Katrina-2005", 
+                  wind_limit = convert_wind_speed(34, "knots", "mps"))
+```
+
+Finally, you can map which counties were exposed to specific types of events, as listed in the NOAA Storm Events Database, using the `map_event_exposure` function.
+For example, you can map counties for which a flood event was listed for Hurricane Floyd (1999) with the call:
+
+```{r fig.width = 7, fig.height = 4}
+map_event_exposure(storm = "Floyd-1999", event_type = "flood")
+```
+
+Similarly, you can map which counties were exposed to tornado events during Hurricane Ivan (2004) with:
+
+```{r fig.width = 7, fig.height = 4}
+map_event_exposure(storm = "Ivan-2004", event_type = "tornado")
+```
+
+When using this function, note that the types of events reported in the NOAA Storm Events Database changed in 1996, so for many event types there will be no listings before 1996.
+Tornado listings do extend earlier, but flood listings do not.
+Further, when using this data, the user should be aware of and take into account the limitations of the database, especially for comparing storms across years.
+Reporting standards likely have changed over time; for example, it is possible that earlier years have more false negatives than more recent years.
+
+Event listings from the NOAA Storm Events database were linked to storms based on the start date of the event being within a five-day window of the date the storm was closest to the county and the county being within 500 km of the county.
+The [`noaastormevents` package](https://github.com/zailchen/noaastormevents) was used to create this dataset.
+For more information on the data, see the help file for the `storm_events` dataset (`?storm_events`) or the vignette for the `noaastormevents` package.
+
+### Plotting storm tracks
+
+If you would like to add a storm's track to one of the maps of county-level exposure, you can do so using the `map_tracks` function.
+This function can be used either to map a storm track on a blank U.S.
+map or to add a track to an existing map, if it is a `ggplot` object (as are all maps generated in the previous section).
+
+First, the `map_tracks` function can be used to create a map of the hurricane tracks for one or more storms. For example, to plot the tracks of Hurricane Floyd in 1999, you can run:
+
+```{r fig.width = 5.5, fig.height = 4}
+map_tracks(storms = "Floyd-1999")
+```
+
+There are some different options you can use for the tracks' appearance.
+For example, if you wanted to plot the tracks of several storms, and plot each point when the track locations were measured (typically every six hours), use some transparency so you can see all the tracks, and show the tracks in blue, you can run:
+
+```{r fig.width = 5.5, fig.height = 4}
+map_tracks(storms = c("Andrew-1992", "Katrina-2005", "Rita-2005"),
+           alpha = 0.5, plot_points = TRUE, color = "blue")
+```
+
+The `map_tracks` function is used in all the mapping functions described in the previous function to add the storm's track to the exposure map.
+However, if you would like to customize the appearance of the storm's track on the map, you can do so by plotting the exposure map without the storm track (`add_track = FALSE`) and then adding the track yourself with the `map_tracks` function.
+For example, to change the color of the storm track in a map of flood event exposure during Hurricane Floyd (1999) and to add points showing the available observations for the storm in the Best Tracks hurricane tracking data, you can run:
+
+```{r fig.width = 7, fig.height = 4}
+floyd_map <- map_event_exposure(storm = "Floyd-1999", event_type = "flood")
+map_tracks(storms = "Floyd-1999", plot_object = floyd_map, plot_points = TRUE, 
+           color = "darkgray")
 ```
@@
 
-There are several functions in this package that help to take the data on rainfall, wind, storm events, and storm tracks stored in the package and use them, in conjunction with user-specified thresholds, to create county-level time series of exposure that can be easily integrated into time series datasets of, for example, health data.
+The `hurricaneexposure` package also has several functions that can input a list of counties and output a list of all of the storms each county was exposed to over a certain period, based on one of the hazard metrics.
+These exposure listings include the storm dates, making them easy to integrate into time series datasets of health or other impact data.
 
-First, the `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds for distance (storm path must come that close or closer to the county) and rainfall (total rainfall summed over the days you specify-- for example, `days_included = c(-1, 0, 1)` would use the total rainfall summed over the day before, day of, and day after the date when the storm was closest to the county).
-The function outputs a dataframe with all of the storms to which each of the counties was exposed.
-
-For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over three days was 100 millimeters or more, you can run:
+For example, the `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds to define rain-based exposure (cumulative rainfall and distance from the storm's track, as explained for the binary rain maps shown in an earlier section) and creates a list of all storms that met these thesholds for the counties.
+For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over a three-day window of the date of closest approach was 100 millimeters or more, you can run:
 
 ```{r}
-county_rain(counties = c("22071", "51700"),
-            start_year = 1995, end_year = 2005,
-            rain_limit = 100, dist_limit = 100,
-            days_included = c(-1, 0, 1))
+county_rain(counties = c("22071", "51700"), start_year = 1995, end_year = 2005,
+            rain_limit = 100, dist_limit = 100, days_included = c(-1, 0, 1))
 ```
 
-In addition to giving you the names and closest dates of each storm for each county (`closest_date`-- note, this is given based on local time), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).
-The returned dataframe also includes columns with the local time (based on the counties timezone using the [`countytimezones`](https://cran.r-project.org/web/packages/countytimezones/index.html) R package) when the storm was closest to that county and the time in UTC (to aid in pairing with other weather data in UTC).
+This function draws on the same dataset (`rain`) from the `hurricaneexposuredata` package as the functions for mapping rain exposure shown above.
+For this and other exposures, see the table earlier in this vignette to see which years of data are available (for rain, data is currently available for storms through 2011).
 
-Some health data is aggregated at the multi-county, rather than county, level (e.g., data is available for New York City as a whole, rather than for each county in the city).
-To get a dataframe listing the relevant storms for multi-county communities, you can use the `multi_county_rain` function in a similar way, passing in a `communities` dataframe that lists all counties associated with each community:
+In addition to giving the names and closest dates of each storm for each county (`closest_date`-- note, this is given based on local time for the county; see documentation for the [`countytimezones` package](https://cran.r-project.org/web/packages/countytimezones/index.html) to find out more about how local time was calculated from UTC for each county), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).
+The returned dataframe also gives the time of closest approach in both local time (`local_time`) and UTC (`closest_time_utc`), based on 15-minute intervals along the storm's track.
+
+Similar functions are available in the package to create listings of county exposures to storms based on wind, distance to the storm track, and events listings from the NOAA Storm Events Database (flood and tornado events).
+For example, to get a listing of all storms between 1988 and 2015 for which Miami-Dade county (FIPS: ) has experienced sustained winds of 34 knots or more (17.5 m / s), you can run:
+
+```{r}
+county_wind(counties = "12086", start_year = 1988, end_year = 2015, wind_limit = 17.5)
+```
+
+The returned dataframe includes the estimated maximum values during the storm for sustained winds (`vmax_sust`) and gust winds (`vmax_gust`), both in m / s.
+It also includes some of the variables on the date and time of closest approach and the distance at the storm's closest approach.
+
+The `county_distance` function can similarly be used to generate a listing of all storms that came within a certain distance of specified counties.
+For example, to get a list of all storms that came within 50 km of Orleans Parish (FIPS 22071) between 1988 and 2015, you can run:
+
+```{r}
+county_distance(counties = "22071", start_year = 1988, end_year = 2015,
+                dist_limit = 50)
+```
+
+Finally, you can use the `county_events` function to get a listing of all storms for which a county had a certain type of NOAA Storm Events listing.
+For example, to get a list of all storms for which Norfolk, Virginia, (FIPS: 51710) had a flood event listing, you can run:
+
+```{r}
+county_events(counties = "51710", start_year = 1996, end_year = 2015, 
+              event_type = "flood")
+```
+
+### Handling multi-county communities
+
+Some datasets covering storm impacts might have communities composed of multiple counties rather than county-specific listings.
+To get a dataframe listing the relevant storms for multi-county communities, you can use the `multi_county_*` family of functions in a similar way to the `county_*` family of functions.
+For example, to get rain exposure listings in the case where one of the communities (New York, NY) is comprised of multiple counties, you can run:
 
@@
                                   "22071", "51700"))
-multi_county_rain(communities = communities,
-                  start_year = 1995, end_year = 2005,
-                  rain_limit = 100, dist_limit = 100,
-                  days_included = c(-1, 0, 1))
+multi_county_rain(communities = communities, start_year = 1995, end_year = 2005,
+                   rain_limit = 100, dist_limit = 100)
 ```
 
-This output includes columns for the average closest distance, averaged across all counties in the community (`mean_dist`), the average precipitation for all community counties (`mean_precip`), the highest precipitation for any community county (`max_rain`), and the smallest distance between the storm track and any of the county population-weighted centers (`min_dist`).
+The output from this function includes columns for the average closest distance for any of the counties in the community (`mean_dist`), the average precipitation for all the counties (`mean_precip`), the highest precipitation for any of the counties (`max_rain`), and the smallest distance between the storm track and any of the county population-weighted centers (`min_dist`).
 
-Similarly, `county_wind` and `multi_county_wind` functions can be used to generate exposure datasets based on wind exposure during tropical storms.
-For example, to determine all storms with sustained winds of 20 m / s or more for Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700) between 1995 and 2005, you can run:
-
-```{r}
-county_wind(counties = c("22071", "51700"),
-            start_year = 1995, end_year = 2005,
-            wind_limit = 20)
-```
-
-This function returns a dataframe that lists all the storms to which each county was exposed.
-The `vmax_sust` column gives the highest sustained wind modeled at the county's population mean center, while the `vmax_gust` gives the highest gust wind speed modeled there.
-
-## Creating and writing time series of exposure
-
-To create time series of hurricane exposure for all of the storms that meet the exposure definition for study communities, you can use the function `rain_exposure`.
-If you have counties for your study, you can run this with a vector of the county FIPS as the `locations` argument (specify the directory path to write the files with `out_dir`):
-
-```{r}
-rain_exposure(locations = c("22071", "51700"),
-              start_year = 1995, end_year = 2005,
-              rain_limit = 100, dist_limit = 100,
-              out_dir = "~/tmp/storms")
-```
-
-If you have multi-county communities, set `locations` instead to be a dataframe with community names (`commun` column) and FIPS codes (`fips` column):
-
-```{r message = FALSE, warning = FALSE}
-communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
-                          fips = c("36005", "36047", "36061",
-                          "36085", "36081", "36119",
-                          "22071", "51700"))
-rain_exposure(locations = communities,
-              start_year = 1995, end_year = 2005,
-              rain_limit = 100, dist_limit = 100, out_dir = "~/tmp/storms")
-```
-
-Both functions will output one file of exposure per county or community into the directory that you specify using the `out_dir` argument.
-
-## Mapping hurricane exposure
-
-This package allows you to create some different maps of hurricane exposures based on distance to the storm track and rainfall.
-
-### Plotting county-level exposure
-
-The `map_counties` function creates county choropleths of different storm exposure variables (right now, rainfall and distance from the storm tracks).
-For example, to plot rain exposure for Hurricane Floyd in 1999 (the ID for this storm is "Floyd-2012"):
-
-```{r fig.width = 7, fig.height = 4, message = FALSE}
-map_1 <- map_counties(storm = "Floyd-1999", metric = "rainfall")
-map_1
-```
-
-You can also use this function to plot the closest distance between the storm and each county. For this, you use the argument `metric = "distance"`.
-
-```{r fig.width = 7, fig.height = 4}
-map_2 <- map_counties(storm = "Sandy-2012", metric = "distance")
-map_2
-```
-
-You can map a binary variable of distance-based exposure using `map_distance_exposure`:
-
-```{r fig.width = 7, fig.height = 4}
-allison_map <- map_distance_exposure(storm = "Allison-2001",
-                                     dist_limit = 75)
-plot(allison_map)
-```
-
-You can also map a binary variable of rain exposure for the communities that were exposed, based on a certain rainfall limit and distance limit:
-
-```{r fig.width = 7, fig.height = 4}
-map_3 <- map_rain_exposure(storm = "Floyd-1999", rain_limit = 125,
-                           dist_limit = 500, 
-                           days_included = c(-1, 0, 1))
-plot(map_3)
-```
-
-### Plotting storm tracks
-
-The `map_tracks` function will map the hurricane tracks for one or more storms. For example, to plot the tracks of Hurricane Floyd in 1999 (the ID for this storm is "Floyd-2012"):
-
-```{r fig.width = 5.5, fig.height = 4}
-map_4 <- map_tracks(storms = "Floyd-1999")
-map_4
-```
-
-There are some different options you can use for the tracks' appearance.
-For example, if you wanted to plot the tracks of several storms, not plot each point when the track locations were measured (typically every six hours), and use some transparency so you can see all the lines, you can use:
-
-```{r fig.width = 5.5, fig.height = 4}
-map_5 <- map_tracks(storms = c("Floyd-1999", "Sandy-2012",
-                               "Katrina-2005"),
-                    plot_points = FALSE,
-                    alpha = 0.5)
-map_5
-```
-
-You can also add these tracks to an existing `ggplot`-created US map.
-You do this through the `plot_object` argument.
-For example, to add the storm track to the plot of distance exposure for Sandy or rain exposure for Floyd, you could run:
-
-```{r fig.width = 7, fig.height = 4}
-map_6 <- map_tracks(storms = "Sandy-2012", plot_object = map_2,
-                    plot_points = FALSE)
-map_6
-```
-
-```{r fig.width = 7, fig.height = 4}
-map_7 <- map_tracks(storms = "Floyd-1999", plot_object = map_3,
-                    plot_points = FALSE)
-map_7
-```
+Similar functions exist for wind exposure (`multi_county_wind`) and distance exposure (`multi_county_distance`).

### 2017-01-27 R/filter_storm_data.R
Update to allow exposure metrics based on other wind values
Including gust winds and wind durations. Also updates to allow the use of wind estimates from the Extended Best Tracks data.

@@ -103,10 +103,28 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 #'
 #' @inheritParams filter_storm_data
 #' @param wind_limit A numeric vector of length one giving the minimum
-#'    wind speed (in meters per second) to use in the filter
+#'    wind speed (in meters per second) or duration of winds of 20 m / s or more
+#'    (in minutes) to use in the filter. The units of this variable will depend on
+#'    the user's choice for the \code{wind_var} parameter. If the Extended Best
+#'    Tracks wind radii are used as the source of the wind data, the sustained
+#'    winds will only be available for cutpoints of 34 knots, 50 knots, and 64
+#'    knots, so these values should be used (e.g., to get all counties with winds of
+#'    34 knots or higher, you could use \code{wind_limit = 17.4}, with the limit
+#'    given as a value just below 34 knots in the units meters per second).
 #' @param wind_var A character string giving the wind variable to use. Choices
-#'    are "vmax_sust" (maximum sustained winds; default) or "vmax_gust" (maximum
-#'    gust winds).
+#'    are \code{"vmax_sust"} (maximum sustained winds; default), \code{"vmax_gust"}
+#'    (maximum gust winds), \code{"sust_dur"} (minutes of sustained winds of 20 m / s
+#'    or higher) and \code{"gust_dur"} (minutes of gust winds of 20 m / s or higher).
+#'    If the Extended Best Tracks wind radii are used as the source of wind data,
+#'    the \code{"gust_dur"} option cannot be selected.
+#' @param wind_source A character string specifying the source to use for the winds. Options
+#'    are \code{"modeled"}, for estimates based on running a wind model from Best
+#'    Tracks data inputs, and \code{"ext_tracks"}, for estimates based on
+#'    the wind radii in the Extended Best Tracks data. See the help files for the
+#'    datasets \code{storm_winds} and \code{ext_tracks_wind} in the
+#'    \code{hurricaneexposuredata} package for more details on each of these sources
+#'    for wind estimates. For the gust wind estimates, these are based on applying
+#'    a gust factor of 1.49 to the sustained wind estimates in both wind data sources.
 #'
 #' @return A dataframe with storms filtered based on the input criteria to the
 #'    function. Columns in the output will vary depending on the user's
@@ -127,11 +145,15 @@ filter_storm_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 #' @export
[…]

### 2017-01-27 R/wind_exposure.R

@@ -21,12 +21,17 @@
 #'                  (FIPS) code}
 #'      \item{\code{max_sust}: }{Maximum sustained wind speed (in m / s)}
 #'      \item{\code{max_gust}: }{Maximum gust wind speed (in m / s)}
+#'      \item{\code{sust_dur}: }{Minutes sustained wind speed was 20 m / s or higher}
+#'      \item{\code{gust_dur}: }{Minutes gust wind speed was 20 m / s or higher (only
+#'          returned if the modeled winds are requested using \code{wind_source})}
 #'    }
 #'
 #' @details For more information on how wind speeds are modeled in this data,
 #'    see the documentation for the \code{stormwindmodel} R package.
 #' @note Only counties in states in the eastern half of the United States can
-#'    be processed by this function.
+#'    be processed by this function. Winds are modeled at 15-minute increments, so
+#'    all duration estimates (\code{sust_dur} and \code{gust_dur} in the output)
+#'    will be divisible by 15.
 #'
 #' @examples
 #' # Ensure that data package is available before running the example.
@@ -42,16 +47,20 @@
 #'
 #' @importFrom dplyr %>%
 county_wind <- function(counties, start_year, end_year, wind_limit,
-                        wind_var = "vmax_sust"){
+                        wind_var = "vmax_sust", wind_source = "modeled"){
 
         hasData()
 
-        wind_df <- filter_wind_data(counties = counties,
+        output_vars <- c("storm_id", "fips", "vmax_sust", "vmax_gust", "sust_dur")
+        if(wind_source == "modeled") {
+                output_vars <- c(output_vars, "gust_dur")
+        }
+
+        wind_df <- filter_wind_data(counties = counties, wind_var = wind_var,
                                          year_range = c(start_year, end_year),
                                          wind_limit = wind_limit,
-                                         output_vars = c("storm_id", "fips",
-                                                         "vmax_sust",
-                                                         "vmax_gust")) %>%
+                                         wind_source = wind_source,
+                                         output_vars = output_vars) %>%
                 dplyr::left_join(hurricaneexposuredata::closest_dist,
                                  by = c("storm_id", "fips"))
 
@@ -71,6 +80,11 @@ county_wind <- function(counties, start_year, end_year, wind_limit,
 #'
 #' @return Returns the same type dataframe as \code{county_rain},
 #'    but with storms listed by community instead of county.
+#'
[…]

### 2017-01-27 R/map_exposure.R
Add wind_var option for wind mapping functions

@@ -364,19 +364,24 @@ map_distance_exposure <- function(storm, dist_limit, add_track = TRUE){
 #' # on installing the required data package.
 #' if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
 #'
-#' beryl_map <- map_wind_exposure(storm = "Beryl-1988", wind_limit = 15)
-#' beryl_map
-#' map_tracks("Beryl-1988", plot_points = FALSE, plot_object = beryl_map)
+#' map_wind_exposure(storm = "Beryl-1988", wind_limit = 15)
 #' }
 #' @importFrom dplyr %>%
 #'
 #' @export
-map_wind_exposure <- function(storm, wind_limit, add_track = TRUE){
+map_wind_exposure <- function(storm, wind_var = "vmax_sust", wind_limit,
+                              add_track = TRUE, wind_source = "modeled"){
 
-        map_data <- filter_wind_data(storm = storm,
-                                      output_vars = c("fips", "vmax_gust",
-                                                      "vmax_sust")) %>%
-                dplyr::mutate_(exposed = ~ vmax_sust >= wind_limit) %>%
+        if(wind_var %in% c("vmax_sust", "vmax_gust")){
+                wind_metric <- "m / s"
+        } else {
+                wind_metric <- "minutes"
+        }
+
+        map_data <- filter_wind_data(storm = storm, wind_source = wind_source,
+                                      output_vars = c("fips", wind_var)) %>%
+                `colnames<-`(c("fips", "wind_value")) %>%
+                dplyr::mutate_(exposed = ~ wind_value >= wind_limit) %>%
                 dplyr::mutate_(value = ~ factor(exposed,
                                                 levels = c("FALSE", "TRUE"))) %>%
                 dplyr::tbl_df()
@@ -401,7 +406,7 @@ map_wind_exposure <- function(storm, wind_limit, add_track = TRUE){
                                  colour = "black", fill = NA, size = 0.2, alpha = 0.5) +
                 ggplot2::theme_void() +
                 ggplot2::coord_map() +
-                ggplot2::scale_fill_manual(name = paste("Wind >", wind_limit, "m/s"),
+                ggplot2::scale_fill_manual(name = paste("Wind >", wind_limit, wind_metric),
                                            values = c("white", "darkorange"),
                                            labels = c("Unexposed", "Exposed"))
 
@@ -495,6 +500,7 @@ map_event_exposure <- function(storm_id, event_type, add_track = TRUE){
 #'    by this function.
[…]

### 2017-01-27 README.Rmd
Finalize vignette / README

@@
 
-This package allows you to explore and map data on county-level exposures to Atlantic-basin tropical storms between 1988 and 2015 for a number of storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
-This package allows the user to map exposures by county for a selected storm or to identify all storms to which selected counties were exposed, based on user-specified thresholds (for example, the package allows the user to identify all storms that brought wind of 34 knots or higher to Miami-Dade County in Florida).
+This package allows users to explore and map data of county-level exposures to Atlantic-basin tropical storms between 1988 and 2015 for a number of storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
+Users can map exposures by county for a selected storm and can also identify all storms to which selected counties were exposed, based on user-specified thresholds (for example, the package allows the user to identify all storms that brought wind of 34 knots or higher to Miami-Dade County in Florida).
 
@@
 To use the `hurricaneexposure` package, you will need to install `hurricaneexposuredata` on your computer.
-You can do that by adding the `drat` repository to your repositories; once you do this, you can install the `hurricaneexposuredata` using the `install.packages` function (and later update it using the `update.packages` function):
+You can do that by adding the `drat` archive to the list of repositories your system will query when adding and updating R packages.
+Once you do this, you can install the `hurricaneexposuredata` using the `install.packages` function (and later update it using the `update.packages` function):
 
@@
 - `storm_winds`: Modeled county wind speeds for Atlantic-basin storms, 1988-2015
-- `storm_events`: Listings from the NOAA Storm Events database that occurred near in time and location to tropical storms, 1988-2015.
+- `storm_events`: Listings from the [NOAA Storm Events database](https://www.ncdc.noaa.gov/stormevents/) that occurred near in time and location to tropical storms, 1988-2015.
 This database changed the types of events it reported in 1996, which should be considered when using the data.
-- `ext_tracks_wind`: Estimated county wind speeds for Atlantic-basin storms, 1988-2015, based on the wind radii listed in the Extended Best Tracks dataset
+- `ext_tracks_wind`: Estimated county wind speeds for Atlantic-basin storms, 1988-2015, based on the wind radii listed in the [Extended Best Tracks dataset](http://rammb.cira.colostate.edu/research/tropical_cyclones/tc_extended_best_track_dataset/)
 
-Once you've installed and loaded `hurricaneexposuredata`, you can load the included data using the `data` function. For example:
+Once you've installed and loaded `hurricaneexposuredata`, you can load the included data using the `data` function. For example, you can access the data on hurricane tracks (`hurr_tracks`) using:
 
@@
 
-For each dataset included in `hurricaneexposuredata`, you can see the helpfiles for the data for more information about the data included in each (e.g., `?hurr_tracks` to read documentation on the hurricane tracks data shown above).
+For each dataset included in `hurricaneexposuredata`, you can see the helpfiles for the data for more information (e.g., `?hurr_tracks` to read documentation on the hurricane tracks data shown above).
 This data documentation includes both explanations of how the dataset was created, definitions of the variables included in each dataset, and the sources of data included.
@@
 In functions throughout the `hurricaneexposure` package, storms are identified based on their name and year (e.g., Hurricane Floyd in 1999 is identified as "Floyd-1999").
-Note that a few storms satisfied this criteria but were unnamed (e.g., "Notnamed" in 1991, "Subtrop" in 1997, etc.).
-These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table (e.g., "Notnamed-1991"), although for many applications you may want to consider excluding these unnamed storms from the full analysis.
+Note that a few storms satisfied the criteria to be included in the data but were unnamed (e.g., "Notnamed" in 1991, "Subtrop" in 1997, etc.).
+These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table (e.g., "Notnamed-1991"), although for some research projects you may want to consider excluding these unnamed storms from the analysis.
 
@@
 The `hurricaneexposure` package has functions to interact with the data in the `hurricaneexposuredata` package.
-First, this package has functions that can be used to create some different maps of hurricane exposures based on distance to the storm track and rainfall.
+First, there are functions that can be used to create some different maps of hurricane exposures based on distance to the storm track and rainfall.
 
@@
 
-First, for several of the storm hazard measurements (currently rain, wind, and distance), you can use the `map_counties` function in `hurricaneexposure` to plot a map for a specific storm showing the continuous value of that metric for all eastern U.S.
+For several of the storm hazard measurements (currently rain, wind, and distance), you can use the `map_counties` function in `hurricaneexposure` to plot a map for a specific storm showing the continuous value of that metric for all eastern U.S.
 counties.
@@
 
-By default, this map shows the cumulative rain in each county for two days before to one day after the date that the storm passed closest to the county (see the help file for the `closest_dist` dataset in `hurricaneexposuredata` to find out more about how this date was determined for each county-storm combination).
+By default, this map shows the cumulative rain in each county for two days before to one day after the date that the storm passed closest to the county (see the help file for the `closest_dist` dataset in `hurricaneexposuredata` to find out more about how this date of closest approach was determined for each county-storm combination).
 The period used to calculate cumulative rainfall for the map can be customized using the `days_included` option in the `map_counties` function.
@@
 
-These county winds are determined based on a model of wind speeds from Best Tracks hurricane data.
+These county winds are determined based on a model of wind speeds, using as input data from the hurricane tracks data.
 See the documentation for the `storm_winds` dataset (`?storm_winds`) to find out more about this data; further details on the wind model are available in the vignettes for the [`stormwindmodel` package](https://cran.r-project.org/web/packages/stormwindmodel/index.html).
 
-[Add details about how to map this from the wind radii estimates, instead]
+You can map a few other variables of wind with this function. For example, you can map the duration of winds of 20 m / s or more by using the argument `wind_var = "sust_dur"`:
 
-Finally, you can also use this function to plot the closest distance between the storm and each county.
+```{r fig.width = 7, fig.height = 4}
+map_counties("Katrina-2005", metric = "wind", wind_var = "sust_dur")
+```
+
+You can map estimated gust winds, rather than sustained winds (the default), using the argument `wind_var = "vmax_gust"`.
+Further, we have included a second source of estimated winds in the data available in `hurricaneexposuredata`.
+These wind estimates are based on the wind radii in the Extended Best Tracks dataset (you can find out more about this data in the helpfile for the `ext_tracks_wind` dataset).
+These data provide estimates of which counties were exposured during a storm to sustained winds in four categories: 0--34 knots; 34--50 knots;  50--64 knots; and 64 knots or higher.
+Therefore, these data provide a categorical rather than continuous estimate of county wind speeds.
+However, they may be preferable to the model wind speeds for some storms, especially storms in extra-tropical transition.
+Further, you may find it interesting to use these data in a sensitivity analysis, to compare if impact study results are sensitive to whether these wind data or the modeled wind data are used for exposure classification.
+You can create a wind map based on this data using the `wind_source = "ext_tracks"` argument in the `map_counties` function.
+For example, to map estimated winds during Hurricane Katrina based on the Extended Best Tracks wind radii, you can run:
+
+```{r fig.width = 7, fig.height = 4}
+map_counties("Katrina-2005", metric = "wind", wind_source = "ext_tracks")
+```
+
+Finally, you can also use the `map_counties` function to plot the closest distance between the storm and each county.
 For this, you use the argument `metric = "distance"`.
@@
 
-For example, you can map a binary variable of distance-based exposure using `map_distance_exposure`, with "exposure" defined as the county being within the number of kilometers given in `dist_limit` of the storm track.
+For example, you can map a binary variable of distance-based exposure using `map_distance_exposure`, with "exposure" defined as that the county was within the number of kilometers given in `dist_limit` of the storm track.
 For example, you can map the counties that were within 75 km of Hurricane Sandy's storm track with the call:
@@
 
-Similarly, you can map a binary variable county exposure based on rain using `map_rain_exposure`.
+Similarly, you can map binary county exposure based on rain using `map_rain_exposure`.
 In the case of rain, you must define exposure thresholds for both rainfall and distance.
 This is because the rain data is observed data, so neglecting to limit "exposed" counties to those close to the storm track may pick up counties far from the storm that experienced a lot of rain over the same period.
-For example, you can use the following call to map rain exposure to Tropical Storm Allison (2001), where exposure is defined as a county where cumulative rainfall from fives days before to three days after the storm's closest approach was 175 mm or more and where the county was within 500 km of the storm's track:
+For example, you can use the following call to map rain exposure to Tropical Storm Allison (2001), where a county is defined as exposed to the storm if cumulative rainfall from five days before to three days after the storm's closest approach was 175 mm or more and the county was within 500 km of the storm's track:
 
@@
 
-You can use the `map_wind_exposure` function to map counties exposed to a certain level of wind during a storm.
+You can use the `map_wind_exposure` function to map counties exposed to a certain winds during a storm.
 For example, to identify all counties exposed to sustained winds of 20 m / s or higher during Hurricane Katrina, you can run:
@@
 
-Event listings from the NOAA Storm Events database were linked to storms based on the start date of the event being within a five-day window of the date the storm was closest to the county and the county being within 500 km of the county.
+Event listings from the NOAA Storm Events database were linked to storms based on the start date of the event being within a five-day window of the date the storm was closest to the county and the storm coming within 500 km of the county.
 The [`noaastormevents` package](https://github.com/zailchen/noaastormevents) was used to create this dataset.
@@
 This function can be used either to map a storm track on a blank U.S.
-map or to add a track to an existing map, if it is a `ggplot` object (as are all maps generated in the previous section).
+map or to add a track to an existing map, if the map is a `ggplot` object (as are all maps generated in the previous section).
 
@@
 
-The `map_tracks` function is used in all the mapping functions described in the previous function to add the storm's track to the exposure map.
+As another example, to map all tracks for storms in 2005, you can run:
+
+```{r fig.width = 5.5, fig.height = 4}
+library(dplyr)
+library(tidyr)
+storms_2005 <- hurr_tracks %>%
+        select(storm_id) %>%
+        distinct() %>%
+        separate(storm_id, c("name", "year"), sep = "-", remove = FALSE) %>%
+        filter(year == "2005")
+map_tracks(storms = storms_2005$storm_id) 
+```
+
+The `map_tracks` function is used in all the mapping functions described in the previous section, to add the storm's track to the exposure maps.
 However, if you would like to customize the appearance of the storm's track on the map, you can do so by plotting the exposure map without the storm track (`add_track = FALSE`) and then adding the track yourself with the `map_tracks` function.
@@
 
-The `hurricaneexposure` package also has several functions that can input a list of counties and output a list of all of the storms each county was exposed to over a certain period, based on one of the hazard metrics.
+The `hurricaneexposure` package also has several functions that can input a list of counties and output a list of all of the storms to which each county was exposed over a certain period, based on one of the hazard metrics.
 These exposure listings include the storm dates, making them easy to integrate into time series datasets of health or other impact data.
 
-For example, the `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds to define rain-based exposure (cumulative rainfall and distance from the storm's track, as explained for the binary rain maps shown in an earlier section) and creates a list of all storms that met these thesholds for the counties.
+The `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds to define rain-based exposure (cumulative rainfall and distance from the storm's track, as explained for the binary rain maps shown in an earlier section) and creates a list of all storms that met these thesholds for the counties.
 For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over a three-day window of the date of closest approach was 100 millimeters or more, you can run:
@@
 
-The returned dataframe includes the estimated maximum values during the storm for sustained winds (`vmax_sust`) and gust winds (`vmax_gust`), both in m / s.
+The returned dataframe includes the estimated maximum values during the storm for sustained winds (`vmax_sust`) and gust winds (`vmax_gust`), both in m / s, as well as the duration for which sustained and gust winds were over 20 m / s in the county (`sust_dur` and `gust_dur`, respectively).
 It also includes some of the variables on the date and time of closest approach and the distance at the storm's closest approach.
+
+The `county_wind` function allows some further specifications when identifying the storms to which a county was exposed based on wind estimates.
+First, it allows the use of alternative wind variables.
+While the default is to use the maximum sustained wind speed experienced in the county during the storm, it is also possible to base the exposure metric on the maximum gust winds (`wind_var = "vmax_gust"`) or on the duration of either sustained or gust winds at or above 20 m / s (`wind_var = "sust_dur"` and `wind_var = "gust_dur"`, respectively).
+For example, to determine which storms brought Orleans Parish sustained winds of 34 knots or higher (17.5 m / s or higher) for an hour or more (`wind_limit = 60`; for wind durations, this wind limit should be specified in minutes), you can run:
+
+```{r}
+county_wind(counties = "12086", start_year = 1988, end_year = 2015, 
+            wind_var = "sust_dur", wind_limit = 60)
+```
+
+Further, the `county_wind` function allows you to pull wind estimates from an alternate source.
+By default, this function uses wind estimates for each county from a wind model (see the vignettes for the `stormwindmodel` package for much more detail on this modeling process).
+However, you can also pull estimates based on the wind radii from the Extended Best Tracks dataset by using the option `wind_source = "ext_tracks"`.
+See the help file for the `ext_tracks_wind` dataset that comes with `hurricaneexposuredata` for more details on this data source and how the county-specific exposures were determined from the data source for use in the `county_wind` function.
+It is important to note that this data results in wind estimates with breaks at 34 knots, 50 knots, and 64 knots, rather than continuous estimates.
+This means that a sustained wind estimate from this data source of 17.4896 m / s (34 knots) is estimating that the county had maximum wind speeds of 34 knots or higher during the storm, but not as high as 50 knots.
+Further, the `dur_sust` value when using this data source is based on number of minutes with winds at or above 34 knots (rather than the 20 m / s value used for durations for the modeled wind data).
+Gust durations can not be determined using this Extended Best Tracks dataset.
+
+The Extended Best Tracks wind radii dataset is more tied to observations during a specific storm than the modeled data.
+In certain cases, especially storms in extratropical transition, this dataset might be preferable to the modeled wind speeds, even though it provides less continuous estimates.
+This source of wind data can be specified using the `wind_source` option in the `county_wind` function.
+For example, to generate a list of all storms in Orleans Parish with maximum sustained winds of 34 knots or more based on the Extended Best Tracks wind radii, you can run:
+
+```{r}
+county_wind(counties = "12086", start_year = 1988, end_year = 2015, 
+            wind_var = "vmax_sust", wind_limit = 17.4, wind_source = "ext_tracks")
+```
 
@@
 Similar functions exist for wind exposure (`multi_county_wind`) and distance exposure (`multi_county_distance`).
+In some cases, these multi-county functions may have less functionality than their `county_*` counterparts.
+For example, the `multi_county_wind` function currently only allows use of the sustained wind speeds for the exposure metric, while the `county_wind` function allows the use of sustained winds, gust winds, or durations of either.

### 2017-01-27 R/filter_storm_data.R

@@ -149,6 +149,11 @@ filter_wind_data <- function(counties = NULL, storm = NULL, year_range = NULL,
 
         hasData()
 
+        if(wind_source == "ext_tracks" & wind_var == "gust_dur"){
+                stop("When using Extended Best Tracks for the wind data source, ",
+                     "you cannot use the\ngust duration as the wind variable.")
+        }
+
         if(wind_source == "modeled"){
                 storm_winds <- data.table::data.table(hurricaneexposuredata::storm_winds)
         } else if (wind_source == "ext_tracks"){

### 2017-01-27 R/map_exposure.R

@@ -517,12 +517,14 @@ map_event_exposure <- function(storm_id, event_type, add_track = TRUE){
 #' map_counties("Katrina-2005", metric = "wind")
 #' map_counties("Katrina-2005", metric = "wind", wind_var = "vmax_gust")
 #' map_counties("Katrina-2005", metric = "wind", wind_var = "sust_dur")
+#' map_counties("Katrina-2005", metric = "wind", wind_source = "ext_tracks")
 #'}
 #' @export
 #'
 #' @importFrom dplyr %>%
 map_counties <- function(storm, metric = "distance", wind_var = "vmax_sust",
-                        days_included = c(-2, -1, 0, 1), add_track = TRUE){
+                        days_included = c(-2, -1, 0, 1), add_track = TRUE,
+                        wind_source = "modeled"){
         if(metric == "distance"){
                 map_data <- filter_storm_data(storm = storm,
                                               output_vars = c("fips",
@@ -544,7 +546,8 @@ map_counties <- function(storm, metric = "distance", wind_var = "vmax_sust",
         }
         map_data <- map_data %>%
                 dplyr::tbl_df()
-        out <- hurr_choropleth(map_data, metric = metric)
+        out <- hurr_choropleth(map_data, metric = metric, wind_var = wind_var,
+                               wind_source = wind_source)
 
         if(add_track){
                 out <- map_tracks(storm, plot_object = out)
@@ -571,7 +574,8 @@ map_counties <- function(storm, metric = "distance", wind_var = "vmax_sust",
 #' to Atlantic basin tropical storms.
 #'
 #' @importFrom dplyr %>%
-hurr_choropleth <- function(map_data, metric = "distance", wind_var = "vmax_sust"){
+hurr_choropleth <- function(map_data, metric = "distance", wind_var = "vmax_sust",
+                            wind_source = "modeled"){
 
         if(metric == "rainfall"){
                 breaks <- seq(0, 200, by = 25)
@@ -583,9 +587,18 @@ hurr_choropleth <- function(map_data, metric = "distance", wind_var = "vmax_sust
                 exposure_legend <- "Distance (km)"
         } else if(metric == "wind"){
                 palette_name <- "Reds"
-                if(wind_var %in% c()){
-                        breaks <- c(0, seq(15, 45, by = 5))
+                if(wind_var %in% c("vmax_gust", "vmax_sust")){
                         exposure_legend <- "Wind speed (m / s)"
[…]

### 2017-01-27 R/wind_exposure.R
Update documentation

@@ -145,6 +145,7 @@ multi_county_wind <- function(communities, start_year, end_year,
 #' @inheritParams county_distance
 #' @inheritParams county_rain
 #' @inheritParams rain_exposure
+#' @inheritParams filter_wind_data
 #'
 #' @return Writes out a directory with rain exposure files for each county or
 #'    community indicated. For more on the columns in this output, see the

### 2017-01-27 R/distance_exposure.R
Final edits of help files

@@ -137,9 +137,9 @@ multi_county_distance <- function(communities, start_year, end_year,
 #'
 #' This function takes an input of locations (either a vector of county FIPS
 #' or a dataframe of multi-county FIPS, with all FIPS listed for each county)
-#' and creates time series dataframes that can be merged with health time series,
-#' giving the dates and exposures for all storms meeting the given
-#' storm distance criteria.
+#' and creates a dataframe with storm listings and dates that can be merged with
+#' time series of health or other outcomes, giving the dates and exposures for all
+#' storms meeting the given storm distance criteria.
 #'
 #' @inheritParams county_distance
 #' @inheritParams county_rain

### 2017-01-27 R/event_exposure.R

@@ -1,5 +1,12 @@
 #' Find events for storms by county
 #'
+#' Takes a list of US counties, based on their 5-digit Federal Information Processing
+#' Standard (FIPS) codes, boundaries on the range of years to be considered, and
+#' the type of event (e.g., "flood", "tornado") and returns a list of all storms with
+#' listings for that type of event for the counties. This output is based on listings in
+#' the NOAA Storm Events database. See the help documentation for the \code{storm_events}
+#' dataset for more information and references on the data source used by this function.
+#'
 #' @param event_type Character string with the type of event to use to identify
 #'    county exposures. Options include \code{"flood"}, \code{"tornado"},
 #'    \code{"wind"}, and \code{"tropical storm"}.
@@ -11,6 +18,28 @@
 #'    event type, you should not use a start year prior to 1996, as events of
 #'    other types were not recorded in the database before 1996.
 #'
+#' #' @return Returns a dataframe with a row for each county-storm pair and with
+#'    columns for:
+#'    \itemize{
+#'      \item{\code{fips}: }{County's 5-digit Federal Information Processing Standard
+#'                  (FIPS) code}
+#'      \item{\code{storm_id}: }{Unique storm identifier with the storm name and year,
+#'                  separated by a hyphen(e.g., "Alberto-1988",
+#'                  "Katrina-2005")}
+#'     \item{\code{closest_time_utc}: }{Time, in UTC, of the closest approach of the
+#'                             storm to the county's population mean center,
+#'                             based on storm tracks linearly interpolated to
+#'                             15-minute increments.}
+#'      \item{\code{storm_dist}: }{Minimum distance (in kilometers) between the storm's
+#'                     track and the county's population mean center.}
+#'      \item{\code{local_time}: }{Local time of the closest approach of the storm to the
+#'                        county's population mean center, based on storm tracks
+#'                        linearly interpolated to 15-minute increments.}
+#'      \item{\code{closest_date}: }{Date (based on local time) of the closest
+#'                          approach of the storm to the county's population
+#'                          mean center.}
+#'    }
+#'
 #' @examples
 #' # Ensure that data package is available before running the example.
 #' #  If it is not, see the `hurricaneexposure` package vignette for details
@@ -18,7 +47,7 @@
 #' if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
 #'
 #' county_events(counties = c("37031", "37053", "37055"),
[…]

### 2017-01-27 R/map_exposure.R

@@ -3,13 +3,13 @@
 #' Creates a \code{ggplot} object with the underlying map of all states in the
 #' eastern section of the US that might be prone to hurricane-related
 #' exposure. Other lines and points can be added to the output using
-#' `ggplot2` plotting functions.
+#' \code{ggplot2} plotting functions.
 #'
 #' @details Only states in the eastern half of the United States (i.e., ones
 #' prone to exposure to Atlantic basin tropical storms) are included on this
 #' map.
 #'
-#' @return A ggplot object that maps the states of the Eastern United States
+#' @return A \code{ggplot} object that maps the states of the Eastern United States
 #'
 #' @export
 #'
@@ -83,14 +83,15 @@ get_eastern_map <- function(map  = "county"){
 #'    outer limits of the plot object (or default map if `plot_object` is
 #'    left as NULL) when cropping hurricane tracks.
 #' @param plot_points TRUE / FALSE indicator of whether to include points,
-#'    as well as lines, when plotting the hurricane tracks.
+#'    as well as lines, when plotting the hurricane tracks. These points show
+#'    the times for which observations were recorded in the hurricane track data.
 #' @param alpha Numerical value designating the amount of transparency to
 #'    use for plotting tracks.
 #' @param color Character string giving the color to use to plot the tracks.
 #'
 #' @return Returns a ggplot object with plotting data for the storm tracks
 #'    of the selected storms. This object can be printed directly or added
-#'    on to with other ggplot commands.
+#'    to with other \code{ggplot2} commands.
 #'
 #' @examples
 #' # Ensure that data package is available before running the example.
@@ -419,11 +420,21 @@ map_wind_exposure <- function(storm, wind_var = "vmax_sust", wind_limit,
 
 #' Map county-level exposure based on reported events
 #'
+#' Map counties as "exposed" or "unexposed" based on the criterion that the county
+#' had an event listing of a specified type in the NOAA Storm Events database.
+#' For more information on the underlying data, see the helpfile for the
+#' \code{storm_events} dataset.
+#'
 #' @param storm_id Character vector with the storm for which to map events
 #'    (e.g., \code{"Katrina-2005"})
 #' @inheritParams county_distance
 #' @inheritParams county_events
 #' @inheritParams map_rain_exposure
+#'
+#' @return A map showing whether eastern US counties were exposed or unexposed
+#'    to a specific storm based on event listings.
+#'
+#' @note Note that flood events are not available for any year before 1996.
 #'
[…]

### 2017-01-27 R/rain_exposure.R

@@ -1,9 +1,9 @@
 #' Hurricane exposure by rain for counties
 #'
-#' This function takes a list of US counties,based on their 5-digit Federal
+#' This function takes a list of US counties, based on their 5-digit Federal
 #' Information Processing Standard (FIPS) codes, boundaries on
 #' the range of years to be considered, thresholds for distance between each
-#' county and the storm track as well as minimum rainfall over a certain time
+#' county and the storm track, as well as minimum rainfall over a certain time
 #' window for the county to be considered "exposed" to the
 #' storm. Based on these inputs, the function returns a dataframe with the
 #' subset of Atlantic basin storms meeting those criteria for each of the listed

### 2017-01-27 R/wind_exposure.R

@@ -1,9 +1,10 @@
 #' Hurricane exposure by wind for counties
 #'
-#' This function takes a list of US counties,based on their 5-digit Federal
+#' This function takes a list of US counties, based on their 5-digit Federal
 #' Information Processing Standard (FIPS) codes, boundaries on
 #' the range of years to be considered, and thresholds for wind speed (in meters
-#' per second) for each  county to be considered "exposed" to the
+#' per second) (or, alternatively, duration of winds at or above 20 m / s in minutes)
+#' for each  county to be considered "exposed" to the
 #' storm. Based on these inputs, the function returns a dataframe with the
 #' subset of Atlantic basin storms meeting those criteria for each of the listed
 #' counties.

### 2017-02-05 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Re-run everything with latest versions of `countyweather` and `hurricaneexposure`

@@
         if(get_data){
-                daily_timeseries(ex_fips, coverage = 0,
+                write_daily_timeseries(ex_fips, coverage = 0,
                                        date_min = "1988-01-01",
@@
 
-        ex_weather <- readRDS(paste0(ex_dir, ex_fips, ".rds"))
-        ex_weather <- as.data.frame(ex_weather$daily_data)
+        ex_weather <- readRDS(paste0(ex_dir, "data/", ex_fips, ".rds"))
 
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-daily_timeseries(allison_fips, coverage = 0,
+write_daily_timeseries(allison_fips, coverage = 0,
                        date_min = format(date_range[1, 1], "%Y-%m-%d"),
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-daily_timeseries(frances_fips, coverage = 0,
+write_daily_timeseries(frances_fips, coverage = 0,
                        date_min = format(date_range[1, 1], "%Y-%m-%d"),
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-daily_timeseries(ike_fips, coverage = 0,
+write_daily_timeseries(ike_fips, coverage = 0,
                        date_min = format(date_range[1, 1], "%Y-%m-%d"),
@@
                   latest_date = max(ymd(closest_date)) + ddays(2))
-daily_timeseries(alex_fips, coverage = 0,
+write_daily_timeseries(alex_fips, coverage = 0,
                        date_min = format(date_range[1, 1], "%Y-%m-%d"),
@@
         counties <- gsub(".rds", "",
-                         list.files(paste0("county_ave/", county_dir)))
+                         list.files(paste0("county_ave/", county_dir, "/data")))
         if(county_dir == "frances"){
@@
         for(x in counties){
-                county_weather <- readRDS(paste0("county_ave/", 
-                                                 county_dir, "/", x,
-                                                 ".rds"))$daily_data %>%
+                county_weather <- readRDS(paste0("county_ave/", county_dir, "/data/",
+                                                 x, ".rds")) %>%
                         dplyr::mutate(date = ymd(date),
@@
                              wind_limit = 0,
-                             wind_var = "max_sust") %>%
+                             wind_var = "vmax_sust") %>%
                 dplyr::filter(storm_id == data$storm_id[1]) %>%
@@
                   highest_wind = max(wind_sust),
-                  model_wind = mean(max_sust))
+                  model_wind = mean(vmax_sust))
 
         plot_mwr <- ggplot(compare_wind, 
-                   aes(x = max_sust, y = wind_sust)) + 
+                   aes(x = vmax_sust, y = wind_sust)) + 
         geom_hline(yintercept = 15, color = "lightgray") + 
@@
 a <- wind_exposure %>%
-        arrange(desc(max_sust)) %>% 
+        arrange(desc(vmax_sust)) %>% 
         slice(1:20) %>%
         left_join(county_centers, by = "fips") %>%
-        select(storm_id, name, max_sust) %>%
+        select(storm_id, name, vmax_sust) %>%
         mutate(storm_id = paste0(gsub("-", " (", storm_id), ")"),
-               max_sust = round(max_sust)) %>%
+               max_sust = round(vmax_sust)) %>%
         rename(Storm = storm_id,
                County = name,
-               "Maximum sustained wind (m / s)" = max_sust) %>%
+               "Maximum sustained wind (m / s)" = vmax_sust) %>%
         xtable(caption = "Top storm-county pairs in terms of most intense sustained winds.", 
@@
 ```{r top_wind_map, echo = FALSE, fig.width = 8, fig.height = 6, fig.align = "center", fig.cap = "Number of storms for which each county experienced one of the top 2,000 intensities for distance"}
-plot_top_x(wind_exposure, exposure_var = "max_sust",
+plot_top_x(wind_exposure, exposure_var = "vmax_sust",
            eastern_states = eastern_states, my_fips = my_fips,

### 2017-07-04 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Add sentence describing drat from our paper

@@
 
-Some studies (e.g., @Zandbergen2009, @Czajkowski2011) assumed symmetrical activity of storm-related factors,which is not a realistic assumption for storm wind and rain patterns, particularly at and following landfall [@Kruk2010, @Halverson2015, @Rezapour2014].
+Some studies (e.g., @Zandbergen2009, @Czajkowski2011) assumed symmetrical activity of storm-related factors, which is not a realistic assumption for storm wind and rain patterns, particularly at and following landfall [@Kruk2010, @Halverson2015, @Rezapour2014].
 
@@
 
-Currently, distance parameters involved with assessing risk of a particular storm have been rather arbitrary, contributing to the necessity in understanding how such parameters influence a county's exposure status.
+To date, distance parameters involved with assessing risk of a particular storm have been rather arbitrary, contributing to the necessity in understanding how such parameters influence a county's exposure status.
 In assessing the risk of a given storm based on distance, recent studies have defined the distance from the storm track affected by a given storm differently.
@@
 
-Distance-based measures that are based on a storm track passing through a county can also miss exposures from storms when they are offshore but near enough to the coast to cause important hazards to health and property (e.g., Agnes in 1972, FLoyd in 1999, Chantal in 1989) [@Czajkowski2011].
+Distance-based measures that are based on a storm track passing through a county can also miss exposures from storms when they are offshore but near enough to the coast to cause important hazards to health and property (e.g., Agnes in 1972, Floyd in 1999, Chantal in 1989) [@Czajkowski2011].
 By basing the distance criteria on distance from the storm's track to county centers, more of these offshore storm exposures could be captured.

### 2017-07-13 papers/final_R_journal_paper/anderson-eddelbuettel-revised.Rmd
Add R Journal final paper draft and UseR slides

---
title: Hosting Data Packages via \pkg{drat}&#58; A Case Study with Hurricane Exposure Data
author:
  - name: G. Brooke Anderson
    affiliation: Colorado State University
    address:
    - Lake Street
    - Fort Collins, CO
    email:  brooke.anderson@colostate.edu
  - name: Dirk Eddelbuettel
    affiliation: Debian and R Projects; Ketchum Trading
    address: 
    - Chicago, IL, USA
    email: edd@debian.org
abstract: >
  Data-only packages offer a way to provide extended functionality for
  other R users. However, such packages can be large enough to exceed
  the package size limit (5 megabytes) for the Comprehensive R Archive Network (CRAN). As an 
  alternative, large data packages can be posted to additional repostiories 
  beyond CRAN itself in a way that allows smaller code packages on CRAN to access 
  and use the data.  The \CRANpkg{drat} package facilitates creation and use of such 
  alternative repositories and makes it particularly simple to host them via GitHub. 
  CRAN packages can draw on packages posted to \pkg{drat} repositories through the use of the
  `Additonal_repositories' field in the DESCRIPTION file. This paper describes how R users
  can create a suite of coordinated packages, in which larger data packages are hosted 
  in an alternative repository created with \pkg{drat}, while a smaller code package that 
  interacts with this data is created that can be submitted to CRAN.
preamble: >
  % Any extra latex you need in the preamble
output: rticles::rjournal_article 
---

# Motivation

"Big data", apart from being a buzzword, also accurately describes the current scale of many
scientific data sets. While the R language and environment \citep{rcore:rbase} enables the creation,
use, and sharing of data packages to support methodology or application packages, the size of these
data packages can be very large. The Bioconductor project has addressed the potentially large size
requirements of data packages through the use of Git Large File Storage, with the package contributor covering costs for extremely
large data packages (over 1 gigabyte) \citep{biocPkgGuide}. The Bioconductor
repository, though, is restricted to topic-specific packages related to bioinformatics. The Comprehensive R Archive Network (CRAN), which archives R packages on any topic, has a
recommended size limit (reasonable for a widely-mirrored repository) of 5 megabytes (MB) for package
data and documentation \citep{rcore:wre}. A twofold need therefore arises for package maintainers
seeking to share large R data packages that are outside the scope of the Bioconductor
project. First, there is a need to create and share such a data package for integration into and extensions of
a given methodology or application package, and, second, there is a need to integrate use of such a
package in a way that makes it seamlessly integrated with smaller CRAN packages that use the 
data package. Here, we outline one possible approach to satisfy these needs by creating a suite of
coordinated packages, in which larger data packages are hosted outside of CRAN but can still be accessed by smaller
code packages that are submitted to CRAN.

The problem of creating CRAN packages that interface with large datasets is not new, and various
approaches have been taken in the past to allow for such an interface. For example, the
\pkg{NMMAPSlite} package (currently available only from the CRAN archive) was built to allow users
to interact with daily data on weather, air pollution, and mortality outcomes for over 100 US
communities over 14 years \citep{peng2008statistical}. To enable interaction with this large dataset
through a CRAN package, the package maintainer posted the data on a server and included functions in
the \pkg{NMMAPSlite} package to create an empty database on the user's computer and fill it
with community-specific datasets as the user queried different communities. This data
interaction was enabled with the \CRANpkg{stashR} package by the same maintainer \citep{stashR}.

More recent packages similarly allow interaction between a web-hosted database and R, in some cases
for a database maintained by an outside entity. For example, the \CRANpkg{rnoaa} package allows
access to weather data posted by the National Oceanic and Atmospheric Administration (NOAA)
\citep{rnoaa}, while the \CRANpkg{tigris} package allows access to spatial data from the United
States Census Bureau \citep{tigris}. Both packages are posted on CRAN and allow R users to work with
a large collection of available online data by creating and sending HTTP requests from R using
the conventions defined in the respective online databases' application program interfaces (APIs). This
approach is a good one for data that is already available through an online database, especially if
the database is outside the control of the R package maintainer or if the potential set of data is
extremely large and it is unlikely that an R user would want to download all of it, since an API allows
an R user to selectively download parts of the data. However, if the data is not already available
through an online database, this approach would require the R package maintainer to create and
maintain an online database, including covering the costs and managing the security of that
database.

Another approach is to create a suite of packages, in which smaller ('code') packages are submitted
to CRAN while larger ('data') packages are posted elsewhere. With this approach, the R user
downloads all the data in the data package when he or she installs the data package, effectively
caching this data locally so that data can be used in different R sessions without reloading or
when the computer is offline. This can be a good approach in cases where the data is not otherwise
available through an online database and when R users are likely to want to download the full set of
data. In this case, this second approach offers several advantages, including: (1) the data can be
documented through package helpfiles that are easily accessible from the R console; (2) if
a user would like to delete all downloaded data from their computer, he or she can easily do so by
removing the entire data package with \code{remove.packages} (as compared to other caching
solutions, in which case the user might need to do some work to determine where a package cached
data on his or her computer); (3) versioning can be used for the data package, which can improve
reproducibility of research using the data \citep{gentleman2004bioconductor}; and (4) the data
package can include the R scripts used to clean the original data (for example, in a \file{data-raw}
directory, with the directory's name excluded from the R package build through a listing in the
\file{.Rbuildignore} file), which will make the package more reproducible for the package author
and, if the full package directory is posted publicly (e.g., through a public GitHub repository),
for other users.

The \pkg{UScensus2000} suite of packages (currently available from the CRAN archive) used this
approach to allow access to U.S. Census Bureau data from the 2000 Decennial Census
\citep{almquist2010us}. This suite of packages included data at a more aggregated spatial level
(e.g., state- and county-level data) through data packages submitted to CRAN, but included the
largest dataset (block-level data) in an R package that was instead posted to the research lab's
website \citep{almquist2010us}. A convenience function was included in one of the CRAN packages in
the suite to facilitate installing this data package from the lab's website \citep{almquist2010us}.

This approach can be facilitated by posting the data package through an online package repository rather than a non-repository website. While
support for repositories outside of CRAN, Bioconductor, and OmegaHat has existed within R for years,
few users appear to have deployed this mechanism to host additional repositories. The \CRANpkg{drat}
package facilitates the creation and use of a user-controlled package repository. Once a package
maintainer has created a repository, he or she can use it to host larger packages like a data
package (and of course also any number of code packages). Use of
a \pkg{drat} repository allows R users to install and update the data package using traditional R functions for managing packages
(e.g., \code{install.packages}, \code{update.packages}) after the user has added the \pkg{drat}
repository through a call to \code{addRepo} \citep{drat}. A \pkg{drat} repository can be published
online through Git Pages, and GitHub repositories have a recommended maximum size of 1 GB, much
larger than the size limit for a CRAN package, with a cap of 100 MB on any single file
(\url{https://help.github.com/articles/what-is-my-disk-quota/}). Even for data packages below the CRAN
size limit, this approach to hosting data packages can help remove some of the burden of hosting and
archiving large data packages from CRAN. Although the \pkg{drat} package is
relatively new, some package maintainers are already taking this approach---for example, the
\CRANpkg{grattan} package facilitates research in R on Australian tax policy, with relevant data
available through the large \pkg{taxstats} data package, posted in a \pkg{drat} repository.

When taking the approach of creating a suite of packages in which the smaller code package or
packages are submitted to CRAN while larger data packages are hosted in \pkg{drat} repositories,
it is necessary to add some infrastructure to the smaller code packages.  CRAN policies do not allow the
submission of a package with mandatory dependencies on packages hosted outside of a mainstream
repository, which means that the smaller package could not be submitted to CRAN if the data package
is included in the smaller package through 'Imports:' or 'Depends:'. The R
package ecosystem offers a solution: a package maintainer can create a weaker relationship between
code and data packages via a 'Suggests:', which makes the data package optional, rather than
mandatory, as either 'Imports:' or 'Depends:' would. Being optional, one then has to condition any
code in the code package that accesses data in the data package on whether that package is
present on the user's system. This approach offers the possibility of posting a data package that is
too large for CRAN on a non-mainstream respository, like a \pkg{drat} repository, while
submitting a package that interacts with the data to CRAN.

This paper outlines in sufficient detail the steps one has to take to use \pkg{drat} to host a
large data package for R and how to properly integrate conditional access to such an optional
data package into a smaller code package. The packaging standard aimed for is the CRAN Repository
Policy and the passing of \samp{R CMD check --as-cran}. Broadly, these steps are:

1. Create a \pkg{drat} repository;
2. Create the data package(s), build it, and post it in the \pkg{drat} repository; and
3. Create / alter the code package(s) to use the data package(s) in a way that complies with CRAN checks.

As a case study, we discuss and illustrate the interaction between the packages \pkg{hurricaneexposure}
\citep{hurricaneexposure} and \pkg{hurricaneexposuredata} \citep{hurricaneexposuredata}.  The latter
package contains data for eastern U.S. counties on rain, wind, and other hurricane exposures,
covering all historical Atlantic-basin tropical storms over a few decades. The total size of the
package's source code is approximately 25 MB, easily exceeding CRAN's package size limit. This
package includes only data, but the companion package, \pkg{hurricaneexposure}, provides functions to
map and interact with this data. The \pkg{hurricaneexposure} package is of standard size and
available from CRAN since the initial version 0.0.1, but to fully utilize all its capabilities
requires access to the data in package \pkg{hurricaneexposuredata}. Here, we highlight specific
elements of code in these packages that allow coordination between the two. The full code for both
packages is available through GitHub repositories
(\url{https://github.com/geanders/hurricaneexposure} and
\url{https://github.com/geanders/hurricaneexposuredata}); this article references code in
version 0.0.2 of both packages.

# Posting a data package to a drat repository

## Creating a drat repository

A package maintainer must first create a \pkg{drat} repository if he or she wishes to host packages through one. Essentially, this repository is a way to store R packages such that it is easy for R users to download and update the packages; the repository can be shared, among other ways, through a GitHub-hosted website. Because a \pkg{drat} repository is controlled by the package maintainer, it allows increased flexibility to package maintainers compared to repositories like CRAN.  A single \pkg{drat} repository can host multiple packages, so a maintainer likely only needs a single \pkg{drat} repository, regardless of how many packages he or she wishes to host on it.

A \pkg{drat} repository is essentially a network-accessible directory structure. The \pkg{drat} repository's directory must include index files (e.g., \file{PACKAGES} and \file{PACKAGES.gz}; Figure \ref{fig:dratstructure}), which have metadata describing the
packages available in the repository. The directory structure should also store files with the source code of one or more packages (e.g., \file{hurricaneexposuredata\_0.0.2.tar.gz}) and can also include operating system-specific package code (e.g., \file{hurricaneexposuredata\_0.0.2.zip}). Multiple versions of a package can be included (e.g., \file{hurricaneexposuredata\_0.0.1.tar.gz} and \file{hurricaneexposuredata\_0.0.2.tar.gz} in Figure \ref{fig:dratstructure}), allowing for archiving of old packages.

A user can create a \pkg{drat} directory with the required structure either by
hand, via functions in the \pkg{drat} package, or by copying an existing \pkg{drat} repository
(e.g., forking one on GitHub, like the original drat repository available at
\url{https://github.com/eddelbuettel/drat}). While this directory can have any name, we suggest the user name the directory \file{drat}, as this allows the easy use of default
variable names (which can of course be overridden as needed) for functions in the \pkg{drat} package, as shown in later code examples.

Second, for other users to be able to install packages from a \pkg{drat} repository, the repository must be available
online. This can be easily achieved via the https protocol using GitHub's Git Pages, which allows
GitHub users to create project webpages by posting content to an \file{gh-pages} branch of the
project's repository. (While there are now ways to publish content from a directory \file{docs/} in
the \file{master} branch, functions in the \pkg{drat} package currently only support use of the
older \file{gh-pages} publishing option). Once this \file{gh-pages} branch is pushed, the content in
that branch will be available as part of the GitHub user's Git Pages website. For example, if the
project is in the GitHub repository \file{https://github.com/username/projectname}, the content from
the \file{gh-pages} branch of that repository will be published at
\file{https://username.github.io/projectname}.

\begin{figure}
\begin{center}
\includegraphics{images/drat_repo_structure.pdf}
\end{center}
\caption{Example of the structure of the directory for a \pkg{drat}
repository for a repository with two versions (0.0.1 and 0.0.2) of
\pkg{hurricaneexposuredata}. Directories and files shown in black are
required while those in blue are optional. This example \pkg{drat} depository
has source code available for the \pkg{hurricaneexposuredata}, as well as the
binaries for Windows (the binaries for Mac OS X could also be included but are 
not shown in this example),
through the optional \file{bin} subdirectory. The top-level \file{index.html}
file can be used to customize the appearance of the webpage a user would see
at \file{https://username.github.io/drat}. Functions from the \pkg{drat}
package automate the insertion of compressed package files (\file{.tar.gz}
for source code files, \file{.tgz} for Mac OS X binaries, and \file{.zip} for
Windows binaries) within this directory structure. The \file{PACKAGES} and
\file{PACKAGES.gz} files serve as index files, with metadata about packages
available in the repository, and are created by \code{drat::insertPackage}
via a call to the R function \code{tools::write\_PACKAGES}.} 
\label{fig:dratstructure}
\end{figure}

Once this \pkg{drat} repository is created and published online through Git Pages,
any source code or binaries for R packages within the repository can be installed and updated by R users through functions in the \pkg{drat} package. An R user
can install a package from a \pkg{drat} repository by first adding the \pkg{drat} respository using the \code{addRepo} function from
\pkg{drat}, with the appropriate GitHub username, and then using \code{install.packages}, as one
would for a package on CRAN. The \pkg{drat} package documentation, including several vignettes, as well as supplementary
webpages, have more detail on this process; see \citet{drat}.

## Creating and building a data package

The next step is to create an R data package that contains the large dataset; this data package will be posted in the \pkg{drat} repository to be accessible to other R
users.  In the case study
example, this package is called \pkg{hurricaneexposuredata} and includes data on rain, wind, flood,
and tornado exposures for all eastern US counties for Atlantic-basin tropical storms between 1988
and 2015. For full details on creating R packages, including data-only packages, see the canonical
reference by
\citet{rcore:wre}; another popular reference is \citet{wickham2015r}.

If a package is hosted in a \pkg{drat} repository rather than posted to CRAN, it does not have to
pass all tests executed by \samp{R CMD check}. However, it is good practice to resolve as many
ERRORS, WARNINGS, and NOTES from CRAN checks as possible for any R package that will be shared with
other users, regardless of how it is shared. Several possibilities exist to build and check a
package so these ERRORS, WARNINGS, and NOTES can be identified and resolved. The standard approach is to
execute \samp{R CMD build} from one directory above the source directory (as discussed below),
followed by \samp{R CMD check} with the resulting tar archive (e.g., \code{hurricaneexposuredata.tar.gz}) as first argument. The optional switch
\samp{--as-cran} is recommended in order to run a wider variety of tests.  Other alternatives for checking the package are to use
the \code{check} function from the \CRANpkg{devtools} package \citep{devtools}, the \code{rcmdcheck}
function of the eponymous \CRANpkg{rcmdcheck} package \citep{rcmdcheck}, the 'Check' button in the
Build pane of the RStudio GUI, or the RStudio keyboard shortcut Ctrl-Shift-E.  For a large data
package, it is desirable to resolve all issues except the NOTE on the package size being large.

Once the code in the data package is finalized, the package can be posted in a \pkg{drat} repository
to be shared with others. Packages are inserted into a \pkg{drat} repository as source code tarballs
(e.g., \file{.tar.gz} files); if desired, package binaries for specific operating systems can also
be inserted (e.g., \file{.zip} or \file{.tgz} files), but this is not required for the application
described here. To build a package into a \file{.tar.gz} file, there are again several possible
approaches. The most convenient one may be to build the package in a temporary directory created
with the \code{tempdir} function, as this directory will be cleaned up when the current R session is
closed. If the current working directory is the package directory, the package can be built to a
temporary directory with the \code{build} function from the \pkg{devtools} package:

```{r eval = FALSE}
tmp <- tempdir()
devtools::build(path = tmp)
```

While this function call assumes that the user is currently using the
directory of the data package as the working directory, the \code{pkg} option of the \code{build} function can be used
to run this call successfully from a different directory.  If the build is successful, a
\file{.tar.gz} file will be created containing the package's source code in the directory specified
by the \code{path} option; this can be checked from R with the call \code{list.files(tmp)}.

## Adding the data package to your drat repository

Once the source code of the package has been built, the data package can be inserted into the \pkg{drat}
repository using the \code{insertPackage} function from the \pkg{drat} package. This function
identifies the package file type (e.g., \file{.tar.gz}, \file{.zip}, or \file{.tgz}), adds the
package to the appropriate spot in the \pkg{drat} directory structure (Figure
\ref{fig:dratstructure}), and adds metadata on the package to the \file{PACKAGES} and
\file{PACKAGES.gz} index files (created by the R function \code{tools::write\_PACKAGES}) in the
appropriate subdirectory. Once this updated version of the \pkg{drat} repository is pushed to
GitHub, the package will be available for other users to install from the repository.

For example, the following code can be used to add version 0.0.2 of \pkg{hurricaneexposuredata} to
the \pkg{drat} repository. This code assumes that the \file{.tar.gz} file for the package was built
into a temporary directory with a path given by the R object \code{tmp}, as would be the case if the
user built the package tarball using the code suggested in the previous subsection, and that the
user is in the same R session as when the package was built (as any temporary directories created
with \code{tempdir} are deleted when an R session is closed). Further, this code assumes that the
user has the \CRANpkg{git2r} package installed and has stored their \pkg{drat} directory within the
parent directory `~/git`; if this is not the case, the correct path to the \pkg{drat} directory
should be specified in the \code{repodir} argument of \code{insertPackage}.

```{r eval = FALSE}
pkg_path <- file.path(tmp, "hurricaneexposuredata_0.0.2.tar.gz", sep = "/")
drat::insertPackage(pkg_path, commit = TRUE)
```

As mentioned before, only material in the \file{gh-pages} branch of the GitHub repository is
published through Git Pages, so it is important that the package be inserted in that branch of the
user's \pkg{drat} repository. The \code{insertPackage} function checks out that branch of the
repository and so ensures that the package file is inserted in the correct branch. However, it is
important that the user be sure to push that specific branch to GitHub to update the online
repository. If unsure, the \code{commit} option can be left at its default value of \code{FALSE},
permitting an inspection of the repository followed by a possible manual commit.

For users who prefer working from the command line, an alternative pipeline for building the data
package and inserting it into the \pkg{drat} repository is to run, from the command line:

```
R CMD build sourcedir/
dratInsert.r pkg_1.2.3.tar.gz
```

\noindent Note that this pipeline requires having
the \CRANpkg{littler} \citep{littler} package installed, as well as the \file{dratInsert.r} helper
for that package.

If desired, operating system-specific binaries of the data package can be built with tools like
win-builder (\url{http://win-builder.r-project.org/}) and rhub (\url{https://builder.r-hub.io}) and
then inserted into the \pkg{drat} repository. However, this step is not necessary, as \samp{R CMD
check --as-cran} will be satisfied for the code package as long as a source package is
available for any suggested packages stored in repositories listed in 'Additional_repositories'
\citep{rcore:wre}.

# Setting up a smaller code package to use the data package

So far, the process described is the same one would use to create and add any R package to a
\pkg{drat} repository.  However, if a package maintainer would like to coordinate a code package
that will be submitted to CRAN with a data package posted in a \pkg{drat} repository, it is necessary
to add some infrastructure to the code package (in our example, \pkg{hurricaneexposure}, which has functions for
exploring and mapping the data in \pkg{hurricaneexposuredata}). These additions
ensure that the code package will pass CRAN checks and also appropriately load and access the
data in the data package posted in the drat repository.

## Add infrastructure to the DESCRIPTION file

First, two additions are needed and a third is suggested in the DESCRIPTION file (Figure \ref{fig:descriptionelements}) of the code package that
will be submitted to CRAN:

```{r descriptionelements, echo = FALSE, fig.align = "center", fig.cap = "Example of the elements that should be added to the DESCRIPTION file of the code package planned to be submitted to CRAN to coordinate it with a data package posted to a \\pkg{drat repository}, showing the DESCRIPTION file for \\pkg{hurricaneexposure}. Elements are: (1) added details on installing the data package in the 'Description' field (suggested but not required); (2) suggested dependency on the data package in the 'Suggests' field; and (3) reference to the drat repository in the 'Additional\\_repositories' field.", out.width = "\\textwidth"}
knitr::include_graphics("images/desc_ex.png")
```

<!-- DE: Should the png file be updated now that the package is at version 0.0.2, the license has been corrected to GPL (>= 2), and the author list expanded? -->
<!-- BA: I've now updated this figure. -->

1. We suggest the 'Description' field of the code package's \file{DESCRIPTION} file be modified to let users
   know how to install the data package and how large it is (this tip is inspired by the
   \CRANpkg{grattan} package; \citet{grattan}). Figure \ref{fig:descriptionelements} (#1) shows an
   example of this added information for the \pkg{hurricaneexposure} \file{DESCRIPTION} file. For a
   CRAN package, this 'Description' field will be posted on the package's CRAN webpage, so
   this field offers an opportunity to inform users about the data package before they install the
   CRAN package. This addition is not required, but is particularly helpful in cases where the data
   package is very large, in which case it would take up a lot of room on a user's computer and take a long time
   to install and load.

2. The 'Suggests' field for the code package must specify the suggested dependency on the data package (Figure
   \ref{fig:descriptionelements}, #2). Because the data package is in a non-mainstream repository,
   this dependency much be specified in the 'Suggests' field rather than the 'Depends' or 'Imports'
   field if the code package is to be submitted to CRAN. The 'Suggests' field allows version requirements, so if the code package requires either a minimum version or an exact version of the data package in the
   \pkg{drat} repository, this requirement can be included in this field.

3. The 'Additional_repositories' field of the code package must give the address of the \pkg{drat} repository that
   stores the data package (Figure \ref{fig:descriptionelements}, #3). This field is necessary if a
   package depends on a package in a non-mainstream repository. Repositories listed here are
   checked by CRAN to confirm their availability \citep{rcore:wre}, but packages from these
   repositories are not installed prior to CRAN checks. This repository address should be listed
   using \file{https:} rather than \file{http:}.

## Customize behavior when the package is loaded

When a package is installed, any packages listed as 'Imports' or 'Depends' in the package
\file{DESCRIPTION} file are guaranteed to be previously installed \citep{rcore:wre}; the same is not
true for packages in 'Suggests'. It is therefore important that the maintainer of any package that
suggests a data package from a \pkg{drat} repository take steps to ensure that the package does not
fail if it is installed without the data package being previously installed. Such steps include
adding code that will be run when the package is loaded (described in this subsection), as well as
ensuring that code in all functions, examples, vignettes, and tests be conditional on whether the data package is installed if the code requires data from
the data package (described in later
subsections).

First, the code package should have code to check whether the data package is installed when the
code package is loaded. This can be achieved through load hooks (\code{.onLoad} and
\code{.onAttach}), saved to a file named, for example, \file{zzz.R} in the \file{R} directory of the code
package (the name \file{zzz.R} dates back to a time when R required this; now any file name
can be chosen). For a concrete example, such a \file{zzz.R} file in the code package might look like
(numbers in comments of this code are used within specific comments later in this section):

\begin{verbatim}
.pkgenv <- new.env(parent=emptyenv())                                             #1

.onLoad  <- function(libname, pkgname) {                                          #2
    has_data <- requireNamespace("hurricaneexposuredata", quietly = TRUE)         #3
    .pkgenv[["has_data"]] <- has_data                                             #4
}

.onAttach <- function(libname, pkgname) {                                         #5
    if (!.pkgenv$has_data) {                                                      #6 
        msg <- paste("To use this package, you must install the",
                     "hurricaneexposuredata package. To install that ",
                     "package, run `install.packages('hurricaneexposuredata',",
                     "repos='https://geanders.github.io/drat/', type='source')`.",
                     "See the `hurricaneexposure` vignette for more details.")
        msg <- paste(strwrap(msg), collapse="\n")
        packageStartupMessage(msg)
    }
}

hasData <- function(has_data = .pkgenv$has_data) {                                #7
    if (!has_data) {
        msg <- paste("To use this function, you must have the",
                     "`hurricaneexposuredata` package installed. See the",
                     "`hurricaneexposure` package vignette for more details.")
        msg <- paste(strwrap(msg), collapse="\n")
        stop(msg)
    }
}

\end{verbatim}

First, this \file{zzz.R} file creates an environment called \code{.pkgenv} (#1 in example
code). This environment will be used to pass a Boolean variable of whether the data package is
available (#4) to other code in the package.

Next, this \file{zzz.R} file defines two functions called \code{.onLoad} and \code{.onAttach}. Both
functions have arguments \code{libname} and \code{pkgname} (#2, #5). The \code{.onLoad()} function,
triggered when code from the package is first accessed, should use \code{requireNamespace} to test
if the user has the data package available (#3). This value is stored in the package environment
(#4). When the package is loaded and attached to the search path, the \code{.onAttach()} function is called
(#5). This function references the stored Boolean and uses this stored value to print a start-up
message (using the \code{packageStartupMessage} function, whose output can be suppressed via
\code{suppressPackageStartupMessages}) for users who lack the data package (#6).

Finally, any functions in the package that use data from the \pkg{drat} data package should check
that the data package is available before running the rest of the code in the function. If the data
package is not available, the function should stop with a useful error message. One way to achieve this is to define a function in an R script file in the package, like the \code{hasData} function defined in the example code above, that checks for availability of the data
package and errors with a helpful message if that package is not installed (#7). This function should then be added within all of
the package's functions that require data from the data package.

## Condition code in the code package on availability of the data package

Next, it is important to ensure that code in the vignettes, examples, and tests of the package will
run without error on CRAN, even without the data package installed. When a package is submitted to
CRAN, the user first creates a tarball of the package source on his or her own computer. In this
local build, the vignette is rendered and the resulting PDF or HTML file is stored within the
\file{inst/doc} directory of the source code \citep{rcore:wre}. This is the version of the rendered
vignette that is available to users when they install the package, and so the vignette is rendered
using the packages and other resources available on the package maintainer's computer during the
package build. However, CRAN runs initial checks on a package submission, and continues to run
regular checks on posted packages, which include testing any executable code within the package
vignette or any examples or tests in the source code that are not explicitly marked to not run on CRAN (e.g.,
with \verb|\donttest{}| within example code). CRAN does not install suggested packages from
non-mainstream repositories before doing these checks. Therefore, code that requires data from a
data package in a \pkg{drat} repository would cause errors in these tests unless it is conditioned
to only run when the data package is available, as is recommended for any example or test code that
uses suggested packages \citep{rcore:wre}.

Therefore, if the code package contains a vignette, the vignette must be coded so that its code  will run without an error on systems that do not have the data package installed.  This can be
done by adding a code chunk to the start of the vignette. This code chunk will check if the data
package is installed on the system. If it is, the vignette will be rendered as usual, and so it will
be rendered correctly on the package maintainer's computer when the package is built for CRAN
submission. However, if the optional data package is not installed, a message about installing the
data package from the \pkg{drat} repository will be printed in the vignette, and all the following
code chunks will be set to not be evaluated using the \code{opts\_chunk} function from \CRANpkg{knitr} \citep{knitr}.

The following code is an example of the code chunk that was added to the beginning of the vignette
in the \pkg{hurricaneexposure} package, for which all code examples require the
\pkg{hurricaneexposuredata} package:

\begin{verbatim}
```{r echo = FALSE, message = FALSE}`r ''`
hasData <- requireNamespace("hurricaneexposuredata", quietly = TRUE)                #1
if (!hasData) {                                                                     #2
    knitr::opts_chunk$set(eval = FALSE)                                             #3
    msg <- paste("Note: Examples in this vignette require that the", 
                 "`hurricaneexposuredata` package be installed. The system",
                 "currently running this vignette does not have that package",
                 "installed, so code examples will not be evaluated.")
    msg <- paste(strwrap(msg), collapse="\n")
    message(msg)                                                                    #4
}
```
\end{verbatim}

In this code, the function \code{requireNamespace} is used to check if \pkg{hurricaneexposuredata}
is installed on the system (#1). It is necessary to run this function in the
vignette with the \code{quietly = TRUE} option; otherwise, this call will cause an error if
\pkg{hurricaneexposuredata} is unavailable. If \pkg{hurricaneexposuredata} is not available, the
result of this call is `FALSE`, in which case (#2) the chunk option `eval` is set to `FALSE` for all
following chunks in the vignette (#3) and a message is printed in the vignette explaining why code
chunks are not evaluated (#4).

Similarly, the code for examples in help files of the CRAN package should be adjusted so they only
run if the data package is available. It may also be helpful to users to include a commented message
on why the example is wrapped in a conditional statement. For example, for a function
that requires the data package the \verb|\examples{}| field
(or \code{@examples} tag if \CRANpkg{roxygen2} is used) might look like:

\begin{verbatim}
\examples{
    # Ensure that data package is available before running the example.
    # If it is not, see the `hurricaneexposure` package vignette for details
    # on installing the required data package.
    if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
        map_counties("Beryl-1988", metric = "wind")
    }
}
\end{verbatim}

As alternatives, the code in the example could also either check the \code{.pkgenv[["has\_data"]]}
object created by code in the \file{zzz.R} file in the conditional statement, or the package
maintainer could create a helper function to in the \file{zzz.R} file to use for conditional running
of examples in the example code. However, the use of the \code{requireNamespace} call, as shown in
the above code example, may be the most transparent for package users to understand when working
through package examples. Code in package tests can similarly be
conditioned to only run if and only if the data package is available.

# Maintaining a suite of \pkg{drat} and CRAN R packages

Once a suite of coordinated CRAN and \pkg{drat} packages have been created, there are a few considerations a
package maintainer should keep in mind for maintaining the packages.

First, unlike a single package that combines data and code, a suite of packages
will require thoughtful coordination between versions of packages in the suite, especially as the packages
are updated. The 'Suggests' field allows the package maintainer to specify an exact version of the
data package required by a code package (e.g., using \samp{(== 0.0.1)} with the data package name in the `Suggests` field) or a minimum
version of the data package (e.g., \samp{(>= 0.0.1)}). If the data package is expected to only change
infrequently, the maintainer may want to use an exact version dependency in 'Suggests' and plan to
submit an updated version of the code package to CRAN any time the data package is updated. However,
CRAN policy recommends that package versions not be submitted more than once every one to two
months. Therefore, if the data package will be updated more frequently, it may make more sense to
use a minimum version dependency. However, the maintainer should be aware that, in this case,
package users may find that any changes in the structure of the data in the data package that
breaks functions in older versions of the code package may cause errors, without notifying
the user that the error can be resolved by updating the code package.

Second, while the maximum size of a GitHub repository is much larger than the recommended maximum
size of an R package, there are still limits on GitHub repository size. If a package maintainer uses
a \pkg{drat} repository to store multiple large data packages, with multiple versions of each, there
may be some cases where the repository approaches or exceeds the maximum allowable GitHub repository
size. In this case, the package maintainer may want to consider, when inserting new versions of the data package into the \pkg{drat} repository, changing the \code{action} option in
the \code{insertPackage} function to free repository space by removing older versions of the package or
packages. This, however, reduces reproducibility of research using
the data package, as older versions would no longer be available through an archive.

Further, while the CRAN maintainers regularly run code from examples and vignettes within packages
posted to CRAN, they do not install any suggested packages from non-mainstream
repositories before doing this. Since
much of the code in the examples and vignette are not run if the data package is not available under the approach we suggest,
regular CRAN checks will provide less code coverage than is typical for a CRAN package. Package
maintainers should keep this in mind, and they may want to consider alternatives for ensuring
regular testing of all examples in the code package. One option may be through use of a regularly
scheduled Cron Job through Travis CI to check the latest stable version of the code package.

It is important to note that, under the suggested approach, proper versioning of any data packages
hosted in a \pkg{drat} repository is entirely the responsibility of the owner of that repository. By
contrast, R users who install packages from CRAN can be confident that a version number of a package
is tied to a unique version of the source code. While versioning of R data packages can improve
research reproducibility \citep{gentleman2004bioconductor}, if the owner of the \pkg{drat}
repository is not vigilant about changing the version number for every change in the source code of
packages posted in the repository, the advantages of packaging the data in terms of facilitating
reproducible research are lost. Similarly, the repository owner is solely responsible for archiving
older versions of the package, unlike a CRAN package, for which archiving is typically ensured by
CRAN. In particular, for very large packages that are updated often, the size limitations of a
GitHub repository may force a repository owner to remove older versions from the archive.

# Conclusion

R packages offer the chance to distribute large datasets while also providing functions for
exploring and working with that data. However, data packages often exceed the suggested size of CRAN
packages, which is a challenge for package maintainers who would like to share their code through
this central and popular repository. Here, we suggest an approach in which the maintainer creates a
smaller code package with the code to interact with the data, which can be submitted to CRAN, and a
separate data package, which can be hosted by the package maintainer through a personal \pkg{drat}
repository. Although \pkg{drat} repositories are not mainstream, and so cannot be listed with an
'Imports' or 'Depends' dependency for a package submitted to CRAN, we suggest a way of including
the data package as a suggested package and incorporating conditional code in the executable code
within vignettes, examples, and tests, as well as conditioning functions in the code package to
check for the availability of the data package. This approach may prove useful for a number of R package maintainers, especially
with the growing trend to the sharing and use of open data in many of the fields in which R is
popular.

# Acknowledgements

We thank Martin Morgan and Roger Peng for comments and suggestion on an earlier draft of this
manuscript. This work was supported in part by a grant from the National Institute of Environmental
Health Sciences (R00ES022631).

\bibliography{anderson-eddelbuettel}

### 2017-07-13 papers/use_r_slides/anderson-eddelbuettel-use_r_talk.Rmd
Add R Journal final paper draft and UseR slides

---
output:
  beamer_presentation:
    keep_tex : true
    includes: 
      in_header: anderson_header.txt
      before_body: anderson_beforebody.txt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Motivation

\vspace{-0.5cm}
\begin{center}
\includegraphics[width = \textwidth]{images/rainexamples.pdf}
\end{center} 
\vspace{-0.8cm}
\textbf{Goal:} Create a publicly-accessible package with historical data (1988--present) on county-level exposure to Atlantic basin tropical storms in the United States (with exposure based on distance to storm track, rain, wind, floods, or tornadoes).

## Motivation

\begin{columns}

\begin{column}{0.45\textwidth}
\begin{block}{``Data" package}
\includegraphics[width = \textwidth]{images/hurrtracks.png} \\
\texttt{hurricaneexposuredata}\\
Hosted with \texttt{drat}. 
\end{block}
\end{column}

\begin{column}{0.45\textwidth}
\begin{block}{``Code" package}
\includegraphics[width = \textwidth]{images/example_hurricane_map.png} \\
\texttt{hurricaneexposure}\\
Available through CRAN. 
\end{block}
\end{column}

\end{columns}

\pause
\bigskip

\begin{block}{Challenges}
\begin{enumerate}
\item \alert<3>{Host a package with \texttt{drat}.} 
\item Connect a CRAN package with a data package hosted with \texttt{drat}.
\end{enumerate}
\end{block}

## Hosting a package with `drat`

\begin{columns}

\begin{column}{0.49\textwidth}

\begin{block}{R repositories and \texttt{drat}}
An R package repository is a network-accessible directory structure, and \texttt{drat} helps you build and maintain your own.
\end{block}

\bigskip

\small

\begin{block}{Extensive resources}
\vspace{-0.3cm}
\begin{itemize}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageUsers.html}{Drat Basics for Package Users}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageAuthors.html}{Drat Basics for Package Authors}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratFAQ.html}{Drat FAQs}
\end{itemize}
\end{block}

\end{column}

\begin{column}{0.49\textwidth}
\includegraphics[width = \textwidth]{images/drat_repo_structure.pdf}
\end{column}

\end{columns}

## Packages hosted on `drat` repositories

### For package users, how is a `drat` package different?

- "Not at all", really.
- Tell R about the additional repository via
    - repos argument directly in `install.packages()` et al.,
    - setting another `repos` field in `options()`, or
    - using `drat` helper functions `addRepo()` or `drat:::add()`
- Full details about `repos` in `help(download.packages)`

### For package maintainers, how is a `drat` package different?

- Need to create a personal repository and make it available to users
- Can use functions in the `drat` package to manage this repository

## Coding a CRAN package to use a `drat` package

### How is GitHub involved?

- Hosting on GitHub particularly easy because 
    - Every GitHub repo has an (optional) web presence
    - A GitHub username plus `drat` is a unique URL
- So if we assume `drat` as the (GitHub) repository name, then we can reference it with just the username:  `addRepo("geanders")`

## Coding a CRAN package to use a `drat` package

\begin{columns}

\begin{column}{0.45\textwidth}
\begin{block}{``Data" package}
\includegraphics[width = \textwidth]{images/hurrtracks.png} \\
\textit{`hurricaneexposuredata`}\\
Host with \texttt{drat}. 
\end{block}
\end{column}

\begin{column}{0.45\textwidth}
\begin{block}{``Code" package}
\includegraphics[width = \textwidth]{images/example_hurricane_map.png} \\
\textit{`hurricaneexposure`}\\
Available through CRAN. 
\end{block}
\end{column}

\end{columns}

\bigskip

\begin{block}{Challenges}
\begin{enumerate}
\item Host a package with \texttt{drat}. 
\item \alert{Connect a CRAN package with a data package hosted with \texttt{drat}.}
\end{enumerate}
\end{block}

## Technical support experience, non-R

\begin{quotation}
``Unfortunately it is out of our area of expertise to do deep level troubleshooting on an individual's computer to find the cause of the issue. In this case I would recommend using a colleague's computer and see if that is successful in creating a submission."\\[0.5ex]
-- Example of recent non-R technical support 
\end{quotation}

## Technical support experience, R

\begin{center}
\begin{enumerate}
\item \begin{minipage}[t]{\linewidth}
          \adjustbox{valign=t}{\includegraphics[width = 0.9\textwidth]{images/github_search.png}}
          \end{minipage} \medskip \pause

\item \begin{minipage}[t]{\linewidth}
          \adjustbox{valign=t}{\includegraphics[width = 0.9\textwidth]{images/the_initial_tweet.png}}
          \end{minipage} \medskip \pause

\item \begin{minipage}[t]{\linewidth}
          \adjustbox{valign=t}{\includegraphics[width = 0.9\textwidth]{images/r_journal_title.png}}
          \end{minipage}

\end{enumerate}
\end{center}

## Coding a CRAN package to use a `drat` package

### How hard is it to code a CRAN package to "depend" on a `drat` package?

- Not very
- The `DESCRIPTION` file has an `Additional_repositories` field
     - Optional packages can reside on `Additional_repositories`
     - Mandatory packages (`Imports:`, `Depends:`, `LinkingTo:`) cannot
     - But `Suggests:` can point there -- feature we use
- Repos created with `drat` are a good use case for `Additional_repositories`
- \alert{But} also ensure you test for presence of optional package

## Coding a CRAN package to use a `drat` package

### How hard is it to code a CRAN package to "depend" on a `drat` package?

- Need to add code in the CRAN package to do three jobs: 
    + Add infrastructure in the DESCRIPTION file
    + Customize behavior when the package is loaded 
    + Condition code based on availability of `drat` package (functions, vignette, examples)

## Coding a CRAN package to use a `drat` data package

\begin{block}{Add infrastructure in the DESCRIPTION file}
\begin{itemize}
\item Include the name of the data package in the ``Suggests" field.
\item Give the address of the \texttt{drat} repository with the data package in the ``Additional\_repositories" field.
\end{itemize}
\end{block}

\bigskip

For example, the `hurricaneexposure` DESCRIPTION file includes: 
\vspace{-1cm}

```
Suggests: hurricaneexposuredata (>= 0.0.2)
Additional_repositories: https://geanders.github.io/drat
```

## Coding a CRAN package to use a `drat` data package

\begin{block}{Customize behavior when the package is loaded }
\begin{itemize}
\item These steps can be included in ``zzz.R" (or similar) file, by writing \texttt{.onLoad} and \texttt{.onAttach} functions.
\item Check if the user has the data package and create an environment that stores the result. 
\item If the user does not have the data package, give a start-up message telling the user how to install it. 
\end{itemize}
\end{block}

## Coding a CRAN package to use a `drat` data package

\begin{block}{Condition code based on availability of data package}
\begin{itemize}
\item Condition code in functions that need access to the data package (e.g., helper function can be written in the package).
\item Condition code in documentation examples to only run if the data package is available.
\item Condition vignette to only evaluate code chunks if the data package is available.
\end{itemize}
\end{block}

## Coding a CRAN package to use a `drat` data package

### Example: Conditioning code in the vignette

A code chunk like the following can be included at the start of the vignette to condition the evaluation of code chunks on having the data package available:

\footnotesize

```{r eval = FALSE, echo = TRUE}
hasData <- requireNamespace("hurricaneexposuredata", quietly = TRUE)
if (!hasData){
  knitr::opts_chunk$set(eval = FALSE)
  msg <- paste("Note: Examples in this vignette require that the",
               "`hurricaneexposuredata` package be installed. The",
               "system currently running this vignette does not",
               "have that package installed, so code examples",
               "will not be evaluated.")
  msg <- paste(strwrap(msg), collapse = "\n")
  message(msg)
}
```

## Find out more

\footnotesize

\begin{block}{Article in The R Journal}
 \href{https://journal.r-project.org/archive/2017/RJ-2017-026/index.html}{Anderson and Eddelbuettel. 2017. Hosting Data Packages via drat: A Case Study with Hurricane Exposure Data. \textit{The R Journal}. 9(1):486--497.}
\end{block}

\begin{block}{Example R packages}
\begin{itemize}
\item \texttt{hurricaneexposure}: \href{www.github.com/geanders/hurricaneexposure}{geanders/hurricaneexposure}
\item \texttt{hurricaneexposuredata}: \href{www.github.com/geanders/hurricaneexposuredata}{geanders/hurricaneexposuredata}
\end{itemize}
\end{block}

\begin{block}{\texttt{drat} documentation}
\begin{itemize}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageUsers.html}{Drat Basics for Package Users}
\item \href{https://cran.r-project.org/web/packages/drat/vignettes/DratForPackageAuthors.html}{Drat Basics for Package Authors}
\end{itemize}
\end{block}

\begin{block}{Acknowledgments}
Thanks to Martin Morgan and Roger Peng for helpful suggestions. This work was supported in part by a grant from the National Institute of Environmental Health Sciences (R00ES022631)
\end{block}

### 2017-07-13 papers/final_R_journal_paper/anderson-eddelbuettel.R

@@ -1,0 +1,62 @@
+####################################################################################
+# This script includes the example R code from Anderson and Eddelbuettel, 'Hosting
+# Data Packages via `drat`: A case study with hurricane exposure data.
+####################################################################################
+
+# Build a package into a temporary directory. This code should be called when working in
+# the directory of the package to be built.
+tmp <- tempdir()
+devtools::build(path = tmp)
+
+# Insert the package into a repository named 'drat' located in '~/git'
+pkg_path <- file.path(tmp, "hurricaneexposuredata_0.0.2.tar.gz", sep = "/")
+drat::insertPackage(pkg_path, commit = TRUE)
+
+# Code to include in the code package's `zzz.R` file
+.pkgenv <- new.env(parent=emptyenv())                                             #1
+
+.onLoad  <- function(libname, pkgname) {                                          #2
+  has_data <- requireNamespace("hurricaneexposuredata", quietly = TRUE)         #3
+  .pkgenv[["has_data"]] <- has_data                                             #4
+}
+
+.onAttach <- function(libname, pkgname) {                                         #5
+  if (!.pkgenv$has_data) {                                                      #6
+    msg <- paste("To use this package, you must install the",
+                 "hurricaneexposuredata package. To install that ",
+                 "package, run `install.packages('hurricaneexposuredata',",
+                 "repos='https://geanders.github.io/drat/', type='source')`.",
+                 "See the `hurricaneexposure` vignette for more details.")
+    msg <- paste(strwrap(msg), collapse="\n")
+    packageStartupMessage(msg)
+  }
+}
+
+hasData <- function(has_data = .pkgenv$has_data) {                                #7
+  if (!has_data) {
+    msg <- paste("To use this function, you must have the",
+                 "`hurricaneexposuredata` package installed. See the",
+                 "`hurricaneexposure` package vignette for more details.")
+    msg <- paste(strwrap(msg), collapse="\n")
+    stop(msg)
+  }
+}
+
+# Code to add at the beginning of the vignette(s) of the code package to ensure vignette
+# code is conditioned to only run if the data package is available
+hasData <- requireNamespace("hurricaneexposuredata", quietly = TRUE)                #1
+if (!hasData) {                                                                     #2
+  knitr::opts_chunk$set(eval = FALSE)                                             #3
+  msg <- paste("Note: Examples in this vignette require that the",
+               "`hurricaneexposuredata` package be installed. The system",
+               "currently running this vignette does not have that package",
[…]

### 2019-03-31 R/distance_exposure.R
Update dplyr / tidyeval usage

@@ -105,30 +105,33 @@ county_distance <- function(counties, start_year, end_year, dist_limit){
 #'                                      dist_limit = 75)
 #'}
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 multi_county_distance <- function(communities, start_year, end_year,
                          dist_limit){
 
         hasData()
 
-        communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
+        communities <- dplyr::mutate(communities, fips = as.character(.data$fips))
 
         distance_df <- hurricaneexposuredata::closest_dist %>%
-                dplyr::mutate_(closest_date =
-                                       ~ lubridate::ymd(closest_date)) %>%
-                dplyr::filter_(~ fips %in% communities$fips &
-                                      lubridate::year(closest_date) >=
-                                       start_year &
-                                      lubridate::year(closest_date) <=
-                                       end_year) %>%
+                dplyr::mutate(closest_date =
+                                       lubridate::ymd(.data$closest_date)) %>%
+                dplyr::filter(.data$fips %in% communities$fips &
+                                    lubridate::year(.data$closest_date) >=
+                                     !!start_year &
+                                    lubridate::year(.data$closest_date) <=
+                                     !!end_year) %>%
                 dplyr::left_join(communities, by = "fips") %>%
-                dplyr::group_by_(~ community_name, ~ storm_id) %>%
-                dplyr::mutate_(min_dist = ~ min(storm_dist)) %>%
-                dplyr::filter_(~ min_dist <= dist_limit) %>%
-                dplyr::summarize_(closest_date = ~ dplyr::first(closest_date),
-                                  local_time = ~ dplyr::first(local_time),
-                                  closest_time_utc = ~ dplyr::first(closest_time_utc),
-                                 mean_dist = ~ mean(storm_dist),
-                                 min_dist = ~ dplyr::first(min_dist))
+                dplyr::group_by(.data$community_name, .data$storm_id) %>%
+                dplyr::mutate(min_dist = min(.data$storm_dist)) %>%
+                dplyr::filter(.data$min_dist <= !!dist_limit) %>%
+                dplyr::summarize(closest_date = dplyr::first(.data$closest_date),
+                                 local_time = dplyr::first(.data$local_time),
+                                 closest_time_utc = dplyr::first(.data$closest_time_utc),
+                                 mean_dist = mean(.data$storm_dist) #,
[…]

### 2019-03-31 R/event_exposure.R

@@ -51,6 +51,7 @@
 #'               event_type = "flood")
 #' }
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 #'
 #' @export
 county_events <- function(counties, start_year, end_year, event_type){
@@ -67,14 +68,14 @@ county_events <- function(counties, start_year, end_year, event_type){
         events <- lapply(events, function(x) subset(x, x$fips %in% counties))
 
         events <- do.call("rbind", events) %>%
-                dplyr::mutate_(flood = ~ grepl("Flood", events),
-                               tornado = ~ grepl("Tornado", events),
-                               tropical_storm = ~ grepl("Hurricane", events) |
-                                      grepl("Tropical Storm", events) |
-                                      grepl("Tropical Depression", events),
-                               wind = ~ tropical_storm | grepl("Wind", events))
+                dplyr::mutate(flood = grepl("Flood", .data$events),
+                               tornado = grepl("Tornado", .data$events),
+                               tropical_storm = grepl("Hurricane", .data$events) |
+                                      grepl("Tropical Storm", .data$events) |
+                                      grepl("Tropical Depression", .data$events),
+                               wind = .data$tropical_storm | grepl("Wind", .data$events))
         events <- events[events[ , event_type], c("fips", "storm_id")] %>%
-                dplyr::mutate_(storm_id = ~ as.character(storm_id)) %>%
+                dplyr::mutate(storm_id = as.character(.data$storm_id)) %>%
                 dplyr::left_join(hurricaneexposuredata::closest_dist,
                                  by = c("storm_id", "fips"))
         return(events)
@@ -101,7 +102,7 @@ county_events <- function(counties, start_year, end_year, event_type){
 #' # on installing the required data package.
 #' if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
 #'
-#' communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
+#' communities <- data.frame(community_name = c(rep("ny", 6), "no", "new"),
 #'                          fips = c("36005", "36047", "36061",
 #'                                   "36085", "36081", "36119",
 #'                                   "22071", "51700"))
@@ -110,13 +111,14 @@ county_events <- function(counties, start_year, end_year, event_type){
 #'                                      event_type = "flood")
 #'}
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 #'
 #' @export
 multi_county_events <- function(communities, start_year, end_year, event_type){
 
         hasData()
 
[…]

### 2019-03-31 R/map_exposure.R

@@ -15,13 +15,15 @@
 #'
 #' @examples
 #' default_map()
+#'
+#' @importFrom rlang .data
 default_map <- function(){
 
         map_data <- get_eastern_map("state")
 
         out <- ggplot2::ggplot(map_data,
-                               ggplot2::aes_(x = ~ long, y = ~ lat,
-                                             group = ~ group)) +
+                               ggplot2::aes(x = .data$long, y = .data$lat,
+                                             group = .data$group)) +
                 ggplot2::geom_polygon(fill = "lightgray", color = "white") +
                 ggmap::theme_nothing(legend = TRUE) +
                 ggplot2::coord_map()
@@ -36,6 +38,7 @@ default_map <- function(){
 #'    \code{ggplot2}, filtered to states in the eastern half of the United States.
 #'
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 get_eastern_map <- function(map  = "county"){
 
         eastern_states <- c("alabama", "arkansas", "connecticut", "delaware",
@@ -49,18 +52,18 @@ get_eastern_map <- function(map  = "county"){
                             "west virginia", "wisconsin")
 
         map_data <- ggplot2::map_data(map = map) %>%
-                dplyr::filter_(~ region %in% eastern_states)
+                dplyr::filter(.data$region %in% eastern_states)
 
         if(map == "county"){
                 county.fips <- maps::county.fips %>%
-                        dplyr::mutate_(polyname = ~ as.character(polyname)) %>%
-                        dplyr::mutate_(polyname = ~ stringr::str_replace(polyname,
+                        dplyr::mutate(polyname = as.character(.data$polyname)) %>%
+                        dplyr::mutate(polyname = stringr::str_replace(.data$polyname,
                                                                          ":.+", ""))
                 map_data <- map_data %>%
-                        tidyr::unite_(col = "polyname", from = c("region", "subregion"),
+                        tidyr::unite(col = "polyname", from = c("region", "subregion"),
                                       sep = ",") %>%
                         dplyr::left_join(county.fips, by = "polyname") %>%
-                        dplyr::mutate_(fips = ~ stringr::str_pad(fips, 5,
+                        dplyr::mutate(fips = stringr::str_pad(.data$fips, 5,
                                                                  side = "left", pad = "0"))
         }
 
@@ -107,6 +110,7 @@ get_eastern_map <- function(map  = "county"){
 #' b
 #' }
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 #'
 #' @export
[…]

### 2019-03-31 R/rain_exposure.R

@@ -73,7 +73,6 @@
 #'             start_year = 1995, end_year = 2005,
 #'             rain_limit = 100, dist_limit = 100)
 #' }
-#' @importFrom dplyr %>%
 #'
 #' @export
 county_rain <- function(counties, start_year, end_year,
@@ -150,7 +149,7 @@ county_rain <- function(counties, start_year, end_year,
 #' # on installing the required data package.
 #' if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
 #'
-#' communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
+#' communities <- data.frame(community_name = c(rep("ny", 6), "no", "new"),
 #'                          fips = c("36005", "36047", "36061",
 #'                                   "36085", "36081", "36119",
 #'                                   "22071", "51700"))
@@ -174,6 +173,7 @@ county_rain <- function(counties, start_year, end_year,
 #' Data and Information Services Center.
 #'
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 #'
 #' @export
 multi_county_rain <- function(communities, start_year, end_year,
@@ -182,40 +182,40 @@ multi_county_rain <- function(communities, start_year, end_year,
 
         hasData()
 
-        communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
+        communities <- dplyr::mutate(communities, fips = as.character(.data$fips))
 
         dots <- stats::setNames(list(lazyeval::interp(~ lubridate::ymd(x),
                                                       x = quote(closest_date))),
                                 "closest_date")
         rain_storm_df <- hurricaneexposuredata::closest_dist %>%
-                dplyr::mutate_(.dots = dots) %>%
-                dplyr::filter_(~ fips %in% communities$fips &
-                                      lubridate::year(closest_date) >= start_year &
-                                      lubridate::year(closest_date) <= end_year) %>%
+                dplyr::mutate(.dots = !!dots) %>%
+                dplyr::filter(.data$fips %in% !!communities$fips &
+                                      lubridate::year(.data$closest_date) >= !!start_year &
+                                      lubridate::year(.data$closest_date) <= !!end_year) %>%
                 dplyr::left_join(communities, by = "fips") %>%
                 dplyr::left_join(hurricaneexposuredata::rain,
                                  by = c("storm_id", "fips")) %>%
-                dplyr::filter_(~ lag %in% days_included) %>%
[…]

### 2019-03-31 R/wind_exposure.R

@@ -95,7 +95,7 @@ county_wind <- function(counties, start_year, end_year, wind_limit,
 #' # on installing the required data package.
 #' if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
 #'
-#' communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
+#' communities <- data.frame(community_name = c(rep("ny", 6), "no", "new"),
 #'                          fips = c("36005", "36047", "36061",
 #'                                   "36085", "36081", "36119",
 #'                                   "22071", "51700"))
@@ -104,32 +104,33 @@ county_wind <- function(counties, start_year, end_year, wind_limit,
 #'                                      wind_limit = 20)
 #' }
 #' @importFrom dplyr %>%
+#' @importFrom rlang .data
 multi_county_wind <- function(communities, start_year, end_year,
                               wind_limit){
 
         hasData()
 
-        communities <- dplyr::mutate_(communities, fips = ~ as.character(fips))
+        communities <- dplyr::mutate(communities, fips = as.character(.data$fips))
 
         wind_df <- hurricaneexposuredata::storm_winds %>%
-                dplyr::mutate_(year = ~ gsub("*.+-", "", storm_id)) %>%
-                dplyr::filter_(~ fips %in% communities$fips &
-                                       year >= start_year &
-                                       year <= end_year) %>%
+                dplyr::mutate(year = gsub("*.+-", "", .data$storm_id)) %>%
+                dplyr::filter(.data$fips %in% !!communities$fips &
+                                       .data$year >= !!start_year &
+                                       .data$year <= !!end_year) %>%
                 dplyr::left_join(communities, by = "fips") %>%
                 dplyr::left_join(hurricaneexposuredata::closest_dist,
                                  by = c("storm_id", "fips")) %>%
-                dplyr::group_by_(~ commun, ~ storm_id) %>%
-                dplyr::mutate_(max_wind = ~ max(vmax_sust),
-                               min_dist = ~ min(storm_dist)) %>%
-                dplyr::filter_(~ max_wind >= wind_limit) %>%
-                dplyr::summarize_(closest_date = ~ dplyr::first(closest_date),
-                                  local_time = ~ dplyr::first(local_time),
-                                  closest_time_utc = ~ dplyr::first(closest_time_utc),
-                                  mean_dist = ~ mean(storm_dist),
-                                  mean_wind = ~ mean(vmax_sust),
-                                  min_dist = ~ dplyr::first(min_dist),
[…]

### 2019-04-01 README.Rmd
Fix some typos

@@
 These wind estimates are based on the wind radii in the Extended Best Tracks dataset (you can find out more about this data in the helpfile for the `ext_tracks_wind` dataset).
-These data provide estimates of which counties were exposured during a storm to sustained winds in four categories: 0--34 knots; 34--50 knots;  50--64 knots; and 64 knots or higher.
+These data provide estimates of which counties were exposed during a storm to sustained winds in four categories: 0--34 knots; 34--50 knots;  50--64 knots; and 64 knots or higher.
 Therefore, these data provide a categorical rather than continuous estimate of county wind speeds.

### 2019-04-01 R/filter_storm_data.R

@@ -8,7 +8,7 @@
 #' @param storm A character string giving the storm ID (e.g., "Floyd-1999")
 #' @param year_range A numeric vector of length two with the starting and
 #'    ending year to subset to.
-#' @param distance_limit A numeric vector of length one giving the maxmimum
+#' @param distance_limit A numeric vector of length one giving the maximum
 #'    distance (in kilometers) to use in the filter
 #' @param rain_limit A numeric vector of length one giving the minimum
 #'    rain (in millimeters) to use in the filter

### 2019-07-12 README.Rmd
Update README

@@
 Note that a few storms satisfied the criteria to be included in the data but were unnamed (e.g., "Notnamed" in 1991, "Subtrop" in 1997, etc.).
-These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table (e.g., "Notnamed-1991"), although for some research projects you may want to consider excluding these unnamed storms from the analysis.
+These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table (e.g., "Notnamed-1991", "Td08-2016"), although for some research projects you may want to consider excluding these unnamed storms from the analysis.
 
@@
 ```{r warning = FALSE}
-communities <- data.frame(commun = c(rep("ny", 6), "no", "new"),
+communities <- data.frame(community_name = c(rep("ny", 6), "no", "new"),
                          fips = c("36005", "36047", "36061",

### 2019-07-12 README.Rmd
Use gfm for README
This replaces markdown_github, which was deprecated
Add a note about updating `hurricaneexposuredata`

@@
   md_document:
-    variant: markdown_github
+    variant: gfm
 ---
@@
 ```
+
+You will want to have the latest version of the `hurricaneexposuredata` package.
+If you have previously installed `hurricaneexposuredata`, you may need to re-run the above code if you update the `hurricaneexposure` package, to update to the latest data.
 

### 2020-02-06 README.Rmd

@@
 <!-- README.md is generated from README.Rmd. Please edit that file -->
-
-```{r eval = FALSE, echo = FALSE}
-[![Travis-CI Build Status](https://travis-ci.org/geanders/hurricaneexposure.svg?branch=master)](https://travis-ci.org/geanders/hurricaneexposure)
-```
 
@@
 - Anderson B, Yan M, Ferreri J, Crosson W, Al-Hamdan M, Schumacher A and
- Eddelbuettel D (2017). _hurricaneexposure: Explore and Map County-Level
- Hurricane Exposure in the United States_. R package version 0.0.1, <URL:
+ Eddelbuettel D (2020). _hurricaneexposure: Explore and Map County-Level
+ Hurricane Exposure in the United States_. R package version 0.1.1, <URL:
  http://CRAN.R-project.org/package=hurricaneexposure>.
@@
 - Anderson B, Schumacher A, Crosson W, Al-Hamdan M, Yan M, Ferreri J, Chen Z,
- Quiring S and Guikema S (2017). _hurricaneexposuredata: Data Characterizing
- Exposure to Hurricanes in United States Counties_. R package version 0.0.1,
+ Quiring S and Guikema S (2020). _hurricaneexposuredata: Data Characterizing
+ Exposure to Hurricanes in United States Counties_. R package version 0.1.0,
  <URL: https://github.com/geanders/hurricaneexposuredata>.
@@
 library(tidyr)
+library(stringr)
 hurr_tracks %>%
@@
         distinct() %>%
-        separate(storm_id, into = c("storm_name", "Year"), sep = "-") %>%
+        mutate(Year = str_extract(storm_id, "-[0-9].+"),
+               Year = str_remove(Year, "-"),
+               storm_name = str_remove(storm_id, "-[0-9].+")) %>% 
         group_by(Year) %>%

### 2020-02-11 README.Rmd
Try to clean up README figure files

@@
 ---
-output:
-  md_document:
-    variant: gfm
+output: github_document
 ---
@@
 In functions throughout the `hurricaneexposure` package, storms are identified based on their name and year (e.g., Hurricane Floyd in 1999 is identified as "Floyd-1999").
-Note that a few storms satisfied the criteria to be included in the data but were unnamed (e.g., "Notnamed" in 1991, "Subtrop" in 1997, etc.).
-These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table (e.g., "Notnamed-1991", "Td08-2016"), although for some research projects you may want to consider excluding these unnamed storms from the analysis.
+Note that a few storms satisfied the criteria to be included in the data but were unnamed (e.g., "AL13" in 1988).
+These storms are identified in functions in the `hurricaneexposure` package based on the identifying name listed in the above table, which is based on the first four characters of the storm's US ATCF code for the storm.
+For some research projects, however, you may want to consider excluding these unnamed storms from the analysis.
 

### 2020-02-11 README.Rmd
Update some of the documentation

@@
 
-This package allows users to explore and map data of county-level exposures to Atlantic-basin tropical storms between 1988 and 2015 for a number of storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
+This package allows users to explore and map data of county-level exposures to Atlantic-basin tropical storms between 1988 and 2018 for a number of storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
 Users can map exposures by county for a selected storm and can also identify all storms to which selected counties were exposed, based on user-specified thresholds (for example, the package allows the user to identify all storms that brought wind of 34 knots or higher to Miami-Dade County in Florida).
@@
 - `county_centers`: Location of United States county centers of population
-- `hurr_tracks`: Storm tracks for Atlantic-basin storms, 1988-2015
-- `closest_dist`: Closest distances between counties and a storm track, for Atlantic-basin storms, 1988-2015
+- `hurr_tracks`: Storm tracks for Atlantic-basin storms, 1988-2018
+- `closest_dist`: Closest distances between counties and a storm track, for Atlantic-basin storms, 1988-2018
 - `rain`: Rainfall for US counties during Atlantic basin tropical storms, 1988-2011; daily rainfall is given from five days before to three days after the storm's closest approach to the county
-- `storm_winds`: Modeled county wind speeds for Atlantic-basin storms, 1988-2015
-- `storm_events`: Listings from the [NOAA Storm Events database](https://www.ncdc.noaa.gov/stormevents/) that occurred near in time and location to tropical storms, 1988-2015.
+- `storm_winds`: Modeled county wind speeds for Atlantic-basin storms, 1988-2018
+- `storm_events`: Listings from the [NOAA Storm Events database](https://www.ncdc.noaa.gov/stormevents/) that occurred near in time and location to tropical storms, 1988-2018.
 This database changed the types of events it reported in 1996, which should be considered when using the data.
-- `ext_tracks_wind`: Estimated county wind speeds for Atlantic-basin storms, 1988-2015, based on the wind radii listed in the [Extended Best Tracks dataset](http://rammb.cira.colostate.edu/research/tropical_cyclones/tc_extended_best_track_dataset/)
+- `ext_tracks_wind`: Estimated county wind speeds for Atlantic-basin storms, 1988-2018, based on the wind radii listed in [HURDAT2](https://www.nhc.noaa.gov/data/)
 
@@
 Further, we have included a second source of estimated winds in the data available in `hurricaneexposuredata`.
-These wind estimates are based on the wind radii in the Extended Best Tracks dataset (you can find out more about this data in the helpfile for the `ext_tracks_wind` dataset).
+These wind estimates are based on the wind radii in the HURDAT2 dataset (you can find out more about this data in the helpfile for the `ext_tracks_wind` dataset).
 These data provide estimates of which counties were exposed during a storm to sustained winds in four categories: 0--34 knots; 34--50 knots;  50--64 knots; and 64 knots or higher.
@@
 You can create a wind map based on this data using the `wind_source = "ext_tracks"` argument in the `map_counties` function.
-For example, to map estimated winds during Hurricane Katrina based on the Extended Best Tracks wind radii, you can run:
+For example, to map estimated winds during Hurricane Katrina based on the HURDAT2 wind radii, you can run:
 
@@
 
-As another example, to map all tracks for storms in 2005, you can run:
+As another example, to map all tracks for storms in 2018, you can run:
 
@@
 library(tidyr)
-storms_2005 <- hurr_tracks %>%
+storms_2018 <- hurr_tracks %>%
         select(storm_id) %>%
         distinct() %>%
-        separate(storm_id, c("name", "year"), sep = "-", remove = FALSE) %>%
-        filter(year == "2005")
-map_tracks(storms = storms_2005$storm_id) 
+        mutate(year = stringr::str_extract(storm_id, "-[0-9].+")) %>%
+        filter(year == "-2018")
+map_tracks(storms = storms_2018$storm_id)  
 ```
@@
 By default, this function uses wind estimates for each county from a wind model (see the vignettes for the `stormwindmodel` package for much more detail on this modeling process).
-However, you can also pull estimates based on the wind radii from the Extended Best Tracks dataset by using the option `wind_source = "ext_tracks"`.
+However, you can also pull estimates based on the wind radii from HURDAT2 by using the option `wind_source = "ext_tracks"`.
 See the help file for the `ext_tracks_wind` dataset that comes with `hurricaneexposuredata` for more details on this data source and how the county-specific exposures were determined from the data source for use in the `county_wind` function.
@@
 Further, the `dur_sust` value when using this data source is based on number of minutes with winds at or above 34 knots (rather than the 20 m / s value used for durations for the modeled wind data).
-Gust durations can not be determined using this Extended Best Tracks dataset.
+Gust durations can not be determined using the HURDAT2 wind radii.
 
-The Extended Best Tracks wind radii dataset is more tied to observations during a specific storm than the modeled data.
+The HURDAT2 wind radii are to observations during a specific storm than the modeled data.
 In certain cases, especially storms in extratropical transition, this dataset might be preferable to the modeled wind speeds, even though it provides less continuous estimates.
 This source of wind data can be specified using the `wind_source` option in the `county_wind` function.
-For example, to generate a list of all storms in Orleans Parish with maximum sustained winds of 34 knots or more based on the Extended Best Tracks wind radii, you can run:
+For example, to generate a list of all storms in Orleans Parish with maximum sustained winds of 34 knots or more based on the HURDAT2 wind radii, you can run:
 

### 2020-02-11 R/event_exposure.R

@@ -9,7 +9,7 @@
 #'
 #' @param event_type Character string with the type of event to use to identify
 #'    county exposures. Options include \code{"flood"}, \code{"tornado"},
-#'    \code{"wind"}, and \code{"tropical storm"}.
+#'    \code{"wind"}, and \code{"tropical_storm"}.
 #' @inheritParams county_distance
 #'
 #' @note Of the event types that this function can pull, only "tornado" and
@@ -18,14 +18,16 @@
 #'    event type, you should not use a start year prior to 1996, as events of
 #'    other types were not recorded in the database before 1996.
 #'
-#' #' @return Returns a dataframe with a row for each county-storm pair and with
+#' @return Returns a dataframe with a row for each county-storm pair and with
 #'    columns for:
 #'    \itemize{
 #'      \item{\code{fips}: }{County's 5-digit Federal Information Processing Standard
 #'                  (FIPS) code}
 #'      \item{\code{storm_id}: }{Unique storm identifier with the storm name and year,
-#'                  separated by a hyphen(e.g., "Alberto-1988",
+#'                  separated by a hyphen (e.g., "Alberto-1988",
 #'                  "Katrina-2005")}
+#'     \item{\code{usa_atcf_id}: }{United States Automated Tropical Cyclone Forecasting
+#'                             storm identifier}
 #'     \item{\code{closest_time_utc}: }{Time, in UTC, of the closest approach of the
 #'                             storm to the county's population mean center,
 #'                             based on storm tracks linearly interpolated to
@@ -47,8 +49,12 @@
 #' if (requireNamespace("hurricaneexposuredata", quietly = TRUE)) {
 #'
 #' county_events(counties = c("37031", "37053", "37055"),
-#'               start_year = 1996, end_year = 2015,
+#'               start_year = 1996, end_year = 2018,
 #'               event_type = "flood")
+#'
+#' county_events(counties = c("37055"),
+#'               start_year = 1996, end_year = 2018,
+#'               event_type = "tropical_storm")
 #' }
 #' @importFrom dplyr %>%
 #' @importFrom rlang .data

### 2020-02-11 R/map_exposure.R

@@ -237,7 +237,7 @@ interp_track <- function(track, tint = 0.25){
 #' map_tracks("Allison-2001", plot_object = allison_map, plot_points = TRUE)
 #'}
 #' @importFrom dplyr %>%
-#' @importFrame rlang .data
+#' @importFrom rlang .data
 #'
 #' @export
 map_rain_exposure <- function(storm, rain_limit, dist_limit,
@@ -472,6 +472,12 @@ map_wind_exposure <- function(storm, wind_var = "vmax_sust", wind_limit,
 #' map_event_exposure(storm_id = "Floyd-1999", event_type = "tornado")
 #' map_event_exposure(storm_id = "Floyd-1999", event_type = "wind")
 #' map_event_exposure(storm_id = "Floyd-1999", event_type = "tropical_storm")
+#'
+#' map_event_exposure(storm_id = "Florence-2018", event_type = "flood")
+#' map_event_exposure(storm_id = "Florence-2018", event_type = "tropical_storm")
+#'
+#' map_event_exposure(storm_id = "Michael-2018", event_type = "wind")
+#' map_event_exposure(storm_id = "Michael-2018", event_type = "tropical_storm")
 #' }
 #' @export
 #'
@@ -564,6 +570,10 @@ map_event_exposure <- function(storm_id, event_type, add_track = TRUE){
 #' map_counties("Katrina-2005", metric = "wind", wind_var = "vmax_gust")
 #' map_counties("Katrina-2005", metric = "wind", wind_var = "sust_dur")
 #' map_counties("Katrina-2005", metric = "wind", wind_source = "ext_tracks")
+#'
+#' #' map_counties("Michael-2018", metric = "wind")
+#' map_counties("Michael-2018", metric = "wind", wind_var = "vmax_gust")
+#' map_counties("Michael-2018", metric = "wind", wind_source = "ext_tracks")
 #'}
 #' @export
 #'

### 2020-02-11 R/wind_exposure.R

@@ -25,6 +25,20 @@
 #'      \item{\code{sust_dur}: }{Minutes sustained wind speed was 20 m / s or higher}
 #'      \item{\code{gust_dur}: }{Minutes gust wind speed was 20 m / s or higher (only
 #'          returned if the modeled winds are requested using \code{wind_source})}
+#'      \item{\code{usa_atcf_id}: }{United States Automated Tropical Cyclone Forecasting
+#'                             storm identifier}
+#'      \item{\code{closest_time_utc}: }{Time, in UTC, of the closest approach of the
+#'                             storm to the county's population mean center,
+#'                             based on storm tracks linearly interpolated to
+#'                             15-minute increments.}
+#'      \item{\code{storm_dist}: }{Minimum distance (in kilometers) between the storm's
+#'                     track and the county's population mean center.}
+#'      \item{\code{local_time}: }{Local time of the closest approach of the storm to the
+#'                        county's population mean center, based on storm tracks
+#'                        linearly interpolated to 15-minute increments.}
+#'      \item{\code{closest_date}: }{Date (based on local time) of the closest
+#'                          approach of the storm to the county's population
+#'                          mean center.}
 #'    }
 #'
 #' @details For more information on how wind speeds are modeled in this data,

### 2020-02-12 README.Rmd
Fix some spelling typos

@@
 
-The `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds to define rain-based exposure (cumulative rainfall and distance from the storm's track, as explained for the binary rain maps shown in an earlier section) and creates a list of all storms that met these thesholds for the counties.
+The `county_rain` function takes a list of county FIPS codes, bounds on the starting and ending years of the analysis, and thresholds to define rain-based exposure (cumulative rainfall and distance from the storm's track, as explained for the binary rain maps shown in an earlier section) and creates a list of all storms that met these thresholds for the counties.
 For example, to get a dataset of all the storms to which Orleans Parish (FIPS 22071), and Newport News, Virginia (FIPS 51700), were exposed between 1995 and 2005, where "exposed" means that the storm passed within 100 kilometers of the county center and the rainfall over a three-day window of the date of closest approach was 100 millimeters or more, you can run:

### 2020-02-12 R/map_exposure.R

@@ -167,11 +167,11 @@ map_tracks <- function(storms, plot_object = NULL, padding = 2, plot_points = FA
 #' Interpolate a storm track
 #'
 #' This function takes a wider-spaced storm track (e.g., every 6 hours) and
-#' interpolates to every 15 minutes. To do this, it
-#' fits GLMs of latitude and longitude regressed on natural cubic splines of
-#' date-time, and then predicts these splines to new intervals. These
-#' splines use degrees of freedom equal to the number of original observations
-#' divided by two.
+#' interpolates to every 15 minutes. To do this, it uses natural cubic
+#' spline interpolation using the `spline` function from the `stats` package.
+#' The track is only interpolated if there are three or more observations
+#' on the central location of the storm (this is almost always the case
+#' for storms tracked in the HURDAT2 dataset).
 #'
 #' @param track A dataframe with hurricane track data for a single storm
 #'
@@ -188,10 +188,12 @@ interp_track <- function(track){
 
                 interp_lat <- stats::spline(x = track$date,
                                             y = track$latitude,
-                                            xout = interp_date)$y
+                                            xout = interp_date,
+                                            method = "natural")$y
                 interp_lon <- stats::spline(x = track$date,
                                             y = track$longitude,
-                                            xout = interp_date)$y
+                                            xout = interp_date,
+                                            method = "natural")$y
 
                 full_track <- data.frame(storm_id = track$storm_id[1],
                                         date = interp_date,

### 2020-06-01 README.Rmd
practice edit

@@
 
-This package allows users to explore and map data of county-level exposures to Atlantic-basin tropical storms between 1988 and 2018 for a number of storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
+This package allows users to explore and map data of county-level exposures to Atlantic-basin tropical storms between 1988 and 2018 for several storm hazards (e.g., wind, rain, flood events, distance from the storm track) for counties in the eastern half of the United States.
 Users can map exposures by county for a selected storm and can also identify all storms to which selected counties were exposed, based on user-specified thresholds (for example, the package allows the user to identify all storms that brought wind of 34 knots or higher to Miami-Dade County in Florida).

### 2020-06-06 README.Rmd
map_wind_exposure edit
changed exposed to a certain winds during a storm to exposed to a certain wind speed during a storm.
add_track = FALSE
In the example code to customize the display of the map_track I added add_track = FALSE in the code above to illustrate this example, although the code doesn't appear to do anything different.

@@
 
-You can use the `map_wind_exposure` function to map counties exposed to a certain winds during a storm.
+You can use the `map_wind_exposure` function to map counties exposed to a certain wind speed during a storm.
 For example, to identify all counties exposed to sustained winds of 20 m / s or higher during Hurricane Katrina, you can run:
@@
 ```{r fig.width = 7, fig.height = 4}
-floyd_map <- map_event_exposure(storm = "Floyd-1999", event_type = "flood")
+floyd_map <- map_event_exposure(storm = "Floyd-1999", event_type = "flood", add_track = FALSE)
 map_tracks(storms = "Floyd-1999", plot_object = floyd_map, plot_points = TRUE, 

### 2020-06-08 README.Rmd
certain winds to certain wind speed close parentheses

@@
 
-In addition to giving the names and closest dates of each storm for each county (`closest_date`-- note, this is given based on local time for the county; see documentation for the [`countytimezones` package](https://cran.r-project.org/web/packages/countytimezones/index.html) to find out more about how local time was calculated from UTC for each county), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).
+In addition to giving the names and closest dates of each storm for each county (`closest_date`) -- note, this is given based on local time for the county; see documentation for the [`countytimezones` package](https://cran.r-project.org/web/packages/countytimezones/index.html) to find out more about how local time was calculated from UTC for each county), this function also gives you the distance between the county and the storm's track at the time when the storm was closest to the county's population weighted center (`storm_dist`, in kilometers) and the total precipitation over the included days (`tot_precip`).
 The returned dataframe also gives the time of closest approach in both local time (`local_time`) and UTC (`closest_time_utc`), based on 15-minute intervals along the storm's track.

### 2020-06-08 README.Rmd
Miami-Dade FIPS code
Added FIPS 12086 to the example in the paragraph before the r chunk that uses it.

--- 2020-06-06 README.Rmd
@@
 Similar functions are available in the package to create listings of county exposures to storms based on wind, distance to the storm track, and events listings from the NOAA Storm Events Database (flood and tornado events).
-For example, to get a listing of all storms between 1988 and 2015 for which Miami-Dade county (FIPS: ) has experienced sustained winds of 34 knots or more (17.5 m / s), you can run:
+For example, to get a listing of all storms between 1988 and 2015 for which Miami-Dade county (FIPS 12086) has experienced sustained winds of 34 knots or more (17.5 m / s), you can run:
 

### 2020-06-15 README.Rmd
The HURDAT2 wind radii are to observations during a specific storm than the modeled data.
I added question marks on either side of this sentence because it doesn't seem like it is complete but I'm not sure what it is supposed to read as.
more closely aligned in readme
removing a colon
removing a colon in one of the FIPS examples for consistency with the others

@@
 
-The HURDAT2 wind radii are to observations during a specific storm than the modeled data.
+The HURDAT2 wind radii are more closely aligned to observations during a specific storm than the modeled data.
 In certain cases, especially storms in extratropical transition, this dataset might be preferable to the modeled wind speeds, even though it provides less continuous estimates.
@@
 Finally, you can use the `county_events` function to get a listing of all storms for which a county had a certain type of NOAA Storm Events listing.
-For example, to get a list of all storms for which Norfolk, Virginia, (FIPS: 51710) had a flood event listing, you can run:
+For example, to get a list of all storms for which Norfolk, Virginia, (FIPS 51710) had a flood event listing, you can run:
 

### 2020-07-06 R/event_exposure.R
period added to out_dir in events_exposure

@@ -167,7 +167,7 @@ multi_county_events <- function(communities, start_year, end_year, event_type){
 #' @inheritParams county_distance
 #' @inheritParams rain_exposure
 #' @param out_dir Character string with the filepath to the directory where the
-#'   data will be saved
+#'   data will be saved.
 #' @param out_type Character string with the type of file to save to. Options are
 #'   "csv" for a comma-separated file (default) and "rds" for an R object file.
 #'

### 2020-07-06 R/filter_storm_data.R
added periods to values of filter_storm_data

@@ -5,15 +5,15 @@
 #'
 #' @param counties A character vector listing all 5-digit county FIPS codes
 #'    for a subset of counties.
-#' @param storm A character string giving the storm ID (e.g., "Floyd-1999")
+#' @param storm A character string giving the storm ID (e.g., "Floyd-1999").
 #' @param year_range A numeric vector of length two with the starting and
 #'    ending year to subset to.
 #' @param distance_limit A numeric vector of length one giving the maximum
-#'    distance (in kilometers) to use in the filter
+#'    distance (in kilometers) to use in the filter.
 #' @param rain_limit A numeric vector of length one giving the minimum
-#'    rain (in millimeters) to use in the filter
+#'    rain (in millimeters) to use in the filter.
 #' @param include_rain A logical specifying whether to pull in rain data to use
-#'    in the filter or give as output (default is FALSE)
+#'    in the filter or give as output (default is FALSE).
 #' @param days_included A numeric vector giving the lag numbers for days to
 #'    include when calculating the total rain over the storm period (e.g.,
 #'    \code{c(-1, 0, 1, 2)} would calculate the rain from the day before the

### 2020-07-06 R/map_exposure.R
period added to storm arg in map_counties

@@ -537,7 +537,7 @@ map_event_exposure <- function(storm_id, event_type, add_track = TRUE){
 #' Map counties
 #'
 #' @param storm Character string giving the name of the storm to plot (e.g.,
-#'    "Floyd-1999")
+#'    "Floyd-1999").
 #' @param metric Character string giving the metric to plot. Current options are
 #'    \code{"distance"}, \code{"wind"}, and \code{"rainfall"}. These options are used
 #'    to customize the color palette and scale of the choropleth map produced

### 2020-07-06 R/map_exposure.R
periods added to storm_id in map_event_exposure

@@ -446,7 +446,7 @@ map_wind_exposure <- function(storm, wind_var = "vmax_sust", wind_limit,
 #' \code{storm_events} dataset.
 #'
 #' @param storm_id Character vector with the storm for which to map events
-#'    (e.g., \code{"Katrina-2005"})
+#'    (e.g., \code{"Katrina-2005"}.)
 #' @inheritParams county_distance
 #' @inheritParams county_events
 #' @inheritParams map_rain_exposure

### 2020-07-06 R/rain_exposure.R
periods added to values in multi_county_rain

@@ -112,10 +112,10 @@ county_rain <- function(counties, start_year, end_year,
 #' @return Returns a dataframe with a row for each county-storm pair and with
 #'    columns for:
 #'    \itemize{
-#'      \item{\code{commun}: }{Each community's unique id}
+#'      \item{\code{commun}: }{Each community's unique id.}
 #'      \item{\code{storm_id}: }{Unique storm identifier with the storm name and year,
 #'                  separated by a hyphen(e.g., "Alberto-1988",
-#'                  "Katrina-2005")}
+#'                  "Katrina-2005").}
 #'      \item{\code{closest_date}: }{Date (based on local time) of the closest
 #'                          approach of the storm to the county's population
 #'                          mean center.}
